### 2020-02-24 scripts/community_final.Rmd
Add files via upload
main script

---
title: "Differential Host Competence of Amphibian Communities for Ranavirus"
output:
  html_document:
    df_print: paged
---

This script is an attempt to bring together the relevant information and scripts from my work up until now. An effort to organize, scrutinize, and finalize. This is formatted to act as a proposal for bringing the project to completion

```{r,load packages, include = FALSE}
library(tidyverse)
library(magrittr)
library(deSolve)
library(knitr)
library(ggridges)
library(reshape2)
```

This project can be split up into three distinct sections
 (A) Data exploration of Ranavirus prevalence in wetlands of amphibians
 (B) Metacommunity analysis of wetlands of amphibians and their propensity for parasite invasion/prevalence
 (C) Theoretical model of transmission for ranavirus in a mixed-host variable-competency community

Composition of hosts within a community plays a role in transmission potential of an introduced parasite. The competency of hosts is an important mechanism behind the relationship between community composition and parasite success. Using empirical data and a theoretical model for disease transmission, we can hypothesize the effects of community composition in a real system. The system being considered is Ranavirus in amphibian communities. Ranavirus, a dsDNA virus that mostly affects amphibians, reptiles, and fish is a known contributor to amphibian declines throughout the globe. Hosts vary in quality for transmission, and environmental transmission is known to occur in addition to contact transmission. These characteristics make this system an excellent opportunity to study the effects of community composition on disease transmission and should allow us to ask further questions about the biological realism that is relevant for studying this relationship.

Ranavirus Summary
```{r, Ranavirus Summary}
data <- read_csv("competence.csv")

#Plot of RV status over time
rv <- data %>% select(Month,WetAltID,RV.Status)
rv$Month <- factor(data$Month,levels=c("Feb","Mar","Apr","May","Jun","Jul"))
rv %>% 
  filter(RV.Status==1) %>% 
  ggplot(., aes(x=Month,
                y=WetAltID,
                group=WetAltID,
                fill=as.factor(WetAltID))) + 
  geom_density_ridges_gradient(show.legend=F) +
  labs(title="# of RV cases per site over time")

#looking at rv of species and drawing a line to divide "low" and "high" competent
rv_species <- data %>% select(vl.4:vl.42) %>% distinct()
rv_species %<>% pivot_longer(.,vl.4:vl.42) %>% filter(value>0)
rv_species %<>% add_column(log_value = log(rv_species$value))
rv_species$name <- factor(rv_species$name, levels = rv_species$name[order(rv_species$log_value)])
rv_species$actual_name <- factor(c("N. viridescens", "S. holbrookii", "A. terrestris", "A. gryllus", "P. crucifer", "H. gratiosa", "H. avivoca",
                                   "P. nigrita", "P. ornata", "G. carolinensis", "L. clamitans", "L. sphenocephalus"))
rv_species$actual_name <- factor(rv_species$actual_name, levels = rv_species$actual_name[order(rv_species$log_value)])

#natural log of values
rv_species %>% ggplot(.,aes(x=actual_name,y=log_value))+geom_point() + xlab("Species") + ylab("ln(Viral Load)") + geom_hline(yintercept = log(20)) + labs(title = "Viral Loads") + theme_classic() + theme(axis.text.x = element_text(angle = 30, hjust = 1))

```
Same summary as before but with a different measure of competency. These measures are calculated using only RV positive individuals
```{r, Ranavirus Summary 2}
data <- read_csv("competence_rv_pos.csv")

#Plot of RV status over time
rv <- data %>% select(Month,WetAltID,RV.Status)
rv$Month <- factor(data$Month,levels=c("Feb","Mar","Apr","May","Jun","Jul"))
rv %>% 
  filter(RV.Status==1) %>% 
  ggplot(., aes(x=Month,
                y=WetAltID,
                group=WetAltID,
                fill=as.factor(WetAltID))) + 
  geom_density_ridges_gradient(show.legend=F) +
  labs(title="# of RV cases per site over time")

#looking at rv of species and drawing a line to divide "low" and "high" competent
rv_species <- data %>% select(vl.9:vl.42) %>% distinct()
rv_species %<>% pivot_longer(.,vl.9:vl.42) %>% filter(value>0)
rv_species %<>% add_column(log_value = log(rv_species$value))
rv_species$name <- factor(rv_species$name, levels = rv_species$name[order(rv_species$log_value)])
rv_species$actual_name <- factor(c("N. viridescens", "S. holbrookii", "A. terrestris", "P. crucifer", "H. gratiosa", "H. avivoca",
                                   "P. nigrita", "P. ornata", "G. carolinensis", "L. clamitans", "L. sphenocephalus"))
rv_species$actual_name <- factor(rv_species$actual_name, levels = rv_species$actual_name[order(rv_species$log_value)])

#natural log of values
rv_species %>% ggplot(.,aes(x=actual_name,y=log_value))+geom_point() + xlab("Species") + ylab("ln(Viral Load)") + geom_hline(yintercept = 5) + labs(title = "Viral Loads") + theme_classic() + theme(axis.text.x = element_text(angle = 30, hjust = 1))

```

Select only site-level data for absolute abundance (species included are only those for which viral load was included)
```{r, Absolute Abundance}
site_time_cc_tot <- data %>% select(WetAltID, Month.1, AB2:AB42, cc) %>% select(-ABAmb,-ABSal) %>% distinct()

site_time_cc_tot <- mutate(site_time_cc_tot, ABHigh = (AB26 + AB42 + AB21 + AB9))
site_time_cc_tot <- mutate(site_time_cc_tot, ABLow = (AB38 + AB34 + AB41 + AB35 + AB20 + AB28 + AB24))
site_time_cc_tot <- mutate(site_time_cc_tot, total = ABHigh + ABLow)

#how to facet this to look at all sites?
site_time_cc_tot %>% filter(WetAltID==8) %>% select(Month.1, ABHigh, ABLow) %>% melt() %>%
  ggplot(., aes(y = value, x = Month.1, fill = variable)) +
  geom_bar(position = "stack", stat = "identity")  +
  labs(title = "Absolute Abundace for high and low competence at wetland 8")
```

Absolute abundance (including all species i.e. more species included in ABLow)
```{r, Absolute Abundance 2}
site_time_cc_tot <- data %>% select(WetAltID, Month.1, AB2:AB42, cc) %>% select(-ABAmb,-ABSal) %>% distinct()

site_time_cc_tot <- mutate(site_time_cc_tot, ABHigh = (AB26 + AB42 + AB21 + AB9))
site_time_cc_tot <- mutate(site_time_cc_tot, 
                           ABLow = (AB2 + AB3 + AB4 + AB5 + AB6 + AB8 + AB20 + AB24 + AB27 + AB28 + AB29 + AB31 + AB34 + AB35 + AB38 + AB39 + AB41))
site_time_cc_tot <- mutate(site_time_cc_tot, total = ABHigh + ABLow)

#how to facet this to look at all sites?
site_time_cc_tot %>% filter(WetAltID==8) %>% select(Month.1, ABHigh, ABLow) %>% melt() %>%
  ggplot(., aes(y = value, x = Month.1, fill = variable)) +
  geom_bar(position = "stack", stat = "identity") +
  labs(title = "Absolute Abundace for high and low competence at wetland 8 (includes all species)")
```

Select only site-level data for relative abundance
```{r, Relative Abundance}
site_time_cc <- data %>% select(WetAltID, Month.1, RA2:RA42, cc) %>% select(-RAAmb,-RASal) %>% distinct()

site_time_cc <- mutate(site_time_cc, RAHigh = (RA9 + RA21 + RA26 + RA42))
site_time_cc <- mutate(site_time_cc, 
                       RALow = (RA2 + RA3 + RA4 + RA5 + RA6 + RA8 + RA20 + RA24 + RA27 + RA28 + RA29 + RA31 + RA34 + RA35 + RA38 + RA39 + RA41))
site_time_cc <- mutate(site_time_cc, total = RAHigh + RALow)

#how to facet this to look at all sites?
site_time_cc %>% filter(WetAltID==8) %>% select(Month.1, RAHigh, RALow) %>% melt() %>%
  ggplot(., aes(y = value, x = Month.1, fill = variable)) +
  geom_bar(position = "stack", stat = "identity") +
  labs(title = "Relative Abundance of high and low competence for wetland 8")

site_time_cc %>% filter(WetAltID==8) %>%
  ggplot(., aes(y=cc, x=Month.1)) +
  geom_point() +
  labs(title = "CC over time for Wetland 8")

data %>% select(WetAltID, Month.1, Prevalence) %>% filter(WetAltID==8) %>%
  ggplot(., aes(y=Prevalence, x=Month.1)) +
  geom_point() +
  labs(title = "Prevalence over time for Wetland 8")
```

I start with a summary of the ranavirus dataset that looks at the variance in ranavirus prevalence throughout a community over space and time and the viral loads of different amphibian species. The viral loads of these species provides us with insight about how to conceptualize these communities in a simplifying way. A simpler way to think about these systems is to split them up into two species with different levels of competency - one high and one low. This framework allows for a manageable transmission model with the ability to alter characteristics such as community composition, viral characteristics, and species characteristics.

Comparing prevalence and cc for sites and time
```{r, Prevalence and Community Competence}
data <- read_csv("competence.csv")

#dataset for looking at prevalence and cc for sitextime
prev_cc <- data %>% select(WetAltID, Month.1, cc, Prevalence, Month)
prev_cc$Month.1 <- factor(data$Month,levels=c("Feb","Mar","Apr","May","Jun","Jul"))
prev_cc %<>% distinct() %>% distinct() %>% remove_missing()

#heatmap of cc for sitextime
ggplot(prev_cc, aes(x=Month.1, y=WetAltID)) + geom_tile(aes(fill=cc)) + labs(title = "Community Competence", x="Month", y = "Wetland") + scale_fill_gradient(low = "white",high = "indianred1")

#heatmap of prevalence for sitextime
ggplot(prev_cc, aes(x=Month.1, y=WetAltID)) + geom_tile(aes(fill=Prevalence)) + labs(title = "Prevalence", x = "Month", y = "Wetland") + scale_fill_gradient(low = "white",high = "indianred1")

#plot cc by Prevalence(this is %prevalence)
prev_cc %>% ggplot(.,aes(x=cc, y=Prevalence)) + 
  geom_point() +
  theme_classic() +
  labs(title = "without lag")

#put everything in order by wetland and month
order <- prev_cc[order(prev_cc$WetAltID, prev_cc$Month.1),]

#this removes months out of sequence
order <- order[-c(4,37,38,45,85),]

#create new column that includes previous month's value for cc
for(n in 2:86){
order$lag[n] <- order$cc[n-1]
}

#plot cc by prevalence with lag
order %>% ggplot(.,aes(x=lag, y=Prevalence)) + 
  geom_point() +
  theme_classic() +
  labs(title = "with lag")

#remove the first entry for each wetland to remove the carryover from the last wetland
clean <- order %>% group_by(WetAltID) %>% filter(duplicated(WetAltID) | n()==1)

#plot cleaned plot with lag
clean %>% filter(Prevalence>0) %>% ggplot(.,aes(x=lag, y=Prevalence)) +
  geom_point() +
  theme_classic() + geom_smooth(method = "lm") +
  labs(title = "", x = "Community Competence", y = "Prevalence")
clean2 <- clean %>% drop_na()

#correlation test
cor.test(clean2$lag,clean2$Prevalence,method="spearman")
```

Community competence is a measure for the overall competency of a community as opposed to individual values of competency for each species of host. This metric is calculated using the relative abundance of a species multiplied by its competency value. This provides a sum for all species within a community and the competency that they contribute to that system. Community competence is a derived characteristic that agrees with our assumption of host competency playing a role in the transmission potential of an introduced parasite. An empirical measure of transmission potential might be the prevalence of ranavirus at a site. We can compare our derived measure and this empirical measure by plotting their correlation (i.e. higher competence relates to higher prevalence), and we consider a time lag of 1 month between an observation of competence and an observation of prevalence. When we do this, a positive relationship appears confirming our hypothesis that community competence correlates with prevalence.

Basic function for developing contour plots using simulated data
```{r, Invasion}
contour <- function(trans_1_min, trans_1_max, trans_3_min, trans_3_max, mort1, mort2, degr) {
data <- expand_grid("trans1" = seq(from = trans_1_min, to = trans_1_max, by = 0.0001),
                    "trans2" = c(0.0001),
                    "trans3" = seq(from = trans_3_min, to = trans_3_max, by = 0.0001),
                    "shed1" = c(rep(1/2, length("trans1"))),
                    "shed2" = c(rep(1/2, length("trans1"))),
                    "reco1" = c(rep(1/10, length("trans1"))),
                    "reco2" = c(rep(1/10, length("trans1"))),
                    "mort1" = c(mort1), 
                    "mort2" = c(mort2), 
                    "degr" = c(degr), 
                    "birth1" = c(rep(5/3, length("trans1"))),
                    "birth2" = c(rep(5/3, length("trans1"))))

data %<>% add_column(dfe1 = (data$birth1/data$mort1),dfe2 = (data$birth2/data$mort2))

mat <- matrix(1, nrow = 3, ncol = 3)

m <- nrow(data)

for(n in 1:m){
  mat[1] = (data$trans1[n]*data$dfe1[n])/(data$reco1[n]+data$mort1[n])
  mat[2] = (data$trans1[n]*data$dfe2[n])/(data$reco2[n]+data$mort2[n])
  mat[3] = (data$shed1[n]/data$degr[n])
  mat[4] = (data$trans2[n]*data$dfe1[n])/(data$reco1[n]+data$mort1[n])
  mat[5] = (data$trans2[n]*data$dfe2[n])/(data$reco2[n]+data$mort2[n])
  mat[6] = (data$shed2[n]/data$degr[n])
  mat[7] = (data$trans3[n]*data$dfe1[n])/(data$reco1[n]+data$mort1[n])
  mat[8] = (data$trans3[n]*data$dfe2[n])/(data$reco2[n]+data$mort2[n])
  mat[9] = 0
  eigen <- max(eigen(mat)$values)
  data$eigen[n] <- eigen
}

data$span1 <- 1/data$mort1
data$span2 <- 1/data$mort2

data$epid <- data$eigen > 1
data$prop1 <- data$trans1/data$trans2
data$prop3 <- data$trans3/data$trans2
data$prop_mort <- data$mort1/data$mort2
data$cap1 <- data$birth1/data$mort1
data$cap2 <- data$birth2/data$mort2
data$cap_prop <- data$cap1/data$cap2
data$tot <- data$cap1 + data$cap2
data$cc <- data$trans1*(data$cap1/data$tot) + data$trans2*(data$cap2/data$tot)
plots <- data %>% select(trans1, trans2, prop1, prop3, trans3, eigen, prop_mort, mort1, mort2, degr, cap1, cap2, cap_prop, tot, cc) %>% distinct()
return(plots)
}

#"default" parameters
plotr <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = 1/50,mort2 = 1/50,degr = 1/2)

plotr %>% ggplot(.,aes(x=prop3,y=prop1))+
  geom_contour(aes(x=prop3,y=prop1,z=eigen,colour = factor(..level.. == 1,levels = c(F,T)),group=mort2), show.legend = FALSE)+
  theme_classic()+
  scale_colour_manual(values = c(NA, "red")) + 
  labs(x = "Environmental", y = "Contact") +
  xlim(2,10) +
  ylim(2,10)
```

Use this to more easily make all the plots for varying different parameters
```{r, Changing Parameters}
plot_mort1 <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = c(1/50,1/60,1/70),mort2 = 1/50,degr = 1/2)

plot_mort1 %>% ggplot(.,aes(x=prop3,y=prop1))+
  geom_contour(aes(x=prop3,y=prop1,z=eigen,colour = factor(..level.. == 1,levels = c(F,T)),group=mort1), show.legend = FALSE)+
  theme_classic()+
  scale_colour_manual(values = c(NA, "red")) + 
  labs(x = "Environmental", y = "Contact") +
  xlim(2,10) +
  ylim(2,10) + 
  labs(title = "vary mort1")

plot_mort2 <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = 1/50,mort2 = c(1/50,1/60,1/70),degr = 1/2)

plot_mort2 %>% ggplot(.,aes(x=prop3,y=prop1))+
  geom_contour(aes(x=prop3,y=prop1,z=eigen,colour = factor(..level.. == 1,levels = c(F,T)),group=mort2), show.legend = FALSE)+
  theme_classic()+
  scale_colour_manual(values = c(NA, "red")) + 
  labs(x = "Environmental", y = "Contact") +
  xlim(2,10) +
  ylim(2,10) + 
  labs(title = "vary mort2")

plot_degr <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = 1/50,mort2 = 1/50,degr = c(1/2,1/3,1/4))

plot_degr %>% ggplot(.,aes(x=prop3,y=prop1))+
  geom_contour(aes(x=prop3,y=prop1,z=eigen,colour = factor(..level.. == 1,levels = c(F,T)),group=degr), show.legend = FALSE)+
  theme_classic()+
  scale_colour_manual(values = c(NA, "red")) + 
  labs(x = "Environmental", y = "Contact") +
  xlim(2,10) +
  ylim(2,10) + 
  labs(title = "vary degr")

```
Assuming that host competency is a relevant metric for evaluating the likelihood of an invasion of a parasite, we can numerically derive R0 from a set of equations. In line with our hypothesis, a greater value of community competency results in a larger range of parameter values in which a parasite can invade. This is represented when comparing the inner lines for plots (vary mort1 and vary mort2). The inner lines correspond and have the same total size for the combined populations but the proportion of highly competent individuals is higher in the "vary mort1" plot. Additionally, we can vary the rate of degradation of the free-infectious virus and change the slope of this curve, altering the ability for environmental transmission to compensate for direct transmission. As we favor environmental transmission, the importance of community composition is reduced. Currently, both species are equally competent in their ability to become infected but they vary in their ability to infect susceptible individuals.

Transmission Dynamics

```{r, Transmission Dynamics}
rm0<-function(t,x,params){
  Sa <- x[1]  
  Ia <- x[2]
  Ra <- x[3]
  Sb <- x[4]  
  Ib <- x[5]
  Rb <- x[6]
  V <- x[7]
  with(as.list(params),{
    dSa <- -beta*Sa*Ia - betab*Sa*Ib - phi*Sa*V + lambda - mu*Sa
    dIa <- beta*Sa*Ia + betab*Sa*Ib + phi*Sa*V - alpha*Ia - mu*Ia
    dRa <- alpha*Ia - mu*Ra

    dSb <- -betab*Sb*Ib - beta*Sb*Ia - phi*Sb*V + lambda - mu*Sb
    dIb <- betab*Sb*Ib + beta*Sb*Ia + phi*Sb*V - alpha*Ib - mu*Ib
    dRb <- alpha*Ib - mu*Rb

    dV <- sigma*Ia + sigma*Ib - epsilon*V
    res<-c(dSa,dIa,dRa,dSb,dIb,dRb,dV)
    list(res)
  })}
maxTime <- 300.0 # time is in years - run model for this time
times<-seq(0,maxTime,by=1) # how long we run the model for
# notes on params
# mu =     <- density-independent mortality rate
# beta =       <- contact transmission rate by infected A
# betab =       <- contact transmission rate by infected B
# phi =      <- environmental transmission rate
# lambda =       <- density-dependent growth rate
# alpha =     <- recovery rate
# sigma =    <- shedding rate
# epsilon =       <- viral degradation rate
params<-c(mu=1/50,
          beta=0.001,
          betab=0.0001,
          phi=0.001,
          lambda=5/3,
          alpha=1/10,
          sigma=1/2,
          epsilon=1/2)  # model parameters

xstart<-c(Sa=116,
          Ia=1,
          Ra=0,
          Sb=83,
          Ib=1,
          Rb=0,
          V=0)  # initial conditions

output<-as.data.frame(lsoda(xstart,times,rm0,params)) # tells computer to solve (integrate) equations

output %>% ggplot(.,aes(x=time))+
  geom_line(aes(y=Sa,col="Susceptible A"))+
  geom_line(aes(y=Ia,col="Infected A"))+
  geom_line(aes(y=Ra,col="Recovered A"))+
  geom_line(aes(y=Sb,col="Susceptible B"))+
  geom_line(aes(y=Ib,col="Infected B"))+
  geom_line(aes(y=Rb,col="Recovered B"))+
  geom_line(aes(y=V,col="Virus"))+
  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
  theme(legend.position = c(0.85,0.5))+
  labs(y="N",x="Time",col="Population")+
  ggtitle("2-Species Transmission Model for Ranavirus in an Amphibian Wetland")+theme(plot.title = element_text(hjust=0, size=10))
```

Transmission Dynamics with greater heterogeneity in species characteristics

```{r, Transmission Dynamics 2}
rm0<-function(t,x,params){
  Sa <- x[1]  
  Ia <- x[2]
  Ra <- x[3]
  Sb <- x[4]  
  Ib <- x[5]
  Rb <- x[6]
  V <- x[7]
  with(as.list(params),{
    dSa <- -betaa*Sa*Ia - betab*Sa*Ib - phi*Sa*V + lambdaa - mua*Sa
    dIa <- betaa*Sa*Ia + betab*Sa*Ib + phi*Sa*V - alphaa*Ia - mua*Ia
    dRa <- alphaa*Ia - mua*Ra

    dSb <- -betab*Sb*Ib - betaa*Sb*Ia - phi*Sb*V + lambdab - mub*Sb
    dIb <- betab*Sb*Ib + betaa*Sb*Ia + phi*Sb*V - alphab*Ib - mub*Ib
    dRb <- alphab*Ib - mub*Rb

    dV <- sigma*Ia + sigma*Ib - epsilon*V
    res<-c(dSa,dIa,dRa,dSb,dIb,dRb,dV)
    list(res)
  })}
maxTime <- 300.0 # time is in years - run model for this time
times<-seq(0,maxTime,by=1) # how long we run the model for
# notes on params
# mu =     <- density-independent mortality rate
# beta =       <- contact transmission rate by infected A
# betab =       <- contact transmission rate by infected B
# phi =      <- environmental transmission rate
# lambda =       <- density-dependent growth rate
# alpha =     <- recovery rate
# sigma =    <- shedding rate
# epsilon =       <- viral degradation rate
params<-c(mua=1/40,
          mub=1/50,
          betaa=0.001,
          betab=0.0001,
          phi=0.001,
          lambdaa=10/3,
          lambdab=5/3,
          alphaa=1/15,
          alphab=1/10,
          sigma=1/2,
          epsilon=1/2)  # model parameters

xstart<-c(Sa=116,
          Ia=1,
          Ra=0,
          Sb=83,
          Ib=1,
          Rb=0,
          V=0)  # initial conditions

output<-as.data.frame(lsoda(xstart,times,rm0,params)) # tells computer to solve (integrate) equations

output %>% ggplot(.,aes(x=time))+
  geom_line(aes(y=Sa,col="Susceptible A"))+
  geom_line(aes(y=Ia,col="Infected A"))+
  geom_line(aes(y=Ra,col="Recovered A"))+
  geom_line(aes(y=Sb,col="Susceptible B"))+
  geom_line(aes(y=Ib,col="Infected B"))+
  geom_line(aes(y=Rb,col="Recovered B"))+
  geom_line(aes(y=V,col="Virus"))+
  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
  theme(legend.position = c(0.85,0.8))+
  labs(y="N",x="Time",col="Population")+
  ggtitle("2-Species Transmission Model for Ranavirus in an Amphibian Wetland")+theme(plot.title = element_text(hjust=0, size=10))
```

These plots look at the transmission dynamics over time for a scenario where the demography of the species are equal and when they are different. In the scenario when they are different we observe a tradeoff between high competency and shorter lifespan and higher birth rate. But this does not consider pathogen-induced mortality which we might expect to be higher for the more competent species that invests less into immunity and dies quicker anyways.

Further questions and directions for this project

 - How might space correlate to changes in community competence or disease prevalence?
     - Can we map out the sites and their distances to each other to see if proximity to highly infected/competent sites are increase the likelihood        of infection/community competence in another?
 - How to apply this model to our data to observe the role that community competence plays in each site?
 - How to consider pathogen-induced mortality and the tradeoffs we might expect from an immunity-reproduction relationship?
 - How to observe community assembly and disassembly over time and the changing risks for disease?

This chunk creates a spatial matrix for all sites and their distances from each other
```{r}
##Measuring distance between points
library(geosphere)
library(biogeo)
library(raster)

latlon <- read_csv("srs_wetland_coordinates.csv")

latlon$dd_x <- dms2dd(latlon$xdd,latlon$xmm,latlon$xss,latlon$xdir)
latlon$dd_y <- dms2dd(latlon$ydd,latlon$ymm,latlon$yss,latlon$ydir)

srs_lat_lon <- tibble(WetAltID = latlon$Wetland, x = latlon$dd_x, y = latlon$dd_y)
#srs_lat_lon$latlon <- paste(srs_lat_lon$x,",", srs_lat_lon$y)

distances <- pointDistance(c(srs_lat_lon$x[1],srs_lat_lon$y[1]),c(srs_lat_lon$x[2],srs_lat_lon$y[2]), lonlat = TRUE)
matrix <- cbind(srs_lat_lon$x,srs_lat_lon$y)
dist_mat <- distm(matrix, fun = distGeo)
colnames(dist_mat) <- c(srs_lat_lon$WetAltID)
rownames(dist_mat) <- c(srs_lat_lon$WetAltID)

srs_lat_lon %>% ggplot(.,aes(x = x, y = y)) + geom_point() + geom_text(label = srs_lat_lon$WetAltID) 
dist_mat
```

Grab climate data from worldclim for each site
```{r}
r <- getData("worldclim",var="bio",res=10)

r <- r[[c(1,12)]]
names(r) <- c("Temp","Prec")

lats <- latlon$dd_x
lons <- latlon$dd_y

coords <- data.frame(x=lons,y=lats)

points <- SpatialPoints(coords, proj4string = r@crs)

values <- raster::extract(r,points)

climate <- cbind.data.frame(coordinates(points),values)
climate
```

This chunk creates a matrix for all amphibian pairwise phylogenetic distances and simulates a random walk for competence as a trait
```{r}
library(ape)
library(rotl)
library(phangorn)
library(geiger)
library(phytools)

#trait1 <- rTraitCont(tree, "OU", theta = 0, alpha=.1, sigma=.01)

names <- c("Notophthalamus viridescens", "Scaphiopus holbrookii", "Anaxyrus terrestris", "Pseudacris crucifer", "Hyla gratiosa", "Hyla avivoca",
                  "Pseudacris nigrita", "Pseudacris ornata", "Gastrophyrne carolinensis", "Lithobates clamitans", "Lithobates sphenocephalus",
           "Acris gryllus", "Hyla chrysoscelis", "Hyla cinerea", "Hyla femoralis", "Lithobates catesbeianus", "Siren intermedia")

resolved <- tnrs_match_names(names)

#tree <- tol_induced_subtree(resolved$ott_id)
#plot(tree, edge.width = "3")

#plot.phylo(tree, edge.width = '3', show.tip.label = TRUE)

species <- resolved$unique_name
species <- gsub(" ", "_", species)

amphibians <- read.nexus("asup_just_tree.txt")
#plot(amphibians, edge.width = "1")

str(amphibians)

setdiff(species,amphibians$tip.label)
species <- gsub("Dryophytes","Hyla",species)
species <- gsub("gratiosus","gratiosa",species)
species <- gsub("cinereus","cinerea",species)
species <- gsub("Anaxyrus","Bufo",species)
setdiff(species,amphibians$tip.label)

#prune
pruned.tree<-drop.tip(amphibians,amphibians$tip.label[-match(species, amphibians$tip.label)])

plot(pruned.tree)

#phylo distances
phy_dist <- cophenetic.phylo(pruned.tree)
phy_dist

#random walk for competence
trait1 <- rTraitCont(pruned.tree, "OU", theta = 0, alpha=.1, sigma=.01)
trait1 <- base::as.data.frame(trait1, row.names = NULL)
trait1 <- rownames_to_column(trait1, "species")
colnames(trait1) <- c("species", "value")
#plotted with squares to remove negative values
ggplot(trait1, aes(x = species, y = value)) + geom_point() + theme_classic() + theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

### 2020-10-30 gradient_analysis.Rmd
changed a lot to try and understand how ordination analysis actually works. focusing on pca for actual ms figure

---
title: "gradient_analysis"
author: "Daniel Suh"
date: "10/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

load packages
```{r}
library(tidyverse)
library(magrittr)
library(ecodist)
library(vegan)
library(here)
```

read data and make species and environmental matrices
```{r}
data <- read_csv(here("data/weighted_prev_competence.csv"))

#make community matrix
community_mat <- data %>% dplyr::select(WetAltID, Month.1, AB2:AB8, AB9, AB20:AB42)
community_mat$Month.1 <- gsub("Month", "", community_mat$Month.1) 
community_mat %<>% mutate(., siteID = paste(WetAltID, Month.1, sep = "_")) %>% 
  distinct() %>% arrange(.,WetAltID) %>% dplyr::select(-WetAltID, -Month.1) %>% dplyr::select(siteID, AB2:AB42)

community_mat %<>% column_to_rownames(., var = "siteID")

#select for environmental variables
env <- data %>% dplyr::select(WetAltID, Month.1, MeanAirT, MeanWaterTempPredC, DryingScore, CanopyCover, Area, Perimeter) %>% distinct() %>% arrange(.,WetAltID) 
env$Month.1 <- gsub("Month", "", env$Month.1)
#make siteID rowname
env$siteID <- paste(env$WetAltID, env$Month.1, sep = "_")
env %<>% dplyr::select(-WetAltID, -Month.1)
#remove duplicated rows (siteID is duplicated but env. variables are not... not sure why this is)
env <- env[-c(53, 90),]
env_mat <- env %>% column_to_rownames(., var = "siteID")

#replace NA's with zeros. Probably not the best option but will allow analysis to run for now
community_mat[is.na(community_mat)] <- 0
env_mat[is.na(env_mat)] <- 0

nrow(community_mat) == nrow(env_mat)

site_labels <- rownames(community_mat)
```

Gradient Analysis - a number of distance-based or rotation-based techniques for ordinating data to reduce the dimensions of data in order to compare two related datasets. A large number of independent variables require many degrees of freedoms and large number of observations. Gradient analysis allows for studying relationship between variables when the number of observations is limited to perform standard statistical procedures. Ordination allows for the reduction of data to ordination scores so that other variables can be correlated to ordination scores.

NMDS: non-metric multidimensional scaling - a distance-based ordination method
```{r}
#bray-curtis dissimilarity scores
spec_dist <- bcdist(community_mat)
spec_nmds <- nmds(spec_dist, mindim = 2, maxdim = 2)
spec_scores <- nmds.min(spec_nmds)

#mahalanobis dissimilarity scoring for sites by environmental variables
env_dist <- ecodist::distance(env_mat, method = "mahalanobis")
env_nmds <- nmds(env_dist, mindim = 2, maxdim = 2)
env_scores <- nmds.min(env_nmds)
```

Direct Gradient Analysis with NMDS
```{r}
####Direct GA w/ NMDS####
#Ordinate independent variable (environment)
#correlate species as vectors to ordination scores
plot(env_scores, col="white") 
vectors=vf(env_scores, env_mat, nperm=100) 
plot(vectors, col="red")

vectors=vf(env_scores, community_mat, nperm=100) 
plot(vectors, ascale=1, col="blue")
#### useful for understanding species composition only in light of the measured independent variables
#### other patterns that might be due to other independent variables are effectively ignored
```

Indirect Gradient Analysis with NMDS
```{r}
####Indirect GA w/ NMDS####
#Ordinate dependent variable (species)
#correlate ordination scores with independent variable (environment)
plot(spec_scores, col="white") 
vectors=vf(spec_scores, community_mat, nperm=100) 
plot(vectors, col="red")

vectors=vf(spec_scores, env_mat, nperm=100) 
plot(vectors, ascale=1, col="blue")
#### useful for understanding the similarity in species composition between sites regardless of independent variable
#### independent variable is then correlated to species but patterns may have been driven by other independent variables not measured
```

PCA: Principal Component Analysis - a rotation-based ordination method
```{r}

pca_env_output <- princomp(env_mat, cor=T) # PCA
biplot(pca_env_output) # Generates a bi-plot with vectors
vec1 <- envfit(pca_env_output$score[,1:2], community_mat, permutations=0) 
plot(vec1, col="blue")

pca_env_scores <- pca_env_output$score[,1:2] # get the PC1 and PC2 scores 
pca_env_spec_scores <- cbind(pca_env_scores,community_mat) # merge with species 
correlations <- cor(pca_env_spec_scores) # calculate correlations
correlations12 <- correlations[3:23,1:2] # the loadings we want

#pca on community matrix and gather scores for first two principal components
pca <- princomp(community_mat, scores = TRUE)
p_spec_scores <- as.data.frame(pca$scores)
p_spec_scores %<>% dplyr::select(Comp.1, Comp.2) %>% rownames_to_column(., var = "siteID")

env_scores <- as.data.frame(pca_env_output$scores)
env_scores %<>% dplyr::select(Comp.1, Comp.2) %>% rownames_to_column(., var = "siteID")

```

What is the difference between the scores for species from envfit and the values from the correlation between environmental ordination scores and the community matrix?

### 2020-10-30 scripts/figures/fig3_code.R

@@ -13,8 +13,10 @@ data <- read_csv(here("data/weighted_prev_competence.csv"))
 
 
 #make community matrix
-community_mat <- data %>% dplyr::select(WetAltID, Month.1, AB2:AB8, AB9, AB20:AB42) %>%
-  mutate(., siteID = paste(WetAltID, Month.1, sep = "_")) %>% distinct() %>% arrange(.,WetAltID) %>% dplyr::select(-WetAltID, -Month.1) %>% dplyr::select(siteID, AB2:AB42)
+community_mat <- data %>% dplyr::select(WetAltID, Month.1, AB2:AB8, AB9, AB20:AB42)
+community_mat$Month.1 <- gsub("Month", "", community_mat$Month.1) 
+community_mat %<>% mutate(., siteID = paste(WetAltID, Month.1, sep = "_")) %>% 
+  distinct() %>% arrange(.,WetAltID) %>% dplyr::select(-WetAltID, -Month.1) %>% dplyr::select(siteID, AB2:AB42)
 
 community_mat %<>% column_to_rownames(., var = "siteID")
 
@@ -33,50 +35,40 @@ env_mat <- env %>% column_to_rownames(., var = "siteID")
 community_mat[is.na(community_mat)] <- 0
 env_mat[is.na(env_mat)] <- 0
 
-nrow(community_mat) == nrow(env_mat)
-
 site_labels <- rownames(community_mat)
 
-#bray-curtis dissimilarity scores
-spec_dist <- bcdist(community_mat)
-spec_nmds <- nmds(spec_dist, mindim = 2, maxdim = 2)
-spec_scores <- nmds.min(spec_nmds)
 
-#mahalanobis dissimilarity scoring for sites by environmental variables
-env_dist <- ecodist::distance(env_mat, method = "mahalanobis")
-env_nmds <- nmds(env_dist, mindim = 2, maxdim = 2)
-env_scores <- nmds.min(env_nmds)
+pca_env_output=princomp(env_mat, cor=T) # PCA
+biplot(pca_env_output) # Generates a bi-plot with vectors
 
-#pca on community matrix and gather scores for first two principal components
-pca <- princomp(community_mat, scores = TRUE)
-scores <- as.data.frame(pca$scores)
-scores %<>% dplyr::select(Comp.1, Comp.2) %>% rownames_to_column(., var = "siteID")
+spec_vec=envfit(pca_env_output$score[,1:2], community_mat, permutations=0) 
[…]

### 2020-11-13 scripts/figures/fig3_code.R
comment out unnecessary sections

@@ -135,18 +135,18 @@ env_scores %>% drop_na(cc) %>% ggplot(.)+geom_point(aes(x=Comp.1,y=Comp.2,fill=C
 #need to look at getting the dots in there and finding a way to make it legible while keeping the original vector lengths
 #I need to be able to better explain what the direction and magnitude of each vector represents
 
-scores12 %>% ggplot(.)+geom_point(aes(x=Comp.1,y=Comp.2),pch=21)+
-  xlab("PC1 (48.2%)")+ylab("PC2 (27.5%)")+
-  scale_fill_manual(values=c("darkgreen","lightgreen"))+
-  scale_color_manual(values=c("darkgreen","black"))+
-  coord_fixed()+
-  geom_segment(data=spp.scrs,
-               aes(x = 0, xend = Comp.1, y = 0, yend = Comp.2),
-               arrow = arrow(length = unit(0.1, "cm")), colour = "darkorange1") +
-  geom_text_repel(data = spp.scrs, aes(x = Comp.1, y = Comp.2, label = Species),
-                  size = 2.5,color="darkorange2")+
-  geom_segment(data=env.scrs,
-               aes(x = 0, xend = Comp.1, y = 0, yend = Comp.2),
-               arrow = arrow(length = unit(0.1, "cm")), colour = "purple") +
-  geom_text_repel(data = env.scrs, aes(x = Comp.1, y = Comp.2, label = Species),
-                  size = 2.5,color="purple")
+# scores12 %>% ggplot(.)+geom_point(aes(x=Comp.1,y=Comp.2),pch=21)+
+#   xlab("PC1 (48.2%)")+ylab("PC2 (27.5%)")+
+#   scale_fill_manual(values=c("darkgreen","lightgreen"))+
+#   scale_color_manual(values=c("darkgreen","black"))+
+#   coord_fixed()+
+#   geom_segment(data=spp.scrs,
+#                aes(x = 0, xend = Comp.1, y = 0, yend = Comp.2),
+#                arrow = arrow(length = unit(0.1, "cm")), colour = "darkorange1") +
+#   geom_text_repel(data = spp.scrs, aes(x = Comp.1, y = Comp.2, label = Species),
+#                   size = 2.5,color="darkorange2")+
+#   geom_segment(data=env.scrs,
+#                aes(x = 0, xend = Comp.1, y = 0, yend = Comp.2),
+#                arrow = arrow(length = unit(0.1, "cm")), colour = "purple") +
[…]

### 2020-11-13 scripts/figures/fig4_code.R

@@ -72,41 +72,41 @@ site_scores %>% ggplot(.,aes(x=pc1Rank,y=cc))+
 #scale_color_manual(values = c( "#4575b4", "#91bfdb", "#e0f3f8", "#fee090", "#fc8d59", "#d73027"))
 
 
-#make same plot with NMDS scores rather than PCA scores
-library(vegan)
-set.seed(3000)
-
-vegdist <- vegdist(community_mat, method = "bray")
-
-#stress test for nmds
-stress <- c()
-for (i in 1:10){
-  output <- metaMDS(community_mat, distance = "bray", k = i, trace = F)
-  stress[i] <- output$stress
-}
-
-stress %<>% as.data.frame(.) %>% rename(., stress = .) %>% mutate(., dim = as.numeric(rownames(.)))
-
-ggplot(stress, aes (x = dim, y = stress)) + geom_point()
-#the difference between NMDS and PCA is that PCA uses euclidean distances while NMDS rank orders observations for ordination (this is why it is non-metric?)
-
-NMDS1 <- metaMDS(community_mat, distance = "bray", k = 6, trace = F, autotransform = FALSE)
-stressplot(NMDS1)
-ordiplot(NMDS1, type = "n")
-orditorp(NMDS1, display = "species", col = "red")
-orditorp(NMDS1, display = "sites", cex = 1.1)
-
-
-points <- as.data.frame(NMDS1$points)
-points %<>% select(MDS1) %>% mutate(., siteID = rownames(.))
-points %<>% left_join(tmp, .)
-points %<>% mutate(rank = dense_rank(desc(MDS1)))
-
-points %>% ggplot(.,aes(x=rank,y=cc))+
-  geom_point(aes(color = Month, shape = factor(WetAltID), size = Size))+
-  labs(x="NMDS Component 1 Rank", y = "Community Competence (CC)") + 
-  scale_shape_manual(values = rep(1:20, len = 20)) +
-  guides(shape=F) +
-  theme_classic() +
-  theme(aspect.ratio=1/1.67, legend.box = "horizontal") +
-  scale_color_brewer(palette = "Paired")
+# #make same plot with NMDS scores rather than PCA scores
+# library(vegan)
+# set.seed(3000)
+# 
+# vegdist <- vegdist(community_mat, method = "bray")
+# 
+# #stress test for nmds
+# stress <- c()
+# for (i in 1:10){
[…]

### 2020-11-20 gradient_analysis.Rmd
change filepath
label

@@
 ```{r}
-data <- read_csv(here("data/weighted_prev_competence.csv"))
+data <- read_csv(here("data/weighted_prev_competence_111220.csv"))
 
@@
 ```{r}
-
+#Direct Gradient Analysis w/ PCA
 pca_env_output <- princomp(env_mat, cor=T) # PCA

### 2021-01-25 gradient_analysis.Rmd
add indirect gradient analysis w/ pca

@@
 
+Indirect Grdient Analysis using PCA
+explanation of how envfit actually works -> https://stackoverflow.com/questions/60953996/how-are-envfit-results-created 
+```{r}
+#sites ordinated by species abundances
+#cor = F to use correlation matrix because variables are all on same scale
+species_pca <- princomp(community_mat, cor = F)
+plot(species_pca$scores, pch = 16)
+summary(species_pca)
+loadings(species_pca)
+scores(species_pca)
+biplot(species_pca)
+
+env_fit_vectors <- envfit(species_pca, env_mat)
+
+biplot(species_pca)
+plot(env_fit_vectors)
+```
+
 What is the difference between the scores for species from envfit and the values from the correlation between environmental ordination scores and the community matrix?

### 2021-01-25 scripts/figures/fig3_code.R
changed scaling for the vectors. we shouldn't scale them

@@ -114,9 +114,10 @@ env.scrs <- as.data.frame(scores(vec2, display = "vectors"))
 env.scrs <- cbind(env.scrs, Species = rownames(env.scrs))
 
 library(ggrepel)
-myScaleUp1 <- 5
-myScaleUp2 <- 2.5
-env_scores %>% drop_na(cc) %>% ggplot(.)+geom_point(aes(x=Comp.1,y=Comp.2,fill=CC,color=CC),pch=21)+
+myScaleUp1 <- 1
+myScaleUp2 <- 1
+env_scores %>% drop_na(cc) %>% ggplot(.)+
+#  geom_point(aes(x=Comp.1,y=Comp.2,fill=CC,color=CC),pch=21)+
   xlab("PC1 (48.2%)")+ylab("PC2 (27.5%)")+
   scale_fill_manual(values=c("darkgreen","lightgreen"))+
   scale_color_manual(values=c("darkgreen","black"))+

### 2021-01-25 scripts/figures/fig6_code.R
fixed join functions

@@ -97,7 +97,8 @@ cor.test(lag_evenness$lag_richness,lag_evenness$Prevalence,method="spearman")
 
 evenness %<>% select(-WetAltID)
 abundances %<>% left_join(.,evenness, by = "siteID")
-abundances %<>% left_join(.,lag_evenness, by = "siteID")
+lag_evenness %<>% select(siteID, lag_cc, lag_richness, lag_J, lag_size)
+abundances %<>% inner_join(.,lag_evenness, by = "siteID")
 abundances <- mutate(abundances, ABHigh = (AB9 + AB21 + AB26 + AB42)) 
 abundances <- mutate(abundances, ABLow = (AB2 + AB3 + AB4 + AB5 + AB6 + AB8 + AB20 + AB24 + AB27 + AB28 + AB29 + AB31 + AB34 + AB35 + AB38 + AB39 + AB41))
 abundances <- mutate(abundances, total = ABHigh + ABLow)
@@ -143,6 +144,7 @@ RA_plot<- z %>% ggplot(.,aes(x=siteID,y=abund,fill=Species))+
   theme(axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+ylab("Rel. Abun.")
 
 #still need to clean this up but final figure should have a similar format to this
+
 final <- AB_plot/evenness_plot/RA_plot/cc_plot
 
 final

### 2021-01-25 gradient_analysis.Rmd
clean up scripts

@@
 ---
+I used this document as my instructions for most of this analysis. https://sites.ualberta.ca/~lkgray/uploads/7/3/6/2/7362679/07-gradientanalysis.pdf s
 

### 2021-01-25 scripts/figures/fig3_code.R

@@ -8,6 +8,8 @@ library(magrittr)
 library(ecodist)
 library(vegan)
 library(here)
+library(ggrepel)
+
 
 data <- read_csv(here("data/weighted_prev_competence_111220.csv"))
 
@@ -48,13 +50,6 @@ plot(spec_vec, col="blue")
 env_var <- as.data.frame(pca_env_output$loadings[,1:2])
 
 
-env_scores <- pca_env_output$score[,1:2] # get the PC1 and PC2 scores 
-env_comm_scores <- cbind(env_scores,community_mat) # merge with species 
-correlations <- cor(env_comm_scores)
-spec_corr <- correlations[3:23,1:2] # the loadings we want
-
-
-
 site_scores <- data %>% dplyr::select(WetAltID, Month.1, cc) 
 site_scores$Month.1 <- gsub("Month", "", site_scores$Month.1)
 site_scores %<>% mutate(., siteID = paste(WetAltID, Month.1, sep = "_")) %>% 
@@ -62,7 +57,7 @@ site_scores %<>% mutate(., siteID = paste(WetAltID, Month.1, sep = "_")) %>%
   arrange(.,WetAltID) %>% 
   dplyr::select(-WetAltID, -Month.1) %>% dplyr::select(siteID, cc)
 
-
+env_scores <- pca_env_output$score[,1:2] # get the PC1 and PC2 scores 
 env_scores %<>% as.data.frame(.)
 env_scores %<>% mutate(siteID=rownames(.))
 tmp <- site_scores %>% select(siteID,cc)
@@ -72,20 +67,18 @@ env_scores %<>% mutate(CC=as.factor(ifelse(cc>45000,"Hi","Lo")))
 spec_scores <- as.data.frame(scores(spec_vec, display = "vectors"))
 
 
-#what is the difference between spec_scores and spec_corr
-
 env_scores %>% drop_na(cc) %>% ggplot(.) +
   geom_point(aes(x=Comp.1, y=Comp.2, shape = CC, color = cc)) +
   xlab("PC1 (48.2%)")+ylab("PC2 (27.5%)")+
   geom_segment(data=spec_scores,
-             aes(x = 0, xend = myScaleUp1*Comp.1, y = 0, yend = myScaleUp1*Comp.2),
+             aes(x = 0, xend = Comp.1, y = 0, yend = Comp.2),
              arrow = arrow(length = unit(0.1, "cm")), colour = "darkorange1") +
[…]

### 2021-02-19 scripts/figures/fig2_code.R
renamed scripts

@@ -1,34 +1,118 @@
 #Daniel Suh and Andrew Park
 #6/3/20
 
-#Figure 2 in rv_cc manuscript
+#Figure 4 in rv_cc manuscript
 
 library(tidyverse)
 library(magrittr)
-library(metacom)
+library(ggforce)
 library(here)
 
 
 data <- read_csv("data/weighted_prev_competence_111220.csv")
 
-#make community matrix
-community_mat <- data %>% dplyr::select(WetAltID, Month.1, AB2:AB8, AB9, AB20:AB42) %>%
-  mutate(., siteID = paste(WetAltID, Month.1, sep = "_")) %>% distinct() %>% arrange(.,WetAltID) %>% dplyr::select(-WetAltID, -Month.1) %>% dplyr::select(siteID, AB2:AB42)
 
+
+
+#select for community variables
+community_mat <- data %>% dplyr::select(WetAltID, Month.1, AB2:AB8, AB9, AB20:AB42) 
+community_mat$Month.1 <- gsub("Month", "", community_mat$Month.1)
+community_mat %<>% mutate(., siteID = paste(WetAltID, Month.1, sep = "_")) %>% distinct() %>% arrange(.,WetAltID) %>% dplyr::select(-WetAltID, -Month.1) %>% dplyr::select(siteID, AB2:AB42)
+#rownames
 community_mat %<>% column_to_rownames(., var = "siteID")
 
-Imagine(t(community_mat), 
-        fill = FALSE, 
-        order = TRUE, 
-        xlab = "", 
-        ylab = "",
-        xline = 5,
-        yline = 4,
-        sitenames = ,
-        speciesnames = ,
-        binary=T)
+#select for environmental variables
+env <- data %>% dplyr::select(WetAltID, Month.1, MeanAirT, MeanWaterTempPredC, DryingScore, CanopyCover, Area, Perimeter) %>% distinct() %>% arrange(.,WetAltID) 
+env$Month.1 <- gsub("Month", "", env$Month.1)
+#make siteID rowname
+env$siteID <- paste(env$WetAltID, env$Month.1, sep = "_")
+env %<>% dplyr::select(-WetAltID, -Month.1)
+#remove duplicated rows (siteID is duplicated but env. variables are not... not sure why this is)
+env <- env[-c(53, 90),]
+env_mat <- env %>% column_to_rownames(., var = "siteID")
 
-ordered_matrix <- OrderMatrix(community_mat)
-rownames(ordered_matrix)
[…]

### 2021-02-19 scripts/figures/fig3_code.R

@@ -1,163 +1,178 @@
-#Daniel Suh and Andrew Park
-#6/3/20
+#Script Originator: Daniel Suh
 
-#Figure 3 in rv_cc manuscript
+#This script creates figure 5
+
 
 library(tidyverse)
 library(magrittr)
-library(ecodist)
 library(vegan)
+library(reshape2)
+library(patchwork)
+library(ggnewscale)
 library(here)
-library(ggrepel)
-
 
 data <- read_csv(here("data/weighted_prev_competence_111220.csv"))
 
+#Get species richness from matrix
 
-#make community matrix
-community_mat <- data %>% dplyr::select(WetAltID, Month.1, AB2:AB8, AB9, AB20:AB42)
-community_mat$Month.1 <- gsub("Month", "", community_mat$Month.1) 
-community_mat %<>% mutate(., siteID = paste(WetAltID, Month.1, sep = "_")) %>% 
-  distinct() %>% arrange(.,WetAltID) %>% dplyr::select(-WetAltID, -Month.1) %>% dplyr::select(siteID, AB2:AB42)
+community_mat <- data %>% dplyr::select(WetAltID, Month.1, AB2:AB8, AB9, AB20:AB42) %>%
+  mutate(., siteID = paste(WetAltID, Month.1, sep = "_")) %>% distinct() 
+
+community_mat <- community_mat[order(community_mat$WetAltID, community_mat$Month.1),]
+
+community_mat %<>% arrange(.,WetAltID) %>% dplyr::select(-WetAltID, -Month.1) %>% dplyr::select(siteID, AB2:AB42)
+
+abundances <- community_mat
 
 community_mat %<>% column_to_rownames(., var = "siteID")
 
+presence_mat <- ifelse(community_mat>0, 1, 0)
 
-#select for environmental variables
-env <- data %>% dplyr::select(WetAltID, Month.1, MeanAirT, MeanWaterTempPredC, DryingScore, CanopyCover, Area, Perimeter) %>% distinct() %>% arrange(.,WetAltID) 
-env$Month.1 <- gsub("Month", "", env$Month.1)
-#make siteID rowname
-env$siteID <- paste(env$WetAltID, env$Month.1, sep = "_")
-env %<>% dplyr::select(-WetAltID, -Month.1)
-#remove duplicated rows (siteID is duplicated but env. variables are not... not sure why this is)
-env <- env[-c(53, 90),]
[…]

### 2021-02-19 scripts/figures/fig4_code.R

@@ -1,118 +1,180 @@
-#Daniel Suh and Andrew Park
-#6/3/20
+#Script Originator: Daniel Suh
 
-#Figure 4 in rv_cc manuscript
+#This script creates figure 6
+
 
 library(tidyverse)
 library(magrittr)
-library(ggforce)
+library(vegan)
+library(reshape2)
+library(patchwork)
+library(ggnewscale)
 library(here)
 
+data <- read_csv(here("data/weighted_prev_competence_111220.csv")) #read data
 
-data <- read_csv("data/weighted_prev_competence_111220.csv")
+#Get species richness from matrix
 
+community_mat <- data %>% dplyr::select(WetAltID, Month.1, AB2:AB8, AB9, AB20:AB42) %>%
+  mutate(., siteID = paste(WetAltID, Month.1, sep = "_")) %>% distinct() #select data and make siteID column
 
+community_mat <- community_mat[order(community_mat$WetAltID, community_mat$Month.1),] #order by wetland and month
 
+community_mat %<>% arrange(.,WetAltID) %>% dplyr::select(-WetAltID, -Month.1) %>% dplyr::select(siteID, AB2:AB42)
 
-#select for community variables
-community_mat <- data %>% dplyr::select(WetAltID, Month.1, AB2:AB8, AB9, AB20:AB42) 
-community_mat$Month.1 <- gsub("Month", "", community_mat$Month.1)
-community_mat %<>% mutate(., siteID = paste(WetAltID, Month.1, sep = "_")) %>% distinct() %>% arrange(.,WetAltID) %>% dplyr::select(-WetAltID, -Month.1) %>% dplyr::select(siteID, AB2:AB42)
-#rownames
-community_mat %<>% column_to_rownames(., var = "siteID")
+abundances <- community_mat #make abundances df
 
-#select for environmental variables
-env <- data %>% dplyr::select(WetAltID, Month.1, MeanAirT, MeanWaterTempPredC, DryingScore, CanopyCover, Area, Perimeter) %>% distinct() %>% arrange(.,WetAltID) 
-env$Month.1 <- gsub("Month", "", env$Month.1)
-#make siteID rowname
-env$siteID <- paste(env$WetAltID, env$Month.1, sep = "_")
-env %<>% dplyr::select(-WetAltID, -Month.1)
-#remove duplicated rows (siteID is duplicated but env. variables are not... not sure why this is)
-env <- env[-c(53, 90),]
[…]

### 2021-02-19 scripts/figures/fig5_code.R

@@ -1,178 +1,73 @@
-#Script Originator: Daniel Suh
+#Daniel Suh and Andrew Park
+#6/3/20
 
-#This script creates figure 5
+#Figure 7 in rv_cc manuscript
 
-
+#install_github("thomasp85/patchwork")
 library(tidyverse)
 library(magrittr)
-library(vegan)
-library(reshape2)
 library(patchwork)
-library(ggnewscale)
 library(here)
 
-data <- read_csv(here("data/weighted_prev_competence_111220.csv"))
 
-#Get species richness from matrix
+data <- read_csv("data/weighted_prev_competence_111220.csv")
 
-community_mat <- data %>% dplyr::select(WetAltID, Month.1, AB2:AB8, AB9, AB20:AB42) %>%
-  mutate(., siteID = paste(WetAltID, Month.1, sep = "_")) %>% distinct() 
+prev_cc <- data %>% dplyr::select(WetAltID, Month.1, cc, Prevalence, Month, AB2:AB42, MeanWaterTempPredC)
+prev_cc %<>% dplyr::select(-ABAmb, -ABSal) %>%  mutate(., total = rowSums(.[6:26])) #get totals for community size
+prev_cc %<>% dplyr::select(-AB2:-AB42)
+prev_cc %<>% distinct()
 
-community_mat <- community_mat[order(community_mat$WetAltID, community_mat$Month.1),]
+order <- prev_cc[order(prev_cc$WetAltID, prev_cc$Month.1),]
 
-community_mat %<>% arrange(.,WetAltID) %>% dplyr::select(-WetAltID, -Month.1) %>% dplyr::select(siteID, AB2:AB42)
+#this removes months out of sequence
+order <- order[-c(4,14,47,90),]
 
-abundances <- community_mat
+order$lag_cc <- order$cc
+order$lag_size <- order$total
+order$lag_temp <- order$MeanWaterTempPredC
 
-community_mat %<>% column_to_rownames(., var = "siteID")
-
-presence_mat <- ifelse(community_mat>0, 1, 0)
-
-siteID <- rownames(presence_mat)
-
-richness_mat <- tibble()
-for (i in 1:length(siteID)){
-  richness_mat[i,1] <- siteID[i]
-  richness_mat[i,2] <- sum(presence_mat[i,])
-}
-richness_mat %<>% rename(siteID = 1, richness = 2)
-
-order <- richness_mat
-
[…]

### 2021-02-19 scripts/figures/fig6_code.R

@@ -1,180 +1,0 @@
-#Script Originator: Daniel Suh
-
-#This script creates figure 6
-
-
-library(tidyverse)
-library(magrittr)
-library(vegan)
-library(reshape2)
-library(patchwork)
-library(ggnewscale)
-library(here)
-
-data <- read_csv(here("data/weighted_prev_competence_111220.csv")) #read data
-
-#Get species richness from matrix
-
-community_mat <- data %>% dplyr::select(WetAltID, Month.1, AB2:AB8, AB9, AB20:AB42) %>%
-  mutate(., siteID = paste(WetAltID, Month.1, sep = "_")) %>% distinct() #select data and make siteID column
-
-community_mat <- community_mat[order(community_mat$WetAltID, community_mat$Month.1),] #order by wetland and month
-
-community_mat %<>% arrange(.,WetAltID) %>% dplyr::select(-WetAltID, -Month.1) %>% dplyr::select(siteID, AB2:AB42)
-
-abundances <- community_mat #make abundances df
-
-community_mat %<>% column_to_rownames(., var = "siteID") #make sitexspecies matrix
-
-presence_mat <- ifelse(community_mat>0, 1, 0) #make pres-abs sitexspecies matrix
-
-siteID <- rownames(presence_mat)
-
-richness_mat <- tibble()
-for (i in 1:length(siteID)){
-  richness_mat[i,1] <- siteID[i]
-  richness_mat[i,2] <- sum(presence_mat[i,])
-}
-richness_mat %<>% rename(siteID = 1, richness = 2) #make df with siteIDs and corresponding species richness
-
-#Get species richness for each site and community competence values
-
-tmp <- data %>% dplyr::select(WetAltID, Month.1, cc, Prevalence,AB2:AB8, AB9, AB20:AB42) %>% 
-    distinct() %>% 
-    mutate(., siteID = paste(WetAltID, Month.1, sep = "_"))
-tmp %<>% mutate(.,size = rowSums(.[5:25])) %>% select(-c(AB2:AB42))
-tmp <- tmp[order(tmp$WetAltID, tmp$Month.1),]
-tmp %<>% dplyr::select(-Month.1)
-
-richness_cc <- full_join(richness_mat, tmp, by = "siteID") #make df with siteIDs, richness, and cc
-
-
-
-#J = H'/ln(S)
-#H' = Shannon diversity index
-#S = Species Richness
-
[…]

### 2021-02-19 scripts/figures/fig7_code.R

@@ -1,73 +1,0 @@
-#Daniel Suh and Andrew Park
-#6/3/20
-
-#Figure 7 in rv_cc manuscript
-
-#install_github("thomasp85/patchwork")
-library(tidyverse)
-library(magrittr)
-library(patchwork)
-library(here)
-
-
-data <- read_csv("data/weighted_prev_competence_111220.csv")
-
-prev_cc <- data %>% dplyr::select(WetAltID, Month.1, cc, Prevalence, Month, AB2:AB42, MeanWaterTempPredC)
-prev_cc %<>% dplyr::select(-ABAmb, -ABSal) %>%  mutate(., total = rowSums(.[6:26])) #get totals for community size
-prev_cc %<>% dplyr::select(-AB2:-AB42)
-prev_cc %<>% distinct()
-
-order <- prev_cc[order(prev_cc$WetAltID, prev_cc$Month.1),]
-
-#this removes months out of sequence
-order <- order[-c(4,14,47,90),]
-
-order$lag_cc <- order$cc
-order$lag_size <- order$total
-order$lag_temp <- order$MeanWaterTempPredC
-
-#create new column that includes previous month's value for cc
-for(n in 2:nrow(order)){
-  order$lag_cc[n] <- order$cc[n-1]
-  order$lag_size[n] <- order$total[n-1]
-  order$lag_temp[n] <- order$MeanWaterTempPredC[n-1]
-}
-
-#remove the first entry for each wetland to remove the carryover from the last wetland
-clean <- order %>% group_by(WetAltID) %>% filter(duplicated(WetAltID) | n()==1)
-
-#plot cleaned plot with lag
-cc_corr <- clean %>% ggplot(.,aes(x=lag_cc, y=Prevalence)) +
-  geom_point() +
-  theme_classic() + geom_smooth(method = "lm") +
-  labs(title = "", x = "Community Competence", y = "Prevalence") +
-  ylim(0,0.65)
-
-#correlation test
-cor.test(clean$lag_cc,clean$Prevalence,method="spearman")
-
-
-size_corr <- clean %>% ggplot(.,aes(x=log(lag_size), y=Prevalence)) +
-  geom_point() +
-  theme_classic() + geom_smooth(method = "lm") +
-  labs(title = "", x = "ln(Community Size)", y = "") +
-  theme(axis.text.y = element_blank(), axis.ticks.y=element_blank()) +
-  ylim(0,0.65)
-
[…]

### 2021-02-19 scripts/figures/supp2_code.R
made scripts for supp figs

@@ -1,0 +1,34 @@
+#Daniel Suh and Andrew Park
+#6/3/20
+
+#Figure 2 in rv_cc manuscript
+
+library(tidyverse)
+library(magrittr)
+library(metacom)
+library(here)
+
+
+data <- read_csv("data/weighted_prev_competence_111220.csv")
+
+#make community matrix
+community_mat <- data %>% dplyr::select(WetAltID, Month.1, AB2:AB8, AB9, AB20:AB42) %>%
+  mutate(., siteID = paste(WetAltID, Month.1, sep = "_")) %>% distinct() %>% arrange(.,WetAltID) %>% dplyr::select(-WetAltID, -Month.1) %>% dplyr::select(siteID, AB2:AB42)
+
+community_mat %<>% column_to_rownames(., var = "siteID")
+
+Imagine(t(community_mat), 
+        fill = FALSE, 
+        order = TRUE, 
+        xlab = "", 
+        ylab = "",
+        xline = 5,
+        yline = 4,
+        sitenames = ,
+        speciesnames = ,
+        binary=T)
+
+ordered_matrix <- OrderMatrix(community_mat)
+rownames(ordered_matrix)
+
+results <- Metacommunity(community_mat, binary = T)

### 2021-02-19 scripts/data_format.Rmd
script for making datasets to be used for making figures

---
title: "Data Formatting"
author: "Daniel Suh"
date: "2/19/2021"
output: html_document
---

This Rmd is used for all of the data formatting that is required for use by the scripts that create figures. Scripts for figures should only include minimal code that manipulates the data.

Packages
```{r}
library(tidyverse)
library(magrittr)
library(here)
```

Data
```{r}
here()
data <- read_csv(here("/data/weighted_prev_competence_111220.csv"))
names <- read_csv(here("/data/species_names_ids.csv"))
```

viral loads
```{r}
vl <- data %>% dplyr::select(vl.4:vl.42) %>% distinct() %>% pivot_longer(vl.4:vl.42)
vl$species_code <- as.double(gsub("vl.","",vl$name))
nrow(vl)
vl %<>% full_join(., names, by="species_code")
sum(vl$value>0)
```
17 total species; 12 species with non-zero viral loads

### 2021-02-19 scripts/figures/gradient_analysis.R
renamed script

@@ -1,0 +1,163 @@
+#Daniel Suh and Andrew Park
+#6/3/20
+
+#Figure 3 in rv_cc manuscript
+
+library(tidyverse)
+library(magrittr)
+library(ecodist)
+library(vegan)
+library(here)
+library(ggrepel)
+
+
+data <- read_csv(here("data/weighted_prev_competence_111220.csv"))
+
+
+#make community matrix
+community_mat <- data %>% dplyr::select(WetAltID, Month.1, AB2:AB8, AB9, AB20:AB42)
+community_mat$Month.1 <- gsub("Month", "", community_mat$Month.1) 
+community_mat %<>% mutate(., siteID = paste(WetAltID, Month.1, sep = "_")) %>% 
+  distinct() %>% arrange(.,WetAltID) %>% dplyr::select(-WetAltID, -Month.1) %>% dplyr::select(siteID, AB2:AB42)
+
+community_mat %<>% column_to_rownames(., var = "siteID")
+
+
+#select for environmental variables
+env <- data %>% dplyr::select(WetAltID, Month.1, MeanAirT, MeanWaterTempPredC, DryingScore, CanopyCover, Area, Perimeter) %>% distinct() %>% arrange(.,WetAltID) 
+env$Month.1 <- gsub("Month", "", env$Month.1)
+#make siteID rowname
+env$siteID <- paste(env$WetAltID, env$Month.1, sep = "_")
+env %<>% dplyr::select(-WetAltID, -Month.1)
+#remove duplicated rows (siteID is duplicated but env. variables are not... not sure why this is)
+env <- env[-c(53, 90),]
+env_mat <- env %>% column_to_rownames(., var = "siteID")
+
+#make presence-absence matrix
+presence_mat <- as.data.frame(ifelse(community_mat>0, 1, 0))
+
+site_labels <- rownames(community_mat)
+
+
+pca_env_output=princomp(env_mat, cor=T) # PCA
+summary(pca_env_output)
+biplot(pca_env_output) # Generates a bi-plot with vectors
+
+spec_vec=envfit(pca_env_output$score[,1:2], community_mat) 
+plot(spec_vec, col="blue")
+
+#get loadings for environmental variables
+env_var <- as.data.frame(pca_env_output$loadings[,1:2])
+
+
+site_scores <- data %>% dplyr::select(WetAltID, Month.1, cc) 
+site_scores$Month.1 <- gsub("Month", "", site_scores$Month.1)
[…]

### 2021-02-19 scripts/amphibian_phylo.R
script for making amphibian phylogeny

@@ -1,0 +1,140 @@
+library(tidyverse)
+library(magrittr)
+library(ape)
+library(rotl)
+library(phangorn)
+library(geiger)
+library(phytools)
+library(here)
+
+#files
+#prev_test.csv
+#asup_just_tree.txt
+
+
+#trait1 <- rTraitCont(tree, "OU", theta = 0, alpha=.1, sigma=.01)
+data <- read_csv(here("data/weighted_prev_competence_111220.csv"))
+
+#Plot of RV status over time - ridges with count of ind. positive for RV at each sitextime
+rv <- data %>% dplyr::select(Month,WetAltID,RV.Status)
+rv$Month <- factor(data$Month,levels=c("Feb","Mar","Apr","May","Jun","Jul"))
+
+#looking at rv of species and drawing a line to divide "low" and "high" competent
+rv_species <- data %>% dplyr::select(vl.4:vl.42) %>% distinct()
+rv_species %<>% pivot_longer(.,vl.4:vl.42) %>% filter(value>0)
+rv_species %<>% add_column(log_value = log10(rv_species$value))
+rv_species$name <- factor(rv_species$name, levels = rv_species$name[order(rv_species$log_value)])
+#adjusted names manually to match the resolved names by tnrs_match_names
+rv_species$actual_name <- factor(c("Notophthalmus viridescens", "Scaphiopus holbrookii", "Bufo terrestris", 
+                                   "Acris gryllus", "Pseudacris crucifer", "Hyla gratiosa", "Hyla avivoca", 
+                                   "Pseudacris nigrita", "Pseudacris ornata", "Gastrophryne carolinensis", 
+                                   "Rana clamitans", "Rana sphenocephala"))
+rv_species$Species <- factor(c("Notophthalmus viridescens", "Scaphiopus holbrookii", "Anaxyrus terrestris", 
+                                   "Acris gryllus", "Pseudacris crucifer", "Hyla gratiosa", "Hyla avivoca", 
+                                   "Pseudacris nigrita", "Pseudacris ornata", "Gastrophryne carolinensis", 
+                                   "Lithobates clamitans", "Lithobates sphenocephalus"))
+
[…]

### 2021-02-19 scripts/data_format.Rmd
break up scripts so that all data prep is done in data_format and then sourced by other scripts

@@
 library(magrittr)
+library(vegan)
+
 library(here)
@@
 vl$species_code <- as.double(gsub("vl.","",vl$name))
-nrow(vl)
 vl %<>% full_join(., names, by="species_code")
-sum(vl$value>0)
+vl %<>% mutate(., ln_value = log(value)) %>% mutate(., log10_value = log10(value))
 ```
-17 total species; 12 species with non-zero viral loads
+`r nrow(vl)` total species; `sum(vl$value>0)`species with non-zero viral loads
+
+Transmission Model for two species - primarily used by fig1
+```{r}
+#define function contour(): inputs are parameters and output is values for grid search for R0
+contour <- function(trans_1_min, trans_1_max, trans_3_min, trans_3_max, mort1, mort2, degr) {
+  #expand data
+  data <- expand_grid("trans1" = seq(from = trans_1_min, to = trans_1_max, by = 0.00001),
+                      "trans2" = c(0.0001),
+                      "trans3" = seq(from = trans_3_min, to = trans_3_max, by = 0.00001),
+                      "shed1" = c(rep(1/2, length("trans1"))),
+                      "shed2" = c(rep(1/2, length("trans1"))),
+                      "reco1" = c(rep(1/10, length("trans1"))),
+                      "reco2" = c(rep(1/10, length("trans1"))),
+                      "mort1" = c(mort1), 
+                      "mort2" = c(mort2), 
+                      "degr" = c(degr), 
+                      "birth1" = c(rep(5/3, length("trans1"))),
+                      "birth2" = c(rep(5/3, length("trans1"))))
+
+  #calculate disease-free equilibria for use in next-gen matrix
+  data %<>% add_column(dfe1 = (data$birth1/data$mort1),dfe2 = (data$birth2/data$mort2)) 
+  #make matrix
+  mat <- matrix(1, nrow = 3, ncol = 3)
+
+  m <- nrow(data)
+  #calculate eigenvalue using next-gen matrix. max eigenvalue is R0
+  for(n in 1:m){
+    mat[1] = (data$trans1[n]*data$dfe1[n])/(data$reco1[n]+data$mort1[n])
+    mat[2] = (data$trans1[n]*data$dfe2[n])/(data$reco2[n]+data$mort2[n])
+    mat[3] = (data$shed1[n]/data$degr[n])
+    mat[4] = (data$trans2[n]*data$dfe1[n])/(data$reco1[n]+data$mort1[n])
+    mat[5] = (data$trans2[n]*data$dfe2[n])/(data$reco2[n]+data$mort2[n])
+    mat[6] = (data$shed2[n]/data$degr[n])
+    mat[7] = (data$trans3[n]*data$dfe1[n])/(data$reco1[n]+data$mort1[n])
+    mat[8] = (data$trans3[n]*data$dfe2[n])/(data$reco2[n]+data$mort2[n])
+    mat[9] = 0
+    eigen <- max(eigen(mat)$values)
+    data$eigen[n] <- eigen
+  }
+
+  data$span1 <- 1/data$mort1 #lifespan for species 1
+  data$span2 <- 1/data$mort2 #lifespan for species 2
+  data$half <- log(1/2)/-data$degr #viral half-life
+  data$epid <- data$eigen > 1 #invasion potential (i.e. R0 > 1)
+  data$prop1 <- data$trans1/data$trans2 #proportion of transmission of species 2 over species 1
+  data$prop3 <- data$trans3/data$trans2 #proportion of environmental transmission over species 1
+  data$prop_mort <- data$mort1/data$mort2 #relative mortality for species 1 and 2
+  data$cap1 <- data$birth1/data$mort1 #carrying capacity for species 1
+  data$cap2 <- data$birth2/data$mort2 #carrying capacity for species 2
+  data$cap_prop <- data$cap1/data$cap2 #relative carrying capacities for species 1 and 2
+  data$tot <- data$cap1 + data$cap2 #total abundance for species 1 and 2
+  data$cc <- data$trans1*(data$cap1/data$tot) + data$trans2*(data$cap2/data$tot) #cc defined as transmission of a species multiplied by its relative abundance
+  output <- data %>% dplyr::select(trans1, trans2, prop1, prop3, trans3, eigen, prop_mort, mort1, mort2, degr, half, cap1, cap2, cap_prop, tot, cc) %>% distinct()
+  return(output)
+}
+
+reference <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = 1/45,mort2 = 1/45,degr = 1/2)
+
+composition <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = 1/60,mort2 = 1/30,degr = 1/2)
+
+size <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = c(1/50),mort2 = c(1/50),degr = 1/2)
+
+halflife <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = c(1/40),mort2 = 1/40,degr = c(1/4))
+
+combined <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = c(1/66.6),mort2 = c(1/33.3),degr = c(1/4))
+
+```
+
+community matrix for PCA - primarily used by fig2
+```{r}
+#select for community variables
+community_mat <- data %>% dplyr::select(WetAltID, Month.1, AB2:AB8, AB9, AB20:AB42) 
+community_mat$Month.1 <- gsub("Month", "", community_mat$Month.1)
+community_mat %<>% mutate(., siteID = paste(WetAltID, Month.1, sep = "_")) %>% distinct() %>% arrange(.,WetAltID) %>% dplyr::select(-WetAltID, -Month.1) %>% dplyr::select(siteID, AB2:AB42)
+#rownames
+community_mat %<>% column_to_rownames(., var = "siteID")
+
+#select for environmental variables
+env <- data %>% dplyr::select(WetAltID, Month.1, MeanAirT, MeanWaterTempPredC, DryingScore, CanopyCover, Area, Perimeter) %>% distinct() %>% arrange(.,WetAltID) 
+env$Month.1 <- gsub("Month", "", env$Month.1)
+#make siteID rowname
+env$siteID <- paste(env$WetAltID, env$Month.1, sep = "_")
+env %<>% dplyr::select(-WetAltID, -Month.1)
+#remove duplicated rows (siteID is duplicated but env. variables are not... not sure why this is)
+env <- env[-c(53, 90),]
+env_mat <- env %>% column_to_rownames(., var = "siteID")
+
+#replace NA's with zeros. Probably not the best option but will allow analysis to run for now
+community_mat[is.na(community_mat)] <- 0
+env_mat[is.na(env_mat)] <- 0
+
+#pca on community matrix and gather scores for first two principal components
+pca <- princomp(community_mat, scores = TRUE)
+summary(pca)
+scores <- as.data.frame(pca$scores)
+scores %<>% dplyr::select(Comp.1, Comp.2) %>% rownames_to_column(., var = "siteID")
+
+#plot pca
+biplot(pca)
+
+tmp <- data %>% dplyr::select(WetAltID, Month.1, cc, Month, AB2:AB42) %>% 
+  dplyr::select(-ABAmb, -ABSal) %>%  mutate(., Size = rowSums(.[5:25])) %>% #get totals for community size
+  dplyr::select(-AB2:-AB42) %>%
+  mutate(., siteID = paste(WetAltID, Month.1, sep = "_")) %>%
+  distinct()
+tmp$siteID <- gsub("Month", "", tmp$siteID)
+
+site_scores  <- inner_join(tmp, scores, by = "siteID")
+
+#order months
+site_scores$Month <- factor(site_scores$Month, levels = c("Feb", "Mar", "Apr", "May", "Jun", "Jul"))
+
+#rank order components and plot in order
+site_scores %<>% mutate(pc1Rank=dense_rank(Comp.1))
+```
+
+cc x evenness data - primarily used by fig3
+```{r}
+#Get species richness from matrix
+
+community_mat <- data %>% dplyr::select(WetAltID, Month.1, AB2:AB8, AB9, AB20:AB42) %>%
+  mutate(., siteID = paste(WetAltID, Month.1, sep = "_")) %>% distinct()
+
+community_mat <- community_mat[order(community_mat$WetAltID, community_mat$Month.1),]
+
+community_mat %<>% arrange(.,WetAltID) %>% dplyr::select(-WetAltID, -Month.1) %>% dplyr::select(siteID, AB2:AB42)
+
+abundances <- community_mat
+
+community_mat %<>% column_to_rownames(., var = "siteID")
+
+presence_mat <- ifelse(community_mat>0, 1, 0)
+
+siteID <- rownames(presence_mat)
+
+richness_mat <- tibble()
+for (i in 1:length(siteID)){
+  richness_mat[i,1] <- siteID[i]
+  richness_mat[i,2] <- sum(presence_mat[i,])
+}
+richness_mat %<>% rename(siteID = 1, richness = 2)
+
+order <- richness_mat
+
+#Get species richness for each site and community competence values
+
+tmp <- data %>% dplyr::select(WetAltID, Month.1, cc, Prevalence,AB2:AB8, AB9, AB20:AB42) %>% 
+  distinct() %>% 
+  mutate(., siteID = paste(WetAltID, Month.1, sep = "_"))
+tmp %<>% mutate(.,size = rowSums(.[5:25])) %>% select(-c(AB2:AB42))
+tmp <- tmp[order(tmp$WetAltID, tmp$Month.1),]
+tmp %<>% dplyr::select(-Month.1)
+
+richness_cc <- full_join(tmp, richness_mat, by = "siteID")
+
+#J = H'/ln(S)
+#H' = Shannon diversity index
+#S = Species Richness
+
+h <- vegan::diversity(community_mat, index = "shannon")
+h <- melt(as.matrix(h))
+h %<>% rename(siteID = Var1, h = value) %>% dplyr::select(-Var2)
+
+evenness <- richness_cc %>% mutate(lnS = log(richness)) %>% rowwise()
+
+evenness %<>% full_join(h, evenness, by = "siteID")
+
+evenness %<>% mutate(J = h/lnS) %>% rowwise()
+
+#create new df for ordered sites to test lagged richness and evenness on prevalence
+lag_evenness <- evenness[-c(4,14,47,90),]
+lag_evenness$lag_richness <- c(0)
+lag_evenness$lag_J <- c(0)
+lag_evenness$lag_cc <- c(0)
+lag_evenness$lag_size <- c(0)
+for(n in 2:91){
+  lag_evenness$lag_richness[n] <- lag_evenness$richness[n-1]
+  lag_evenness$lag_J[n] <- lag_evenness$J[n-1]
+  lag_evenness$lag_cc[n] <- lag_evenness$cc[n-1]
+  lag_evenness$lag_size[n] <- lag_evenness$size[n-1]
+}
+
+#remove the first entry for each wetland to remove the carryover from the last wetland
+lag_evenness %<>% group_by(WetAltID) %>% filter(duplicated(WetAltID) | n()==1)
+
+#fit two lines to cc_evenness_prev plot
+lag_evenness$high <- c(0)
+lag_evenness$low <- c(0)
+for (i in 1:nrow(lag_evenness)){
+  if (lag_evenness$lag_J[i] > 0.6){
+    lag_evenness$high[i] <- 1
+    lag_evenness$low[i] <- 1
+  }
+  else if (lag_evenness$lag_cc[i] < 45000){
+    lag_evenness$low[i] <- 1
+  }
+  else if (lag_evenness$lag_cc[i] > 45000){
+    lag_evenness$high[i] <- 1
+  }
+  else {
+    lag_evenness$high[i] <- 0
+    lag_evenness$low[i] <- 0
+  }
+}
+low_cc <- lag_evenness %>% filter(low==1)
+cor.test(x=low_cc$lag_J, y=low_cc$lag_cc,method = "spearman")
+high_cc <- lag_evenness %>% filter(high==1)
+cor.test(x=high_cc$lag_J, y=high_cc$lag_cc,method = "spearman")
+```
+
+data for bar plots - primarily used by fig4
+```{r}
+
+#Get species richness from matrix
+
+community_mat <- data %>% dplyr::select(WetAltID, Month.1, AB2:AB8, AB9, AB20:AB42) %>%
+  mutate(., siteID = paste(WetAltID, Month.1, sep = "_")) %>% distinct() #select data and make siteID column
+
+community_mat <- community_mat[order(community_mat$WetAltID, community_mat$Month.1),] #order by wetland and month
+
+community_mat %<>% arrange(.,WetAltID) %>% dplyr::select(-WetAltID, -Month.1) %>% dplyr::select(siteID, AB2:AB42)
+
+abundances <- community_mat #make abundances df
+
+community_mat %<>% column_to_rownames(., var = "siteID") #make sitexspecies matrix
+
+presence_mat <- ifelse(community_mat>0, 1, 0) #make pres-abs sitexspecies matrix
+
+siteID <- rownames(presence_mat)
+
+richness_mat <- tibble()
+for (i in 1:length(siteID)){
+  richness_mat[i,1] <- siteID[i]
+  richness_mat[i,2] <- sum(presence_mat[i,])
+}
+richness_mat %<>% rename(siteID = 1, richness = 2) #make df with siteIDs and corresponding species richness
+
+#Get species richness for each site and community competence values
+
+tmp <- data %>% dplyr::select(WetAltID, Month.1, cc, Prevalence,AB2:AB8, AB9, AB20:AB42) %>% 
+    distinct() %>% 
+    mutate(., siteID = paste(WetAltID, Month.1, sep = "_"))
+tmp %<>% mutate(.,size = rowSums(.[5:25])) %>% select(-c(AB2:AB42))
+tmp <- tmp[order(tmp$WetAltID, tmp$Month.1),]
+tmp %<>% dplyr::select(-Month.1)
+
+richness_cc <- full_join(richness_mat, tmp, by = "siteID") #make df with siteIDs, richness, and cc
+
+#J = H'/ln(S)
+#H' = Shannon diversity index
+#S = Species Richness
+
+h <- vegan::diversity(community_mat, index = "shannon") #calculate shannons index for each site-month
+h <- melt(as.matrix(h)) #melt into df
+h %<>% rename(siteID = Var1, h = value) %>% dplyr::select(-Var2) #get df with shannons index for each site-month
+
+evenness <- richness_cc %>% mutate(lnS = log(richness)) %>% rowwise() 
+evenness %<>% full_join(h, evenness, by = "siteID")
+evenness %<>% mutate(J = h/lnS) %>% rowwise() #calculate Pielou's J
+
+#test for prevalence-richness relationship
+
+evenness %>% ggplot(.,aes(x=richness, y=Prevalence))+
+  geom_smooth(method="lm")
+cor.test(evenness$richness,evenness$Prevalence,method="spearman")
+
+#create new df for ordered sites to test lagged richness and evenness on prevalence
+lag_evenness <- evenness[-c(4,14,47,90),] #remove site-months that are out of sequence
+lag_evenness$lag_richness <- c(0)
+lag_evenness$lag_J <- c(0)
+lag_evenness$lag_cc <- c(0)
+lag_evenness$lag_size <- c(0)
+for(n in 2:91){ #lag values by one month
+  lag_evenness$lag_richness[n] <- lag_evenness$richness[n-1]
+  lag_evenness$lag_J[n] <- lag_evenness$J[n-1]
+  lag_evenness$lag_cc[n] <- lag_evenness$cc[n-1]
+  lag_evenness$lag_size[n] <- lag_evenness$size[n-1]
+}
+#remove the first entry for each wetland to remove the carryover from the last wetland
+lag_evenness %<>% group_by(WetAltID) %>% filter(duplicated(WetAltID) | n()==1)
+#lag_evenness %<>% select(siteID, lag_richness:lag_size)
+
+lag_evenness %>% ggplot(.,aes(x=lag_richness, y=Prevalence))+
+  geom_smooth(method="lm")
+cor.test(lag_evenness$lag_richness,lag_evenness$Prevalence,method="spearman")
+
+#barplots comparing community composition
+
+evenness %<>% select(-WetAltID)
+abundances %<>% left_join(.,evenness, by = "siteID")
+lag_evenness %<>% select(siteID, lag_cc, lag_richness, lag_J, lag_size)
+abundances %<>% inner_join(.,lag_evenness, by = "siteID")
+abundances <- mutate(abundances, ABHigh = (AB9 + AB21 + AB26 + AB42)) 
+abundances <- mutate(abundances, ABLow = (AB2 + AB3 + AB4 + AB5 + AB6 + AB8 + AB20 + AB24 + AB27 + AB28 + AB29 + AB31 + AB34 + AB35 + AB38 + AB39 + AB41))
+abundances <- mutate(abundances, total = ABHigh + ABLow)
+#final df for plots. includes data for abundance, prevalence, cc, diversity measure, lagged values
+
+abundances$siteID <- reorder(abundances$siteID, -abundances$total) #order by total community size
+#abundances$siteID <- reorder(abundances$siteID, -abundances$cc) #order by community competence
+#abundances$siteID <- reorder(abundances$siteID, -abundances$J) #order by Pielou's J
+
+```
+
+correlations - primarily used by fig5
+```{r}
+prev_cc <- data %>% dplyr::select(WetAltID, Month.1, cc, Prevalence, Month, AB2:AB42, MeanWaterTempPredC)
+prev_cc %<>% dplyr::select(-ABAmb, -ABSal) %>%  mutate(., total = rowSums(.[6:26])) #get totals for community size
+prev_cc %<>% dplyr::select(-AB2:-AB42)
+prev_cc %<>% distinct()
+
+order <- prev_cc[order(prev_cc$WetAltID, prev_cc$Month.1),]
+
+#this removes months out of sequence
+order <- order[-c(4,14,47,90),]
+
+order$lag_cc <- order$cc
+order$lag_size <- order$total
+order$lag_temp <- order$MeanWaterTempPredC
+
+#create new column that includes previous month's value for cc
+for(n in 2:nrow(order)){
+  order$lag_cc[n] <- order$cc[n-1]
+  order$lag_size[n] <- order$total[n-1]
+  order$lag_temp[n] <- order$MeanWaterTempPredC[n-1]
+}
+
+#remove the first entry for each wetland to remove the carryover from the last wetland
+clean <- order %>% group_by(WetAltID) %>% filter(duplicated(WetAltID) | n()==1)
+
+```

### 2021-02-19 data_format.R

@@ -1,0 +1,20 @@
+## -----------------------------------------------------------------------------------
+library(tidyverse)
+library(magrittr)
+library(here)
+
+
+## -----------------------------------------------------------------------------------
+here()
+data <- read_csv(here("/data/weighted_prev_competence_111220.csv"))
+names <- read_csv(here("/data/species_names_ids.csv"))
+
+
+## -----------------------------------------------------------------------------------
+vl <- data %>% dplyr::select(vl.4:vl.42) %>% distinct() %>% pivot_longer(vl.4:vl.42)
+vl$species_code <- as.double(gsub("vl.","",vl$name))
+nrow(vl)
+vl %<>% full_join(., names, by="species_code")
+vl %<>% mutate(., ln_value = log(value)) %>% mutate(., log10_value = log10(value))
+sum(vl$value>0)
+

### 2021-02-19 scripts/figures/fig1_code.R

@@ -11,62 +11,6 @@ library(patchwork)
 library(ggnewscale)
 
 
-#define function contour(): inputs are parameters and output is values for grid search for R0
-contour <- function(trans_1_min, trans_1_max, trans_3_min, trans_3_max, mort1, mort2, degr) {
-  #expand data
-  data <- expand_grid("trans1" = seq(from = trans_1_min, to = trans_1_max, by = 0.00001),
-                      "trans2" = c(0.0001),
-                      "trans3" = seq(from = trans_3_min, to = trans_3_max, by = 0.00001),
-                      "shed1" = c(rep(1/2, length("trans1"))),
-                      "shed2" = c(rep(1/2, length("trans1"))),
-                      "reco1" = c(rep(1/10, length("trans1"))),
-                      "reco2" = c(rep(1/10, length("trans1"))),
-                      "mort1" = c(mort1), 
-                      "mort2" = c(mort2), 
-                      "degr" = c(degr), 
-                      "birth1" = c(rep(5/3, length("trans1"))),
-                      "birth2" = c(rep(5/3, length("trans1"))))
-  
-  #calculate disease-free equilibria for use in next-gen matrix
-  data %<>% add_column(dfe1 = (data$birth1/data$mort1),dfe2 = (data$birth2/data$mort2)) 
-  #make matrix
-  mat <- matrix(1, nrow = 3, ncol = 3)
-  
-  
-  m <- nrow(data)
-  #calculate eigenvalue using next-gen matrix. max eigenvalue is R0
-  for(n in 1:m){
-    mat[1] = (data$trans1[n]*data$dfe1[n])/(data$reco1[n]+data$mort1[n])
-    mat[2] = (data$trans1[n]*data$dfe2[n])/(data$reco2[n]+data$mort2[n])
-    mat[3] = (data$shed1[n]/data$degr[n])
-    mat[4] = (data$trans2[n]*data$dfe1[n])/(data$reco1[n]+data$mort1[n])
-    mat[5] = (data$trans2[n]*data$dfe2[n])/(data$reco2[n]+data$mort2[n])
-    mat[6] = (data$shed2[n]/data$degr[n])
-    mat[7] = (data$trans3[n]*data$dfe1[n])/(data$reco1[n]+data$mort1[n])
[…]

### 2021-02-19 scripts/figures/fig2_code.R

@@ -1,7 +1,7 @@
 #Daniel Suh and Andrew Park
 #6/3/20
 
-#Figure 4 in rv_cc manuscript
+#Figure 2 in rv_cc manuscript
 
 library(tidyverse)
 library(magrittr)
@@ -9,56 +9,7 @@ library(ggforce)
 library(here)
 
 
-data <- read_csv("data/weighted_prev_competence_111220.csv")
 
-
-
-
-#select for community variables
-community_mat <- data %>% dplyr::select(WetAltID, Month.1, AB2:AB8, AB9, AB20:AB42) 
-community_mat$Month.1 <- gsub("Month", "", community_mat$Month.1)
-community_mat %<>% mutate(., siteID = paste(WetAltID, Month.1, sep = "_")) %>% distinct() %>% arrange(.,WetAltID) %>% dplyr::select(-WetAltID, -Month.1) %>% dplyr::select(siteID, AB2:AB42)
-#rownames
-community_mat %<>% column_to_rownames(., var = "siteID")
-
-#select for environmental variables
-env <- data %>% dplyr::select(WetAltID, Month.1, MeanAirT, MeanWaterTempPredC, DryingScore, CanopyCover, Area, Perimeter) %>% distinct() %>% arrange(.,WetAltID) 
-env$Month.1 <- gsub("Month", "", env$Month.1)
-#make siteID rowname
-env$siteID <- paste(env$WetAltID, env$Month.1, sep = "_")
-env %<>% dplyr::select(-WetAltID, -Month.1)
-#remove duplicated rows (siteID is duplicated but env. variables are not... not sure why this is)
-env <- env[-c(53, 90),]
-env_mat <- env %>% column_to_rownames(., var = "siteID")
-
-#replace NA's with zeros. Probably not the best option but will allow analysis to run for now
-community_mat[is.na(community_mat)] <- 0
-env_mat[is.na(env_mat)] <- 0
-
-#pca on community matrix and gather scores for first two principal components
-pca <- princomp(community_mat, scores = TRUE)
-summary(pca)
-scores <- as.data.frame(pca$scores)
-scores %<>% dplyr::select(Comp.1, Comp.2) %>% rownames_to_column(., var = "siteID")
-
-#plot pca
-biplot(pca)
-
-
-tmp <- data %>% dplyr::select(WetAltID, Month.1, cc, Month, AB2:AB42) %>% 
[…]

### 2021-02-19 scripts/figures/fig3_code.R

@@ -1,6 +1,6 @@
 #Script Originator: Daniel Suh
 
-#This script creates figure 5
+#This script creates figure 3
 
 
 library(tidyverse)
@@ -11,100 +11,7 @@ library(patchwork)
 library(ggnewscale)
 library(here)
 
-data <- read_csv(here("data/weighted_prev_competence_111220.csv"))
 
-#Get species richness from matrix
-
-community_mat <- data %>% dplyr::select(WetAltID, Month.1, AB2:AB8, AB9, AB20:AB42) %>%
-  mutate(., siteID = paste(WetAltID, Month.1, sep = "_")) %>% distinct() 
-
-community_mat <- community_mat[order(community_mat$WetAltID, community_mat$Month.1),]
-
-community_mat %<>% arrange(.,WetAltID) %>% dplyr::select(-WetAltID, -Month.1) %>% dplyr::select(siteID, AB2:AB42)
-
-abundances <- community_mat
-
-community_mat %<>% column_to_rownames(., var = "siteID")
-
-presence_mat <- ifelse(community_mat>0, 1, 0)
-
-siteID <- rownames(presence_mat)
-
-richness_mat <- tibble()
-for (i in 1:length(siteID)){
-  richness_mat[i,1] <- siteID[i]
-  richness_mat[i,2] <- sum(presence_mat[i,])
-}
-richness_mat %<>% rename(siteID = 1, richness = 2)
-
-order <- richness_mat
-
-#Get species richness for each site and community competence values
-
-tmp <- data %>% dplyr::select(WetAltID, Month.1, cc, Prevalence,AB2:AB8, AB9, AB20:AB42) %>% 
-  distinct() %>% 
-  mutate(., siteID = paste(WetAltID, Month.1, sep = "_"))
-tmp %<>% mutate(.,size = rowSums(.[5:25])) %>% select(-c(AB2:AB42))
-tmp <- tmp[order(tmp$WetAltID, tmp$Month.1),]
-tmp %<>% dplyr::select(-Month.1)
-
-richness_cc <- full_join(tmp, richness_mat, by = "siteID")
-
-
-
-#J = H'/ln(S)
-#H' = Shannon diversity index
-#S = Species Richness
-
-h <- vegan::diversity(community_mat, index = "shannon")
-h <- melt(as.matrix(h))
[…]

### 2021-02-19 scripts/figures/fig4_code.R

@@ -1,6 +1,6 @@
 #Script Originator: Daniel Suh
 
-#This script creates figure 6
+#This script creates figure 4
 
 
 library(tidyverse)
@@ -11,103 +11,6 @@ library(patchwork)
 library(ggnewscale)
 library(here)
 
-data <- read_csv(here("data/weighted_prev_competence_111220.csv")) #read data
-
-#Get species richness from matrix
-
-community_mat <- data %>% dplyr::select(WetAltID, Month.1, AB2:AB8, AB9, AB20:AB42) %>%
-  mutate(., siteID = paste(WetAltID, Month.1, sep = "_")) %>% distinct() #select data and make siteID column
-
-community_mat <- community_mat[order(community_mat$WetAltID, community_mat$Month.1),] #order by wetland and month
-
-community_mat %<>% arrange(.,WetAltID) %>% dplyr::select(-WetAltID, -Month.1) %>% dplyr::select(siteID, AB2:AB42)
-
-abundances <- community_mat #make abundances df
-
-community_mat %<>% column_to_rownames(., var = "siteID") #make sitexspecies matrix
-
-presence_mat <- ifelse(community_mat>0, 1, 0) #make pres-abs sitexspecies matrix
-
-siteID <- rownames(presence_mat)
-
-richness_mat <- tibble()
-for (i in 1:length(siteID)){
-  richness_mat[i,1] <- siteID[i]
-  richness_mat[i,2] <- sum(presence_mat[i,])
-}
-richness_mat %<>% rename(siteID = 1, richness = 2) #make df with siteIDs and corresponding species richness
-
-#Get species richness for each site and community competence values
-
-tmp <- data %>% dplyr::select(WetAltID, Month.1, cc, Prevalence,AB2:AB8, AB9, AB20:AB42) %>% 
-    distinct() %>% 
-    mutate(., siteID = paste(WetAltID, Month.1, sep = "_"))
-tmp %<>% mutate(.,size = rowSums(.[5:25])) %>% select(-c(AB2:AB42))
-tmp <- tmp[order(tmp$WetAltID, tmp$Month.1),]
-tmp %<>% dplyr::select(-Month.1)
-
-richness_cc <- full_join(richness_mat, tmp, by = "siteID") #make df with siteIDs, richness, and cc
-
-
-
-#J = H'/ln(S)
-#H' = Shannon diversity index
-#S = Species Richness
-
[…]

### 2021-02-19 scripts/figures/fig5_code.R

@@ -1,7 +1,7 @@
 #Daniel Suh and Andrew Park
 #6/3/20
 
-#Figure 7 in rv_cc manuscript
+#Figure 5 in rv_cc manuscript
 
 #install_github("thomasp85/patchwork")
 library(tidyverse)
@@ -12,29 +12,6 @@ library(here)
 
 data <- read_csv("data/weighted_prev_competence_111220.csv")
 
-prev_cc <- data %>% dplyr::select(WetAltID, Month.1, cc, Prevalence, Month, AB2:AB42, MeanWaterTempPredC)
-prev_cc %<>% dplyr::select(-ABAmb, -ABSal) %>%  mutate(., total = rowSums(.[6:26])) #get totals for community size
-prev_cc %<>% dplyr::select(-AB2:-AB42)
-prev_cc %<>% distinct()
-
-order <- prev_cc[order(prev_cc$WetAltID, prev_cc$Month.1),]
-
-#this removes months out of sequence
-order <- order[-c(4,14,47,90),]
-
-order$lag_cc <- order$cc
-order$lag_size <- order$total
-order$lag_temp <- order$MeanWaterTempPredC
-
-#create new column that includes previous month's value for cc
-for(n in 2:nrow(order)){
-  order$lag_cc[n] <- order$cc[n-1]
-  order$lag_size[n] <- order$total[n-1]
-  order$lag_temp[n] <- order$MeanWaterTempPredC[n-1]
-}
-
-#remove the first entry for each wetland to remove the carryover from the last wetland
-clean <- order %>% group_by(WetAltID) %>% filter(duplicated(WetAltID) | n()==1)
 
 #plot cleaned plot with lag
 cc_corr <- clean %>% ggplot(.,aes(x=lag_cc, y=Prevalence)) +

### 2021-02-19 scripts/figures/supp1_code.R

@@ -1,0 +1,12 @@
+## Script originator: Daniel Suh
+## Date created: Feb. 19, 2021
+
+#This script creates supplementary figure 1: visual of the estimated viral load for each species
+
+library(tidyverse)
+library(magrittr)
+library(here)
+
+
+source(knitr::purl(here("/scripts/data_format.Rmd"), quiet=TRUE))
+

### 2021-02-19 scripts/data_format.Rmd
cleaning up code

@@
 library(vegan)
-
+library(reshape)
 library(here)
@@
 
+community summary - includes each site_month as rows and variables cc, size, prevalenceetc... as columns
+```{r}
+comm_summ <- data %>% dplyr::select(WetAltID, Month.1, cc, Month, AB2:AB42, Prevalence) %>% 
+  dplyr::select(-ABAmb, -ABSal) %>%  mutate(., Size = rowSums(.[5:25])) %>% #get totals for community size
+  dplyr::select(-AB2:-AB42) %>%
+  mutate(., siteID = paste(WetAltID, Month.1, sep = "_")) %>%
+  distinct()
+comm_summ$siteID <- gsub("Month", "", comm_summ$siteID)
+```
+
 viral loads
@@
 `r nrow(vl)` total species; `sum(vl$value>0)`species with non-zero viral loads
+
+community matrices - abundance matrix and presence-absence matrix
+```{r}
+community_mat <- data %>% dplyr::select(WetAltID, Month.1, AB2:AB8, AB9, AB20:AB42) 
+community_mat$Month <- gsub("Month", "", community_mat$Month.1)
+community_mat %<>% mutate(., siteID = paste(WetAltID, Month, sep = "_")) %>% distinct() %>% arrange(.,WetAltID) %>% dplyr::select(siteID, AB2:AB42)
+
+#make abundance and pres/abs matrices with rownames as siteIDs
+abundance_mat <- community_mat %>% column_to_rownames(., var = "siteID")
+presence_mat <- as.data.frame(ifelse(abundance_mat>0, 1, 0))
+```
+
+environmental variables matrix
+```{r}
+#select for environmental variables
+env <- data %>% dplyr::select(WetAltID, Month.1, MeanAirT, MeanWaterTempPredC, DryingScore, CanopyCover, Area, Perimeter) %>% distinct() %>% arrange(.,WetAltID) 
+env$Month <- gsub("Month", "", env$Month.1)
+#make siteID rowname
+env %<>% mutate(., siteID = paste(env$WetAltID, env$Month, sep = "_")) %>% distinct() %>% arrange(.,WetAltID) %>% dplyr::select(siteID, MeanAirT:Perimeter)
+
+#need to remove duplicated rows and determine what to do with NA's
+env_mat <- env %>% column_to_rownames(., var = "siteID")
+```
+environmental variables are `r colnames(env_mat)`
+
+site 13, month 1 and site 19, month 4 are duplicated. not sure why
+
+richness values for each site
+```{r}
+#Get species richness from pres/abs matrix
+richness <- tibble()
+for (i in 1:length(rownames(presence_mat))){
+  richness[i,1] <- rownames(presence_mat)[i]
+  richness[i,2] <- sum(presence_mat[i,])
+}
+richness %<>% rename(siteID = 1, richness = 2)
+```
 
@@
 
-community matrix for PCA - primarily used by fig2
+PCA for community matrix - primarily used by fig2
 ```{r}
-#select for community variables
-community_mat <- data %>% dplyr::select(WetAltID, Month.1, AB2:AB8, AB9, AB20:AB42) 
-community_mat$Month.1 <- gsub("Month", "", community_mat$Month.1)
-community_mat %<>% mutate(., siteID = paste(WetAltID, Month.1, sep = "_")) %>% distinct() %>% arrange(.,WetAltID) %>% dplyr::select(-WetAltID, -Month.1) %>% dplyr::select(siteID, AB2:AB42)
-#rownames
-community_mat %<>% column_to_rownames(., var = "siteID")
+#pca on community matrix and gather scores for first two principal components
+pca <- princomp(abundance_mat, scores = TRUE)
+summary(pca)
+pca_scores <- as.data.frame(pca$scores)
+pca_scores %<>% dplyr::select(Comp.1, Comp.2) %>% rownames_to_column(., var = "siteID")
 
-#select for environmental variables
-env <- data %>% dplyr::select(WetAltID, Month.1, MeanAirT, MeanWaterTempPredC, DryingScore, CanopyCover, Area, Perimeter) %>% distinct() %>% arrange(.,WetAltID) 
-env$Month.1 <- gsub("Month", "", env$Month.1)
-#make siteID rowname
-env$siteID <- paste(env$WetAltID, env$Month.1, sep = "_")
-env %<>% dplyr::select(-WetAltID, -Month.1)
-#remove duplicated rows (siteID is duplicated but env. variables are not... not sure why this is)
-env <- env[-c(53, 90),]
-env_mat <- env %>% column_to_rownames(., var = "siteID")
-
-#replace NA's with zeros. Probably not the best option but will allow analysis to run for now
-community_mat[is.na(community_mat)] <- 0
-env_mat[is.na(env_mat)] <- 0
-
-#pca on community matrix and gather scores for first two principal components
-pca <- princomp(community_mat, scores = TRUE)
-summary(pca)
-scores <- as.data.frame(pca$scores)
-scores %<>% dplyr::select(Comp.1, Comp.2) %>% rownames_to_column(., var = "siteID")
-
-#plot pca
-biplot(pca)
-
-tmp <- data %>% dplyr::select(WetAltID, Month.1, cc, Month, AB2:AB42) %>% 
-  dplyr::select(-ABAmb, -ABSal) %>%  mutate(., Size = rowSums(.[5:25])) %>% #get totals for community size
-  dplyr::select(-AB2:-AB42) %>%
-  mutate(., siteID = paste(WetAltID, Month.1, sep = "_")) %>%
-  distinct()
-tmp$siteID <- gsub("Month", "", tmp$siteID)
-
-site_scores  <- inner_join(tmp, scores, by = "siteID")
+site_scores  <- full_join(comm_summ, pca_scores, by = "siteID")
 
@@
 ```{r}
-#Get species richness from matrix
-
-community_mat <- data %>% dplyr::select(WetAltID, Month.1, AB2:AB8, AB9, AB20:AB42) %>%
-  mutate(., siteID = paste(WetAltID, Month.1, sep = "_")) %>% distinct()
-
-community_mat <- community_mat[order(community_mat$WetAltID, community_mat$Month.1),]
-
-community_mat %<>% arrange(.,WetAltID) %>% dplyr::select(-WetAltID, -Month.1) %>% dplyr::select(siteID, AB2:AB42)
-
-abundances <- community_mat
-
-community_mat %<>% column_to_rownames(., var = "siteID")
-
-presence_mat <- ifelse(community_mat>0, 1, 0)
-
-siteID <- rownames(presence_mat)
-
-richness_mat <- tibble()
-for (i in 1:length(siteID)){
-  richness_mat[i,1] <- siteID[i]
-  richness_mat[i,2] <- sum(presence_mat[i,])
-}
-richness_mat %<>% rename(siteID = 1, richness = 2)
-
-order <- richness_mat
-
-#Get species richness for each site and community competence values
-
-tmp <- data %>% dplyr::select(WetAltID, Month.1, cc, Prevalence,AB2:AB8, AB9, AB20:AB42) %>% 
-  distinct() %>% 
-  mutate(., siteID = paste(WetAltID, Month.1, sep = "_"))
-tmp %<>% mutate(.,size = rowSums(.[5:25])) %>% select(-c(AB2:AB42))
-tmp <- tmp[order(tmp$WetAltID, tmp$Month.1),]
-tmp %<>% dplyr::select(-Month.1)
-
-richness_cc <- full_join(tmp, richness_mat, by = "siteID")
+richness_cc <- full_join(comm_summ, richness_mat, by = "siteID")
 
@@
 
-h <- vegan::diversity(community_mat, index = "shannon")
-h <- melt(as.matrix(h))
-h %<>% rename(siteID = Var1, h = value) %>% dplyr::select(-Var2)
+shannon <- vegan::diversity(abundance_mat, index = "shannon")
+shannon <- melt(as.matrix(shannon))
+shannon %<>% mutate(siteID = X1, h = value) %>% dplyr::select(siteID, h)
 
@@
 
-evenness %<>% full_join(h, evenness, by = "siteID")
+evenness %<>% full_join(shannon, evenness, by = "siteID")
 
 evenness %<>% mutate(J = h/lnS) %>% rowwise()
-
+evenness <- evenness[order(evenness$WetAltID),]
 #create new df for ordered sites to test lagged richness and evenness on prevalence
-lag_evenness <- evenness[-c(4,14,47,90),]
+lag_evenness <- evenness[-c(4,14,47,90),] #brute force remove site_months out of sequence
 lag_evenness$lag_richness <- c(0)
@@
 
-#fit two lines to cc_evenness_prev plot
-lag_evenness$high <- c(0)
-lag_evenness$low <- c(0)
-for (i in 1:nrow(lag_evenness)){
-  if (lag_evenness$lag_J[i] > 0.6){
-    lag_evenness$high[i] <- 1
-    lag_evenness$low[i] <- 1
-  }
-  else if (lag_evenness$lag_cc[i] < 45000){
-    lag_evenness$low[i] <- 1
-  }
-  else if (lag_evenness$lag_cc[i] > 45000){
-    lag_evenness$high[i] <- 1
-  }
-  else {
-    lag_evenness$high[i] <- 0
-    lag_evenness$low[i] <- 0
-  }
-}
-low_cc <- lag_evenness %>% filter(low==1)
-cor.test(x=low_cc$lag_J, y=low_cc$lag_cc,method = "spearman")
-high_cc <- lag_evenness %>% filter(high==1)
-cor.test(x=high_cc$lag_J, y=high_cc$lag_cc,method = "spearman")
 ```

### 2021-02-19 scripts/figures/fig3_code.R

@@ -11,7 +11,30 @@ library(patchwork)
 library(ggnewscale)
 library(here)
 
+#fit two lines to cc_evenness_prev plot
+lag_evenness$high <- c(0)
+lag_evenness$low <- c(0)
+for (i in 1:nrow(lag_evenness)){
+  if (lag_evenness$lag_J[i] > 0.6){
+    lag_evenness$high[i] <- 1
+    lag_evenness$low[i] <- 1
+  }
+  else if (lag_evenness$lag_cc[i] < 45000){
+    lag_evenness$low[i] <- 1
+  }
+  else if (lag_evenness$lag_cc[i] > 45000){
+    lag_evenness$high[i] <- 1
+  }
+  else {
+    lag_evenness$high[i] <- 0
+    lag_evenness$low[i] <- 0
+  }
+}
 
+# low_cc <- lag_evenness %>% filter(low==1)
+# cor.test(x=low_cc$lag_J, y=low_cc$lag_cc,method = "spearman")
+# high_cc <- lag_evenness %>% filter(high==1)
+# cor.test(x=high_cc$lag_J, y=high_cc$lag_cc,method = "spearman")
 
 p1 <- evenness %>% remove_missing() %>% ggplot(aes(x=richness, y=cc)) + 
   geom_smooth(method ="lm") +

### 2021-02-19 scripts/data_format.Rmd
clean up code

@@
 ```{r}
-comm_summ <- data %>% dplyr::select(WetAltID, Month.1, cc, Month, AB2:AB42, Prevalence) %>% 
-  dplyr::select(-ABAmb, -ABSal) %>%  mutate(., Size = rowSums(.[5:25])) %>% #get totals for community size
+comm_summ <- data %>% dplyr::select(WetAltID, Month.1, Month, cc, Month, AB2:AB42, Prevalence, MeanWaterTempPredC) %>% 
+  dplyr::select(-ABAmb, -ABSal) %>%  mutate(., size = rowSums(.[5:25])) %>% #get totals for community size
   dplyr::select(-AB2:-AB42) %>%
@@
 
-environmental variables matrix
+environmental variables matrix - needs to be fixed!!
 ```{r}
@@
 #need to remove duplicated rows and determine what to do with NA's
-env_mat <- env %>% column_to_rownames(., var = "siteID")
+#env_mat <- env %>% column_to_rownames(., var = "siteID")
 ```
@@
 #Get species richness from pres/abs matrix
-richness <- tibble()
-for (i in 1:length(rownames(presence_mat))){
+richness <- tibble(siteID="", richness=1:nrow(presence_mat))
+for (i in 1:nrow(presence_mat)){
   richness[i,1] <- rownames(presence_mat)[i]
@@
 }
-richness %<>% rename(siteID = 1, richness = 2)
 ```
@@
   #calculate disease-free equilibria for use in next-gen matrix
-  data %<>% add_column(dfe1 = (data$birth1/data$mort1),dfe2 = (data$birth2/data$mort2)) 
+  data %<>% add_column(dfe1 = (data$birth1/data$mort1),dfe2 = (data$birth2/data$mort2), eigen = NA) 
   #make matrix
@@
   m <- nrow(data)
+
   #calculate eigenvalue using next-gen matrix. max eigenvalue is R0
@@
 ```{r}
-richness_cc <- full_join(comm_summ, richness_mat, by = "siteID")
+richness_cc <- full_join(comm_summ, richness, by = "siteID")
 
@@
 lag_evenness %<>% group_by(WetAltID) %>% filter(duplicated(WetAltID) | n()==1)
-
 ```
@@
 ```{r}
-
-#Get species richness from matrix
-
-community_mat <- data %>% dplyr::select(WetAltID, Month.1, AB2:AB8, AB9, AB20:AB42) %>%
-  mutate(., siteID = paste(WetAltID, Month.1, sep = "_")) %>% distinct() #select data and make siteID column
-
-community_mat <- community_mat[order(community_mat$WetAltID, community_mat$Month.1),] #order by wetland and month
-
-community_mat %<>% arrange(.,WetAltID) %>% dplyr::select(-WetAltID, -Month.1) %>% dplyr::select(siteID, AB2:AB42)
-
-abundances <- community_mat #make abundances df
-
-community_mat %<>% column_to_rownames(., var = "siteID") #make sitexspecies matrix
-
-presence_mat <- ifelse(community_mat>0, 1, 0) #make pres-abs sitexspecies matrix
-
-siteID <- rownames(presence_mat)
-
-richness_mat <- tibble()
-for (i in 1:length(siteID)){
-  richness_mat[i,1] <- siteID[i]
-  richness_mat[i,2] <- sum(presence_mat[i,])
-}
-richness_mat %<>% rename(siteID = 1, richness = 2) #make df with siteIDs and corresponding species richness
-
-#Get species richness for each site and community competence values
-
-tmp <- data %>% dplyr::select(WetAltID, Month.1, cc, Prevalence,AB2:AB8, AB9, AB20:AB42) %>% 
-    distinct() %>% 
-    mutate(., siteID = paste(WetAltID, Month.1, sep = "_"))
-tmp %<>% mutate(.,size = rowSums(.[5:25])) %>% select(-c(AB2:AB42))
-tmp <- tmp[order(tmp$WetAltID, tmp$Month.1),]
-tmp %<>% dplyr::select(-Month.1)
-
-richness_cc <- full_join(richness_mat, tmp, by = "siteID") #make df with siteIDs, richness, and cc
-
-#J = H'/ln(S)
-#H' = Shannon diversity index
-#S = Species Richness
-
-h <- vegan::diversity(community_mat, index = "shannon") #calculate shannons index for each site-month
-h <- melt(as.matrix(h)) #melt into df
-h %<>% rename(siteID = Var1, h = value) %>% dplyr::select(-Var2) #get df with shannons index for each site-month
-
-evenness <- richness_cc %>% mutate(lnS = log(richness)) %>% rowwise() 
-evenness %<>% full_join(h, evenness, by = "siteID")
-evenness %<>% mutate(J = h/lnS) %>% rowwise() #calculate Pielou's J
-
-#test for prevalence-richness relationship
-
-evenness %>% ggplot(.,aes(x=richness, y=Prevalence))+
-  geom_smooth(method="lm")
-cor.test(evenness$richness,evenness$Prevalence,method="spearman")
-
-#create new df for ordered sites to test lagged richness and evenness on prevalence
-lag_evenness <- evenness[-c(4,14,47,90),] #remove site-months that are out of sequence
-lag_evenness$lag_richness <- c(0)
-lag_evenness$lag_J <- c(0)
-lag_evenness$lag_cc <- c(0)
-lag_evenness$lag_size <- c(0)
-for(n in 2:91){ #lag values by one month
-  lag_evenness$lag_richness[n] <- lag_evenness$richness[n-1]
-  lag_evenness$lag_J[n] <- lag_evenness$J[n-1]
-  lag_evenness$lag_cc[n] <- lag_evenness$cc[n-1]
-  lag_evenness$lag_size[n] <- lag_evenness$size[n-1]
-}
-#remove the first entry for each wetland to remove the carryover from the last wetland
-lag_evenness %<>% group_by(WetAltID) %>% filter(duplicated(WetAltID) | n()==1)
-#lag_evenness %<>% select(siteID, lag_richness:lag_size)
-
-lag_evenness %>% ggplot(.,aes(x=lag_richness, y=Prevalence))+
-  geom_smooth(method="lm")
-cor.test(lag_evenness$lag_richness,lag_evenness$Prevalence,method="spearman")
-
 #barplots comparing community composition
-
-evenness %<>% select(-WetAltID)
-abundances %<>% left_join(.,evenness, by = "siteID")
+abundances <- community_mat %>% left_join(.,evenness, by = "siteID")
 lag_evenness %<>% select(siteID, lag_cc, lag_richness, lag_J, lag_size)
@@
 ```{r}
-prev_cc <- data %>% dplyr::select(WetAltID, Month.1, cc, Prevalence, Month, AB2:AB42, MeanWaterTempPredC)
-prev_cc %<>% dplyr::select(-ABAmb, -ABSal) %>%  mutate(., total = rowSums(.[6:26])) #get totals for community size
-prev_cc %<>% dplyr::select(-AB2:-AB42)
-prev_cc %<>% distinct()
+prev_cc <- comm_summ
 
@@
 
-order$lag_cc <- order$cc
-order$lag_size <- order$total
-order$lag_temp <- order$MeanWaterTempPredC
+order %<>% add_column(lag_cc = NA, lag_size = NA, lag_temp = NA)
 
@@
   order$lag_cc[n] <- order$cc[n-1]
-  order$lag_size[n] <- order$total[n-1]
+  order$lag_size[n] <- order$size[n-1]
   order$lag_temp[n] <- order$MeanWaterTempPredC[n-1]
@@
 clean <- order %>% group_by(WetAltID) %>% filter(duplicated(WetAltID) | n()==1)
-
 ```

### 2021-02-19 scripts/data_format.Rmd
deleted some white space

@@
 ```
+
 `r nrow(vl)` total species; `sum(vl$value>0)`species with non-zero viral loads
@@
 ```
+
 environmental variables are `r colnames(env_mat)`

### 2021-02-19 scripts/data_format.Rmd
make sure all figure scripts run

@@
 comm_summ <- data %>% dplyr::select(WetAltID, Month.1, Month, cc, Month, AB2:AB42, Prevalence, MeanWaterTempPredC) %>% 
-  dplyr::select(-ABAmb, -ABSal) %>%  mutate(., size = rowSums(.[5:25])) %>% #get totals for community size
-  dplyr::select(-AB2:-AB42) %>%
+  dplyr::select(-ABAmb, -ABSal) %>%  
+  mutate(., size = rowSums(.[5:25]), 
+    ABHigh = (AB9 + AB21 + AB26 + AB42), 
+    ABLow = (AB2 + AB3 + AB4 + AB5 + AB6 + AB8 + AB20 + AB24 + AB27 + AB28 + AB29 + AB31 + AB34 + AB35 + AB38 + AB39 + AB41)) %>% #get totals for community size
   mutate(., siteID = paste(WetAltID, Month.1, sep = "_")) %>%
@@
 
-Transmission Model for two species - primarily used by fig1
-```{r}
-#define function contour(): inputs are parameters and output is values for grid search for R0
-contour <- function(trans_1_min, trans_1_max, trans_3_min, trans_3_max, mort1, mort2, degr) {
-  #expand data
-  data <- expand_grid("trans1" = seq(from = trans_1_min, to = trans_1_max, by = 0.00001),
-                      "trans2" = c(0.0001),
-                      "trans3" = seq(from = trans_3_min, to = trans_3_max, by = 0.00001),
-                      "shed1" = c(rep(1/2, length("trans1"))),
-                      "shed2" = c(rep(1/2, length("trans1"))),
-                      "reco1" = c(rep(1/10, length("trans1"))),
-                      "reco2" = c(rep(1/10, length("trans1"))),
-                      "mort1" = c(mort1), 
-                      "mort2" = c(mort2), 
-                      "degr" = c(degr), 
-                      "birth1" = c(rep(5/3, length("trans1"))),
-                      "birth2" = c(rep(5/3, length("trans1"))))
-
-  #calculate disease-free equilibria for use in next-gen matrix
-  data %<>% add_column(dfe1 = (data$birth1/data$mort1),dfe2 = (data$birth2/data$mort2), eigen = NA) 
-  #make matrix
-  mat <- matrix(1, nrow = 3, ncol = 3)
-
-  m <- nrow(data)
-
-  #calculate eigenvalue using next-gen matrix. max eigenvalue is R0
-  for(n in 1:m){
-    mat[1] = (data$trans1[n]*data$dfe1[n])/(data$reco1[n]+data$mort1[n])
-    mat[2] = (data$trans1[n]*data$dfe2[n])/(data$reco2[n]+data$mort2[n])
-    mat[3] = (data$shed1[n]/data$degr[n])
-    mat[4] = (data$trans2[n]*data$dfe1[n])/(data$reco1[n]+data$mort1[n])
-    mat[5] = (data$trans2[n]*data$dfe2[n])/(data$reco2[n]+data$mort2[n])
-    mat[6] = (data$shed2[n]/data$degr[n])
-    mat[7] = (data$trans3[n]*data$dfe1[n])/(data$reco1[n]+data$mort1[n])
-    mat[8] = (data$trans3[n]*data$dfe2[n])/(data$reco2[n]+data$mort2[n])
-    mat[9] = 0
-    eigen <- max(eigen(mat)$values)
-    data$eigen[n] <- eigen
-  }
-
-  data$span1 <- 1/data$mort1 #lifespan for species 1
-  data$span2 <- 1/data$mort2 #lifespan for species 2
-  data$half <- log(1/2)/-data$degr #viral half-life
-  data$epid <- data$eigen > 1 #invasion potential (i.e. R0 > 1)
-  data$prop1 <- data$trans1/data$trans2 #proportion of transmission of species 2 over species 1
-  data$prop3 <- data$trans3/data$trans2 #proportion of environmental transmission over species 1
-  data$prop_mort <- data$mort1/data$mort2 #relative mortality for species 1 and 2
-  data$cap1 <- data$birth1/data$mort1 #carrying capacity for species 1
-  data$cap2 <- data$birth2/data$mort2 #carrying capacity for species 2
-  data$cap_prop <- data$cap1/data$cap2 #relative carrying capacities for species 1 and 2
-  data$tot <- data$cap1 + data$cap2 #total abundance for species 1 and 2
-  data$cc <- data$trans1*(data$cap1/data$tot) + data$trans2*(data$cap2/data$tot) #cc defined as transmission of a species multiplied by its relative abundance
-  output <- data %>% dplyr::select(trans1, trans2, prop1, prop3, trans3, eigen, prop_mort, mort1, mort2, degr, half, cap1, cap2, cap_prop, tot, cc) %>% distinct()
-  return(output)
-}
-
-reference <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = 1/45,mort2 = 1/45,degr = 1/2)
-
-composition <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = 1/60,mort2 = 1/30,degr = 1/2)
-
-size <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = c(1/50),mort2 = c(1/50),degr = 1/2)
-
-halflife <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = c(1/40),mort2 = 1/40,degr = c(1/4))
-
-combined <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = c(1/66.6),mort2 = c(1/33.3),degr = c(1/4))
-
-```
-
 PCA for community matrix - primarily used by fig2
@@
 shannon <- melt(as.matrix(shannon))
-shannon %<>% mutate(siteID = X1, h = value) %>% dplyr::select(siteID, h)
+shannon %<>% mutate(siteID = Var1, h = value) %>% dplyr::select(siteID, h)
 
@@
 #barplots comparing community composition
-abundances <- community_mat %>% left_join(.,evenness, by = "siteID")
-lag_evenness %<>% select(siteID, lag_cc, lag_richness, lag_J, lag_size)
-abundances %<>% inner_join(.,lag_evenness, by = "siteID")
-abundances <- mutate(abundances, ABHigh = (AB9 + AB21 + AB26 + AB42)) 
-abundances <- mutate(abundances, ABLow = (AB2 + AB3 + AB4 + AB5 + AB6 + AB8 + AB20 + AB24 + AB27 + AB28 + AB29 + AB31 + AB34 + AB35 + AB38 + AB39 + AB41))
-abundances <- mutate(abundances, total = ABHigh + ABLow)
-#final df for plots. includes data for abundance, prevalence, cc, diversity measure, lagged values
+abundances <- comm_summ %>% inner_join(.,evenness)
 
-abundances$siteID <- reorder(abundances$siteID, -abundances$total) #order by total community size
+abundances$siteID <- reorder(abundances$siteID, -abundances$size) #order by total community size
 #abundances$siteID <- reorder(abundances$siteID, -abundances$cc) #order by community competence

### 2021-02-19 data_format.R

@@ -1,20 +1,144 @@
-## -----------------------------------------------------------------------------------
+## ---------------------------------------------------------------------------------------------------------------------------------
 library(tidyverse)
 library(magrittr)
+library(vegan)
+library(reshape)
 library(here)
 
 
-## -----------------------------------------------------------------------------------
+## ---------------------------------------------------------------------------------------------------------------------------------
 here()
 data <- read_csv(here("/data/weighted_prev_competence_111220.csv"))
 names <- read_csv(here("/data/species_names_ids.csv"))
 
 
-## -----------------------------------------------------------------------------------
+## ---------------------------------------------------------------------------------------------------------------------------------
+comm_summ <- data %>% dplyr::select(WetAltID, Month.1, Month, cc, Month, AB2:AB42, Prevalence, MeanWaterTempPredC) %>% 
+  dplyr::select(-ABAmb, -ABSal) %>%  
+  mutate(., size = rowSums(.[5:25]), 
+    ABHigh = (AB9 + AB21 + AB26 + AB42), 
+    ABLow = (AB2 + AB3 + AB4 + AB5 + AB6 + AB8 + AB20 + AB24 + AB27 + AB28 + AB29 + AB31 + AB34 + AB35 + AB38 + AB39 + AB41)) %>% #get totals for community size
+  mutate(., siteID = paste(WetAltID, Month.1, sep = "_")) %>%
+  distinct()
+comm_summ$siteID <- gsub("Month", "", comm_summ$siteID)
+
+
+## ---------------------------------------------------------------------------------------------------------------------------------
 vl <- data %>% dplyr::select(vl.4:vl.42) %>% distinct() %>% pivot_longer(vl.4:vl.42)
 vl$species_code <- as.double(gsub("vl.","",vl$name))
-nrow(vl)
 vl %<>% full_join(., names, by="species_code")
 vl %<>% mutate(., ln_value = log(value)) %>% mutate(., log10_value = log10(value))
-sum(vl$value>0)
+
+
+## ---------------------------------------------------------------------------------------------------------------------------------
+community_mat <- data %>% dplyr::select(WetAltID, Month.1, AB2:AB8, AB9, AB20:AB42) 
+community_mat$Month <- gsub("Month", "", community_mat$Month.1)
[…]

### 2021-02-19 scripts/figures/fig1_code.R

@@ -10,6 +10,72 @@ library(patchwork)
 #install.packages("ggnewscale")
 library(ggnewscale)
 
+#define function contour(): inputs are parameters and output is values for grid search for R0
+contour <- function(trans_1_min, trans_1_max, trans_3_min, trans_3_max, mort1, mort2, degr) {
+  #expand data
+  data <- expand_grid("trans1" = seq(from = trans_1_min, to = trans_1_max, by = 0.00001),
+                      "trans2" = c(0.0001),
+                      "trans3" = seq(from = trans_3_min, to = trans_3_max, by = 0.00001),
+                      "shed1" = c(rep(1/2, length("trans1"))),
+                      "shed2" = c(rep(1/2, length("trans1"))),
+                      "reco1" = c(rep(1/10, length("trans1"))),
+                      "reco2" = c(rep(1/10, length("trans1"))),
+                      "mort1" = c(mort1), 
+                      "mort2" = c(mort2), 
+                      "degr" = c(degr), 
+                      "birth1" = c(rep(5/3, length("trans1"))),
+                      "birth2" = c(rep(5/3, length("trans1"))))
+  
+  #calculate disease-free equilibria for use in next-gen matrix
+  data %<>% add_column(dfe1 = (data$birth1/data$mort1),dfe2 = (data$birth2/data$mort2), eigen = NA) 
+  #make matrix
+  mat <- matrix(1, nrow = 3, ncol = 3)
+  
+  
+  m <- nrow(data)
+  
+  #calculate eigenvalue using next-gen matrix. max eigenvalue is R0
+  for(n in 1:m){
+    mat[1] = (data$trans1[n]*data$dfe1[n])/(data$reco1[n]+data$mort1[n])
+    mat[2] = (data$trans1[n]*data$dfe2[n])/(data$reco2[n]+data$mort2[n])
+    mat[3] = (data$shed1[n]/data$degr[n])
+    mat[4] = (data$trans2[n]*data$dfe1[n])/(data$reco1[n]+data$mort1[n])
+    mat[5] = (data$trans2[n]*data$dfe2[n])/(data$reco2[n]+data$mort2[n])
+    mat[6] = (data$shed2[n]/data$degr[n])
+    mat[7] = (data$trans3[n]*data$dfe1[n])/(data$reco1[n]+data$mort1[n])
[…]

### 2021-02-19 scripts/figures/fig2_code.R

@@ -8,6 +8,7 @@ library(magrittr)
 library(ggforce)
 library(here)
 
+source(knitr::purl(here("/scripts/data_format.Rmd"), quiet=TRUE))
 
 
 site_scores %>% ggplot(.,aes(x=pc1Rank,y=cc))+
@@ -17,7 +18,7 @@ site_scores %>% ggplot(.,aes(x=pc1Rank,y=cc))+
   facet_zoom(xlim= c(5,25))
 
 site_scores %>% ggplot(.,aes(x=pc1Rank,y=cc))+
-  geom_point(aes(color = Month, shape = factor(WetAltID), size = Size))+
+  geom_point(aes(color = Month, shape = factor(WetAltID), size = size))+
   labs(x="Principal Component 1 Rank", y = "Community Competence (CC)", size = "Size") + 
   scale_shape_manual(values = rep(1:20, len = 20)) +
   scale_size_continuous(range = c(2,10)) +

### 2021-02-19 scripts/figures/fig3_code.R

@@ -11,6 +11,9 @@ library(patchwork)
 library(ggnewscale)
 library(here)
 
+source(knitr::purl(here("/scripts/data_format.Rmd"), quiet=TRUE))
+
+
 #fit two lines to cc_evenness_prev plot
 lag_evenness$high <- c(0)
 lag_evenness$low <- c(0)
@@ -31,9 +34,9 @@ for (i in 1:nrow(lag_evenness)){
   }
 }
 
-# low_cc <- lag_evenness %>% filter(low==1)
+low_cc <- lag_evenness %>% filter(low==1)
 # cor.test(x=low_cc$lag_J, y=low_cc$lag_cc,method = "spearman")
-# high_cc <- lag_evenness %>% filter(high==1)
+high_cc <- lag_evenness %>% filter(high==1)
 # cor.test(x=high_cc$lag_J, y=high_cc$lag_cc,method = "spearman")
 
 p1 <- evenness %>% remove_missing() %>% ggplot(aes(x=richness, y=cc)) + 

### 2021-02-19 scripts/figures/fig4_code.R

@@ -11,6 +11,8 @@ library(patchwork)
 library(ggnewscale)
 library(here)
 
+source(knitr::purl(here("/scripts/data_format.Rmd"), quiet=TRUE))
+
 
 cc_plot <- abundances %>% dplyr::select(siteID, cc) %>% distinct() %>%
   ggplot(., aes(y=cc, x=siteID)) +
@@ -28,11 +30,11 @@ evenness_plot <- abundances %>% dplyr::select(siteID, J, richness) %>% distinct(
 #labs(title = "Values of evenness for sites ordered by community size (descending)") +
 
 z <- abundances
-z %<>% select(siteID,ABLow,AB26,AB42,AB21,AB9,J,total)
-z %<>% rename(Low=ABLow)
+z %<>% select(siteID,ABLow,AB26,AB42,AB21,AB9,J,size)
+z %<>% mutate(Low=ABLow)
 z %<>% pivot_longer(cols=2:6,names_to="Species",values_to="abund") #every row is a site-month-species?
 #z %<>% filter(J<0.6)
-z$siteID <- reorder(z$siteID, -z$total) #order by total community size
+z$siteID <- reorder(z$siteID, -z$size) #order by total community size
 
 AB_plot <- z %>% ggplot(.,aes(x=siteID,y=abund))+
   geom_bar(stat="identity", width = 0.8)+

### 2021-02-19 scripts/figures/fig5_code.R

@@ -10,7 +10,7 @@ library(patchwork)
 library(here)
 
 
-data <- read_csv("data/weighted_prev_competence_111220.csv")
+source(knitr::purl(here("/scripts/data_format.Rmd"), quiet=TRUE))
 
 
 #plot cleaned plot with lag

### 2021-02-22 scripts/data_format.Rmd
change joining function
update for making phylogeny with vl data

@@
 vl <- data %>% dplyr::select(vl.4:vl.42) %>% distinct() %>% pivot_longer(vl.4:vl.42)
-vl$species_code <- as.double(gsub("vl.","",vl$name))
+vl %<>% mutate(species_code = as.double(gsub("vl.","",vl$name)))
 vl %<>% full_join(., names, by="species_code")
+vl %<>% mutate(name = gsub("_"," ",vl$species_name))
+vl$value %<>% na_if(.,0)
 vl %<>% mutate(., ln_value = log(value)) %>% mutate(., log10_value = log10(value))
+vl %<>% replace_na(.,list(value=0,ln_value=0,log10_value=0))
 ```
@@
 #barplots comparing community composition
-abundances <- comm_summ %>% inner_join(.,evenness)
+abundances <- comm_summ %>% full_join(.,evenness)
 

### 2021-02-22 data_format.R

@@ -112,7 +112,7 @@ lag_evenness %<>% group_by(WetAltID) %>% filter(duplicated(WetAltID) | n()==1)
 
 ## ---------------------------------------------------------------------------------------------------------------------------------
 #barplots comparing community composition
-abundances <- comm_summ %>% inner_join(.,evenness)
+abundances <- comm_summ %>% full_join(.,evenness)
 
 
 abundances$siteID <- reorder(abundances$siteID, -abundances$size) #order by total community size

### 2021-02-22 scripts/amphibian_phylo.R

@@ -5,11 +5,38 @@ library(rotl)
 library(phangorn)
 library(geiger)
 library(phytools)
+library(patchwork)
+library(ggtree)
 library(here)
 
 #files
 #prev_test.csv
 #asup_just_tree.txt
+source(knitr::purl(here("/scripts/data_format.Rmd"), quiet=TRUE))
+
+names <- c("Notophthalamus viridescens", "Scaphiopus holbrookii", "Anaxyrus terrestris", "Pseudacris crucifer", "Hyla gratiosa", "Hyla avivoca",
+           "Pseudacris nigrita", "Pseudacris ornata", "Gastrophyrne carolinensis", "Lithobates clamitans", "Lithobates sphenocephalus",
+           "Acris gryllus", "Hyla chrysoscelis", "Hyla cinerea", "Hyla femoralis", "Lithobates catesbeianus", "Siren intermedia")
+resolved <- tnrs_match_names(names)
+tree_1 <- tol_induced_subtree(resolved$ott_id)
+tree_1$tip.label <- strip_ott_ids(tree_1$tip.label,remove_underscores=T)
+vl %<>% select(tnrs_name,value) %>% mutate(.,label = tnrs_name) %>% select(-tnrs_name)
+tree_1 %<>% full_join(.,vl)
+#horizontal tree
+p1.h <- ggtree(tree_1,color="gray")+
+  geom_tippoint(aes(color=value))+
+  scale_colour_viridis_c("Viral load",direction = -1)+
+  geom_tiplab(size=3,hjust=0,offset=0.2,fontface="italic")+
+  xlim(0, 10)
+#vertical tree
+p1.v <- ggtree(tree_1,color="gray")+
+  geom_tippoint(aes(color=value))+
+  scale_colour_viridis_c("Viral load")+
+  geom_tiplab(size=2.2,hjust=1,offset=-0.6,fontface="italic",angle=90)+
+  scale_x_reverse(limits=c(20,0))+
+  coord_flip() 
+
+
 
 
 #trait1 <- rTraitCont(tree, "OU", theta = 0, alpha=.1, sigma=.01)
@@ -101,6 +128,38 @@ ggtree(tree_vl) +geom_tiplab(hjust=0,offset =5)+
   geom_tippoint(aes(color=value,size=10)) +
   scale_color_gradient(low = "#fdc70c",high = "#e93e3a") +
   theme(legend.position = c(0.10,0.75))
+
+
[…]

### 2021-02-22 scripts/figures/fig4_code.R
add note about species 32

@@ -68,7 +68,8 @@ RA_plot<- z %>% ggplot(.,aes(x=siteID,y=abund,fill=Species))+
   geom_bar(position="fill",stat="identity", width = 0.8)+
   scale_fill_manual(values = c("#238443", "#78C679", "#C2E699", "#FFFFB2", "black"))+
   theme_minimal()+
-  theme(axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+ylab("Rel. Abun.")
+  theme(axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
+  ylab("Rel. Abun.")
 
 #still need to clean this up but final figure should have a similar format to this
 

### 2021-02-24 scripts/data_format.Rmd
add note about species 32
remove melt function and add as_tibble

@@
 library(vegan)
-library(reshape)
 library(here)
@@
 `r nrow(vl)` total species; `sum(vl$value>0)`species with non-zero viral loads
+
+Note: Species 32 - Hyla avivoca is the bird-voiced tree frog. Data exists for this species' viral load but there is no data regarding its abundance.
 
@@
 
-shannon <- vegan::diversity(abundance_mat, index = "shannon")
-shannon <- melt(as.matrix(shannon))
-shannon %<>% mutate(siteID = Var1, h = value) %>% dplyr::select(siteID, h)
+shannon <- as_tibble(vegan::diversity(abundance_mat, index = "shannon"),rownames= "siteID")
+
+shannon %<>% mutate(h = value) %>% dplyr::select(siteID, h)
 

### 2021-02-26 scripts/figures/fig1_code.R
clean up packages

@@ -3,11 +3,10 @@
 
 #Figure 1 in rv_cc manuscript
 
-#install_github("thomasp85/patchwork")
 library(tidyverse)
 library(magrittr)
+#install_github("thomasp85/patchwork")
 library(patchwork)
-#install.packages("ggnewscale")
 library(ggnewscale)
 library(here)
 

### 2021-02-26 scripts/figures/fig2_code.R

@@ -3,9 +3,6 @@
 
 #Figure 2 in rv_cc manuscript
 
-library(tidyverse)
-library(magrittr)
-library(ggforce)
 library(here)
 
 source(knitr::purl(here("/scripts/data_format.Rmd"), quiet=TRUE))

### 2021-02-26 scripts/figures/fig3_code.R

@@ -2,12 +2,6 @@
 
 #This script creates figure 3
 
-
-library(tidyverse)
-library(magrittr)
-library(vegan)
-library(patchwork)
-library(ggnewscale)
 library(here)
 
 source(knitr::purl(here("/scripts/data_format.Rmd"), quiet=TRUE))

### 2021-02-26 scripts/figures/fig4_code.R

@@ -3,13 +3,7 @@
 #This script creates figure 4
 
 
-library(tidyverse)
-library(magrittr)
-library(vegan)
-library(reshape2)
-library(patchwork)
-library(ggnewscale)
-library(ggtree)
+
 library(rotl)
 library(here)
 

### 2021-02-26 data_format.R
final formatting for figure 5

@@ -2,6 +2,8 @@
 library(tidyverse)
 library(magrittr)
 library(vegan)
+library(patchwork)
+library(ggnewscale)
 library(here)
 
 

### 2021-02-26 scripts/figures/fig5_code.R

@@ -3,10 +3,6 @@
 
 #Figure 5 in rv_cc manuscript
 
-#install_github("thomasp85/patchwork")
-library(tidyverse)
-library(magrittr)
-library(patchwork)
 library(here)
 
 
@@ -20,8 +16,7 @@ cc_corr <- clean %>% ggplot(.,aes(x=lag_cc, y=Prevalence)) +
   labs(title = "", x = "Community Competence", y = "Prevalence") +
   ylim(0,0.65)
 
-#correlation test
-cor.test(clean$lag_cc,clean$Prevalence,method="spearman")
+#cor.test(clean$lag_cc,clean$Prevalence,method="spearman")
 
 
 size_corr <- clean %>% ggplot(.,aes(x=log(lag_size), y=Prevalence)) +
@@ -31,7 +26,7 @@ size_corr <- clean %>% ggplot(.,aes(x=log(lag_size), y=Prevalence)) +
   theme(axis.text.y = element_blank(), axis.ticks.y=element_blank()) +
   ylim(0,0.65)
 
-cor.test(clean$lag_size,clean$Prevalence,method="spearman")
+#cor.test(clean$lag_size,clean$Prevalence,method="spearman")
 
 
 temp_corr <- clean %>% ggplot(.,aes(x=lag_temp, y=Prevalence)) +
@@ -41,10 +36,10 @@ temp_corr <- clean %>% ggplot(.,aes(x=lag_temp, y=Prevalence)) +
   theme(axis.text.y = element_blank(), axis.ticks.y=element_blank()) +
   ylim(0,0.65)
 
-
-cor.test(clean$lag_temp,clean$Prevalence,method="spearman")
+#cor.test(clean$lag_temp,clean$Prevalence,method="spearman")
 
 
 #plot everything together with patchwork
-corr_plots <- cc_corr| size_corr| temp_corr
-corr_plots
+figure_5 <- cc_corr| size_corr| temp_corr
+ggsave("fig_5.png",plot=figure_5,device="png",path=here("/figures"))
+

### 2021-03-29 scripts/data_format.Rmd
clean up packages
make updates for supp fig 3 and 4

@@
 library(vegan)
+library(patchwork)
+library(ggnewscale)
 library(here)
@@
 
+Make matrix for environmental variables
+```{r}
+#select for environmental variables
+env <- data %>% dplyr::select(WetAltID, Month.1, MeanAirT, MeanWaterTempPredC, DryingScore, CanopyCover, Area, Perimeter) %>% distinct() %>% arrange(.,WetAltID) 
+env$Month.1 <- gsub("Month", "", env$Month.1)
+#make siteID rowname
+env$siteID <- paste(env$WetAltID, env$Month.1, sep = "_")
+env %<>% dplyr::select(-WetAltID, -Month.1)
+#remove duplicated rows (siteID is duplicated but env. variables are not... not sure why this is)
+env <- env[-c(53, 90),]
+env_mat <- env %>% column_to_rownames(., var = "siteID")
+```
+
 cc x evenness data - primarily used by fig3

### 2021-03-29 data_format.R

@@ -1,4 +1,4 @@
-## ------------------------------------------------------------------------
+## -------------------------------------------------------------------------------------------------------------------------------------------------
 library(tidyverse)
 library(magrittr)
 library(vegan)
@@ -7,13 +7,13 @@ library(ggnewscale)
 library(here)
 
 
-## ------------------------------------------------------------------------
+## -------------------------------------------------------------------------------------------------------------------------------------------------
 here()
 data <- read_csv(here("/data/weighted_prev_competence_111220.csv"))
 names <- read_csv(here("/data/species_names_ids.csv"))
 
 
-## ------------------------------------------------------------------------
+## -------------------------------------------------------------------------------------------------------------------------------------------------
 comm_summ <- data %>% dplyr::select(WetAltID, Month.1, Month, cc, Month, AB2:AB42, Prevalence, MeanWaterTempPredC) %>% 
   dplyr::select(-ABAmb, -ABSal) %>%  
   mutate(., size = rowSums(.[5:25]), 
@@ -24,7 +24,7 @@ comm_summ <- data %>% dplyr::select(WetAltID, Month.1, Month, cc, Month, AB2:AB4
 comm_summ$siteID <- gsub("Month", "", comm_summ$siteID)
 
 
-## ------------------------------------------------------------------------
+## -------------------------------------------------------------------------------------------------------------------------------------------------
 vl <- data %>% dplyr::select(vl.4:vl.42) %>% distinct() %>% pivot_longer(vl.4:vl.42)
 vl %<>% mutate(species_code = as.double(gsub("vl.","",vl$name)))
 vl %<>% full_join(., names, by="species_code")
@@ -34,7 +34,7 @@ vl %<>% mutate(., ln_value = log(value)) %>% mutate(., log10_value = log10(value
 vl %<>% replace_na(.,list(value=0,ln_value=0,log10_value=0))
 
 
-## ------------------------------------------------------------------------
+## -------------------------------------------------------------------------------------------------------------------------------------------------
 community_mat <- data %>% dplyr::select(WetAltID, Month.1, AB2:AB8, AB9, AB20:AB42) 
 community_mat$Month <- gsub("Month", "", community_mat$Month.1)
 community_mat %<>% mutate(., siteID = paste(WetAltID, Month, sep = "_")) %>% distinct() %>% arrange(.,WetAltID) %>% dplyr::select(siteID, AB2:AB42)
[…]

### 2021-03-29 scripts/figures/gradient_analysis.R

@@ -117,7 +117,7 @@ comm_scores %>% ggplot(.) +
   coord_fixed()
 
 comm_scores %>% ggplot(.) +
-  xlab("PC1")+ylab("PC2")+
+  xlab("PC1 (71.2% Var. Explained)")+ylab("PC2 (26% Var. Explained)")+
   geom_segment(data=comm_scores,
                aes(x = 0, xend = Comp.1, y = 0, yend = Comp.2),
                arrow = arrow(length = unit(0.1, "cm")), colour = "darkorange1") +

### 2021-03-29 scripts/figures/supp1_code.R

@@ -3,8 +3,6 @@
 
 #This script creates supplementary figure 1: visual of the estimated viral load for each species
 
-library(tidyverse)
-library(magrittr)
 library(here)
 
 

### 2021-03-29 scripts/figures/supp3_code.R

@@ -1,0 +1,32 @@
+## Script originator: Daniel Suh
+## Date created: Mar. 29, 2021
+
+#This script creates supplementary figure 3: PCA
+
+library(here)
+
+
+source(knitr::purl(here("/scripts/data_format.Rmd"), quiet=TRUE))
+
+
+
+community_pca <- princomp(abundance_mat, scores = TRUE)
+summary(community_pca)
+
+comm_scores <- as.data.frame(community_pca$loadings[,1:2])
+env_vec <- envfit(community_pca, env_mat, na.rm = T) 
+env_vec <- as.data.frame(scores(env_vec, display = "vectors"))
+
+comm_scores %>% ggplot(.) +
+  xlab("PC1 (71.2% Var. Explained)")+ylab("PC2 (26% Var. Explained)")+
+  geom_segment(data=comm_scores,
+               aes(x = 0, xend = Comp.1, y = 0, yend = Comp.2),
+               arrow = arrow(length = unit(0.1, "cm")), colour = "darkorange1") +
+  geom_label_repel(data = comm_scores, aes(x = Comp.1, y = Comp.2, label = row.names(comm_scores)),
+                   size = 2,color="darkorange1", segment.colour = 'black')+
+  geom_segment(data=env_vec,
+               aes(x = 0, xend = Comp.1, y = 0, yend = Comp.2),
+               arrow = arrow(length = unit(0.1, "cm")), colour = "purple") +
+  geom_label_repel(data = env_vec, aes(x = Comp.1, y = Comp.2, label = rownames(env_vec)),
+                   size = 2,color="purple", segment.colour = 'black')+
+  coord_fixed()

### 2021-03-29 scripts/figures/supp4_code.R

@@ -1,0 +1,5 @@
+## Script originator: Daniel Suh
+## Date created: Mar. 29, 2021
+
+#This script creates supplementary figure 4: prevalence ~ richness
+

### 2021-03-31 scripts/data_format.Rmd
fix lagged data

@@
 here()
-data <- read_csv(here("/data/weighted_prev_competence_111220.csv"))
-names <- read_csv(here("/data/species_names_ids.csv"))
+data <- read_csv(here("data/weighted_prev_competence_111220.csv"))
+names <- read_csv(here("data/species_names_ids.csv"))
 ```
@@
 lag_evenness$lag_size <- c(0)
-for(n in 2:91){
+for(n in 2:92){
   lag_evenness$lag_richness[n] <- lag_evenness$richness[n-1]

### 2021-03-31 data_format.R

@@ -9,8 +9,8 @@ library(here)
 
 ## -------------------------------------------------------------------------------------------------------------------------------------------------
 here()
-data <- read_csv(here("/data/weighted_prev_competence_111220.csv"))
-names <- read_csv(here("/data/species_names_ids.csv"))
+data <- read_csv(here("data/weighted_prev_competence_111220.csv"))
+names <- read_csv(here("data/species_names_ids.csv"))
 
 
 ## -------------------------------------------------------------------------------------------------------------------------------------------------

### 2021-05-18 data_format.R
update supp figs

@@ -1,4 +1,4 @@
-## ---------------------------------------------------------------------------------------------------------------------------------------------
+## ------------------------------------------------------------------------------------------------------------------
 library(tidyverse)
 library(magrittr)
 library(vegan)
@@ -7,13 +7,13 @@ library(ggnewscale)
 library(here)
 
 
-## ---------------------------------------------------------------------------------------------------------------------------------------------
+## ------------------------------------------------------------------------------------------------------------------
 here()
 data <- read_csv(here("data/weighted_prev_competence_111220.csv"))
 names <- read_csv(here("data/species_names_ids.csv"))
 
 
-## ---------------------------------------------------------------------------------------------------------------------------------------------
+## ------------------------------------------------------------------------------------------------------------------
 comm_summ <- data %>% dplyr::select(WetAltID, Month.1, Month, cc, Month, AB2:AB42, Prevalence, MeanWaterTempPredC) %>% 
   dplyr::select(-ABAmb, -ABSal) %>%  
   mutate(., size = rowSums(.[5:25]), 
@@ -24,7 +24,7 @@ comm_summ <- data %>% dplyr::select(WetAltID, Month.1, Month, cc, Month, AB2:AB4
 comm_summ$siteID <- gsub("Month", "", comm_summ$siteID)
 
 
-## ---------------------------------------------------------------------------------------------------------------------------------------------
+## ------------------------------------------------------------------------------------------------------------------
 vl <- data %>% dplyr::select(vl.4:vl.42) %>% distinct() %>% pivot_longer(vl.4:vl.42)
 vl %<>% mutate(species_code = as.double(gsub("vl.","",vl$name)))
 vl %<>% full_join(., names, by="species_code")
@@ -34,7 +34,7 @@ vl %<>% mutate(., ln_value = log(value)) %>% mutate(., log10_value = log10(value
 vl %<>% replace_na(.,list(value=0,ln_value=0,log10_value=0))
 
 
-## ---------------------------------------------------------------------------------------------------------------------------------------------
+## ------------------------------------------------------------------------------------------------------------------
 community_mat <- data %>% dplyr::select(WetAltID, Month.1, AB2:AB8, AB9, AB20:AB42) 
 community_mat$Month <- gsub("Month", "", community_mat$Month.1)
 community_mat %<>% mutate(., siteID = paste(WetAltID, Month, sep = "_")) %>% distinct() %>% arrange(.,WetAltID) %>% dplyr::select(siteID, AB2:AB42)
[…]

### 2021-05-18 scripts/figures/supp5_code.R

@@ -11,12 +11,12 @@ source(knitr::purl(here("scripts/data_format.Rmd"), quiet=TRUE))
 
 evenness$Month <- factor(evenness$Month,levels=c("Feb","Mar","Apr","May","Jun","Jul"))
 
-supp5 <- evenness %>% ggplot(.,aes(x=Month,y=cc)) + geom_boxplot() + geom_point() + ylab("Community Competence")
+supp5a <- evenness %>% ggplot(.,aes(x=Month,y=cc)) + geom_boxplot() + geom_point() + ylab("Community Competence")
 
 evenness %>% ggplot(.,aes(x=Month,y=cc)) + geom_boxplot() + geom_point() + geom_line(aes(group=WetAltID))
 
-supp6 <- evenness %>% ggplot(.,aes(x=log10(size),y=cc)) + geom_point() + geom_smooth(method="lm")
+supp5b <- evenness %>% ggplot(.,aes(x=log10(size),y=cc)) + geom_point() + geom_smooth(method="lm")
 cor.test(log10(evenness$size), evenness$cc, method = "spearman")
   
-ggsave("supp5.png",plot=supp5,device="png",path=here("/figures"))
-ggsave("supp6.png",plot=supp6,device="png",path=here("/figures"))
+ggsave("supp5a.png",plot=supp5a,device="png",path=here("figures"))
+ggsave("supp5b.png",plot=supp5b,device="png",path=here("figures"))

### 2021-05-18 scripts/figures/supp6_code.R

@@ -1,5 +1,14 @@
+#Script Originator: Daniel Suh
+
+#This script creates supplemental figure 6
+
+
 library(ape)
-tr2 <- ape::read.nexus("./data/asup_just_tree.txt")
+library(here)
+
+source(knitr::purl(here("scripts/data_format.Rmd"), quiet=TRUE))
+
+tr2 <- ape::read.nexus(here("data/asup_just_tree.txt"))
 mySpp <- c(union(names$tnrs_name,names$species_name),"Bufo terrestris")
 mySpp <- gsub(" ","_",mySpp)
 getRid <- setdiff(tr2$tip.label,mySpp)
@@ -47,6 +56,8 @@ for (k in 1:dim(y)[1]){
   y$min.pdAbund[k] <- min.pdAbund
 }
 y %>% filter(relAbund>0) %>% ggplot(.,aes(x=mpd,y=relAbund))+geom_point()+geom_smooth(method="loess",span=1.4)
-y %>% filter(relAbund>0) %>% ggplot(.,aes(x=min.pd,y=relAbund))+geom_point()+geom_smooth(method="loess",span=1.4)
+supp6 <- y %>% filter(relAbund>0) %>% ggplot(.,aes(x=min.pd,y=relAbund))+geom_point()+geom_smooth(method="loess",span=1.4)
 y %>% filter(relAbund>0) %>% ggplot(.,aes(x=mpdAbund,y=relAbund))+geom_point()+geom_smooth(method="loess",span=1.4)
 y %>% filter(relAbund>0) %>% ggplot(.,aes(x=min.pdAbund,y=relAbund))+geom_point()+geom_smooth(method="loess",span=1.4)
+
+ggsave("supp6.png",plot=supp6,device="png",path=here("figures"))

### 2021-05-18 scripts/data_format.Rmd
update supp figs
clean up code

@@
 library(ggnewscale)
+library(ape)
+library(rotl)
 library(here)
@@
 data <- read_csv(here("data/weighted_prev_competence_111220.csv"))
+tree <- ape::read.nexus(here("data/asup_just_tree.txt"))
 names <- read_csv(here("data/species_names_ids.csv"))
@@
 ```
+
+Phylogeny
+```{r}
+resolved <- tnrs_match_names(names$tnrs_name)
+resolved$unique_name <- gsub(" ", "_",resolved$unique_name)
+resolved$search_string <- paste(toupper(substr(resolved$search_string, 1, 1)), 
+                                  substr(resolved$search_string,2,nchar(resolved$search_string)),sep = "")
+
+resolved %<>% rename("species_name" = "search_string") %>% select(species_name, unique_name, ott_id)
+
+names %<>% left_join(.,resolved)
+
+names %<>% filter(tnrs_name!="Dryophytes avivoca") #drop species because there is no abundance data
+mySpp <- c(union(names$tnrs_name,names$species_name),"Bufo terrestris") #make vector for all species names
+mySpp <- gsub(" ","_",mySpp)
+getRid <- setdiff(tree$tip.label,mySpp)
+pruned_tree <- drop.tip(tree,getRid)
+
+```

### 2021-05-18 data_format.R

@@ -4,12 +4,15 @@ library(magrittr)
 library(vegan)
 library(patchwork)
 library(ggnewscale)
+library(ape)
+library(rotl)
 library(here)
 
 
 ## ------------------------------------------------------------------------------------------------------------------
 here()
 data <- read_csv(here("data/weighted_prev_competence_111220.csv"))
+tree <- ape::read.nexus(here("data/asup_just_tree.txt"))
 names <- read_csv(here("data/species_names_ids.csv"))
 
 
@@ -157,4 +160,22 @@ for(n in 2:nrow(order)){
 
 #remove the first entry for each wetland to remove the carryover from the last wetland
 clean <- order %>% group_by(WetAltID) %>% filter(duplicated(WetAltID) | n()==1)
+
+
+## ------------------------------------------------------------------------------------------------------------------
+resolved <- tnrs_match_names(names$tnrs_name)
+resolved$unique_name <- gsub(" ", "_",resolved$unique_name)
+resolved$search_string <- paste(toupper(substr(resolved$search_string, 1, 1)), 
+                                  substr(resolved$search_string,2,nchar(resolved$search_string)),sep = "")
+
+resolved %<>% rename("species_name" = "search_string") %>% select(species_name, unique_name, ott_id)
+
+names %<>% left_join(.,resolved)
+
+names %<>% filter(tnrs_name!="Dryophytes avivoca") #drop species because there is no abundance data
+mySpp <- c(union(names$tnrs_name,names$species_name),"Bufo terrestris") #make vector for all species names
+mySpp <- gsub(" ","_",mySpp)
+getRid <- setdiff(tree$tip.label,mySpp)
+pruned_tree <- drop.tip(tree,getRid)
+
 

### 2021-05-18 scripts/figures/fig4_code.R

@@ -6,20 +6,15 @@
 library(ggnewscale)
 library(ggtree)
 library(rotl)
-library(ape)
 library(here)
-here()
+
 source(knitr::purl(here("scripts/data_format.Rmd"), quiet=TRUE))
 vl %<>% filter(tnrs_name!="Dryophytes avivoca") # as doesn't appear sampled in any communities? (AP)
+#H. avivoca (ID=32) has no abundance measured but has one recording for viral load
+#viral load for H. avivoca is actually the highest but should not be included because there is no abundance data
 
 
-tr2 <- ape::read.nexus("./data/asup_just_tree.txt")
-names %<>% filter(tnrs_name!="Dryophytes avivoca")
-mySpp <- c(union(names$tnrs_name,names$species_name),"Bufo terrestris")
-mySpp <- gsub(" ","_",mySpp)
-getRid <- setdiff(tr2$tip.label,mySpp)
-tr2.pruned <- drop.tip(tr2,getRid)
-tr2.pruned$tip.label <- gsub("_"," ",tr2.pruned$tip.label)
+pruned_tree$tip.label <- gsub("_"," ",pruned_tree$tip.label)
 
 
 vl %<>% select(species_name,value,ln_value,log10_value) %>% mutate(.,label = species_name) %>% select(-species_name)
@@ -36,7 +31,7 @@ top4sampled <- c("Pseudacris crucifer","Rana sphenocephala","Bufo terrestris","N
 vl %<>% mutate(labelAlt=as_factor(if_else(labelAlt %in% top4sampled,labelAlt,"Low competence")))
 vl %<>% mutate(labelAlt=fct_relevel(labelAlt,c(top4sampled,"Low competence")))
 
-tree <- tr2.pruned %<>% full_join(.,vl)
+vl_tree <- pruned_tree %>% full_join(.,vl)
 
 
 abundances$siteID <- reorder(abundances$siteID, -abundances$cc)
@@ -71,7 +66,7 @@ RA_plot<- z %>% ggplot(.,aes(x=siteID,y=abund,fill=Species))+
 #   geom_tiplab(size=2.2,hjust=,offset=7,angle=90)+#fontface="italic",
 #   scale_size_continuous(name="log10(Viral load)")+
 #   coord_flip()+guides(color=F)+xlim(0,300)
-newTree <- tree %>% ggtree(.,color="gray")+
+newTree <- vl_tree %>% ggtree(.,color="gray")+
[…]

### 2021-05-18 scripts/figures/supp6_code.R

@@ -3,18 +3,13 @@
 #This script creates supplemental figure 6
 
 
-library(ape)
 library(here)
 
 source(knitr::purl(here("scripts/data_format.Rmd"), quiet=TRUE))
 
-tr2 <- ape::read.nexus(here("data/asup_just_tree.txt"))
-mySpp <- c(union(names$tnrs_name,names$species_name),"Bufo terrestris")
-mySpp <- gsub(" ","_",mySpp)
-getRid <- setdiff(tr2$tip.label,mySpp)
-tr2.pruned <- drop.tip(tr2,getRid)
-plot(tr2.pruned)
-pd <- cophenetic(tr2.pruned)
+
+plot(pruned_tree)
+pd <- cophenetic(pruned_tree)
 rownames(pd) <- gsub("_"," ",rownames(pd))
 #mpd and min.pd
 y <- abundances
@@ -36,6 +31,12 @@ y %<>% mutate(mpd=-999,min.pd=-999,mpdAbund=-999,min.pdAbund=-999)
 #top4sampledSyn <- c(top4sampled,"Bufo terrestris")
 #y %<>% filter(Species %in% top4sampledSyn)
 #end optional focus
+
+#y is a df where every row is one species in one site-month
+#so row numbers should equal number of distinct species * number of site-months
+#i.e. y = 16 species * 96 site-months = 1536 rows
+
+
 for (k in 1:dim(y)[1]){
   tmp <- y %>% filter(siteID==y$siteID[k])
   pd.row <- which(rownames(pd)==y$Species[k])
@@ -55,6 +56,7 @@ for (k in 1:dim(y)[1]){
   y$mpdAbund[k] <- mpdAbund
   y$min.pdAbund[k] <- min.pdAbund
 }
+
 y %>% filter(relAbund>0) %>% ggplot(.,aes(x=mpd,y=relAbund))+geom_point()+geom_smooth(method="loess",span=1.4)
 supp6 <- y %>% filter(relAbund>0) %>% ggplot(.,aes(x=min.pd,y=relAbund))+geom_point()+geom_smooth(method="loess",span=1.4)
 y %>% filter(relAbund>0) %>% ggplot(.,aes(x=mpdAbund,y=relAbund))+geom_point()+geom_smooth(method="loess",span=1.4)

### 2021-05-19 scripts/data_format.Rmd
redo with pairwise distances

@@
 resolved <- tnrs_match_names(names$tnrs_name)
-resolved$unique_name <- gsub(" ", "_",resolved$unique_name)
-resolved$search_string <- paste(toupper(substr(resolved$search_string, 1, 1)), 
-                                  substr(resolved$search_string,2,nchar(resolved$search_string)),sep = "")
+resolved %<>% rename("tnrs_name" = "unique_name")
+#resolved$search_string <- paste(toupper(substr(resolved$search_string, 1, 1)), 
+#                                  substr(resolved$search_string,2,nchar(resolved$search_string)),sep = "")
 
-resolved %<>% rename("species_name" = "search_string") %>% select(species_name, unique_name, ott_id)
+resolved %<>% select(tnrs_name, ott_id)
 

### 2021-05-19 data_format.R

@@ -164,11 +164,11 @@ clean <- order %>% group_by(WetAltID) %>% filter(duplicated(WetAltID) | n()==1)
 
 ## ------------------------------------------------------------------------------------------------------------------
 resolved <- tnrs_match_names(names$tnrs_name)
-resolved$unique_name <- gsub(" ", "_",resolved$unique_name)
-resolved$search_string <- paste(toupper(substr(resolved$search_string, 1, 1)), 
-                                  substr(resolved$search_string,2,nchar(resolved$search_string)),sep = "")
+resolved %<>% rename("tnrs_name" = "unique_name")
+#resolved$search_string <- paste(toupper(substr(resolved$search_string, 1, 1)), 
+#                                  substr(resolved$search_string,2,nchar(resolved$search_string)),sep = "")
 
-resolved %<>% rename("species_name" = "search_string") %>% select(species_name, unique_name, ott_id)
+resolved %<>% select(tnrs_name, ott_id)
 
 names %<>% left_join(.,resolved)
 

### 2021-05-19 scripts/figures/supp6_code.R

@@ -84,7 +84,7 @@ y %>% filter(relAbund>0) %>% ggplot(.,aes(x=mpdAbund,y=relAbund))+
 y %>% filter(relAbund>0) %>% ggplot(.,aes(x=min.pdAbund,y=relAbund))+
   geom_point()+geom_smooth(method="loess",span=1.4)
 
-
+#calculate mean pairwise distance for each community and plot community competence against it
 
 
 

### 2021-05-19 scripts/phylo_diversity.R

@@ -1,69 +1,43 @@
 
-library(tidyverse)
-library(magrittr)
-library(phytools)
-library(phylobase)
-library(treeio)
-library(phanghorn)
-library(ape)
-library(ggtree)
-library(tidytree)
-library(picante)
-library(rotl)
-library(ggnewscale)
+
 library(here)
 
 
-data <- read_csv(here("data/weighted_prev_competence_111220.csv"))
+source(knitr::purl(here("scripts/data_format.Rmd"), quiet=TRUE))
 
-tree <- read.nexus("data/asup_just_tree.txt")
 
-names <- read_csv("data/species_names_ids.csv")
 
-resolved <- tnrs_match_names(names$tnrs_name)
-resolved$unique_name <- gsub(" ", "_",resolved$unique_name)
-resolved$search_string <- paste(toupper(substr(resolved$search_string, 1, 1)), 
-                                  substr(resolved$search_string,2,nchar(resolved$search_string)),sep = "")
-
-resolved %<>% rename("species_name" = "search_string") %>% select(species_name, unique_name, ott_id)
-
-names %<>% left_join(.,resolved)
-
-species <- unique(c(names$species_name[which(names$species_name %in% tree$tip.label)], 
-                    names$unique_name[which(names$unique_name %in% tree$tip.label)]))
-
-tree %<>% drop.tip(.,tree$tip.label[-match(species, tree$tip.label)])
-
-subset <- names[unique(c(which(names$species_name %in% species),which(names$unique_name %in% species))),]
-
+#subset <- names[unique(c(which(names$species_name %in% species),which(names$unique_name %in% species))),]
+subset <- names %>% dplyr::select(species_name, species_code, tnrs_name)
 subset <- subset[order(subset$species_code),]
-print(subset$species_code)
+#print(subset$species_code)
 
-subset %<>% filter(species_code!=c(31,32))
+#subset %<>% filter(species_code!=c(31,32))
 IDs <- c(subset$species_code)
-community_mat <- data %>% dplyr::select(WetAltID, Month.1, AB4:AB42) %>% dplyr::select(-c(ABAmb, ABSal, AB31, AB5, AB6, AB8, AB21)) %>%
[…]

### 2021-05-23 scripts/data_format.Rmd
get mpd for each site and plot hist

@@
 library(rotl)
+library(picante)
 library(here)

### 2021-05-23 scripts/phylo_diversity.R

@@ -70,3 +70,10 @@ for (i in 1:nrow(subset_1)){
 phy_comm_1 %<>% select(-c("AB8")) #remove because we don't know the identity of this species
 
 mean_pairwise_distances <- mpd(phy_comm_1,dist_mat)
+
+site_IDs <- row.names(phy_comm_1) #make sure these rownames match properly
+amphib_mpd <- tibble("siteID" = site_IDs, "mpd" = mean_pairwise_distances)
+
+amphib_mpd <- inner_join(pd,amphib_mpd,by="siteID")
+
+ggplot(amphib_mpd, aes(x=mpd)) + geom_histogram()

### 2021-05-23 scripts/phylo_diversity.R
play with number of bins

@@ -76,4 +76,4 @@ amphib_mpd <- tibble("siteID" = site_IDs, "mpd" = mean_pairwise_distances)
 
 amphib_mpd <- inner_join(pd,amphib_mpd,by="siteID")
 
-ggplot(amphib_mpd, aes(x=mpd)) + geom_histogram()
+ggplot(amphib_mpd, aes(x=mpd)) + geom_histogram(bins="50")

### 2021-05-27 scripts/data_format.Rmd
working on code for eeid poster

@@
 ```{r}
+rv_sampled <- data %>% select(Species) %>% distinct() #species codes for those sampled for rv
+rv_sampled %<>% filter(Species!=32) #remove species 32 because there is no abundance data
+keep <- c(rv_sampled$Species)
+
 resolved <- tnrs_match_names(names$tnrs_name)
@@
 
+names %<>% filter(species_code %in% keep)
+
 names %<>% filter(tnrs_name!="Dryophytes avivoca") #drop species because there is no abundance data
@@
 pruned_tree <- drop.tip(tree,getRid)
-
 ```

### 2021-05-27 data_format.R

@@ -6,6 +6,7 @@ library(patchwork)
 library(ggnewscale)
 library(ape)
 library(rotl)
+library(picante)
 library(here)
 
 
@@ -163,6 +164,10 @@ clean <- order %>% group_by(WetAltID) %>% filter(duplicated(WetAltID) | n()==1)
 
 
 ## ------------------------------------------------------------------------------------------------------------------
+rv_sampled <- data %>% select(Species) %>% distinct() #species codes for those sampled for rv
+rv_sampled %<>% filter(Species!=32) #remove species 32 because there is no abundance data
+keep <- c(rv_sampled$Species)
+
 resolved <- tnrs_match_names(names$tnrs_name)
 resolved %<>% rename("tnrs_name" = "unique_name")
 #resolved$search_string <- paste(toupper(substr(resolved$search_string, 1, 1)), 
@@ -172,10 +177,11 @@ resolved %<>% select(tnrs_name, ott_id)
 
 names %<>% left_join(.,resolved)
 
+names %<>% filter(species_code %in% keep)
+
 names %<>% filter(tnrs_name!="Dryophytes avivoca") #drop species because there is no abundance data
 mySpp <- c(union(names$tnrs_name,names$species_name),"Bufo terrestris") #make vector for all species names
 mySpp <- gsub(" ","_",mySpp)
 getRid <- setdiff(tree$tip.label,mySpp)
 pruned_tree <- drop.tip(tree,getRid)
-
 

### 2021-05-27 scripts/figures/fig5_code.R

@@ -43,7 +43,7 @@ temp_corr <- clean %>% ggplot(.,aes(x=lag_temp, y=Prevalence)) +
 
 #plot everything together with patchwork
 figure_5 <- cc_corr| size_corr| temp_corr
-ggsave("fig_5.png",plot=figure_5,device="png",path=here("/figures"))
+#ggsave("fig_5.png",plot=figure_5,device="png",path=here("/figures"))
 
 
 

### 2021-05-27 scripts/figures/supp4_code.R

@@ -18,8 +18,10 @@ cor.test(evenness$richness, evenness$Prevalence, method="spearm")
 
 supp4 <- lag_evenness %>% ggplot(.,aes(x=lag_richness, y=Prevalence)) +
   geom_point() +
-  geom_smooth(method="lm")
+  geom_smooth(method="lm") +
+  theme_classic() +
+  labs(x="Richness", y="Prevalence (t+1)")
 
 cor.test(lag_evenness$lag_richness, lag_evenness$Prevalence, method="spearm")
 
-ggsave("supp4.png",plot=supp4,device="png",path=here("/figures"))
+#ggsave("supp4.png",plot=supp4,device="png",path=here("/figures"))

### 2021-05-27 scripts/figures/supp6_code.R

@@ -1,7 +1,12 @@
 #Script Originator: Daniel Suh
 
 #This script creates supplemental figure 6
-
+#Objective: A plot that models the cophenetic distance of a community and relative abundances of species
+#This provides a look at patterns related to species abundance and phylogenetic relatedness
+#We hypothesize that the species in our data are most abundant when they are moderately related to their neighbor
+#If their closest neighbor is very phylogenetically close then competitive exclusion is likely to take place
+#If their closest neighbor is very phylogenetically distant then environmental filtering may reduce coexistence
+#If their closest neighbor is moderately phylogenetically close then they may coexist at high abundance
 
 library(here)
 
@@ -26,7 +31,7 @@ y %<>% left_join(.,names2)
 # relabundo
 y %<>% rowwise() %>% mutate(relAbund=abund/size)
 y %<>% drop_na(Species)
-y %<>% mutate(mpd=-999,min.pd=-999,mpdAbund=-999,min.pdAbund=-999)
+y %<>% mutate(mpd=-999,min.pd=-999,mpdAbund=-999,min.pdAbund=-999,neighbor=-999)
 #optional focus on top4
 #top4sampledSyn <- c(top4sampled,"Bufo terrestris")
 #y %<>% filter(Species %in% top4sampledSyn)
@@ -54,8 +59,16 @@ for (k in 1:dim(y)[1]){ #loop through each row
   myDists <- pd[pd.row,pd.cols] #get distances between focal species and other species
   mpd <- mean(myDists) #calculate mean pd
   min.pd <- min(myDists) #calculate min pd
+  neighbor <- names(which(myDists==min.pd))
+  if(is.null(neighbor)){
+    neighbor <- otherSpp
+  }
+  if(identical(neighbor, character(0))){
+    neighbor <- "no neighbor"
+  }
   y$mpd[k] <- mpd #add to df
   y$min.pd[k] <- min.pd #add to df
+  y$neighbor[k] <- neighbor
   #do same thing for relative abundances of each species in each site-month
   tmp2 <- tmp %>% filter(Species!=y$Species[k]) #make tmp df for species that are not focal species
   tmp2 %<>% filter(relAbund == max(relAbund)) #filter for max relative abundance
@@ -73,11 +86,14 @@ y %>% filter(relAbund>0) %>% ggplot(.,aes(x=mpd,y=relAbund))+
   geom_point()+geom_smooth(method="loess",span=1.4)
[…]

### 2021-05-27 scripts/phylo_diversity.R

@@ -67,7 +67,8 @@ for (i in 1:nrow(subset_1)){
   phy_comm_1 <- rename(phy_comm_1, !!subset_1$tnrs_name[i] := col_name)
 }
 
-phy_comm_1 %<>% select(-c("AB8")) #remove because we don't know the identity of this species
+phy_comm_1 %<>% select(-c("AB8","AB2","AB3","AB5","AB6")) 
+#remove because we don't know the identity of this species or species is not sampled for ranavirus
 
 mean_pairwise_distances <- mpd(phy_comm_1,dist_mat)
 
@@ -76,4 +77,5 @@ amphib_mpd <- tibble("siteID" = site_IDs, "mpd" = mean_pairwise_distances)
 
 amphib_mpd <- inner_join(pd,amphib_mpd,by="siteID")
 
-ggplot(amphib_mpd, aes(x=mpd)) + geom_histogram(bins="50")
+ggplot(amphib_mpd, aes(x=mpd)) + geom_histogram(bins="40") + theme_minimal()+
+  labs(x="Mean Pairwise Distances", y = "Frequency")

### 2021-06-16 data_format.R
make tables

@@ -1,4 +1,4 @@
-## ----------------------------------------------------------------------------------------------------------------------
+## -----------------------------------------------------------------------------
 library(tidyverse)
 library(magrittr)
 library(vegan)
@@ -10,14 +10,14 @@ library(picante)
 library(here)
 
 
-## ----------------------------------------------------------------------------------------------------------------------
+## -----------------------------------------------------------------------------
 here()
 data <- read_csv(here("data/weighted_prev_competence_111220.csv"))
 tree <- ape::read.nexus(here("data/asup_just_tree.txt"))
 names <- read_csv(here("data/species_names_ids.csv"))
 
 
-## ----------------------------------------------------------------------------------------------------------------------
+## -----------------------------------------------------------------------------
 comm_summ <- data %>% dplyr::select(WetAltID, Month.1, Month, cc, Month, AB2:AB42, Prevalence, MeanWaterTempPredC) %>% 
   dplyr::select(-ABAmb, -ABSal) %>%  
   mutate(., size = rowSums(.[5:25]), 
@@ -28,7 +28,7 @@ comm_summ <- data %>% dplyr::select(WetAltID, Month.1, Month, cc, Month, AB2:AB4
 comm_summ$siteID <- gsub("Month", "", comm_summ$siteID)
 
 
-## ----------------------------------------------------------------------------------------------------------------------
+## -----------------------------------------------------------------------------
 vl <- data %>% dplyr::select(vl.4:vl.42) %>% distinct() %>% pivot_longer(vl.4:vl.42)
 vl %<>% mutate(species_code = as.double(gsub("vl.","",vl$name)))
 vl %<>% full_join(., names, by="species_code")
@@ -38,7 +38,7 @@ vl %<>% mutate(., ln_value = log(value)) %>% mutate(., log10_value = log10(value
 vl %<>% replace_na(.,list(value=0,ln_value=0,log10_value=0))
 
 
-## ----------------------------------------------------------------------------------------------------------------------
+## -----------------------------------------------------------------------------
 community_mat <- data %>% dplyr::select(WetAltID, Month.1, AB2:AB8, AB9, AB20:AB42) 
 community_mat$Month <- gsub("Month", "", community_mat$Month.1)
 community_mat %<>% mutate(., siteID = paste(WetAltID, Month, sep = "_")) %>% distinct() %>% arrange(.,WetAltID) %>% dplyr::select(siteID, AB2:AB42)
[…]

### 2021-06-16 scripts/figures/supp2_code.R

@@ -26,5 +26,8 @@ ordered_matrix <- OrderMatrix(community_mat)
 rownames(ordered_matrix)
 
 results <- Metacommunity(community_mat, binary = T)
+coherence <- results$Coherence
+turnover <- results$Turnover
+boundary <- results$Boundary
 
 #wasn't able to save output from Imagine as a plot and had to save using dialog within RStudio

### 2022-09-09 scripts/data_format.Rmd
fix filepath

@@
 here()
-data <- read_csv(here("data/weighted_prev_competence_111220.csv"))
-tree <- ape::read.nexus(here("data/asup_just_tree.txt"))
-names <- read_csv(here("data/species_names_ids.csv"))
+data <- read_csv(here("raw_data/weighted_prev_competence_111220.csv"))
+tree <- ape::read.nexus(here("raw_data/asup_just_tree.txt"))
+names <- read_csv(here("raw_data/species_names_ids.csv"))
 ```

### 2022-09-09 scripts/figures/fig2_code.R
comment out save

@@ -23,4 +23,4 @@ figure_2 <- comm_data %>% ggplot(.,aes(x=pc1Rank,y=cc))+
 
 figure_2
   
-ggsave("fig_2.png",plot=figure_2,device="png",path=here("figures"))
+#ggsave("fig_2.png",plot=figure_2,device="png",path=here("figures"))

### 2022-09-09 scripts/figures/fig3_code.R

@@ -69,7 +69,7 @@ figure_3 <- ggplot(data=lag_evenness,aes(x=lag_J, y=lag_cc)) +
         legend.text = element_text(size=10))
 
 figure_3
-ggsave("fig_3.png",plot=figure_3,device="png",path=here("figures"))
+#ggsave("fig_3.png",plot=figure_3,device="png",path=here("figures"))
 
 
 

### 2022-09-09 scripts/figures/fig4_code.R

@@ -83,5 +83,5 @@ newTree <- vl_tree +
 figure_4 <- cc_plot/RA_plot/newTree
 figure_4
 
-ggsave("fig_4.png",plot=figure_4,device="png",path=here("figures"))
+#ggsave("fig_4.png",plot=figure_4,device="png",path=here("figures"))
 

### 2022-09-09 scripts/figures/fig5_code.R

@@ -78,7 +78,7 @@ p.adjust(pvals,method="holm")
 #plot everything together with patchwork
 figure_5 <- cc_corr| size_corr| temp_corr
 figure_5
-ggsave("fig_5.png",plot=figure_5,device="png",path=here("figures"))
+#ggsave("fig_5.png",plot=figure_5,device="png",path=here("figures"))
 
 
 

### 2022-09-12 ms/manuscript.Rmd
add rmd for manuscript

---
title: "manuscript"
author: "Daniel Suh"
date: "2022-09-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2022-09-12 ms/manuscript.Rmd
make template

@@
 author: "Daniel Suh"
-date: "2022-09-12"
-output: html_document
+date: "`r format(Sys.time(), '%B %d, %Y')`"
+output: pdf_document
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
+# Introduction
 
-## R Markdown
+# Methods
 
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+# Results
 
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
+# Discussion
 
-```{r cars}
-summary(cars)
-```
-
-## Including Plots
-
-You can also embed plots, for example:
-
-```{r pressure, echo=FALSE}
-plot(pressure)
-```
-
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+# References

### 2022-09-12 scripts/workshop.Rmd
bray curtis dissimilarity

---
title: "workshop"
author: "Daniel Suh"
date: "2022-09-12"
output: html_document
---

```{r}
library(here)

source(here("base","src.R"))
```

```{r}
data <- read_csv(here("raw_data/weighted_prev_competence_111220.csv"))
tree <- ape::read.nexus(here("raw_data/asup_just_tree.txt"))
names <- read_csv(here("raw_data/species_names_ids.csv"))
```

```{r}
comm_summ <- data %>% dplyr::select(WetAltID, Month.1, Month, cc, Month, AB2:AB42, Prevalence, MeanWaterTempPredC) %>% 
  dplyr::select(-ABAmb, -ABSal) %>%  
  mutate(., size = rowSums(.[5:25]), 
    ABHigh = (AB9 + AB21 + AB26 + AB42), 
    ABLow = (AB2 + AB3 + AB4 + AB5 + AB6 + AB8 + AB20 + AB24 + AB27 + AB28 + AB29 + AB31 + AB34 + AB35 + AB38 + AB39 + AB41)) %>% #get totals for community size
  mutate(., siteID = paste(WetAltID, Month.1, sep = "_")) %>%
  distinct()
comm_summ$siteID <- gsub("Month", "", comm_summ$siteID)
```

```{r}
community_mat <- data %>% dplyr::select(WetAltID, Month.1, AB2:AB8, AB9, AB20:AB42) 
community_mat$Month <- gsub("Month", "", community_mat$Month.1)
community_mat %<>% mutate(., siteID = paste(WetAltID, Month, sep = "_")) %>% distinct() %>% arrange(.,WetAltID) %>% dplyr::select(siteID, AB2:AB42)

#make abundance and pres/abs matrices with rownames as siteIDs
abundance_mat <- community_mat %>% column_to_rownames(., var = "siteID")
presence_mat <- as.data.frame(ifelse(abundance_mat>0, 1, 0))
```

```{r}
log_abun_mat <- log1p(abundance_mat)

bc_dist <- vegdist(log_abun_mat, method = "bray") #bray-curtis dissimilarity is default

bc_dist_mat <- as.matrix(bc_dist)
bc_dist_df <- melt(bc_dist_mat, varnames=c("row", "col"))

ggplot(bc_dist_df, aes(x = col, y = row, fill = value)) + geom_tile()
```

### 2022-09-12 scripts/figures/fig2_code.R

@@ -21,6 +21,17 @@ figure_2 <- comm_data %>% ggplot(.,aes(x=pc1Rank,y=cc))+
   scale_color_viridis_d(direction = -1) +
   theme(text = element_text(size=18))
 
+comm_data %>% ggplot(., aes(x=pc1Rank, y=cc)) +
+  geom_point(aes(color = factor(WetAltID), shape = factor(WetAltID), size = size)) + 
+  facet_wrap(vars(Month.1), nrow = 6)+
+  scale_color_manual(values = rep(1:20, len = 20)) +
+  scale_shape_manual(values = rep(1:20, len = 20)) +
+  scale_size_continuous(range = c(2,10)) +
+  guides(shape=F) +
+  labs(x="Principal Component 1 Rank (Community Similarity)", y = "Community Competence (CC)", size = "Host\nAbundance") + 
+  theme_classic() +
+  theme(text = element_text(size=18))
+  
 figure_2
   
 #ggsave("fig_2.png",plot=figure_2,device="png",path=here("figures"))

### 2022-09-12 scripts/workshop.Rmd
really very confusing stuff

@@
 ```
+
+# Questions
+
+Do communities that are similar in community composition have similar community competence values?
+
+Can communities with different community compositions have similar community competence values?
+
+How are these communities distributed across space and time?
+i.e. how much spatial and temporal autocorrelation is there between sites in terms of community composition and community competence?
 
@@
 
+Use Bray-Curtis distances to analyze how similar sites are to each other
+Higher BC distance means more dissimilar
+1 means perfect mismatch
 ```{r}
@@
 bc_dist_mat <- as.matrix(bc_dist)
-bc_dist_df <- melt(bc_dist_mat, varnames=c("row", "col"))
+bc_dist_df <- melt(bc_dist_mat, varnames=c("row", "col"), value.name = "bc_dist")
 
-ggplot(bc_dist_df, aes(x = col, y = row, fill = value)) + geom_tile()
+ggplot(bc_dist_df, aes(x = col, y = row, fill = bc_dist)) + geom_tile()
 ```
+
+Manhattan method is the absolute value of the difference between two values
+So higher distance means larger difference
+Normalized to have range 0 to 1
+```{r}
+cc_diff <- comm_summ %>% dplyr::select(siteID, cc) %>% mutate(log_cc = log1p(cc)) %>% dplyr::select(-cc)
+cc_dist <- dist(cc_diff, method = "manhattan")
+cc_diff_mat <- as.matrix(cc_dist, labels = T)
+colnames(cc_diff_mat) <- cc_diff[[1]]
+rownames(cc_diff_mat) <- cc_diff[[1]]
+cc_diff_df <- melt(cc_diff_mat, varnames=c("row", "col"), value.name = "cc_dist") %>% 
+  mutate(cc_norm = (cc_dist-min(.$cc_dist))/(max(.$cc_dist)-min(.$cc_dist)))
+
+ggplot(cc_diff_df, aes(x = col, y = row, fill = cc_norm)) + geom_tile()
+```
+
+```{r}
+bc_cc_df <- left_join(cc_diff_df, bc_dist_df)
+bc_cc_df %<>% mutate(diff = abs(bc_dist-cc_norm))
+
+ggplot(bc_cc_df, aes(x = col, y = row, fill = diff)) + geom_tile()
+
+bc_cc_df %>% filter(bc_dist!=1) %>% ggplot(.,aes(x=cc_norm, y=bc_dist)) + geom_point() + geom_smooth(method = "lm")
+```
+
+Principal Component Ranking may not be reliable as a proxy for community similarity because of double-zeroes
+
+If two sites both have zeroes for a species then that does not show that a site is similar but only shows that both sites may be unsuitable for the species
+
+Bray-Curtis dissimilarity does not face this limitation
+
+But the question is how similar are sites with the similar community competence?
+
+The manhattan distance is equivalent to bray-curtis distance

### 2022-09-12 scripts/workshop.Rmd
cc and bc distances

@@
 
+bc_dist_df %>% filter(bc_dist!=0) %>% ggplot(., aes(x=bc_dist)) + geom_histogram()
 ggplot(bc_dist_df, aes(x = col, y = row, fill = bc_dist)) + geom_tile()
+```
+
+quick check on normality of community competence data
+```{r}
+comm_summ %>% ggplot(.,aes(x=cc)) + geom_histogram()
 ```
@@
 So higher distance means larger difference
-Normalized to have range 0 to 1
+Normalized distances to have range 0 to 1
 ```{r}
@@
 
+cc_diff_df %>% filter(cc_norm != 0) %>% ggplot(., aes(x=cc_norm)) + geom_histogram()
 ggplot(cc_diff_df, aes(x = col, y = row, fill = cc_norm)) + geom_tile()
@@
 bc_cc_df <- left_join(cc_diff_df, bc_dist_df)
-bc_cc_df %<>% mutate(diff = abs(bc_dist-cc_norm))
+```
+
+Each point is a pairwise interaction
+
+The null hypothesis is that community competence is achieved through specific community compositions. i.e. you cannot achieve the same community competence with different community compositions
+
+So if a pair of sites have very different community competence then they should also have very different community composition
+
+But if community competence can be achieved through a variety of different community compositions then a pair of sites with similar community competence can have a high level of dissimilarity
+
+So we are looking for low cc distance between two sites but high bc distance
+An upward trend is expected because more dissimilar sites in terms of competence probably have different composition
+```{r}
+bc_cc_df %>% filter(diff!=0) %>% ggplot(.,aes(x=cc_norm, y=bc_dist)) + geom_point() + geom_smooth(method = "lm")
+```
+
+If diff is set as community competence similarity minus community composition similarity then a low value will mean the difference between composition similarity is large compared to competence similarity.
+Either competence similarity is very high (cc dist is low) and composition similarity (bc dist) is moderate or high OR composition similarity is low (i.e.
+BC dist is high) and competence similarity (cc dist is moderate or low)
+
+```{r}
+bc_cc_df %<>% mutate(diff = cc_norm-bc_dist)
 
@@
 
-bc_cc_df %>% filter(bc_dist!=1) %>% ggplot(.,aes(x=cc_norm, y=bc_dist)) + geom_point() + geom_smooth(method = "lm")
+bc_cc_df %>% filter(diff!=0) %>% ggplot(.,aes(x=diff)) + geom_histogram()
 ```

### 2022-09-13 scripts/workshop.Rmd
knit to html
change date to sys date

@@
 author: "Daniel Suh"
-date: "2022-09-12"
+date: "`r format(Sys.time(), '%d %B, %Y')`"
 output: html_document
@@
 source(here("base","src.R"))
+
+library(reshape2)
 ```
@@
 bc_cc_df <- left_join(cc_diff_df, bc_dist_df)
+
+bc_cc_df %<>% mutate(diff = cc_norm-bc_dist)
+
 ```
@@
 ```{r}
-bc_cc_df %<>% mutate(diff = cc_norm-bc_dist)
 

### 2022-09-13 scripts/workshop.Rmd
update text
update with fig 2
add text about next steps

@@
 ```
+
+![motivating figure](https://dcsuh.github.io/rana-competence/figures/fig_2.png)
 
@@
 
-The manhattan distance is equivalent to bray-curtis distance
+sometimes similar and sometimes not
+
+## next step
+
+do PCA after Hellinger transformation of data.
+PCA has the double zero problem but Hellinger distance avoids that problem and avoids overweighting of abundance data by turning absolute species abundances into relative abundances.
+more info [here](https://davidzeleny.net/blog/2022/03/17/euclidean-distance-is-sensitive-to-double-zero-problem-while-hellinger-is-not-visualization/)

### 2022-09-13 scripts/workshop.Rmd
do transformation based pca

@@
 ```{r}
-bc_cc_df %>% filter(diff!=0) %>% ggplot(.,aes(x=cc_norm, y=bc_dist)) + geom_point() + geom_smooth(method = "lm")
+bc_cc_df %>% filter(diff!=0) %>% ggplot(.,aes(x=bc_dist, y=cc_norm)) + geom_point() + geom_smooth(method = "lm")
 ```
@@
 more info [here](https://davidzeleny.net/blog/2022/03/17/euclidean-distance-is-sensitive-to-double-zero-problem-while-hellinger-is-not-visualization/)
+
+```{r}
+pca <- princomp(abundance_mat)
+
+biplot(pca)
+
+abundance_hellinger <- decostand(log1p(abundance_mat), 'hellinger')
+
+tb_pca <- princomp(abundance_hellinger)
+
+biplot(tb_pca)
+```
+
+```{r}
+tb_pca_scores <- as.data.frame(tb_pca$scores) %>% dplyr::select(Comp.1, Comp.2) %>% rownames_to_column(., var = "siteID")
+
+tb_pca_scores_summ <- left_join(comm_summ, tb_pca_scores, by = "siteID")
+
+tb_pca_scores_summ$Month <- factor(tb_pca_scores_summ$Month, levels = c("Feb", "Mar", "Apr", "May", "Jun", "Jul"))
+
+tb_pca_scores_summ %<>% mutate(pc1Rank=dense_rank(Comp.1))
+
+tb_pca_scores_summ %>% ggplot(.,aes(x=pc1Rank, y=cc)) + geom_point()
+
+tb_pca_scores_summ %>% ggplot(.,aes(x=Comp.1, y=cc)) + geom_point() + geom_smooth(method = "loess")
+
+```
+
+### Let's try to explain the residuals
+
+The residuals from the cc_norm ~ bc_dist represent when the community competence is more or less similar than expected given the similarity of the community.
+If community composition is not explaining similarity in community competence then this must be due to the two communities having

### 2022-09-13 scripts/workshop.Rmd
tb_pca version of fig 2

@@
 
+tb_pca_scores_summ %>% ggplot(.,aes(x=pc1Rank, y=cc)) + geom_point(aes(color = Month, shape = factor(WetAltID), size = size))+
+  labs(x="Principal Component 1", y = "Community Competence (CC)", size = "Host\nAbundance") + 
+  scale_shape_manual(values = rep(1:20, len = 20)) +
+  scale_size_continuous(range = c(2,10)) +
+  guides(shape=F) +
+  theme_classic() +
+  theme(aspect.ratio=1/1.618, legend.box = "horizontal") +
+  scale_color_viridis_d(direction = -1) +
+  theme(text = element_text(size=18))
+
 ```

### 2022-09-13 scripts/workshop.Rmd
analysis of residuals

@@
 The residuals from the cc_norm ~ bc_dist represent when the community competence is more or less similar than expected given the similarity of the community.
-If community composition is not explaining similarity in community competence then this must be due to the two communities having
+If community composition is not explaining similarity in community competence then this must be due to the communities having a tradeoff where competence is derived from a different class of hosts (low competence vs.
+high competence; abundant low competence hosts can result in community competence that is similar to rare high competent host) or a substitution of hosts (proportion of low vs.
+high competence hosts may be similar but the identity of those hosts may be different).
+These situations will result in community competence that is unexplained by community composition (i.e.
+high or low residual).
+
+plot residuals on y-axis and pairwise difference in evenness (tradeoff) and different in proportion high-low competence (substitution) on x-axis
+```{r}
+comm_data <- readRDS(here("processed_data","comm_data.rds"))
+
+delta_j <- comm_data %>% select(siteID, J)
+
+delta_p <- comm_data %>% select(siteID, ABHigh, ABLow) %>% mutate(P = ABHigh/(ABHigh+ABLow)) %>% select(-c(ABHigh, ABLow))
+
+```
+
+```{r}
+j_dist <- dist(delta_j, method = "manhattan")
+j_dist_mat <- as.matrix(j_dist, labels = T)
+colnames(j_dist_mat) <- delta_j[[1]]
+rownames(j_dist_mat) <- delta_j[[1]]
+
+j_dist_df <- melt(j_dist_mat, varnames=c("row", "col"), value.name = "j_dist")
+```
+
+```{r}
+p_dist <- dist(delta_p, method = "manhattan")
+p_dist_mat <- as.matrix(p_dist, labels = T)
+colnames(p_dist_mat) <- delta_p[[1]]
+rownames(p_dist_mat) <- delta_p[[1]]
+
+p_dist_df <- melt(p_dist_mat, varnames=c("row", "col"), value.name = "p_dist")
+```
+
+```{r}
+dist_df <- bc_cc_df %>% left_join(.,j_dist_df)
+dist_df %<>% left_join(.,p_dist_df)
+```
+
+```{r}
+model <- lm(cc_norm ~ bc_dist, data = dist_df)
+
+cc_bc_resid <- tibble(row = dist_df[[1]], col = dist_df[[2]], resid = residuals(model))
+
+dist_df %<>% left_join(.,cc_bc_resid)
+```
+
+```{r}
+dist_df %>% ggplot(.,aes(x=j_dist, y=resid)) + geom_point()
+
+dist_df %>% ggplot(.,aes(x=p_dist, y=resid)) + geom_point()
+```

### 2022-09-14 scripts/gradient_analysis.Rmd
move files

--- 2020-11-20 gradient_analysis.Rmd
@@
 
+Indirect Grdient Analysis using PCA
+explanation of how envfit actually works -> https://stackoverflow.com/questions/60953996/how-are-envfit-results-created 
+```{r}
+#sites ordinated by species abundances
+#cor = F to use correlation matrix because variables are all on same scale
+species_pca <- princomp(community_mat, cor = F)
+#plot(species_pca$scores, pch = 16)
+summary(species_pca)
+#loadings(species_pca)
+#scores(species_pca)
+biplot(species_pca)
+
+env_fit_vectors <- envfit(species_pca, env_mat)
+
+biplot(species_pca)
+plot(env_fit_vectors)
+```
+
 What is the difference between the scores for species from envfit and the values from the correlation between environmental ordination scores and the community matrix?

### 2022-09-14 process/competencyCode.R
update to fix temperature

@@ -1,5 +1,8 @@
 #Script originated by Andrew Park and edited by Daniel Suh
 
+
+#Run this script after 'relative_abundance.R'
+
 library(tidyverse)
 library(magrittr)
 library(here)

### 2022-09-14 process/relative_abundance.R

@@ -2,6 +2,10 @@
 ##Script Originator: Daniel Suh
 ##Date: 11/12/20
 
+
+#Process this script first
+
+
 #load packages
 library(tidyverse)
 library(magrittr)

### 2022-09-14 process/weighted_prev.R

@@ -1,5 +1,8 @@
 #Script originator: Daniel Suh
 
+#Run this script after competencyCode.R
+
+
 #This script recalculates prevalence of ranavirus at each site by estimating a prevalence value that is weighted by overall abundances at that site
 #i.e. sum of ((known prevalence in species n) X (relative abundance of species n)) / (total individuals in community) = weighted prevalence
 

### 2022-09-14 scripts/figures/fig5_code.R

@@ -78,7 +78,7 @@ p.adjust(pvals,method="holm")
 #plot everything together with patchwork
 figure_5 <- cc_corr| size_corr| temp_corr
 figure_5
-#ggsave("fig_5.png",plot=figure_5,device="png",path=here("figures"))
+ggsave("fig_5.png",plot=figure_5+ proj_theme,width = outwidth[1], height = outwidth[1]/golden,device="png",path=here("figures"))
 
 
 

### 2022-09-14 scripts/figures/fig5_code.R
Update fig5_code.R

@@ -78,7 +78,7 @@ p.adjust(pvals,method="holm")
 #plot everything together with patchwork
 figure_5 <- cc_corr| size_corr| temp_corr
 figure_5
-ggsave("fig_5.png",plot=figure_5+ proj_theme,width = outwidth[1], height = outwidth[1]/golden,device="png",path=here("figures"))
+ggsave("fig_5.png",plot=figure_5,width = outwidth[1], height = outwidth[1]/golden,device="png",path=here("figures"))
 
 
 

### 2022-09-14 scripts/figures/fig1_code.R
update fig1

@@ -155,4 +155,4 @@ p5 <- combined %>%
 
 figure_1 <- (p2 | p3)/(p4 | p5)
 figure_1
-#ggsave("fig_1.png",plot=figure_1,device="png",path=here("figures"),width = 11,height = 11.6,units = "cm")
+ggsave("fig_1.png",plot=figure_1+ proj_theme,width = outwidth[1], height = outwidth[1]/golden,device="png",path=here("figures"))

### 2022-09-14 scripts/figures/fig4_code.R
update fig4

@@ -83,5 +83,5 @@ newTree <- vl_tree +
 figure_4 <- cc_plot/RA_plot/newTree
 figure_4
 
-ggsave("fig_4.png",plot=figure_4+ proj_theme,width = outwidth[1], height = outwidth[1]/golden,device="png",path=here("figures"))
+ggsave("fig_4.png",plot=figure_4,width = outwidth[1], height = outwidth[1]/golden,device="png",path=here("figures"))
 

### 2022-09-14 scripts/workshop/alt_fig1.Rmd
move files around

---
title: "alternate fig 1"
author: "Daniel Suh"
date: "2/22/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = F}
library(tidyverse)
library(magrittr)
#install_github("thomasp85/patchwork")
library(patchwork)
library(ggnewscale)

```

```{r}
#define function contour(): inputs are parameters and output is values for grid search for R0
contour <- function(trans_1_min, trans_1_max, trans_3_min, trans_3_max, mort1, mort2, degr) {
  #expand data
  data <- expand_grid("trans1" = seq(from = trans_1_min, to = trans_1_max, by = 0.00001),
                      "trans2" = c(0.0001),
                      "trans3" = seq(from = trans_3_min, to = trans_3_max, by = 0.00001),
                      "shed1" = c(rep(1/2, length("trans1"))),
                      "shed2" = c(rep(1/2, length("trans1"))),
                      "reco1" = c(rep(1/10, length("trans1"))),
                      "reco2" = c(rep(1/10, length("trans1"))),
                      "mort1" = c(mort1), 
                      "mort2" = c(mort2), 
                      "degr" = c(degr), 
                      "birth1" = c(rep(5/3, length("trans1"))),
                      "birth2" = c(rep(5/3, length("trans1"))))

  #calculate disease-free equilibria for use in next-gen matrix
  data %<>% add_column(dfe1 = (data$birth1/data$mort1),dfe2 = (data$birth2/data$mort2), eigen = NA) 
  #make matrix
  mat <- matrix(1, nrow = 3, ncol = 3)

  m <- nrow(data)

  #calculate eigenvalue using next-gen matrix. max eigenvalue is R0
  for(n in 1:m){
    mat[1] = (data$trans1[n]*data$dfe1[n])/(data$reco1[n]+data$mort1[n])
    mat[2] = (data$trans1[n]*data$dfe2[n])/(data$reco2[n]+data$mort2[n])
    mat[3] = (data$shed1[n]/data$degr[n])
    mat[4] = (data$trans2[n]*data$dfe1[n])/(data$reco1[n]+data$mort1[n])
    mat[5] = (data$trans2[n]*data$dfe2[n])/(data$reco2[n]+data$mort2[n])
    mat[6] = (data$shed2[n]/data$degr[n])
    mat[7] = (data$trans3[n]*data$dfe1[n])/(data$reco1[n]+data$mort1[n])
    mat[8] = (data$trans3[n]*data$dfe2[n])/(data$reco2[n]+data$mort2[n])
    mat[9] = 0
    eigen <- max(eigen(mat)$values)
    data$eigen[n] <- eigen
  }

  data$span1 <- 1/data$mort1 #lifespan for species 1
  data$span2 <- 1/data$mort2 #lifespan for species 2
  data$half <- log(1/2)/-data$degr #viral half-life
  data$epid <- data$eigen > 1 #invasion potential (i.e. R0 > 1)
  data$prop1 <- data$trans1/data$trans2 #proportion of transmission of species 2 over species 1
  data$prop3 <- data$trans3/data$trans2 #proportion of environmental transmission over species 1
  data$prop_mort <- data$mort1/data$mort2 #relative mortality for species 1 and 2
  data$cap1 <- data$birth1/data$mort1 #carrying capacity for species 1
  data$cap2 <- data$birth2/data$mort2 #carrying capacity for species 2
  data$cap_prop <- data$cap1/data$cap2 #relative carrying capacities for species 1 and 2
  data$tot <- data$cap1 + data$cap2 #total abundance for species 1 and 2
  data$cc <- data$trans1*(data$cap1/data$tot) + data$trans2*(data$cap2/data$tot) #cc defined as transmission of a species multiplied by its relative abundance
  output <- data %>% dplyr::select(trans1, trans2, prop1, prop3, trans3, eigen, prop_mort, mort1, mort2, degr, half, cap1, cap2, cap_prop, tot, cc) %>% distinct()
  return(output)
}
```

```{r}
reference <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0,trans_3_max = 0.001,
                     mort1 = 1/45,mort2 = 1/45,degr = 1/1.947799)

composition <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0,trans_3_max = 0.001,
                       mort1 = 1/60,mort2 = 1/30,degr = 1/1.947799)

size <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0,trans_3_max = 0.001,
                mort1 = c(1/52.5),mort2 = c(1/52.5),degr = 1/1.947799)

halflife <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0,trans_3_max = 0.001,
                    mort1 = c(1/40),mort2 = 1/40,degr = 1/3.895598)

combined <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0,trans_3_max = 0.001,
                    mort1 = c(1/70),mort2 = c(1/35),degr = 1/3.895598)
```

```{r}
axis_text_size = 8
plot_label_size = 20

p1 <- reference %>% ggplot(.,aes(x=prop3,y=prop1))+
  geom_contour(aes(x=prop3,y=prop1,z=eigen,colour = factor(..level.. == 1,levels = c(F,T)),group=mort1), show.legend = FALSE)+
  theme_classic()+
  scale_colour_manual(na.value = NA, values = c(NA, "red")) + 
  labs(x = "Environmental", y = "Contact") +
  xlim(1,8) +
  ylim(1,14) + 
  labs(title = "Reference")

p2 <- composition %>% ggplot(.,aes(x=prop3,y=prop1))+
  geom_contour(aes(x=prop3,y=prop1,z=eigen,colour = factor(..level.. == 1,levels = c(F,T)),group=mort1), show.legend = FALSE)+
  theme_classic()+
  scale_colour_manual(na.value = NA, values = c(NA, "red")) + 
  new_scale_color()+
  geom_contour(data = reference, 
               aes(x=prop3,y=prop1,z=eigen,colour = factor(..level.. == 1,levels = c(F,T)),group=mort1), show.legend = F)+
  scale_colour_manual(na.value = NA, values = c(NA, "gray")) +
  labs(x = "", y = "") +
  xlim(1,8) +
  ylim(1,14) + 
  labs(title = "A") + 
  theme(axis.title = element_text(size=axis_text_size),
        plot.title = element_text(size=plot_label_size))
#p2

p3 <- size %>% filter(tot != 150) %>% 
  ggplot(.,aes(x=prop3,y=prop1))+
  geom_contour(aes(x=prop3,y=prop1,z=eigen,colour = factor(..level.. == 1,levels = c(F,T)),group=mort1), show.legend = FALSE)+
  theme_classic()+
  scale_colour_manual(na.value = NA, values = c(NA, "orange")) + 
  new_scale_color()+
  geom_contour(data = reference, 
               aes(x=prop3,y=prop1,z=eigen,colour = factor(..level.. == 1,levels = c(F,T)),group=mort1), show.legend = F)+
  scale_colour_manual(na.value = NA, values = c(NA, "gray")) +
  labs(x = "", y = "") +
  xlim(1,8) +
  ylim(1,14) + 
  labs(title = "B") + 
  theme(plot.title = element_text(size=plot_label_size))
#p3

p4 <- halflife %>% ggplot(.,aes(x=prop3,y=prop1))+
  geom_contour(aes(x=prop3,y=prop1,z=eigen,colour = factor(..level.. == 1,levels = c(F,T)),group=mort1), show.legend = FALSE)+
  scale_colour_manual(na.value = NA, values = c(NA, "forestgreen")) + 
  new_scale_color()+
  geom_contour(data = reference, 
               aes(x=prop3,y=prop1,z=eigen,colour = factor(..level.. == 1,levels = c(F,T)),group=mort1), show.legend = F)+
  scale_colour_manual(na.value = NA, values = c(NA, "gray")) +
  labs(x="Relative Environmental", y = "                                                               Relative Contact Transmission Rate", title = "C") +
  xlim(1,8) +
  ylim(1,14) + 
  theme_classic() +
  theme(axis.title = element_text(size=axis_text_size),
        plot.title = element_text(size=plot_label_size),
        axis.title.x=element_text(hjust=1.05))
#p4

p5 <- combined %>%
  ggplot(.,aes(x=prop3,y=prop1)) +
  geom_contour(aes(x=prop3,y=prop1,z=eigen,colour = factor(..level.. == 1,levels = c(F,T)),group=mort1), show.legend = FALSE) +
  scale_colour_manual(na.value = NA, values = c(NA, "purple")) + 
  new_scale_color()+
  geom_contour(data = reference, 
               aes(x=prop3,y=prop1,z=eigen,colour = factor(..level.. == 1,levels = c(F,T)),group=mort1), show.legend = F) +
  scale_colour_manual(na.value = NA, values = c(NA, "gray")) +
  labs(x = "Transmission Rate", y = "", title = "D") +
  xlim(1,8) +
  ylim(1,14) + 
  theme_classic() +
  theme(axis.title = element_text(size=axis_text_size),
        plot.title = element_text(size=plot_label_size),
        axis.title.x=element_text(hjust=-0.625))
#p5
```

```{r}
figure_1 <- (p2 | p3)/(p4 | p5)
figure_1 + theme(title = "Original fig w/ lines for R0==1")
```

```{r}
#heatmaps and contour plots of difference between control and treatment

ref_sel <- reference %>% select(prop3,prop1,eigen) %>% rename(.,ref_eigen = eigen)
comp_sel <- composition %>% select(prop3,prop1,eigen) %>% rename(.,comp_eigen = eigen)
size_sel <- size %>% select(prop3,prop1,eigen) %>% rename(.,size_eigen = eigen)
halflife_sel <- halflife %>% select(prop3,prop1,eigen) %>% rename(.,half_eigen = eigen)
combined_sel <- combined %>% select(prop3,prop1,eigen) %>% rename(.,all_eigen = eigen)

ref_comp <- inner_join(ref_sel,comp_sel, by = c("prop3","prop1"))
ref_comp %<>% inner_join(.,size_sel, by =c("prop3","prop1"))
ref_comp %<>% inner_join(.,halflife_sel, by =c("prop3","prop1"))
ref_comp %<>% inner_join(.,combined_sel, by =c("prop3","prop1"))

ref_comp %<>% mutate(.,comp_diff = comp_eigen-ref_eigen,
                     size_diff = size_eigen-ref_eigen,
                     half_diff = half_eigen-ref_eigen,
                     all_diff = all_eigen-ref_eigen)

#heatmaps with contours

comp_fig <- ref_comp %>% ggplot(.,aes(x=prop3,y=prop1,fill=comp_diff)) + 
  geom_tile() +
  scale_fill_gradient2() + 
  geom_contour(aes(z=comp_diff))

size_fig <- ref_comp %>% ggplot(.,aes(x=prop3,y=prop1,fill=size_diff)) + 
  geom_tile() +
  scale_fill_gradient2() + 
  geom_contour(aes(z=size_diff))

half_fig <- ref_comp %>% ggplot(.,aes(x=prop3,y=prop1,fill=half_diff)) + 
  geom_tile() +
  scale_fill_gradient2() + 
  geom_contour(aes(z=half_diff))

combined_fig <- ref_comp %>% ggplot(.,aes(x=prop3,y=prop1,fill=all_diff)) + 
  geom_tile() +
  scale_fill_gradient2() + 
  geom_contour(aes(z=all_diff))
```

```{r}
(comp_fig | size_fig)/(half_fig | combined_fig) + theme(title = "heatmap of difference between treatment and reference")
```

```{r}
#filled contour with standardized breaks

breaks <- c(-0.2,-0.1,0,0.1,0.2,0.3,0.4,0.5,2)

comp_fig <- ref_comp %>% ggplot(.,aes(x=prop3,y=prop1)) + 
  geom_contour_filled(aes(z=comp_diff),breaks = breaks) +
  scale_fill_viridis_d(drop=F) +
  theme(legend.position = "none") + 
  labs(x="environmental transmission", y = "contact transmission", title = "composition")

size_fig <- ref_comp %>% ggplot(.,aes(x=prop3,y=prop1)) + 
  geom_contour_filled(aes(z=size_diff),breaks = breaks) +
  scale_fill_viridis_d(drop=F) +
  theme(legend.position = "none") + 
  labs(x="environmental transmission", y = "contact transmission", title = "abundance")

half_fig <- ref_comp %>% ggplot(.,aes(x=prop3,y=prop1)) + 
  geom_contour_filled(aes(z=half_diff),breaks = breaks) +
  scale_fill_viridis_d(drop=F) +
  theme(legend.position = "none") + 
  labs(x="environmental transmission", y = "contact transmission", title = "halflife")

combined_fig <- ref_comp %>% ggplot(.,aes(x=prop3,y=prop1)) + 
  geom_contour_filled(aes(z=all_diff),breaks = breaks) +
  scale_fill_viridis_d(drop=F) +   
  labs(x="environmental transmission", y = "contact transmission", title = "combined")
```

```{r}
(comp_fig | size_fig)/(half_fig | combined_fig) + theme("filled contours of diff between treatment and reference")
```

```{r}
#comparison between combined treatment and the additive effect

ref_comp %<>% rowwise %>% mutate(diff_sum = comp_diff + size_diff + half_diff)
ref_comp %<>% rowwise %>% mutate(effect = all_diff-diff_sum)

# ref_comp %>% ggplot(.,aes(x=prop3,y=prop1,fill=effect)) + 
#   geom_tile() +
#   scale_fill_gradient2() + geom_contour(aes(z=effect))
# 
# ref_comp %>% ggplot(.,aes(x=prop3,y=prop1)) + 
#   geom_contour_filled(aes(z=effect))

breaks <- c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9)

comb <- ref_comp %>% ggplot(.,aes(x=prop3,y=prop1)) + 
  geom_contour_filled(aes(z=all_diff),breaks = breaks) +
  scale_fill_viridis_d(drop=F) +   
  theme(legend.position = "none") + 
  labs(x="", y = "", title = "combined")

sum <- ref_comp %>% ggplot(.,aes(x=prop3,y=prop1)) + 
  geom_contour_filled(aes(z=diff_sum),breaks = breaks) +
  scale_fill_viridis_d(drop=F) + 
  labs(x = "", y = "contact transmission", title = "sum of its parts")

diff <- ref_comp %>% ggplot(.,aes(x=prop3,y=prop1)) + 
  geom_contour_filled(aes(z=effect),breaks = breaks) +
  scale_fill_viridis_d(drop=F) + 
  theme(legend.position = "none") +
  labs(x = "environmental", y = "", title = "difference (combined - sum of its parts)")

comb / sum / diff 
```

```{r}
#comparison between combined treament and additive effect by percent change

ref_percent <- ref_comp %>% mutate(., comp_per = comp_diff/ref_eigen,
                                   size_per = size_diff/ref_eigen,
                                   half_per = half_diff/ref_eigen,
                                   all_per = all_eigen/ref_eigen,
                                   sum_per = diff_sum/ref_eigen)

breaks <- c(0:20)*0.1

comb_per <- ref_percent %>% ggplot(.,aes(x=prop3,y=prop1)) + 
  geom_contour_filled(aes(z=all_per),breaks = breaks) +
  scale_fill_viridis_d(drop=F) +   
  theme(legend.position = "none") + 
  labs(x="", y = "", title = "combined")

sum_per <- ref_percent %>% ggplot(.,aes(x=prop3,y=prop1)) + 
  geom_contour_filled(aes(z=sum_per),breaks = breaks) +
  scale_fill_viridis_d(drop=F) + 
  labs(x = "", y = "contact transmission", title = "sum of its parts")

diff_per <- ref_percent %>% ggplot(.,aes(x=prop3,y=prop1)) + 
  geom_contour_filled(aes(z=all_per-sum_per),breaks = breaks) +
  scale_fill_viridis_d(drop=F) + 
  theme(legend.position = "none") +
  labs(x = "environmental", y = "", title = "difference (combined - sum of its parts)")

comb_per / sum_per / diff_per 
```

### 2022-09-14 scripts/data_format.R

@@ -1,148 +1,0 @@
-## ------------------------------------------------------------------------
-library(tidyverse)
-library(magrittr)
-library(vegan)
-library(patchwork)
-library(ggnewscale)
-library(here)
-
-
-## ------------------------------------------------------------------------
-here()
-data <- read_csv(here("/data/weighted_prev_competence_111220.csv"))
-names <- read_csv(here("/data/species_names_ids.csv"))
-
-
-## ------------------------------------------------------------------------
-comm_summ <- data %>% dplyr::select(WetAltID, Month.1, Month, cc, Month, AB2:AB42, Prevalence, MeanWaterTempPredC) %>% 
-  dplyr::select(-ABAmb, -ABSal) %>%  
-  mutate(., size = rowSums(.[5:25]), 
-    ABHigh = (AB9 + AB21 + AB26 + AB42), 
-    ABLow = (AB2 + AB3 + AB4 + AB5 + AB6 + AB8 + AB20 + AB24 + AB27 + AB28 + AB29 + AB31 + AB34 + AB35 + AB38 + AB39 + AB41)) %>% #get totals for community size
-  mutate(., siteID = paste(WetAltID, Month.1, sep = "_")) %>%
-  distinct()
-comm_summ$siteID <- gsub("Month", "", comm_summ$siteID)
-
-
-## ------------------------------------------------------------------------
-vl <- data %>% dplyr::select(vl.4:vl.42) %>% distinct() %>% pivot_longer(vl.4:vl.42)
-vl %<>% mutate(species_code = as.double(gsub("vl.","",vl$name)))
-vl %<>% full_join(., names, by="species_code")
-vl %<>% mutate(name = gsub("_"," ",vl$species_name))
-vl$value %<>% na_if(.,0)
-vl %<>% mutate(., ln_value = log(value)) %>% mutate(., log10_value = log10(value))
-vl %<>% replace_na(.,list(value=0,ln_value=0,log10_value=0))
-
-
-## ------------------------------------------------------------------------
-community_mat <- data %>% dplyr::select(WetAltID, Month.1, AB2:AB8, AB9, AB20:AB42) 
-community_mat$Month <- gsub("Month", "", community_mat$Month.1)
-community_mat %<>% mutate(., siteID = paste(WetAltID, Month, sep = "_")) %>% distinct() %>% arrange(.,WetAltID) %>% dplyr::select(siteID, AB2:AB42)
-
-#make abundance and pres/abs matrices with rownames as siteIDs
[…]

### 2022-09-15 ms/manuscript.Rmd
test references for markdown to pdf

@@
 output: pdf_document
+bibliography: refs.bib
 ---
@@
 # Introduction
+Effects of diversity on disease exist [@ostfeld_effects_2012].
 

### 2022-09-15 ms/manuscript.Rmd
try knitting to word

@@
 date: "`r format(Sys.time(), '%B %d, %Y')`"
-output: pdf_document
+output:   
+  word_document:
+    reference_docx: rv_ms_081921.docx
+  pdf_document: default
 bibliography: refs.bib

### 2022-09-16 ms/manuscript.Rmd
knit files

--- 2022-09-15 ms/manuscript.Rmd
@@
 date: "`r format(Sys.time(), '%B %d, %Y')`"
-output: pdf_document
+output:   
+  pdf_document: default
+  word_document:
+    reference_docx: rv_ms_081921.docx
 bibliography: refs.bib
 ---
+
+# Notes
+
+Community composition is a primary driver of epidemics caused by generalist parasites.
+Dilution effects are an example of community composition influencing parasite transmission and species richness can be predictive of dilution effects when the necessary conditions are present.
+These conditions include a correlation between relative abundance of hosts, the rarity of hosts, and predictable assembly/disassembly of communities.
+For example, if host competence correlates positively with relative abundance and if more competent hosts are less likely to be lost as communities disassemble, then species richness can be predictive of parasite transmission via these mechanisms.
+Another way to understand this relationship is to use community competence as a community-level metric for how competent a community is.
+When dilution effect conditions exist, community competence will predictably increase as species are lost in a community.
+However, similar community competence values can be achieved via alternative community compositions that may not follow a dilution effect pattern.
+If host species still vary in their competence and if there is still a positive correlation between relative abundance and competence BUT if assembly/disassembly does not follow a dilution effect pattern, then species richness is not a useful predictor for parasite transmission but community competence can still be predictive.
+If assembly/disassembly is random, then high competence communities will exist simply by chance.
+Alternatively, if there is not a correlation between relative abundance and host competence but there are dilution effect patterns in assembly/disassembly, then species richness is still not useful, but community competence is.
+In a situation where the rare host is most competent, high species richness communities may be the most competent.
+Also, if competence is distributed randomly with high variance, then community competence may vary unpredictably with species richness, but community competence can still be predictive of parasite transmission.
+
+* Dilution effects depend on
+  + a positive correlation between relative abundance and competence
+  + a correlation between relative abundance and assembly pattern
+    + i.e. higher relative abundance means less likely to be lost from a community
+* Community competence should correlate with species richness if these conditions hold
+* Community competence can still be high even if these conditions do not hold
+* In empirical data, do we see these dilution effect patterns?
+* If we do not see these dilution effect patterns, then are we still able to see high community competence values?
+  + What drives these community competence values?
+    + Is it a rare but highly competent species, abundant but moderately competent species?
+    + How does evenness help to explain community competence?
+
+* Other things to check
+  + Variation in competence within species - is competence a robust trait or highly plastic?
+  + Variation in qPCR values - how reliable are our estimates?
 

### 2022-09-16 scripts/workshop/workshop.Rmd
Update workshop.Rmd

--- 2022-09-13 scripts/workshop.Rmd
@@
 ```{r}
-data <- read_csv(here("raw_data/weighted_prev_competence_111220.csv"))
-tree <- ape::read.nexus(here("raw_data/asup_just_tree.txt"))
-names <- read_csv(here("raw_data/species_names_ids.csv"))
+data <- read_csv(here("data/weighted_prev_competence_111220.csv"))
+tree <- ape::read.nexus(here("data/raw_data/asup_just_tree.txt"))
+names <- read_csv(here("data/raw_data/species_names_ids.csv"))
 ```
@@
 ```
+
+## Variation in competence for each species
+
+```{r}
+species_comp <- data %>% group_by(Species) %>% summarize(mean = mean(SQMean),
+                                                         var = var(SQMean),
+                                                         se = sqrt(var(SQMean)/n()))
+
+species_comp %>% ggplot(.,aes(x=Species, y=mean)) + geom_point() + geom_linerange(aes(ymin=mean-se,ymax=mean+se))
+```

### 2022-09-16 scripts/workshop/workshop.Rmd
update supp 1

@@
                                                          var = var(SQMean),
-                                                         se = sqrt(var(SQMean)/n()))
+                                                         se = sqrt(var(SQMean)/n()),
+                                                         n = n()) %>% mutate(species_code = Species)
 
-species_comp %>% ggplot(.,aes(x=Species, y=mean)) + geom_point() + geom_linerange(aes(ymin=mean-se,ymax=mean+se))
+species_comp %<>% left_join(.,names)
+
+species_comp %>% ggplot(.,aes(x=reorder(abb_name, mean), y=mean)) + geom_point() + geom_linerange(aes(ymin=mean-se,ymax=mean+se)) + theme(axis.text.x = element_text(angle = 30, hjust = 1))
+
 ```

### 2022-09-16 scripts/workshop/workshop.Rmd
updating supp 1

@@
 
-species_comp %>% ggplot(.,aes(x=reorder(abb_name, mean), y=mean)) + geom_point() + geom_linerange(aes(ymin=mean-se,ymax=mean+se)) + theme(axis.text.x = element_text(angle = 30, hjust = 1))
+vl %>% ggplot(.,aes(x=reorder(abb_name, mean), y=mean)) + geom_point() + geom_linerange(aes(ymin=mean-se,ymax=mean+se)) + theme(axis.text.x = element_text(angle = 30, hjust = 1))
 

### 2022-09-16 process/data_format.R

@@ -90,13 +90,19 @@ comm_data$Month <- factor(comm_data$Month, levels = c("Feb", "Mar", "Apr", "May"
 
 #includes average viral load data for each species
 #each row is one host species
-vl <- data %>% dplyr::select(vl.4:vl.42) %>% distinct() %>% pivot_longer(vl.4:vl.42)
-vl %<>% mutate(species_code = as.double(gsub("vl.","",vl$name)))
+# vl <- data %>% dplyr::select(vl.4:vl.42) %>% distinct() %>% pivot_longer(vl.4:vl.42)
+# vl %<>% mutate(species_code = as.double(gsub("vl.","",vl$name)))
+# vl %<>% full_join(., names, by="species_code")
+# vl %<>% mutate(name = gsub("_"," ",vl$species_name))
+# vl$value %<>% na_if(.,0)
+# vl %<>% mutate(., ln_value = log(value)) %>% mutate(., log10_value = log10(value))
+# vl %<>% replace_na(.,list(value=0,ln_value=0,log10_value=0))
+
+vl <- data %>% group_by(Species) %>% summarize(mean = mean(SQMean),
+                                               var = var(SQMean),
+                                               se = sqrt(var(SQMean)/n()),
+                                               n = n()) %>% mutate(species_code = Species)
 vl %<>% full_join(., names, by="species_code")
-vl %<>% mutate(name = gsub("_"," ",vl$species_name))
-vl$value %<>% na_if(.,0)
-vl %<>% mutate(., ln_value = log(value)) %>% mutate(., log10_value = log10(value))
-vl %<>% replace_na(.,list(value=0,ln_value=0,log10_value=0))
 
 
 

### 2022-09-16 scripts/figures/supp1_code.R

@@ -10,11 +10,12 @@ source(here("base","src.R"))
 
 vl <- readRDS(here("processed_data","vl.rds"))
 
-supp1 <- vl %>% arrange(log10_value) %>% 
+supp1 <- vl %>% arrange(mean) %>% 
   mutate(abb_name = factor(abb_name, levels = abb_name)) %>% 
-  ggplot(.,aes(x=abb_name,y=log10_value)) +
+  ggplot(.,aes(x=abb_name,y=mean)) +
   geom_point() + 
-  xlab("Species") + ylab("log10(Viral Load)") + 
+  xlab("Species") + ylab("Viral Load") + 
+  geom_linerange(aes(ymin=mean-se,ymax=mean+se)) +
   geom_hline(yintercept = 1) + 
   labs(title = "Viral Loads") + 
   theme_classic() + 

### 2022-09-16 ms/manuscript.Rmd
update supp 1
more to do items

@@
 ---
+
+# To do
+
+* update supp1 with standard error bars
+* update all data by using log-transformed raw data rather than log-transforming means of raw data
+  + confusing thing is that the rank order of the raw means is different than the means of the log-transformed data
+  + is this correct or did I do something wrong?
+
+```{r}
+m <- c(4, 8 , 15, 16, 23, 42)
+tmp <- tibble(m, l = log10(x))
+tmp_s <- tmp %>% summarize(mean = mean(m), log_mean = mean(l))
+tmp_s %<>% mutate(log_m = log(mean))
+```
 

### 2022-09-16 process/data_format.R
check to make sure sample sizes are same

@@ -106,7 +106,8 @@ vl_log <- data %>%
   group_by(Species) %>% 
   summarize(log10_mean = mean(log10_SQ), 
             log10_var = var(log10_SQ),
-            log10_se = sqrt(var(log10_SQ)/n())) %>% 
+            log10_se = sqrt(var(log10_SQ)/n()),
+            m = n()) %>% 
   mutate(species_code = Species)
 vl %<>% left_join(.,vl_log)
 vl %<>% full_join(., names, by="species_code")

### 2022-09-16 ms/manuscript.Rmd
log-transformed mean puzzle

@@
   + is this correct or did I do something wrong?
+  + this must be because I am dealing with values smaller than 1 which become negative when log-transformed which can bring down the average
+    + hmm I am still getting this result with log1p. not sure what is up
 
 ```{r}
-m <- c(4, 8 , 15, 16, 23, 42)
-tmp <- tibble(m, l = log10(x))
+m <- c(0.1, 0.5, 4, 8 , 15, 16, 23, 42)
+tmp <- tibble(m, l = log10(m))
 tmp_s <- tmp %>% summarize(mean = mean(m), log_mean = mean(l))

### 2022-09-16 process/competencyCode.R

@@ -15,7 +15,8 @@ library(here)
 q <- read_csv("data/rv_data_111220.csv")
 
 q %<>% mutate(log10_SQ = ifelse(is.infinite(log10(SQMean)), 0, log10(SQMean)),
-              ln_SQ = ifelse(is.infinite(log(SQMean)), 0, log(SQMean)))
+              ln_SQ = ifelse(is.infinite(log(SQMean)), 0, log(SQMean)),
+              log1p_SQ = log1p(SQMean))
 
 
 

### 2022-09-16 process/data_format.R

@@ -104,9 +104,9 @@ vl <- data %>% group_by(Species) %>% summarize(mean = mean(SQMean),
                                                n = n())
 vl_log <- data %>% 
   group_by(Species) %>% 
-  summarize(log10_mean = mean(log10_SQ), 
-            log10_var = var(log10_SQ),
-            log10_se = sqrt(var(log10_SQ)/n()),
+  summarize(log1p_mean = mean(log1p_SQ), 
+            log1p_var = var(log1p_SQ),
+            log1p_se = sqrt(var(log1p_SQ)/n()),
             m = n()) %>% 
   mutate(species_code = Species)
 vl %<>% left_join(.,vl_log)

### 2022-09-16 scripts/figures/supp1_code.R

@@ -24,10 +24,10 @@ supp1_raw <- vl %>% arrange(mean) %>%
 
 supp1_log10 <- vl %>% arrange(mean) %>% 
   mutate(abb_name = factor(abb_name, levels = abb_name)) %>% 
-  ggplot(.,aes(x=abb_name,y=log10_mean)) +
+  ggplot(.,aes(x=abb_name,y=log1p_mean)) +
   geom_point() + 
   xlab("Species") + ylab("log10(Viral Load)") + 
-  geom_linerange(aes(ymin=log10_mean-log10_se,ymax=log10_mean+log10_se)) +
+  geom_linerange(aes(ymin=log1p_mean-log1p_se,ymax=log1p_mean+log1p_se)) +
   geom_hline(yintercept = 1) + 
   labs(title = "Viral Loads") + 
   theme_classic() + 

### 2022-09-19 scripts/workshop/workshop.Rmd
update workshop

--- 2022-09-16 scripts/workshop/workshop.Rmd
@@
 
+### comparing ranked means with and without log-transformation of raw data
+
 ```{r}
-species_comp <- data %>% group_by(Species) %>% summarize(mean = mean(SQMean),
-                                                         var = var(SQMean),
-                                                         se = sqrt(var(SQMean)/n()))
+vl <- readRDS(here("processed_data","vl.rds"))
 
-species_comp %>% ggplot(.,aes(x=Species, y=mean)) + geom_point() + geom_linerange(aes(ymin=mean-se,ymax=mean+se))
+supp1_raw <- vl %>% arrange(mean) %>% 
+  mutate(abb_name = factor(abb_name, levels = abb_name)) %>% 
+  ggplot(.,aes(x=abb_name,y=mean)) +
+  geom_point() + 
+  xlab("Species") + ylab("Viral Load") + 
+  geom_linerange(aes(ymin=mean-se,ymax=mean+se)) +
+  geom_hline(yintercept = 1) + 
+  labs(title = "Viral Loads") + 
+  theme_classic() + 
+  theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
+  scale_y_continuous(breaks=seq(0,6,1))
+
+supp1_log10 <- vl %>% arrange(mean) %>% 
+  mutate(abb_name = factor(abb_name, levels = abb_name)) %>% 
+  ggplot(.,aes(x=abb_name,y=log1p_mean)) +
+  geom_point() + 
+  xlab("Species") + ylab("log1p(Viral Load)") + 
+  geom_linerange(aes(ymin=log1p_mean-log1p_se,ymax=log1p_mean+log1p_se)) +
+  geom_hline(yintercept = 1) + 
+  labs(title = "Viral Loads") + 
+  theme_classic() + 
+  theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
+  scale_y_continuous(breaks=seq(0,6,1))
+
+supp1_raw
+supp1_log10
 ```

### 2022-09-19 scripts/workshop/workshop.Rmd
add nmds

@@
 
+NMDS with bray-curtis
+```{r}
+set.seed(8878896)
+nmds <- metaMDS(abundance_mat, distance = "bray")
+
+plot(nmds)
+
+nmds_scores <- as_tibble(scores(nmds))
+nmds_scores$siteID <- community_mat$siteID
+
+nmds_scores_summ <- left_join(comm_summ, nmds_scores, by = "siteID")
+nmds_scores_summ$Month <- factor(nmds_scores_summ$Month, levels = c("Feb", "Mar", "Apr", "May", "Jun", "Jul"))
+nmds_scores_summ %<>% mutate(pc1Rank=dense_rank(NMDS1))
+
+nmds_scores_summ %>% ggplot(.,aes(x=NMDS1, y = cc)) + geom_point() + geom_smooth(method = "loess")
+```
+
 ### Let's try to explain the residuals

### 2022-09-21 ms/manuscript.Rmd
update manuscript rmd and knitted files

--- 2022-09-16 ms/manuscript.Rmd
@@
 output:   
-  pdf_document: default
   word_document:
     reference_docx: rv_ms_081921.docx
+  pdf_document: default
 bibliography: refs.bib
 ---
+
+# To do
 
@@
 # Introduction
-Effects of diversity on disease exist [@ostfeld_effects_2012].
+Most parasites must navigate a complex environment in order to complete their life cycle and are constrained by the availability and suitability of hosts as well as environmental conditions.
+Generalist parasites contend with changes in the relative and absolute abundance of their various hosts, and parasites that are transmitted environmentally must rely on favorable conditions for persistence until they are taken up by a new host.
+Much attention has been paid to anthropogenic-induced changes in the biotic and abiotic conditions experienced by parasites, including introduction and extirpation of species, climate warming, pollution, and land conversion (Altizer et al., 2013; Patz et al., 2004; Young et al., 2017).
+Many of these changes feed into the dilution effect hypothesis, which posits that transmission (in general, or relative to a focal host species) is augmented as host species richness declines, often in response to anthropogenic influence (Pieter T.J.
+Johnson et al., 2015; Keesing et al., 2006).
+However, biotic and abiotic change also occurs naturally over time and space.
+Moreover, the joint effects of abiotic and biotic drivers of epizootics are rarely considered together (Bienentreu & Lesbarrères, 2020; Dillon & Meentemeyer, 2019; Hall et al., 2010; North et al., 2015; Scott-Baumann & Morgan, 2015), and theoretical prediction for such interaction is underdeveloped.
+In part, this is due to the notion that transmission is influenced by abiotic factors at large spatial scales and by biotic factors at smaller scales (Halliday & Rohr, 2019; Rohr et al., 2019).
+However, host mobility often means transmission is occurring at intermediate spatial scales, where communities of hosts are connected, and in these metacommunities, the joint influence of host community structure and environmental conditions on parasite transmission is not well understood.
+In response to this knowledge gap, we combine theoretical and empirical approaches to study how biotic and abiotic factors interact to promote transmission of ranavirus in amphibian communities.
+We demonstrate that focusing on the relative abundance of species, in concert with total community size and environmental conditions, provides a more detailed explanation for spatiotemporal variation in outbreak risk that cannot be solely explained by biotic or abiotic variables alone.
+In addition, we build upon existing theory regarding the relationship between host biodiversity and parasite transmission by demonstrating that host species evenness, rather than richness, can be used to relate community composition to outbreak risk.
+
+Generalist parasites, such as ranaviruses, are commonly found in nature and often use a range of alternative host species to reproduce (Park et al., 2018; Poulin et al., 2011).
+The transmission potential of a generalist parasite for a given host species may vary due to differences in host species abundances (Fenton et al., 2002; Hopkins et al., 2020), host competence (i.e.
+the ability for an individual host to become infected and pass infection on to others (Martin et al., 2019)), and the environmental conditions in which the interaction takes place (Altizer et al., 2013; Cable et al., 2017; Rohr et al., 2011).
+The combination of these factors makes it difficult to assess how multi-species host communities are differentially at risk of parasite invasion and onward transmission to other communities (P T J Johnson et al., 2015).
+One method to separately quantify the effects of host abundance and competence on parasite transmission is to use community competence: the weighted average of competence values of host species in the community (Pieter T.J.
+Johnson et al., 2013a).
+In addition, community competence is advantageous because it considers the relative abundance of a species rather than the presence of a species alone.
+For example, a commonly occurring, low-competence host could be responsible for more transmission than a rare, high-competence host (Holt et al., 2003).
+In contrast to host species richness, which is the typical predictor variable in the dilution effect hypothesis literature, community competence and host abundance are more directly related to the parasite basic reproductive number, R0 (Dobson, 2004; Rohr et al., 2019).
+
+Ranaviruses (genus Ranavirus, family Iridoviridae) are well-suited for studying host-parasite-environment interactions at the community scale and are associated with worldwide amphibian declines, making them a parasite of global concern (Lesbarrères et al., 2012).
+The Ranavirus genus of double-stranded DNA viruses (Gray et al., 2009), affects most amphibian species (Jesse L Brunner et al., 2015) and host species exhibit a wide range of competence for the parasite (Hoverman et al., 2011).
+At the Savannah River Site (South Carolina, USA), a spatially-distributed set of 20 wetlands has been extensively sampled in the context of ranavirus infection over a 6-month period.
+We focus on larval amphibian communities which are bound by their aquatic habitat but connected through the overland movement of their adult stages (Mihaljevic et al., 2018).
+Ranavirus is transmitted via intra- and interspecific direct transmission, which exhibits a positive, but ultimately saturating, relationship with host abundance (Jesse L Brunner et al., 2017).
+Additionally, ranavirus can be transmitted environmentally via free-living infectious virions shed into water by infected live individuals or decaying carcasses (Brenes et al., 2014).
+Ranavirus persistence times in water have been found to diminish at high ambient temperatures (Nazir et al., 2012) suggesting that the relative importance of environmental transmission, versus contact-based transmission, should depend on environmental conditions.
+We hypothesize that community-scale prevalence of ranavirus will have a positive relationship with host abundance and community competence but a negative relationship with increasing water temperatures.
+
+We begin with a theoretical model, where we use the community basic reproductive number, R0, to estimate the transmission potential of ranavirus in multi-host communities (Dobson, 2004).
+R0 provides a valuable lens into transmission potential, clarifying conditions necessary for parasite invasion and persistence.
+Calculation of R0 at the community scale explicitly considers the role of community composition in the transmission of a multi-host parasite along with host abundance.
+In our case it also includes the relative force of infection from free-living infectious virions, versus infected hosts.
+Multiple transmission routes can increase overall transmission potential, as evidenced by contact and environmental transmission in models of avian influenza virus (Breban et al., 2009) and causal agents of diarrheal diseases (Tien & Earn, 2010) as well as contact and sexual transmission in models of Ebola transmission (Abbate et al., 2016; Vinson et al., 2016).
+From our model, we analyzed the relative contribution of direct and environmental transmission in allowing the parasite to exceed the critical value of R0=1.
+These two routes are linked because environmental transmission relies on the shedding of viral material from infected individuals and higher environmental transmission rates will generate more opportunities for subsequent direct transmission.
+Specifically, we determine how transmission from each route facilitates parasite invasion under different scenarios of community composition, host abundance, and virus decay rate (half-life) in water, both as separate factors and acting in concert.
+
+Using empirical data, we then apply metacommunity theory and dimensionality reduction to explain how biotic and abiotic factors contribute to community composition (Leibold et al., 2004), and then we evaluate these communities in their ability to differentially facilitate transmission.
+Community compositions change over space or time and can include dynamic, predictable patterns of assembly or disassembly (Pieter T.J.
+Johnson et al., 2019; Ostfeld & LoGiudice, 2003).
+These changes in community composition can then influence transmission potential of a parasite and may map to dilution or amplification effects (Luis et al., 2018; North et al., 2015).
+According to theory, a dilution effect may occur whenever hosts vary in competence and reductions in diversity are associated with increases in mean competence (Pieter T.J.
+Johnson et al., 2015).
+In contrast, an amplification effect can occur when increased host diversity causes an increase in mean competence (Keesing et al., 2006).
+Considering these conflicting patterns, we relate species evenness, in addition to species richness, as a predictor for both community competence and infection prevalence.
+In the ranavirus system, we measure competence as a species’ average viral load, which estimates the differential propensity for hosts to generate new infections.
+While the amphibian community comprises several species, their viral loads are bimodal and communities can be simplified into two groups (high and low competence), which is especially useful in model building and analysis.
+
+Modeling results show that the effect on R0 by community composition, host abundance, and environmental persistence of virus are independently important such that weak contribution to R0 by a factor can be compensated by appropriately strong contributions from other factors.
+In addition, we find that each of these factors combined can have an appreciably greater than additive effect, meaning some host communities can be disproportionately at risk of epizootics.
+The empirical data confirm that outbreaks are frequently associated with ‘perfect storms’ of large communities dominated by competent species at relatively low temperatures.
+Importantly, neither species richness nor species evenness were found to be powerful predictors for ranavirus prevalence.
+However, species evenness and community competence emerged as informative metrics for identifying mechanism in the diversity-disease relationship.
+Our results improve our understanding of the location and timing of outbreaks of multi-host parasites and help to predict when host communities may be most vulnerable, supporting the study of zoonotic diseases by identifying potential disease hot spots that may form due to natural or anthropogenic causes.
+We contend that the study of abiotic factors directly influencing transmission and indirectly driving community composition holds great promise to advance diversity-disease research.
 
@@
 
+## Transmission Model
+
+According to known transmission, ranaviruses can infect a wide range of amphibian hosts and infectious periods can range from a few days up to weeks (Gray et al., 2009).
+Transmission can occur both directly and indirectly and exposure appears to induce an adaptive immune response in surviving hosts.
+Accordingly, we modeled a host community using an SIR framework for each species (Table 1).
+For model tractability, we limited the community to two host types that could vary in key traits including abundance and competence.
+As well as facilitating model analysis, this choice also reflects the empirical observation that viral loads of host species are bimodal (Supplementary Figure 1).
+In addition to intra- and inter-specific direct transmission, we included environmental transmission, whereby infectious host individuals shed virus into the environment where it persists for some finite time and can cause new infections without host-to-host contact (Gray et al., 2009).
+We included host demography via a constant birth rate and a constant per capita mortality rate, resulting in a disease-free equilibrium for each host species given by the ratio of the birth rate and mortality rate.
+For the model to reflect the viral load-based definition of host competence, infectivity, but not susceptibility, varied between the high and low competence host.
+This means that the rate of transmission from infectious individuals did not depend on whether transmission was to an intra- or inter-specific host, but rather on the high or low competence status of the infectious animal.
+For example, a high competence individual is likely to infect many individuals of either species when infectious and a low competence individual is expected to infect fewer individuals.
+In keeping with the assumption of no difference in host susceptibility between species, the environmental transmission rate was set to be the same for both species.
+
+Using the next-generation matrix method (Diekmann et al., 1990, 2009), we calculated the community basic reproductive number (Dobson, 2004), hereafter referred to as R0, for our community of hosts to determine the conditions necessary for parasite invasion (R0>1).
+Consequently, we determined how the boundary R0=1 operates as a function of parameters for communities with varying characteristics, specifically the effects of community composition, total host abundance, and viral half-life.
+To illustrate these effects, we created a reference community and 4 manipulated communities that were designed to make the communities more suitable for parasite invasion.
+The reference community had an equal number of both species, a total host abundance of 150 individuals, and a viral half-life of 1.35 days.
+The composition-manipulated community was altered to be dominated by the more competent species by a ratio of 2:1.
+The abundance-manipulated community was altered only in abundance, to a value of 175 individuals, and the half-life-manipulated community was altered by doubling the viral half-life to 2.7 days.
+The final manipulated community combined each of these single-factor manipulations.
+For each community, we calculated R0 over a range of values for environmental and contact transmission rates of the more competent species while holding the contact transmission rate for the less competent species constant.
+This allowed us to characterize the extent to which parasite invasion was more likely, i.e., occurring for an increased set of transmission parameters that included combinations previously associated with failure to invade (R0<1).
+
+## Data Collection and Calculation of Community Competence
+
+All empirical data were previously collected (Coleman, 2018) at the Savannah River Site (South Carolina, USA).
+20 wetlands were sampled monthly for 6 months from February to July in 2016 at the Savannah River Site.
+Of the 120 sampling events, 96 produced data, with the others being discounted due to wetlands being dry at the time of sampling.
+Each sampling event included an estimate of larval amphibian abundance ascertained from minnow traps and dip-net sweeps around the perimeter and center of the wetland.
+In addition to abundance, a single individual per species was collected per dip-net sweep or minnow trap.
+These individuals were tested for ranavirus load using qPCR in triplicate which were averaged to determine the viral load for an individual.
+At the species-level, all individuals that were analyzed for viral load were then averaged to provide a species level estimate of viral load, which was used as a proxy for competence.
+Using species-level competence, we then calculated community competence as the weighted average of each species’ competence, with weights given by the relative abundance of each species.
+Each site-month combination was treated as a distinct community in these calculations.
+Empirical data on community composition, viral loads, and infection prevalence enabled us to analyze host communities according to their community competence and relate these to diversity metrics, as detailed below.
+
+## Host Community Analysis
+
+First, using a metacommunity framework, we created a site-by-species matrix that included species presence or absence, and where ‘site’ referred to a unique wetland-month combination, referred to herein as a community.
+Within this matrix, communities were ordinated by minimizing the number of embedded absences and were then tested for key metacommunity patterns including nestedness and block replacement using the “metacom” package in R (Dallas, 2014; Mathew A Leibold & Mikkelson, 2002).
+Following this, we used principal component analysis to order communities according to their similarity in community composition.
+We did this by ranking each community according to the score of their first principal component (PC1), which captured 71% of the variance.
+We then used this list of ranked communities to ascertain the congruence between host community composition and community competence, testing the assumption that similarly composed communities will be similarly competent.
+We used a generalized additive model with a geo-additive term to test the relationship between community competence and time (sampling month), space (wetland coordinates), and host abundance to establish the effects of these variables on community competence.
+
+Following this, we treated community competence and infection prevalence as separate response variables and used species richness and species evenness as predictor variables.
+Community competence is used to measure an instantaneous property of the community in terms of its propensity to support a parasite, but infection prevalence is expected to manifest after sufficient transmission has occurred and so was lagged by a month (the period of time between measurements and a plausible estimate of the parasite generation time (Brunner & Collins, 2009)).
+We estimated infection prevalence for the community by combining the observed host abundances and the raw infection prevalence of individual host organisms that were sampled for parasite presence using qPCR.
+The sampled hosts provided a percentage for each species that were found to be infected and this percentage was then weighted by the observed abundance of that species in that community.
+Species richness was calculated as the sum of unique species present and species evenness was measured by Pielou’s J (Pielou, 1966).
+Species richness was used to test if basic patterns of assembly and disassembly were able to predict transmission potential via dilution or amplification effects, whereas species evenness was used to detect if relative abundances of species were controlling transmission via community competence.
+In an effort to more closely examine the effects of species evenness on community competence, we ordered sites according to community competence and identified trends in the relative abundance of species and the phylogenetic relationships of our host species.
+We expected phylogenetically clustered or overdispersed species to co-occur less frequently, and we additionally determined the extent to which competence was clustered (versus dispersed) within the phylogeny so that we could evaluate the extent to which multiple competent species may co-occur and potentially dominate a community.
+Finally, we tested for statistically significant relationships between infection prevalence and community competence, abundance, and mean water temperature using both multivariate (generalized linear model) and univariate tests (Spearman Rank correlation test with a correction for multiple comparisons).
+Mean water temperature was included to serve as a proxy for environmental persistence time of free-living infectious virions; as mean water temperature increases the viral half-life is expected to decrease (Nazir et al., 2012).
+
 # Results
+
+## Transmission Model
+
+Model analysis determined the criteria for an epizootic to occur (R0>1) (Figure 1).
+In the relative transmission rate space (Figure 1), each community exhibits minimum conditions for invasion (R0=1, contour lines).
+The linear shape of each line indicates that the direct and environmental transmission routes are always substitutive; a reduction in one parameter can be compensated by a proportional increase in the other.
+However, increasing each factor (Figure 1; A: proportion of high-competence hosts=red line, B: total host abundance=gold line, C: viral half-life=green line) affects the invasion contour differently when compared to the reference, reflecting differences in which transmission route can better tolerate a reduction while still allowing parasite invasion in the altered community.
+When community composition is altered to an increased proportion of the highly competent species, the slope of the invasion threshold is reduced relative to the reference community, meaning that contact transmission can be reduced appreciably while maintaining R0>1 (Figure 1A; red vs.
+gray line).
+However, when viral half-life is increased, it is environmental transmission that can be reduced to a greater extent (Figure 1C; green vs.
+gray line).
+Increasing total host abundance does not change the slope of the invasion threshold, but does reduce the x- and y-intercept, meaning both transmission rates can be lowered while maintaining R0>1, but neither rate is more impactful than the other (Figure 1B; gold vs.
+gray line).
+Finally, when all manipulations are combined, there is a lowering of the invasion threshold that exceeds the changes of any condition alone, indicating synergistic interaction between factors affecting overall transmission potential (Figure 1D; purple vs.
+gray line).
+
+## Host Community Analysis
+
+Using the sampling unit of community (a unique wetland-month combination), metacommunity analysis revealed a block-replacement pattern, meaning groups of species tended to turn over together across communities (coherence: P < 0.00001; turnover: P < 0.00001; boundary clumping: P < 0.00001; Supplementary Figure 2).
+When ordinated according to community composition using rank order along PC1 we found that similar communities exhibited similar values of community competence (Figure 2).
+Additionally, we note that communities of different compositions had similar values for community competence (e.g., Figure 2, communities ranked ~5 and ~30) in keeping with the fact that several combinations of weighted competence scores can evaluate to the same community competence.
+Analysis of the relationship between community competence and the predictors of sampling month, wetland coordinates, and host abundance using a generalized additive model revealed significant effects for sampling month (P < 0.00001) and host abundance (P < 0.0001) but not for wetland coordinates (P = 0.34).
+Overall, communities with higher values of competence occurred in February through May, whereas in June and July, several communities showed low values of community competence (Supplementary Figure 3a).
+In terms of absolute host abundance, both high and low values could result in high community competence (Figure 2, shape size), but the trend was generally positive (Supplementary Figure 3b).
+While temperature is predicted to influence environmental transmission via viral half-life, community structure is not strongly dictated by temperature (Figure 2, color); air and water temperature at sites align with PC2, which explained 26% of the variance (Supplementary Figure 4).
+
+Using evenness as our predictor variable, we found that community competence had a bimodal relationship at low evenness that converged at higher evenness (Figure 3).
+At low evenness, communities were generally found to have either high or low values of community competence with only a few communities exhibiting moderate values of community competence.
+At higher values of evenness (Pielou’s J > 0.6), most communities had either low or moderate values of community competence.
+Intuitively, we expected higher evenness communities to have moderate values of community competence because our set of species were made up of both high and low competence hosts.
+However, at lower evenness, the competence of dominant species will drive community competence which led to the bimodal relationship.
+Furthermore, we found that communities of low evenness and high community competence were generally of higher abundance and had higher future infection prevalence.
+There were also exceptions to this pattern where communities with low or moderate community competence still had high future infection prevalence.
+This can be caused by high host abundance driving transmission in some communities (Figure 3; i and ii) or low sample size resulting in relatively high infection prevalence in others (Figure 3).
+When ordered by community competence, we found that the most competent communities were dominated by either one or a few highly competent species (Figure 4).
+In the phylogeny, we see that these highly competent species are moderately dispersed, suggesting that limiting similarity may not be hindering coexistence (Figure 4, bottom panel; Supplementary Figure 6).
+
+Finally, we found that community competence and host abundance correlated positively with prevalence while mean water temperature correlated negatively with prevalence.
+In separate univariate analyses when using a Spearman rank correlation test we found significant results for community competence (P < 0.0001; figure 5), abundance (P < 0.01; figure 5), and mean water temperature (P = 0.021; figure 5) after adjusting for multiple comparisons using the “Holm” method (Aickin & Gensler, 1996; Holm, 1979).
+Due to the inherently noisy data on parasite prevalence, a multivariate model that predicted infection prevalence in response to community competence, total host abundance, and mean water temperature was not able to detect any significant relationships.
+Overall, these results suggest that the prevalence of ranavirus in a community may be driven by multiple factors working in conjunction and support our predictions regarding the directional effects of community competence, host abundance, and virus half-life (Figure 5).
+Species richness was not predictive of infection prevalence meaning we did not detect a classic dilution effect pattern given by a negative relationship; rather the non-significant trend was slightly positive (Supplementary Figure 5; Spearman’s rho = 0.15, P = 0.14).
 
@@
 
+Using a blend of theoretical model predictions and empirical data, we have shown that the effects of host abundance, community composition, and the environment can simultaneously affect the transmission potential of a generalist parasite in communities of hosts.
+In our mathematical model, we demonstrated that these independent factors can all enhance transmission and that when combined, can have a more than additive effect.
+Using empirical data, we confirmed that both biotic and abiotic factors can have detectable effects on parasite transmission at an intermediate spatial scale.
+Importantly, we also found that species evenness was a more informative predictor for parasite transmission potential compared to species richness.
+Moreover, our results inform the risk of amphibian communities to ranavirus epizootics and extend theory regarding the relationship between biodiversity and the transmission of generalist parasites by including environmental transmission and the factors that promote it.
+
+ From our theoretical model, we learned that the inclusion of both contact-based and environmental transmission allowed for each transmission mode to compensate for the other; a low strength of one transmission mode could be supplemented by the other transmission mode, resulting in a broader range of favorable conditions for the parasite (Vinson et al., 2016).
+In other studies, it has been noted that the incorporation of environmental transmission can provide dramatically different estimates of R0 and allow for persistence of parasites at population sizes that would not be supported by contact transmission alone (Breban et al., 2009, 2010; Tien & Earn, 2010).
+We build upon these findings by demonstrating how the relative strength of transmission modes can emphasize the importance of different factors relevant to parasite transmission.
+In a community where the strength of contact-based transmission is high and environmental transmission strength is low we expect community composition to have a significant impact on transmission.
+Conversely, when environmental transmission is high and contact-based transmission is low, we expect community composition to be less important for overall transmission.
+This suggests that in natural systems where both modes of transmission are available, the influence of both may be present but obscured by the other.
+When we enhanced absolute host abundance, we anticipated that contact-based transmission would be favored, but our analysis showed that the increase in available hosts can augment both transmission modes equally.
+We understood this to be because as the number of hosts increases, the quantity of free-living infectious virions also increases proportionally due to a greater number of infected hosts shedding viral material as well as simply increasing host density.
+This results in an increase in the force of infection for both contact-based transmission and environmental transmission.
+
+From the empirical data, we found that community composition and evenness were far more informative in predicting transmission potential compared to species richness, in spite of the fact that the diversity-disease literature includes many correlative studies that investigate the relationship between species richness and infection prevalence.
+Indeed, recent research has emphasized the importance of a more mechanistic understanding of diversity-disease relationships, and host species evenness may aid in this effort (Merrill & Johnson, 2020; Rohr et al., 2019).
+Evenness, rather than richness, can more directly relate the composition of a community to its transmission potential.
+At low evenness, the identity of the dominant species drives community competence, and we found that communities with the highest absolute abundance were usually of lower evenness and dominated by a few highly competent species.
+This pattern is supported in theory connecting life history traits to host competence, where hosts with a faster pace-of-life invest heavily into rapid growth and reproduction and less into immunity, resulting in highly abundant, highly competent hosts (Joseph et al., 2013).
+Together, host abundance, evenness, and community competence have the potential to offer a more mechanistic lens into community compositions and how they will affect infection prevalence.
+
+Detection of diversity-disease relationships is facilitated by predictable community structures (Faust et al., 2017).
+Classic dilution-effect systems typically operate by nested structures influencing parasite transmission through biodiversity loss (LoGiudice et al., 2003).
+In these examples, species losses follow a pattern where less competent hosts are lost before more competent hosts (Pieter T.J.
+Johnson et al., 2013b).
+In contrast, our study revealed block-replacement of sets of host species, where transmission potential was not changing due to species richness, but rather by species evenness and identity; often, ranavirus attained high prevalence due to correlated competence and abundance of certain key species.
+However, the generality of this depends on how competence is typically distributed among a community of host species and how relationships among those species affect their probability of co-occurring and abundance.
+The amphibian species in our study showed evidence of phylogenetic repulsion and environmental filtering, as has been detected, for example, in avian communities (Weinstein et al., 2017).
+When co-occurring, species that were closely related or distantly related to their neighbors had low relative abundance (Supplementary Figure 6).
+In contrast, species that were intermediately related to others on average attained high relative abundance.
+Combined with the fact that high competence was a relatively well-dispersed trait in the studied amphibian clade, this created the conditions for competent species to co-occur and attain high relative abundance.
+By corollary, if competent species were clustered phylogenetically, then high community competence would be enhanced by the presence of an environmental filter and absence of phylogenetic repulsion.
+Future work may benefit from considering host communities simultaneously in terms of the rules that govern species co-occurrence and abundance and the distribution of competence levels among those species.
+
+The study of individual host competence and how it scales to a community can provide us with valuable insight into how well host communities can support generalist parasites (Becker et al., 2020).
+However, host competence can be a challenging trait to measure because it is multifactorial, relying on components such as exposure, susceptibility, suitability, and transmissibility (Downs et al., 2019).
+In our study, we use mean viral load as a proxy for host competence which is useful because it informs shedding rate, infectivity, and host suitability.
+However, viral load is not a holistic measure of host competence because it does not address, for example, the infectious period or susceptibility.
+Further research on host competence should incorporate more robust measures of competence such as those derived from controlled infection assays (Merrill et al., 2019).
+Factors within host species such as age and sex will also likely change the competence of a host and while there may be the ability to measure competence with genetic methods, there is likely to be phenotypic plasticity within the trait (Gervasi et al., 2015).
+
+Ranaviruses in amphibians are a useful and important system for studying diversity-disease relationships because of their highly generalist nature and role in global amphibian biodiversity declines (Lesbarrères et al., 2012; North et al., 2015).
+Environmental transmission of ranavirus and the connection of wetlands by amphibian adults made this system especially useful for our study of the effects of host composition, host abundance, and environmental factors.
+However, these features are not limited to this study system and other multi-host generalist parasite systems can benefit from our findings.
+Furthermore, our research focused on natural variations in biodiversity, but it is important to recognize how anthropogenic influence will alter these patterns as well (Altizer et al., 2013; Rohr et al., 2011).
+Amphibian species, in particular, have been extremely sensitive to anthropogenic disturbance and have been considered proxies for overall environmental health (Scheele et al., 2019).
+Land use and climate change are also factors that affect amphibian communities globally and it is important to understand the implications of these changes (Wilkinson et al., 2018).
+Altogether, the ecology of ranavirus in amphibians and the global relevance of amphibian biodiversity make this system a valuable case study for the joint influence of biotic and abiotic factors on parasite transmission under natural and artificial settings.
+
+By focusing on parasites with multiple transmission modes, we have highlighted how biotic or abiotic factors may differentially exacerbate the impact of parasites introduced into a host community.
+In addition, we have expanded theory regarding diversity-disease relationships to advocate the use of species evenness and community competence as informative predictors for the impact of generalist parasites in multi-host communities.
+These metrics can facilitate analysis that is more focused on the mechanics of dilution or amplification rather than correlative association between disease presence and underlying assembly rules that may be inconsistent across systems, space, and time.
+Furthermore, we advocate for deeper analysis of host community phylogeny in the context of diversity-disease relationships which can provide insight into patterns of host species co-occurrence and host traits that may correlate with competence.
+Together, these developments combined with the contextual understanding of specific host-parasite ecology should allow for mechanistic understanding of many generalist parasite systems.
+
 # References

### 2022-09-21 ms/manuscript.Rmd
update manuscript

@@
 output:   
+  pdf_document: default
   word_document:
     reference_docx: rv_ms_081921.docx
-  pdf_document: default
 bibliography: refs.bib
@@
 Generalist parasites contend with changes in the relative and absolute abundance of their various hosts, and parasites that are transmitted environmentally must rely on favorable conditions for persistence until they are taken up by a new host.
-Much attention has been paid to anthropogenic-induced changes in the biotic and abiotic conditions experienced by parasites, including introduction and extirpation of species, climate warming, pollution, and land conversion (Altizer et al., 2013; Patz et al., 2004; Young et al., 2017).
-Many of these changes feed into the dilution effect hypothesis, which posits that transmission (in general, or relative to a focal host species) is augmented as host species richness declines, often in response to anthropogenic influence (Pieter T.J.
-Johnson et al., 2015; Keesing et al., 2006).
+Much attention has been paid to anthropogenic-induced changes in the biotic and abiotic conditions experienced by parasites, including introduction and extirpation of species, climate warming, pollution, and land conversion [@altizer_climate_2013; @patz_unhealthy_2004; @young_introduced_2017].
+Many of these changes feed into the dilution effect hypothesis, which posits that transmission (in general, or relative to a focal host species) is augmented as host species richness declines, often in response to anthropogenic influence [@johnson_why_2015; @keesing_effects_2006].
 However, biotic and abiotic change also occurs naturally over time and space.
-Moreover, the joint effects of abiotic and biotic drivers of epizootics are rarely considered together (Bienentreu & Lesbarrères, 2020; Dillon & Meentemeyer, 2019; Hall et al., 2010; North et al., 2015; Scott-Baumann & Morgan, 2015), and theoretical prediction for such interaction is underdeveloped.
-In part, this is due to the notion that transmission is influenced by abiotic factors at large spatial scales and by biotic factors at smaller scales (Halliday & Rohr, 2019; Rohr et al., 2019).
+Moreover, the joint effects of abiotic and biotic drivers of epizootics are rarely considered together [@bienentreu_amphibian_2020; @dillon_direct_2019; @hall_why_2010; @north_anthropogenic_2015; @scott-baumann_review_2015], and theoretical prediction for such interaction is underdeveloped.
+In part, this is due to the notion that transmission is influenced by abiotic factors at large spatial scales and by biotic factors at smaller scales [@halliday_past_2019; @rohr_towards_2019].
 However, host mobility often means transmission is occurring at intermediate spatial scales, where communities of hosts are connected, and in these metacommunities, the joint influence of host community structure and environmental conditions on parasite transmission is not well understood.

### 2022-09-23 ms/manuscript.Rmd
update ms

@@
 
-Generalist parasites, such as ranaviruses, are commonly found in nature and often use a range of alternative host species to reproduce (Park et al., 2018; Poulin et al., 2011).
-The transmission potential of a generalist parasite for a given host species may vary due to differences in host species abundances (Fenton et al., 2002; Hopkins et al., 2020), host competence (i.e.
-the ability for an individual host to become infected and pass infection on to others (Martin et al., 2019)), and the environmental conditions in which the interaction takes place (Altizer et al., 2013; Cable et al., 2017; Rohr et al., 2011).
-The combination of these factors makes it difficult to assess how multi-species host communities are differentially at risk of parasite invasion and onward transmission to other communities (P T J Johnson et al., 2015).
-One method to separately quantify the effects of host abundance and competence on parasite transmission is to use community competence: the weighted average of competence values of host species in the community (Pieter T.J.
-Johnson et al., 2013a).
+Generalist parasites, such as ranaviruses, are commonly found in nature and often use a range of alternative host species to reproduce [@park_characterizing_2018; @poulin_comparative_2011].
+The transmission potential of a generalist parasite for a given host species may vary due to differences in host species abundances [@fenton_parasite_2002; @hopkins_systematic_2020], host competence (i.e.
+the ability for an individual host to become infected and pass infection on to others [@martin_burrows_2019]), and the environmental conditions in which the interaction takes place [@altizer_climate_2013; @cable_global_2017; @rohr_frontiers_2011].
+The combination of these factors makes it difficult to assess how multi-species host communities are differentially at risk of parasite invasion and onward transmission to other communities [@johnson_why_2015].
+One method to separately quantify the effects of host abundance and competence on parasite transmission is to use community competence: the weighted average of competence values of host species in the community [@johnson_biodiversity_2013].
 In addition, community competence is advantageous because it considers the relative abundance of a species rather than the presence of a species alone.
-For example, a commonly occurring, low-competence host could be responsible for more transmission than a rare, high-competence host (Holt et al., 2003).
-In contrast to host species richness, which is the typical predictor variable in the dilution effect hypothesis literature, community competence and host abundance are more directly related to the parasite basic reproductive number, R0 (Dobson, 2004; Rohr et al., 2019).
+For example, a commonly occurring, low-competence host could be responsible for more transmission than a rare, high-competence host [@holt_parasite_2003].
+In contrast to host species richness, which is the typical predictor variable in the dilution effect hypothesis literature, community competence and host abundance are more directly related to the parasite basic reproductive number, R0 [@dobson_population_2004; @rohr_towards_2019].
 

### 2022-09-23 ms/manuscript.Rmd
Update manuscript.Rmd

@@
 
-Ranaviruses (genus Ranavirus, family Iridoviridae) are well-suited for studying host-parasite-environment interactions at the community scale and are associated with worldwide amphibian declines, making them a parasite of global concern (Lesbarrères et al., 2012).
-The Ranavirus genus of double-stranded DNA viruses (Gray et al., 2009), affects most amphibian species (Jesse L Brunner et al., 2015) and host species exhibit a wide range of competence for the parasite (Hoverman et al., 2011).
+Ranaviruses (genus Ranavirus, family Iridoviridae) are well-suited for studying host-parasite-environment interactions at the community scale and are associated with worldwide amphibian declines, making them a parasite of global concern [@lesbarreres_ranavirus_2012].
+The Ranavirus genus of double-stranded DNA viruses [@gray_ecology_2009], affects most amphibian species [@brunner_ranavirus_2015] and host species exhibit a wide range of competence for the parasite [@hoverman_phylogeny_2011].
 At the Savannah River Site (South Carolina, USA), a spatially-distributed set of 20 wetlands has been extensively sampled in the context of ranavirus infection over a 6-month period.
-We focus on larval amphibian communities which are bound by their aquatic habitat but connected through the overland movement of their adult stages (Mihaljevic et al., 2018).
-Ranavirus is transmitted via intra- and interspecific direct transmission, which exhibits a positive, but ultimately saturating, relationship with host abundance (Jesse L Brunner et al., 2017).
-Additionally, ranavirus can be transmitted environmentally via free-living infectious virions shed into water by infected live individuals or decaying carcasses (Brenes et al., 2014).
-Ranavirus persistence times in water have been found to diminish at high ambient temperatures (Nazir et al., 2012) suggesting that the relative importance of environmental transmission, versus contact-based transmission, should depend on environmental conditions.
+We focus on larval amphibian communities which are bound by their aquatic habitat but connected through the overland movement of their adult stages [@mihaljevic_parasite_2018].
+Ranavirus is transmitted via intra- and interspecific direct transmission, which exhibits a positive, but ultimately saturating, relationship with host abundance [@brunner_heterogeneities_2017].
+Additionally, ranavirus can be transmitted environmentally via free-living infectious virions shed into water by infected live individuals or decaying carcasses [@brenes_transmission_2014].
+Ranavirus persistence times in water have been found to diminish at high ambient temperatures [@nazir_environmental_2012] suggesting that the relative importance of environmental transmission, versus contact-based transmission, should depend on environmental conditions.
 We hypothesize that community-scale prevalence of ranavirus will have a positive relationship with host abundance and community competence but a negative relationship with increasing water temperatures.
 
-We begin with a theoretical model, where we use the community basic reproductive number, R0, to estimate the transmission potential of ranavirus in multi-host communities (Dobson, 2004).
+We begin with a theoretical model, where we use the community basic reproductive number, R0, to estimate the transmission potential of ranavirus in multi-host communities [@dobson_population_2004].
 R0 provides a valuable lens into transmission potential, clarifying conditions necessary for parasite invasion and persistence.
@@
 In our case it also includes the relative force of infection from free-living infectious virions, versus infected hosts.
-Multiple transmission routes can increase overall transmission potential, as evidenced by contact and environmental transmission in models of avian influenza virus (Breban et al., 2009) and causal agents of diarrheal diseases (Tien & Earn, 2010) as well as contact and sexual transmission in models of Ebola transmission (Abbate et al., 2016; Vinson et al., 2016).
+Multiple transmission routes can increase overall transmission potential, as evidenced by contact and environmental transmission in models of avian influenza virus [@breban_role_2009] and causal agents of diarrheal diseases [@tien_multiple_2010] as well as contact and sexual transmission in models of Ebola transmission [@abbate_potential_2016; @vinson_potential_2016].
 From our model, we analyzed the relative contribution of direct and environmental transmission in allowing the parasite to exceed the critical value of R0=1.
@@
 
-Using empirical data, we then apply metacommunity theory and dimensionality reduction to explain how biotic and abiotic factors contribute to community composition (Leibold et al., 2004), and then we evaluate these communities in their ability to differentially facilitate transmission.
-Community compositions change over space or time and can include dynamic, predictable patterns of assembly or disassembly (Pieter T.J.
-Johnson et al., 2019; Ostfeld & LoGiudice, 2003).
-These changes in community composition can then influence transmission potential of a parasite and may map to dilution or amplification effects (Luis et al., 2018; North et al., 2015).
-According to theory, a dilution effect may occur whenever hosts vary in competence and reductions in diversity are associated with increases in mean competence (Pieter T.J.
-Johnson et al., 2015).
-In contrast, an amplification effect can occur when increased host diversity causes an increase in mean competence (Keesing et al., 2006).
+Using empirical data, we then apply metacommunity theory and dimensionality reduction to explain how biotic and abiotic factors contribute to community composition [@leibold_metacommunity_2004], and then we evaluate these communities in their ability to differentially facilitate transmission.
+Community compositions change over space or time and can include dynamic, predictable patterns of assembly or disassembly [@johnson_community_2019; @ostfeld_community_2003].
+These changes in community composition can then influence transmission potential of a parasite and may map to dilution or amplification effects [@luis_species_2018; @north_anthropogenic_2015].
+According to theory, a dilution effect may occur whenever hosts vary in competence and reductions in diversity are associated with increases in mean competence [@johnson_why_2015].
+In contrast, an amplification effect can occur when increased host diversity causes an increase in mean competence [@keesing_effects_2006].
 Considering these conflicting patterns, we relate species evenness, in addition to species richness, as a predictor for both community competence and infection prevalence.
@@
 
-According to known transmission, ranaviruses can infect a wide range of amphibian hosts and infectious periods can range from a few days up to weeks (Gray et al., 2009).
+According to known transmission, ranaviruses can infect a wide range of amphibian hosts and infectious periods can range from a few days up to weeks[@gray_ecology_2009].
 Transmission can occur both directly and indirectly and exposure appears to induce an adaptive immune response in surviving hosts.
@@
 As well as facilitating model analysis, this choice also reflects the empirical observation that viral loads of host species are bimodal (Supplementary Figure 1).
-In addition to intra- and inter-specific direct transmission, we included environmental transmission, whereby infectious host individuals shed virus into the environment where it persists for some finite time and can cause new infections without host-to-host contact (Gray et al., 2009).
+In addition to intra- and inter-specific direct transmission, we included environmental transmission, whereby infectious host individuals shed virus into the environment where it persists for some finite time and can cause new infections without host-to-host contact [@gray_ecology_2009].
 We included host demography via a constant birth rate and a constant per capita mortality rate, resulting in a disease-free equilibrium for each host species given by the ratio of the birth rate and mortality rate.
@@
 
-Using the next-generation matrix method (Diekmann et al., 1990, 2009), we calculated the community basic reproductive number (Dobson, 2004), hereafter referred to as R0, for our community of hosts to determine the conditions necessary for parasite invasion (R0>1).
+Using the next-generation matrix method [@diekmann_definition_1990; @diekmann_construction_2009], we calculated the community basic reproductive number [@dobson_population_2004], hereafter referred to as R0, for our community of hosts to determine the conditions necessary for parasite invasion (R0>1).
 Consequently, we determined how the boundary R0=1 operates as a function of parameters for communities with varying characteristics, specifically the effects of community composition, total host abundance, and viral half-life.
@@
 
-All empirical data were previously collected (Coleman, 2018) at the Savannah River Site (South Carolina, USA).
+All empirical data were previously collected [@coleman_incorporating_2018] at the Savannah River Site (South Carolina, USA).
 20 wetlands were sampled monthly for 6 months from February to July in 2016 at the Savannah River Site.
@@
 First, using a metacommunity framework, we created a site-by-species matrix that included species presence or absence, and where ‘site’ referred to a unique wetland-month combination, referred to herein as a community.
-Within this matrix, communities were ordinated by minimizing the number of embedded absences and were then tested for key metacommunity patterns including nestedness and block replacement using the “metacom” package in R (Dallas, 2014; Mathew A Leibold & Mikkelson, 2002).
+Within this matrix, communities were ordinated by minimizing the number of embedded absences and were then tested for key metacommunity patterns including nestedness and block replacement using the “metacom” package in R [@dallas_metacom_2014; @leibold_coherence_2002].
 Following this, we used principal component analysis to order communities according to their similarity in community composition.
@@
 Following this, we treated community competence and infection prevalence as separate response variables and used species richness and species evenness as predictor variables.
-Community competence is used to measure an instantaneous property of the community in terms of its propensity to support a parasite, but infection prevalence is expected to manifest after sufficient transmission has occurred and so was lagged by a month (the period of time between measurements and a plausible estimate of the parasite generation time (Brunner & Collins, 2009)).
+Community competence is used to measure an instantaneous property of the community in terms of its propensity to support a parasite, but infection prevalence is expected to manifest after sufficient transmission has occurred and so was lagged by a month (the period of time between measurements and a plausible estimate of the parasite generation time [@brunner_testing_2009]).
 We estimated infection prevalence for the community by combining the observed host abundances and the raw infection prevalence of individual host organisms that were sampled for parasite presence using qPCR.
 The sampled hosts provided a percentage for each species that were found to be infected and this percentage was then weighted by the observed abundance of that species in that community.
-Species richness was calculated as the sum of unique species present and species evenness was measured by Pielou’s J (Pielou, 1966).
+Species richness was calculated as the sum of unique species present and species evenness was measured by Pielou’s J [@pielou_interpretation_1984].
 Species richness was used to test if basic patterns of assembly and disassembly were able to predict transmission potential via dilution or amplification effects, whereas species evenness was used to detect if relative abundances of species were controlling transmission via community competence.
@@
 Finally, we tested for statistically significant relationships between infection prevalence and community competence, abundance, and mean water temperature using both multivariate (generalized linear model) and univariate tests (Spearman Rank correlation test with a correction for multiple comparisons).
-Mean water temperature was included to serve as a proxy for environmental persistence time of free-living infectious virions; as mean water temperature increases the viral half-life is expected to decrease (Nazir et al., 2012).
+Mean water temperature was included to serve as a proxy for environmental persistence time of free-living infectious virions; as mean water temperature increases the viral half-life is expected to decrease [@nazir_environmental_2012].
 
@@
 Finally, we found that community competence and host abundance correlated positively with prevalence while mean water temperature correlated negatively with prevalence.
-In separate univariate analyses when using a Spearman rank correlation test we found significant results for community competence (P < 0.0001; figure 5), abundance (P < 0.01; figure 5), and mean water temperature (P = 0.021; figure 5) after adjusting for multiple comparisons using the “Holm” method (Aickin & Gensler, 1996; Holm, 1979).
+In separate univariate analyses when using a Spearman rank correlation test we found significant results for community competence (P < 0.0001; figure 5), abundance (P < 0.01; figure 5), and mean water temperature (P = 0.021; figure 5) after adjusting for multiple comparisons using the “Holm” method [@2] (Aickin & Gensler, 1996; Holm, 1979).
 Due to the inherently noisy data on parasite prevalence, a multivariate model that predicted infection prevalence in response to community competence, total host abundance, and mean water temperature was not able to detect any significant relationships.
@@
 
- From our theoretical model, we learned that the inclusion of both contact-based and environmental transmission allowed for each transmission mode to compensate for the other; a low strength of one transmission mode could be supplemented by the other transmission mode, resulting in a broader range of favorable conditions for the parasite (Vinson et al., 2016).
-In other studies, it has been noted that the incorporation of environmental transmission can provide dramatically different estimates of R0 and allow for persistence of parasites at population sizes that would not be supported by contact transmission alone (Breban et al., 2009, 2010; Tien & Earn, 2010).
+ From our theoretical model, we learned that the inclusion of both contact-based and environmental transmission allowed for each transmission mode to compensate for the other; a low strength of one transmission mode could be supplemented by the other transmission mode, resulting in a broader range of favorable conditions for the parasite [@vinson_potential_2016].
+In other studies, it has been noted that the incorporation of environmental transmission can provide dramatically different estimates of R0 and allow for persistence of parasites at population sizes that would not be supported by contact transmission alone [@breban_role_2009; @breban_general_2010; @tien_multiple_2010].
 We build upon these findings by demonstrating how the relative strength of transmission modes can emphasize the importance of different factors relevant to parasite transmission.
@@
 From the empirical data, we found that community composition and evenness were far more informative in predicting transmission potential compared to species richness, in spite of the fact that the diversity-disease literature includes many correlative studies that investigate the relationship between species richness and infection prevalence.
-Indeed, recent research has emphasized the importance of a more mechanistic understanding of diversity-disease relationships, and host species evenness may aid in this effort (Merrill & Johnson, 2020; Rohr et al., 2019).
+Indeed, recent research has emphasized the importance of a more mechanistic understanding of diversity-disease relationships, and host species evenness may aid in this effort [@merrill_towards_2020; @rohr_towards_2019].
 Evenness, rather than richness, can more directly relate the composition of a community to its transmission potential.
 At low evenness, the identity of the dominant species drives community competence, and we found that communities with the highest absolute abundance were usually of lower evenness and dominated by a few highly competent species.
-This pattern is supported in theory connecting life history traits to host competence, where hosts with a faster pace-of-life invest heavily into rapid growth and reproduction and less into immunity, resulting in highly abundant, highly competent hosts (Joseph et al., 2013).
+This pattern is supported in theory connecting life history traits to host competence, where hosts with a faster pace-of-life invest heavily into rapid growth and reproduction and less into immunity, resulting in highly abundant, highly competent hosts [@joseph_does_2013].
 Together, host abundance, evenness, and community competence have the potential to offer a more mechanistic lens into community compositions and how they will affect infection prevalence.
 
-Detection of diversity-disease relationships is facilitated by predictable community structures (Faust et al., 2017).
-Classic dilution-effect systems typically operate by nested structures influencing parasite transmission through biodiversity loss (LoGiudice et al., 2003).
-In these examples, species losses follow a pattern where less competent hosts are lost before more competent hosts (Pieter T.J.
+Detection of diversity-disease relationships is facilitated by predictable community structures [@faust_null_2017].
+Classic dilution-effect systems typically operate by nested structures influencing parasite transmission through biodiversity loss [@logiudice_ecology_2003].
+In these examples, species losses follow a pattern where less competent hosts are lost before more competent hosts [@2] (Pieter T.J.
 Johnson et al., 2013b).
@@
 However, the generality of this depends on how competence is typically distributed among a community of host species and how relationships among those species affect their probability of co-occurring and abundance.
-The amphibian species in our study showed evidence of phylogenetic repulsion and environmental filtering, as has been detected, for example, in avian communities (Weinstein et al., 2017).
+The amphibian species in our study showed evidence of phylogenetic repulsion and environmental filtering, as has been detected, for example, in avian communities [@2] (Weinstein et al., 2017).
 When co-occurring, species that were closely related or distantly related to their neighbors had low relative abundance (Supplementary Figure 6).
@@
 
-The study of individual host competence and how it scales to a community can provide us with valuable insight into how well host communities can support generalist parasites (Becker et al., 2020).
-However, host competence can be a challenging trait to measure because it is multifactorial, relying on components such as exposure, susceptibility, suitability, and transmissibility (Downs et al., 2019).
+The study of individual host competence and how it scales to a community can provide us with valuable insight into how well host communities can support generalist parasites [@2] (Becker et al., 2020).
+However, host competence can be a challenging trait to measure because it is multifactorial, relying on components such as exposure, susceptibility, suitability, and transmissibility [@2] (Downs et al., 2019).
 In our study, we use mean viral load as a proxy for host competence which is useful because it informs shedding rate, infectivity, and host suitability.
 However, viral load is not a holistic measure of host competence because it does not address, for example, the infectious period or susceptibility.
-Further research on host competence should incorporate more robust measures of competence such as those derived from controlled infection assays (Merrill et al., 2019).
-Factors within host species such as age and sex will also likely change the competence of a host and while there may be the ability to measure competence with genetic methods, there is likely to be phenotypic plasticity within the trait (Gervasi et al., 2015).
+Further research on host competence should incorporate more robust measures of competence such as those derived from controlled infection assays [@2] (Merrill et al., 2019).
+Factors within host species such as age and sex will also likely change the competence of a host and while there may be the ability to measure competence with genetic methods, there is likely to be phenotypic plasticity within the trait [@2] (Gervasi et al., 2015).
 
-Ranaviruses in amphibians are a useful and important system for studying diversity-disease relationships because of their highly generalist nature and role in global amphibian biodiversity declines (Lesbarrères et al., 2012; North et al., 2015).
+Ranaviruses in amphibians are a useful and important system for studying diversity-disease relationships because of their highly generalist nature and role in global amphibian biodiversity declines [@2] (Lesbarrères et al., 2012; North et al., 2015).
 Environmental transmission of ranavirus and the connection of wetlands by amphibian adults made this system especially useful for our study of the effects of host composition, host abundance, and environmental factors.
 However, these features are not limited to this study system and other multi-host generalist parasite systems can benefit from our findings.
-Furthermore, our research focused on natural variations in biodiversity, but it is important to recognize how anthropogenic influence will alter these patterns as well (Altizer et al., 2013; Rohr et al., 2011).
-Amphibian species, in particular, have been extremely sensitive to anthropogenic disturbance and have been considered proxies for overall environmental health (Scheele et al., 2019).
-Land use and climate change are also factors that affect amphibian communities globally and it is important to understand the implications of these changes (Wilkinson et al., 2018).
+Furthermore, our research focused on natural variations in biodiversity, but it is important to recognize how anthropogenic influence will alter these patterns as well [@2] (Altizer et al., 2013; Rohr et al., 2011).
+Amphibian species, in particular, have been extremely sensitive to anthropogenic disturbance and have been considered proxies for overall environmental health [@2] (Scheele et al., 2019).
+Land use and climate change are also factors that affect amphibian communities globally and it is important to understand the implications of these changes [@2] (Wilkinson et al., 2018).
 Altogether, the ecology of ranavirus in amphibians and the global relevance of amphibian biodiversity make this system a valuable case study for the joint influence of biotic and abiotic factors on parasite transmission under natural and artificial settings.

### 2022-09-23 ms/manuscript.Rmd
Update manuscript.Rmd

@@
 Classic dilution-effect systems typically operate by nested structures influencing parasite transmission through biodiversity loss [@logiudice_ecology_2003].
-In these examples, species losses follow a pattern where less competent hosts are lost before more competent hosts [@2] (Pieter T.J.
-Johnson et al., 2013b).
+In these examples, species losses follow a pattern where less competent hosts are lost before more competent hosts [@johnson_biodiversity_2013].
 In contrast, our study revealed block-replacement of sets of host species, where transmission potential was not changing due to species richness, but rather by species evenness and identity; often, ranavirus attained high prevalence due to correlated competence and abundance of certain key species.
 However, the generality of this depends on how competence is typically distributed among a community of host species and how relationships among those species affect their probability of co-occurring and abundance.
-The amphibian species in our study showed evidence of phylogenetic repulsion and environmental filtering, as has been detected, for example, in avian communities [@2] (Weinstein et al., 2017).
+The amphibian species in our study showed evidence of phylogenetic repulsion and environmental filtering, as has been detected, for example, in avian communities [@weinstein_role_2017].
 When co-occurring, species that were closely related or distantly related to their neighbors had low relative abundance (Supplementary Figure 6).
@@
 
-The study of individual host competence and how it scales to a community can provide us with valuable insight into how well host communities can support generalist parasites [@2] (Becker et al., 2020).
-However, host competence can be a challenging trait to measure because it is multifactorial, relying on components such as exposure, susceptibility, suitability, and transmissibility [@2] (Downs et al., 2019).
+The study of individual host competence and how it scales to a community can provide us with valuable insight into how well host communities can support generalist parasites [@becker_beyond_2020].
+However, host competence can be a challenging trait to measure because it is multifactorial, relying on components such as exposure, susceptibility, suitability, and transmissibility [@downs_scaling_2019].
 In our study, we use mean viral load as a proxy for host competence which is useful because it informs shedding rate, infectivity, and host suitability.
 However, viral load is not a holistic measure of host competence because it does not address, for example, the infectious period or susceptibility.
-Further research on host competence should incorporate more robust measures of competence such as those derived from controlled infection assays [@2] (Merrill et al., 2019).
-Factors within host species such as age and sex will also likely change the competence of a host and while there may be the ability to measure competence with genetic methods, there is likely to be phenotypic plasticity within the trait [@2] (Gervasi et al., 2015).
+Further research on host competence should incorporate more robust measures of competence such as those derived from controlled infection assays [@merrill_variation_2019].
+Factors within host species such as age and sex will also likely change the competence of a host and while there may be the ability to measure competence with genetic methods, there is likely to be phenotypic plasticity within the trait [@gervasi_context_2015].
 
-Ranaviruses in amphibians are a useful and important system for studying diversity-disease relationships because of their highly generalist nature and role in global amphibian biodiversity declines [@2] (Lesbarrères et al., 2012; North et al., 2015).
+Ranaviruses in amphibians are a useful and important system for studying diversity-disease relationships because of their highly generalist nature and role in global amphibian biodiversity declines [@lesbarreres_ranavirus_2012; @north_anthropogenic_2015].
 Environmental transmission of ranavirus and the connection of wetlands by amphibian adults made this system especially useful for our study of the effects of host composition, host abundance, and environmental factors.
 However, these features are not limited to this study system and other multi-host generalist parasite systems can benefit from our findings.
-Furthermore, our research focused on natural variations in biodiversity, but it is important to recognize how anthropogenic influence will alter these patterns as well [@2] (Altizer et al., 2013; Rohr et al., 2011).
-Amphibian species, in particular, have been extremely sensitive to anthropogenic disturbance and have been considered proxies for overall environmental health [@2] (Scheele et al., 2019).
-Land use and climate change are also factors that affect amphibian communities globally and it is important to understand the implications of these changes [@2] (Wilkinson et al., 2018).
+Furthermore, our research focused on natural variations in biodiversity, but it is important to recognize how anthropogenic influence will alter these patterns as well [@altizer_climate_2013; @rohr_frontiers_2011].
+Amphibian species, in particular, have been extremely sensitive to anthropogenic disturbance and have been considered proxies for overall environmental health [@scheele_amphibian_2019].
+Land use and climate change are also factors that affect amphibian communities globally and it is important to understand the implications of these changes [@wilkinson_habitat_2018].
 Altogether, the ecology of ranavirus in amphibians and the global relevance of amphibian biodiversity make this system a valuable case study for the joint influence of biotic and abiotic factors on parasite transmission under natural and artificial settings.

### 2022-09-23 ms/manuscript.Rmd
update ms with citations

@@
 Finally, we found that community competence and host abundance correlated positively with prevalence while mean water temperature correlated negatively with prevalence.
-In separate univariate analyses when using a Spearman rank correlation test we found significant results for community competence (P < 0.0001; figure 5), abundance (P < 0.01; figure 5), and mean water temperature (P = 0.021; figure 5) after adjusting for multiple comparisons using the “Holm” method [@2] (Aickin & Gensler, 1996; Holm, 1979).
+In separate univariate analyses when using a Spearman rank correlation test we found significant results for community competence (P < 0.0001; figure 5), abundance (P < 0.01; figure 5), and mean water temperature (P = 0.021; figure 5) after adjusting for multiple comparisons using the “Holm” method [@aickin_adjusting_1996; @holm_simple_1979].
 Due to the inherently noisy data on parasite prevalence, a multivariate model that predicted infection prevalence in response to community competence, total host abundance, and mean water temperature was not able to detect any significant relationships.

### 2022-09-23 ms/manuscript.Rmd
update ms

@@
 output:   
-  pdf_document: default
   word_document:
     reference_docx: rv_ms_081921.docx
+  pdf_document: default
 bibliography: refs.bib

### 2022-10-03 scripts/comm_similarity.Rmd
break up workshop script

--- 2022-09-16 scripts/workshop/workshop.Rmd
@@
 ---
-title: "workshop"
+title: "distances"
 author: "Daniel Suh"
-date: "`r format(Sys.time(), '%d %B, %Y')`"
+date: "2022-10-03"
 output: html_document
@@
 
-```{r}
+```{r, message = F}
 library(here)
@@
 
-```{r}
+```{r, message = F}
 data <- read_csv(here("data/weighted_prev_competence_111220.csv"))
@@
 
-Principal Component Ranking may not be reliable as a proxy for community similarity because of double-zeroes
-
-If two sites both have zeroes for a species then that does not show that a site is similar but only shows that both sites may be unsuitable for the species
-
-Bray-Curtis dissimilarity does not face this limitation
-
-But the question is how similar are sites with the similar community competence?
-
-sometimes similar and sometimes not
-
-## next step
-
-do PCA after Hellinger transformation of data.
-PCA has the double zero problem but Hellinger distance avoids that problem and avoids overweighting of abundance data by turning absolute species abundances into relative abundances.
-more info [here](https://davidzeleny.net/blog/2022/03/17/euclidean-distance-is-sensitive-to-double-zero-problem-while-hellinger-is-not-visualization/)
-
-```{r}
-pca <- princomp(abundance_mat)
-
-biplot(pca)
-
-abundance_hellinger <- decostand(log1p(abundance_mat), 'hellinger')
-
-tb_pca <- princomp(abundance_hellinger)
-
-biplot(tb_pca)
-```
-
-```{r}
-tb_pca_scores <- as.data.frame(tb_pca$scores) %>% dplyr::select(Comp.1, Comp.2) %>% rownames_to_column(., var = "siteID")
-
-tb_pca_scores_summ <- left_join(comm_summ, tb_pca_scores, by = "siteID")
-
-tb_pca_scores_summ$Month <- factor(tb_pca_scores_summ$Month, levels = c("Feb", "Mar", "Apr", "May", "Jun", "Jul"))
-
-tb_pca_scores_summ %<>% mutate(pc1Rank=dense_rank(Comp.1))
-
-tb_pca_scores_summ %>% ggplot(.,aes(x=pc1Rank, y=cc)) + geom_point()
-
-tb_pca_scores_summ %>% ggplot(.,aes(x=Comp.1, y=cc)) + geom_point() + geom_smooth(method = "loess")
-
-tb_pca_scores_summ %>% ggplot(.,aes(x=pc1Rank, y=cc)) + geom_point(aes(color = Month, shape = factor(WetAltID), size = size))+
-  labs(x="Principal Component 1", y = "Community Competence (CC)", size = "Host\nAbundance") + 
-  scale_shape_manual(values = rep(1:20, len = 20)) +
-  scale_size_continuous(range = c(2,10)) +
-  guides(shape=F) +
-  theme_classic() +
-  theme(aspect.ratio=1/1.618, legend.box = "horizontal") +
-  scale_color_viridis_d(direction = -1) +
-  theme(text = element_text(size=18))
-
-```
-
 ### Let's try to explain the residuals
@@
 
-## Variation in competence for each species
+```{r}
+dist_df %>% ggplot(.,aes(x=j_dist,y=p_dist,color=resid)) + geom_point() + scale_color_viridis_c()
 
-```{r}
-species_comp <- data %>% group_by(Species) %>% summarize(mean = mean(SQMean),
-                                                         var = var(SQMean),
-                                                         se = sqrt(var(SQMean)/n()),
-                                                         n = n()) %>% mutate(species_code = Species)
+dist_df %<>% mutate(resid_sign = ifelse(resid>0,1,0))
 
-species_comp %<>% left_join(.,names)
-
-vl %>% ggplot(.,aes(x=reorder(abb_name, mean), y=mean)) + geom_point() + geom_linerange(aes(ymin=mean-se,ymax=mean+se)) + theme(axis.text.x = element_text(angle = 30, hjust = 1))
-
+dist_df %>% ggplot(.,aes(x=j_dist,y=p_dist,color=resid_sign)) + geom_point() + scale_color_viridis_b()
 ```

### 2022-10-03 scripts/workshop/gradient_analysis.Rmd
break up workshop script

--- 2022-09-14 scripts/gradient_analysis.Rmd
@@
 What is the difference between the scores for species from envfit and the values from the correlation between environmental ordination scores and the community matrix?
+
+```{r}
+comm_summ <- data %>% dplyr::select(WetAltID, Month.1, Month, cc, Month, AB2:AB42, Prevalence, MeanWaterTempPredC) %>% 
+  dplyr::select(-ABAmb, -ABSal) %>%  
+  mutate(., size = rowSums(.[5:25]), 
+    ABHigh = (AB9 + AB21 + AB26 + AB42), 
+    ABLow = (AB2 + AB3 + AB4 + AB5 + AB6 + AB8 + AB20 + AB24 + AB27 + AB28 + AB29 + AB31 + AB34 + AB35 + AB38 + AB39 + AB41)) %>% #get totals for community size
+  mutate(., siteID = paste(WetAltID, Month.1, sep = "_")) %>%
+  distinct()
+comm_summ$siteID <- gsub("Month", "", comm_summ$siteID)
+```
+
+```{r}
+community_mat <- data %>% dplyr::select(WetAltID, Month.1, AB2:AB8, AB9, AB20:AB42) 
+community_mat$Month <- gsub("Month", "", community_mat$Month.1)
+community_mat %<>% mutate(., siteID = paste(WetAltID, Month, sep = "_")) %>% distinct() %>% arrange(.,WetAltID) %>% dplyr::select(siteID, AB2:AB42)
+
+#make abundance and pres/abs matrices with rownames as siteIDs
+abundance_mat <- community_mat %>% column_to_rownames(., var = "siteID")
+presence_mat <- as.data.frame(ifelse(abundance_mat>0, 1, 0))
+```
+
+Principal Component Ranking may not be reliable as a proxy for community similarity because of double-zeroes
+
+If two sites both have zeroes for a species then that does not show that a site is similar but only shows that both sites may be unsuitable for the species
+
+Bray-Curtis dissimilarity does not face this limitation
+
+But the question is how similar are sites with the similar community competence?
+
+sometimes similar and sometimes not
+
+## next step
+
+do PCA after Hellinger transformation of data.
+PCA has the double zero problem but Hellinger distance avoids that problem and avoids overweighting of abundance data by turning absolute species abundances into relative abundances.
+more info [here](https://davidzeleny.net/blog/2022/03/17/euclidean-distance-is-sensitive-to-double-zero-problem-while-hellinger-is-not-visualization/)
+
+```{r}
+pca <- princomp(abundance_mat)
+
+biplot(pca)
+
+abundance_hellinger <- decostand(log1p(abundance_mat), 'hellinger')
+
+tb_pca <- princomp(abundance_hellinger)
+
+biplot(tb_pca)
+```
+
+```{r}
+tb_pca_scores <- as.data.frame(tb_pca$scores) %>% dplyr::select(Comp.1, Comp.2) %>% rownames_to_column(., var = "siteID")
+
+tb_pca_scores_summ <- left_join(comm_summ, tb_pca_scores, by = "siteID")
+
+tb_pca_scores_summ$Month <- factor(tb_pca_scores_summ$Month, levels = c("Feb", "Mar", "Apr", "May", "Jun", "Jul"))
+
+tb_pca_scores_summ %<>% mutate(pc1Rank=dense_rank(Comp.1))
+
+tb_pca_scores_summ %>% ggplot(.,aes(x=pc1Rank, y=cc)) + geom_point()
+
+tb_pca_scores_summ %>% ggplot(.,aes(x=Comp.1, y=cc)) + geom_point() + geom_smooth(method = "loess")
+
+tb_pca_scores_summ %>% ggplot(.,aes(x=pc1Rank, y=cc)) + geom_point(aes(color = Month, shape = factor(WetAltID), size = size))+
+  labs(x="Principal Component 1", y = "Community Competence (CC)", size = "Host\nAbundance") + 
+  scale_shape_manual(values = rep(1:20, len = 20)) +
+  scale_size_continuous(range = c(2,10)) +
+  guides(shape=F) +
+  theme_classic() +
+  theme(aspect.ratio=1/1.618, legend.box = "horizontal") +
+  scale_color_viridis_d(direction = -1) +
+  theme(text = element_text(size=18))
+
+```
+
+NMDS with bray-curtis
+```{r}
+set.seed(8878896)
+nmds <- metaMDS(abundance_mat, distance = "bray")
+
+plot(nmds)
+
+nmds_scores <- as_tibble(scores(nmds))
+nmds_scores$siteID <- community_mat$siteID
+
+nmds_scores_summ <- left_join(comm_summ, nmds_scores, by = "siteID")
+nmds_scores_summ$Month <- factor(nmds_scores_summ$Month, levels = c("Feb", "Mar", "Apr", "May", "Jun", "Jul"))
+nmds_scores_summ %<>% mutate(pc1Rank=dense_rank(NMDS1))
+
+nmds_scores_summ %>% ggplot(.,aes(x=NMDS1, y = cc)) + geom_point() + geom_smooth(method = "loess")
+```

### 2022-10-03 scripts/workshop/workshop.Rmd
break up workshop script

--- 2022-09-19 scripts/workshop/workshop.Rmd
@@
 names <- read_csv(here("data/raw_data/species_names_ids.csv"))
-```
-
-![motivating figure](https://dcsuh.github.io/rana-competence/figures/fig_2.png)
-
-# Questions
-
-Do communities that are similar in community composition have similar community competence values?
-
-Can communities with different community compositions have similar community competence values?
-
-How are these communities distributed across space and time?
-i.e. how much spatial and temporal autocorrelation is there between sites in terms of community composition and community competence?
-
-```{r}
-comm_summ <- data %>% dplyr::select(WetAltID, Month.1, Month, cc, Month, AB2:AB42, Prevalence, MeanWaterTempPredC) %>% 
-  dplyr::select(-ABAmb, -ABSal) %>%  
-  mutate(., size = rowSums(.[5:25]), 
-    ABHigh = (AB9 + AB21 + AB26 + AB42), 
-    ABLow = (AB2 + AB3 + AB4 + AB5 + AB6 + AB8 + AB20 + AB24 + AB27 + AB28 + AB29 + AB31 + AB34 + AB35 + AB38 + AB39 + AB41)) %>% #get totals for community size
-  mutate(., siteID = paste(WetAltID, Month.1, sep = "_")) %>%
-  distinct()
-comm_summ$siteID <- gsub("Month", "", comm_summ$siteID)
-```
-
-```{r}
-community_mat <- data %>% dplyr::select(WetAltID, Month.1, AB2:AB8, AB9, AB20:AB42) 
-community_mat$Month <- gsub("Month", "", community_mat$Month.1)
-community_mat %<>% mutate(., siteID = paste(WetAltID, Month, sep = "_")) %>% distinct() %>% arrange(.,WetAltID) %>% dplyr::select(siteID, AB2:AB42)
-
-#make abundance and pres/abs matrices with rownames as siteIDs
-abundance_mat <- community_mat %>% column_to_rownames(., var = "siteID")
-presence_mat <- as.data.frame(ifelse(abundance_mat>0, 1, 0))
-```
-
-Use Bray-Curtis distances to analyze how similar sites are to each other
-Higher BC distance means more dissimilar
-1 means perfect mismatch
-```{r}
-log_abun_mat <- log1p(abundance_mat)
-
-bc_dist <- vegdist(log_abun_mat, method = "bray") #bray-curtis dissimilarity is default
-
-bc_dist_mat <- as.matrix(bc_dist)
-bc_dist_df <- melt(bc_dist_mat, varnames=c("row", "col"), value.name = "bc_dist")
-
-bc_dist_df %>% filter(bc_dist!=0) %>% ggplot(., aes(x=bc_dist)) + geom_histogram()
-ggplot(bc_dist_df, aes(x = col, y = row, fill = bc_dist)) + geom_tile()
-```
-
-quick check on normality of community competence data
-```{r}
-comm_summ %>% ggplot(.,aes(x=cc)) + geom_histogram()
-```
-
-Manhattan method is the absolute value of the difference between two values
-So higher distance means larger difference
-Normalized distances to have range 0 to 1
-```{r}
-cc_diff <- comm_summ %>% dplyr::select(siteID, cc) %>% mutate(log_cc = log1p(cc)) %>% dplyr::select(-cc)
-cc_dist <- dist(cc_diff, method = "manhattan")
-cc_diff_mat <- as.matrix(cc_dist, labels = T)
-colnames(cc_diff_mat) <- cc_diff[[1]]
-rownames(cc_diff_mat) <- cc_diff[[1]]
-cc_diff_df <- melt(cc_diff_mat, varnames=c("row", "col"), value.name = "cc_dist") %>% 
-  mutate(cc_norm = (cc_dist-min(.$cc_dist))/(max(.$cc_dist)-min(.$cc_dist)))
-
-cc_diff_df %>% filter(cc_norm != 0) %>% ggplot(., aes(x=cc_norm)) + geom_histogram()
-ggplot(cc_diff_df, aes(x = col, y = row, fill = cc_norm)) + geom_tile()
-```
-
-```{r}
-bc_cc_df <- left_join(cc_diff_df, bc_dist_df)
-
-bc_cc_df %<>% mutate(diff = cc_norm-bc_dist)
-
-```
-
-Each point is a pairwise interaction
-
-The null hypothesis is that community competence is achieved through specific community compositions. i.e. you cannot achieve the same community competence with different community compositions
-
-So if a pair of sites have very different community competence then they should also have very different community composition
-
-But if community competence can be achieved through a variety of different community compositions then a pair of sites with similar community competence can have a high level of dissimilarity
-
-So we are looking for low cc distance between two sites but high bc distance
-An upward trend is expected because more dissimilar sites in terms of competence probably have different composition
-```{r}
-bc_cc_df %>% filter(diff!=0) %>% ggplot(.,aes(x=bc_dist, y=cc_norm)) + geom_point() + geom_smooth(method = "lm")
-```
-
-If diff is set as community competence similarity minus community composition similarity then a low value will mean the difference between composition similarity is large compared to competence similarity.
-Either competence similarity is very high (cc dist is low) and composition similarity (bc dist) is moderate or high OR composition similarity is low (i.e.
-BC dist is high) and competence similarity (cc dist is moderate or low)
-
-```{r}
-
-ggplot(bc_cc_df, aes(x = col, y = row, fill = diff)) + geom_tile()
-
-bc_cc_df %>% filter(diff!=0) %>% ggplot(.,aes(x=diff)) + geom_histogram()
-```
-
-Principal Component Ranking may not be reliable as a proxy for community similarity because of double-zeroes
-
-If two sites both have zeroes for a species then that does not show that a site is similar but only shows that both sites may be unsuitable for the species
-
-Bray-Curtis dissimilarity does not face this limitation
-
-But the question is how similar are sites with the similar community competence?
-
-sometimes similar and sometimes not
-
-## next step
-
-do PCA after Hellinger transformation of data.
-PCA has the double zero problem but Hellinger distance avoids that problem and avoids overweighting of abundance data by turning absolute species abundances into relative abundances.
-more info [here](https://davidzeleny.net/blog/2022/03/17/euclidean-distance-is-sensitive-to-double-zero-problem-while-hellinger-is-not-visualization/)
-
-```{r}
-pca <- princomp(abundance_mat)
-
-biplot(pca)
-
-abundance_hellinger <- decostand(log1p(abundance_mat), 'hellinger')
-
-tb_pca <- princomp(abundance_hellinger)
-
-biplot(tb_pca)
-```
-
-```{r}
-tb_pca_scores <- as.data.frame(tb_pca$scores) %>% dplyr::select(Comp.1, Comp.2) %>% rownames_to_column(., var = "siteID")
-
-tb_pca_scores_summ <- left_join(comm_summ, tb_pca_scores, by = "siteID")
-
-tb_pca_scores_summ$Month <- factor(tb_pca_scores_summ$Month, levels = c("Feb", "Mar", "Apr", "May", "Jun", "Jul"))
-
-tb_pca_scores_summ %<>% mutate(pc1Rank=dense_rank(Comp.1))
-
-tb_pca_scores_summ %>% ggplot(.,aes(x=pc1Rank, y=cc)) + geom_point()
-
-tb_pca_scores_summ %>% ggplot(.,aes(x=Comp.1, y=cc)) + geom_point() + geom_smooth(method = "loess")
-
-tb_pca_scores_summ %>% ggplot(.,aes(x=pc1Rank, y=cc)) + geom_point(aes(color = Month, shape = factor(WetAltID), size = size))+
-  labs(x="Principal Component 1", y = "Community Competence (CC)", size = "Host\nAbundance") + 
-  scale_shape_manual(values = rep(1:20, len = 20)) +
-  scale_size_continuous(range = c(2,10)) +
-  guides(shape=F) +
-  theme_classic() +
-  theme(aspect.ratio=1/1.618, legend.box = "horizontal") +
-  scale_color_viridis_d(direction = -1) +
-  theme(text = element_text(size=18))
-
-```
-
-### Let's try to explain the residuals
-
-The residuals from the cc_norm ~ bc_dist represent when the community competence is more or less similar than expected given the similarity of the community.
-If community composition is not explaining similarity in community competence then this must be due to the communities having a tradeoff where competence is derived from a different class of hosts (low competence vs.
-high competence; abundant low competence hosts can result in community competence that is similar to rare high competent host) or a substitution of hosts (proportion of low vs.
-high competence hosts may be similar but the identity of those hosts may be different).
-These situations will result in community competence that is unexplained by community composition (i.e.
-high or low residual).
-
-plot residuals on y-axis and pairwise difference in evenness (tradeoff) and different in proportion high-low competence (substitution) on x-axis
-```{r}
-comm_data <- readRDS(here("processed_data","comm_data.rds"))
-
-delta_j <- comm_data %>% select(siteID, J)
-
-delta_p <- comm_data %>% select(siteID, ABHigh, ABLow) %>% mutate(P = ABHigh/(ABHigh+ABLow)) %>% select(-c(ABHigh, ABLow))
-
-```
-
-```{r}
-j_dist <- dist(delta_j, method = "manhattan")
-j_dist_mat <- as.matrix(j_dist, labels = T)
-colnames(j_dist_mat) <- delta_j[[1]]
-rownames(j_dist_mat) <- delta_j[[1]]
-
-j_dist_df <- melt(j_dist_mat, varnames=c("row", "col"), value.name = "j_dist")
-```
-
-```{r}
-p_dist <- dist(delta_p, method = "manhattan")
-p_dist_mat <- as.matrix(p_dist, labels = T)
-colnames(p_dist_mat) <- delta_p[[1]]
-rownames(p_dist_mat) <- delta_p[[1]]
-
-p_dist_df <- melt(p_dist_mat, varnames=c("row", "col"), value.name = "p_dist")
-```
-
-```{r}
-dist_df <- bc_cc_df %>% left_join(.,j_dist_df)
-dist_df %<>% left_join(.,p_dist_df)
-```
-
-```{r}
-model <- lm(cc_norm ~ bc_dist, data = dist_df)
-
-cc_bc_resid <- tibble(row = dist_df[[1]], col = dist_df[[2]], resid = residuals(model))
-
-dist_df %<>% left_join(.,cc_bc_resid)
-```
-
-```{r}
-dist_df %>% ggplot(.,aes(x=j_dist, y=resid)) + geom_point()
-
-dist_df %>% ggplot(.,aes(x=p_dist, y=resid)) + geom_point()
 ```

### 2022-10-03 scripts/workshop/comm_similarity.Rmd
update community similarity analysis rmd

--- 2022-10-03 scripts/comm_similarity.Rmd
@@
 ---
-title: "distances"
+title: "Community Similarity Analysis"
 author: "Daniel Suh"
@@
 
-dist_df %<>% mutate(resid_sign = ifelse(resid>0,1,0))
+dist_df %<>% mutate(resid_sign = as.factor(ifelse(resid>0,1,0)))
 
-dist_df %>% ggplot(.,aes(x=j_dist,y=p_dist,color=resid_sign)) + geom_point() + scale_color_viridis_b()
+dist_df %>% ggplot(.,aes(x=j_dist,y=p_dist,color=resid_sign)) + geom_point() + scale_color_viridis_d()
 ```

### 2022-10-03 ms/manuscript.Rmd
update scripts with date

@@
 output:
+  html_document:
+    df_print: paged
+  pdf_document: default
   word_document:
     reference_docx: rv_ms_081921.docx
-  pdf_document: default
 bibliography: refs.bib

### 2022-10-03 scripts/workshop/gradient_analysis.Rmd
update scripts with date

@@
 author: "Daniel Suh"
-date: "10/30/2020"
+date: "`r format(Sys.time(), '%d %B, %Y')`"
 output: html_document

### 2022-10-03 scripts/workshop/comm_similarity.Rmd
update scripts with date
update distances

@@
 author: "Daniel Suh"
-date: "2022-10-03"
+date: "`r format(Sys.time(), '%d %B, %Y')`"
 output: html_document
@@
 cc_diff <- comm_summ %>% dplyr::select(siteID, cc) %>% mutate(log_cc = log1p(cc)) %>% dplyr::select(-cc)
-cc_dist <- dist(cc_diff, method = "manhattan")
+cc_dist <- dist(cc_diff$log_cc, method = "manhattan")
 cc_diff_mat <- as.matrix(cc_dist, labels = T)
@@
 ```{r}
-j_dist <- dist(delta_j, method = "manhattan")
+j_dist <- dist(delta_j$J, method = "manhattan")
 j_dist_mat <- as.matrix(j_dist, labels = T)
@@
 ```{r}
-p_dist <- dist(delta_p, method = "manhattan")
+p_dist <- dist(delta_p$P, method = "manhattan")
 p_dist_mat <- as.matrix(p_dist, labels = T)

### 2022-12-06 scripts/workshop/sem.Rmd
Create sem.Rmd

---
title: "sem"
output: html_document
date: "2022-12-06"
---

```{r, message = F}
library(here)

library(tidyverse)
library(magrittr)
library(piecewiseSEM)

comm_data <- readRDS(here("processed_data","comm_data.rds"))
```

cc ~ mean water temp; size ~ mean water temp; prevalence ~ cc + size + mean water temp
```{r}
cc_temp <- glm(cc ~ MeanWaterTempPredC, data = comm_data)
size_temp <- glm(size ~ MeanWaterTempPredC, data = comm_data)
prev_cc <- glm(Prevalence ~ cc + size, data = comm_data)

sem <- psem(cc_temp, size_temp, prev_cc)
```

```{r}
summary(sem)
```

### 2022-12-07 scripts/workshop/sem.Rmd
Update sem.Rmd

@@
 size_temp <- glm(size ~ MeanWaterTempPredC, data = comm_data)
-prev_cc <- glm(Prevalence ~ cc + size, data = comm_data)
+prev_cc <- glm(Prevalence ~ cc + MeanWaterTempPredC + size, data = comm_data)
 
@@
 ```{r}
+dSep(sem)
 summary(sem)

### 2022-12-07 scripts/workshop/sem.Rmd
wroking on lefchk tutorial

@@
 library(piecewiseSEM)
+library(nlme)
+library(lmerTest)
 
@@
 ```
+Lefcheck example
+
+```{r}
+library(devtools) #Install from CRAN if you do not already have
+install_github("jslefche/piecewiseSEM")
+library(piecewiseSEM)
+```
+
+```{r}
+shipley2009.modlist = list(
+  lme(DD~lat, random = ~1|site/tree, na.action = na.omit, 
+    data = Shipley),
+  lme(Date~DD, random = ~1|site/tree, na.action = na.omit, 
+    data = Shipley),
+  lme(Growth~Date, random = ~1|site/tree, na.action = na.omit, 
+    data = Shipley),
+  glmer(Live~Growth+(1|site)+(1|tree), 
+    family=binomial(link = "logit"), data = shipley2009) )
+```
+
+```{r}
+sem.fit(shipley2009.modlist, shipley2009)
+```

### 2022-12-07 scripts/workshop/sem.Rmd
Update sem.Rmd
shipley example

--- 2022-12-07 scripts/workshop/sem.Rmd
@@
 ```
+Lefcheck example
+
+```{r}
+data(shipley)
+```
+
+```{r}
+dd_lat <-  lme(DD~lat, random = ~1|site/tree, na.action = na.omit, 
+    data = shipley)
+
+date_dd <-lme(Date~DD, random = ~1|site/tree, na.action = na.omit, 
+    data = shipley)
+
+growth_date <-  lme(Growth~Date, random = ~1|site/tree, na.action = na.omit, 
+    data = shipley)
+
+live_growth <-  glmer(Live~Growth+(1|site)+(1|tree), 
+    family=binomial(link = "logit"), data = shipley)
+```
+
+```{r}
+shipley_sem <- psem(dd_lat, date_dd, growth_date, live_growth)
+```
+
+```{r}
+summary(shipley_sem)
+```

### 2022-12-09 scripts/workshop/sem.Rmd
look at sem more in depth

@@
 library(lmerTest)
+library(dagitty)
 
 comm_data <- readRDS(here("processed_data","comm_data.rds"))
+```
+
+```{r}
+prev_mod <- dagitty("dag{
+  temp -> cc
+  temp -> size
+  cc -> prev
+  size -> prev
+  temp -> prev
+             }")
+
+plot(graphLayout(prev_mod))
+
 ```
@@
 ```{r}
-dSep(sem)
+basisSet(sem)
+dSep(sem, conditioning = T)
+fisherC(sem)
+coefs(sem, intercepts = T)
+rsquared(sem)
+```
+
+```{r}
 summary(sem)
+plot(sem)
+plot(sem, return = T)
 ```
+
 Lefcheck example

### 2022-12-09 scripts/workshop/sem.Rmd
log transform variables

@@
 
+histograms
+```{r}
+comm_data %>% ggplot(.,aes(x=MeanWaterTempPredC)) + geom_histogram()
+comm_data %>% ggplot(.,aes(x=size)) + geom_histogram()
+comm_data %>% ggplot(.,aes(x=log10(size))) + geom_histogram()
+comm_data %>% ggplot(.,aes(x=cc)) + geom_histogram()
+comm_data %>% ggplot(.,aes(x=log10(cc))) + geom_histogram()
+comm_data %>% ggplot(.,aes(x=Prevalence)) + geom_histogram()
+```
+plots
+```{r}
+comm_data %>% ggplot(.,aes(x=MeanWaterTempPredC, y = log10(cc))) + geom_point() + geom_smooth(method = "lm")
+comm_data %>% ggplot(.,aes(x=MeanWaterTempPredC, y = log10(size))) + geom_point() + geom_smooth(method = "lm")
+comm_data %>% ggplot(.,aes(x=MeanWaterTempPredC, y = Prevalence)) + geom_point() + geom_smooth(method = "lm")
+comm_data %>% ggplot(.,aes(x=log10(cc), y = Prevalence)) + geom_point() + geom_smooth(method = "lm")
+comm_data %>% ggplot(.,aes(x=log10(size), y = Prevalence)) + geom_point() + geom_smooth(method = "lm")
+```
+
 cc ~ mean water temp; size ~ mean water temp; prevalence ~ cc + size + mean water temp
 ```{r}
-cc_temp <- glm(cc ~ MeanWaterTempPredC, data = comm_data)
-size_temp <- glm(size ~ MeanWaterTempPredC, data = comm_data)
-prev_cc <- glm(Prevalence ~ cc + MeanWaterTempPredC + size, data = comm_data)
+cc_temp <- glm(log10(cc) ~ MeanWaterTempPredC, data = comm_data)
+size_temp <- glm(log10(size) ~ MeanWaterTempPredC, data = comm_data)
+prev_cc <- glm(Prevalence ~ log10(cc) + MeanWaterTempPredC + log10(size), data = comm_data)
 

### 2022-12-09 scripts/workshop/sem.Rmd
trying to use gams

@@
 library(lmerTest)
+library(mgcv)
 library(dagitty)
@@
 ```{r}
-cc_temp <- glm(log10(cc) ~ MeanWaterTempPredC, data = comm_data)
-size_temp <- glm(log10(size) ~ MeanWaterTempPredC, data = comm_data)
-prev_cc <- glm(Prevalence ~ log10(cc) + MeanWaterTempPredC + log10(size), data = comm_data)
+cc_temp <- gam(log10(cc) ~ s(MeanWaterTempPredC), family = gaussian, data = comm_data)
+size_temp <- gam(log10(size) ~ s(MeanWaterTempPredC), family = gaussian, data = comm_data)
+prev_cc <- glm(Prevalence ~ log10(cc) + log10(size), data = comm_data)
 

### 2022-12-09 scripts/workshop/sem.Rmd
add significant relationship from dsep test

@@
 
+How to model nonlinear effect of temperature?
+- hump-shaped relationship may be expected with temperature
+
 cc ~ mean water temp; size ~ mean water temp; prevalence ~ cc + size + mean water temp
 ```{r}
-cc_temp <- gam(log10(cc) ~ s(MeanWaterTempPredC), family = gaussian, data = comm_data)
-size_temp <- gam(log10(size) ~ s(MeanWaterTempPredC), family = gaussian, data = comm_data)
+cc_temp <- glm(log10(cc) ~ (MeanWaterTempPredC), data = comm_data)
+size_temp <- glm(log10(size) ~ (MeanWaterTempPredC) + log10(cc), data = comm_data)
 prev_cc <- glm(Prevalence ~ log10(cc) + log10(size), data = comm_data)

### 2022-12-12 scripts/workshop/sem.Rmd
add lagged data

@@
 
+Make lagged data
+```{r}
+prev_cc <- comm_data
+
+order <- prev_cc[order(prev_cc$WetAltID, prev_cc$Month.1),]
+
+#this removes months out of sequence
+order <- order[-c(4,14,47,90),]
+
+order %<>% add_column(lag_cc = NA, lag_size = NA, lag_temp = NA)
+
+#create new column that includes previous month's value for cc
+for(n in 2:nrow(order)){
+  order$lag_cc[n] <- order$cc[n-1]
+  order$lag_size[n] <- order$size[n-1]
+  order$lag_temp[n] <- order$MeanWaterTempPredC[n-1]
+}
+
+#remove the first entry for each wetland to remove the carryover from the last wetland
+clean <- order %>% group_by(WetAltID) %>% filter(duplicated(WetAltID) | n()==1)
+
+```
+
 ```{r}
@@
   temp -> size
+  temp -> prev
   cc -> prev
   size -> prev
-  temp -> prev
              }")
@@
 
+Covariance-based SEM
+```{r}
+
+```
+
 How to model nonlinear effect of temperature?
@@
 ```{r}
-cc_temp <- glm(log10(cc) ~ (MeanWaterTempPredC), data = comm_data)
-size_temp <- glm(log10(size) ~ (MeanWaterTempPredC) + log10(cc), data = comm_data)
-prev_cc <- glm(Prevalence ~ log10(cc) + log10(size), data = comm_data)
+cc_temp <- lm(log10(cc) ~ (MeanWaterTempPredC), data = comm_data)
+size_temp <- lm(log10(size) ~ (MeanWaterTempPredC), data = comm_data)
+prev_cc <- lm(Prevalence ~ log10(cc) + log10(size), data = comm_data)
+prev_temp_cc <- glm(Prevalence ~ log10(cc) + log10(size) + MeanWaterTempPredC, data = comm_data)
 
 sem <- psem(cc_temp, size_temp, prev_cc)
+sem_1 <- psem(cc_temp, size_temp, prev_temp_cc)
+```
+
+Using lagged data
+```{r}
+lagged_cc_temp <- lm(lag_cc ~ lag_temp, data = clean)
+lagged_size_temp <- lm(log10(lag_size) ~ lag_temp + lag_cc, data = clean)
+lagged_prev_cc <- lm(Prevalence ~ lag_cc + log10(lag_size), data = clean)
+lagged_prev_temp_cc <- glm(Prevalence ~ lag_cc + log10(lag_size) + lag_temp, data = clean)
+
+sem_lagged <- psem(lagged_cc_temp, lagged_size_temp, lagged_prev_cc)
+sem_1_lagged <- psem(lagged_cc_temp, lagged_size_temp, lagged_prev_temp_cc)
 ```
@@
 ```{r}
+basisSet(sem_1)
+dSep(sem_1, conditioning = T)
+fisherC(sem_1)
+coefs(sem_1, intercepts = T)
+rsquared(sem_1)
+```
+
+```{r}
 summary(sem)
@@
 plot(sem, return = T)
+```
+
+```{r}
+summary(sem_1)
+plot(sem_1)
+plot(sem_1, return = T)
+```
+
+```{r}
+summary(sem_lagged)
+plot(sem_lagged)
+plot(sem_lagged, return = T)
 ```

### 2022-12-12 scripts/workshop/sem.Rmd
with direct effect temp

@@
 lagged_prev_cc <- lm(Prevalence ~ lag_cc + log10(lag_size), data = clean)
-lagged_prev_temp_cc <- glm(Prevalence ~ lag_cc + log10(lag_size) + lag_temp, data = clean)
+lagged_prev_temp_cc <- lm(Prevalence ~ lag_cc + log10(lag_size) + lag_temp, data = clean)
 
@@
 
+```{r}
+summary(sem_1_lagged)
+plot(sem_1_lagged)
+plot(sem_1_lagged, return = T)
+```
+
 Lefcheck example

### 2022-12-29 scripts/workshop/sem.Rmd
Update sem.Rmd

@@
 comm_data %>% ggplot(.,aes(x=cc)) + geom_histogram()
-comm_data %>% ggplot(.,aes(x=log10(cc))) + geom_histogram()
+comm_data %>% ggplot(.,aes(x=log(cc))) + geom_histogram()
 comm_data %>% ggplot(.,aes(x=Prevalence)) + geom_histogram()
+comm_data %>% ggplot(.,aes(x=log10(Prevalence+1))) + geom_histogram()
+
 ```
@@
 ```{r}
-comm_data %>% ggplot(.,aes(x=MeanWaterTempPredC, y = log10(cc))) + geom_point() + geom_smooth(method = "lm")
-comm_data %>% ggplot(.,aes(x=MeanWaterTempPredC, y = log10(size))) + geom_point() + geom_smooth(method = "lm")
-comm_data %>% ggplot(.,aes(x=MeanWaterTempPredC, y = Prevalence)) + geom_point() + geom_smooth(method = "lm")
+comm_data %>% ggplot(.,aes(x=MeanWaterTempPredC, y = cc)) + geom_point() + geom_smooth()
+comm_data %>% ggplot(.,aes(x=MeanWaterTempPredC, y = log10(size))) + geom_point() + geom_smooth()
+comm_data %>% ggplot(.,aes(x=MeanWaterTempPredC, y = Prevalence)) + geom_point() + geom_smooth()
 comm_data %>% ggplot(.,aes(x=log10(cc), y = Prevalence)) + geom_point() + geom_smooth(method = "lm")
@@
 prev_temp_cc <- glm(Prevalence ~ log10(cc) + log10(size) + MeanWaterTempPredC, data = comm_data)
+
+sem <- psem(cc_temp, size_temp, prev_cc)
+sem_1 <- psem(cc_temp, size_temp, prev_temp_cc)
+```
+
+```{r}
+comm_data %<>% filter(MeanWaterTempPredC>0)
+
+cc_temp <- gam(log10(cc) ~ s(MeanWaterTempPredC), data = comm_data)
+size_temp <- lm(log10(size) ~ (MeanWaterTempPredC), data = comm_data)
+prev_cc <- lm(Prevalence ~ log10(cc) + log10(size), data = comm_data)
+prev_temp_cc <- glm(Prevalence ~ log10(cc) + log10(size) + MeanWaterTempPredC, data = comm_data)
+
+summary(cc_temp)
 

### 2023-01-31 scripts/workshop/sem.Rmd
troubleshooting error
add more notes

@@
 
-Make lagged data
+Make lagged data - I'll use this later to compare results using lagged and not lagged data
 ```{r}
@@
 
+This is just a way to visualize the SEM
+
 ```{r}
@@
 
-histograms
+Just some histograms to check the distributions of the data
+
 ```{r}
@@
 ```
-plots
+
+And some basic plots
+
 ```{r}
@@
 
-sem <- psem(cc_temp, size_temp, prev_cc)
-sem_1 <- psem(cc_temp, size_temp, prev_temp_cc)
+#fix error
+#sem <- psem(cc_temp, size_temp, prev_cc)
+#sem_1 <- psem(cc_temp, size_temp, prev_temp_cc)
 ```
@@
 Lefcheck example
+This is example data and code that I got from Lefcheck's tutorial
+https://cran.r-project.org/web/packages/piecewiseSEM/vignettes/piecewiseSEM.html#worked-example
 

### 2023-01-31 scripts/workshop/sem.Rmd
more text

@@
 library(dagitty)
+```
 
+Read in data
+```{r, message = F}
 comm_data <- readRDS(here("processed_data","comm_data.rds"))
 ```
+
+The purpose of this script was to try out SEM with the ranavirus data to answer whether there were direct or indirect effects of temperature on infection prevalence that were possible being mediated by abundance or community competence.
+SEM can help us to understand relationships between multiple variables that may be causally related to each other.
 
@@
 
-This is just a way to visualize the SEM
+This is just a way to visualize the SEM. I think it must have a method for customizing the orientation of things to make it look nicer but I haven't tried to figure out how to do that.
 
@@
 
-And some basic plots
+And some basic plots for all the univariate relationships I am considering
 
 ```{r}
-comm_data %>% ggplot(.,aes(x=MeanWaterTempPredC, y = cc)) + geom_point() + geom_smooth()
-comm_data %>% ggplot(.,aes(x=MeanWaterTempPredC, y = log10(size))) + geom_point() + geom_smooth()
-comm_data %>% ggplot(.,aes(x=MeanWaterTempPredC, y = Prevalence)) + geom_point() + geom_smooth()
-comm_data %>% ggplot(.,aes(x=log10(cc), y = Prevalence)) + geom_point() + geom_smooth(method = "lm")
-comm_data %>% ggplot(.,aes(x=log10(size), y = Prevalence)) + geom_point() + geom_smooth(method = "lm")
-```
-
-Covariance-based SEM
-```{r}
-
+comm_data %>% ggplot(.,aes(x=MeanWaterTempPredC, y = cc)) + geom_point() + geom_smooth() + 
+  labs(title = "cc ~ temp")
+comm_data %>% ggplot(.,aes(x=MeanWaterTempPredC, y = log10(size))) + geom_point() + geom_smooth() + 
+  labs(title = "log(size) ~ temp")
+comm_data %>% ggplot(.,aes(x=MeanWaterTempPredC, y = Prevalence)) + geom_point() + geom_smooth() + 
+  labs(title = "Prev ~ temp")
+comm_data %>% ggplot(.,aes(x=log10(cc), y = Prevalence)) + geom_point() + geom_smooth(method = "lm") + 
+  labs(title = "Prev ~ log(cc)")
+comm_data %>% ggplot(.,aes(x=log10(size), y = Prevalence)) + geom_point() + geom_smooth(method = "lm") + 
+  labs(title = "Prev ~ log(size)")
 ```
@@
 
-cc ~ mean water temp; size ~ mean water temp; prevalence ~ cc + size + mean water temp
+# SEMs
+
+In this first iteration, I first propose the relationships independently and then combine them later to generate the SEM.
+Each relationship is modeled as a linear regression and is evaluated separately.
+The 'piecewiseSEM' package generates a test statistic (Fisher's C) and and AIC score through a method known as "tests of directed seperation".
+This method evalutes the relationships between variables that are *un*specified in the model and asks whether these relationships are truly independent or whether some causal relationship may exist.
+piecewiseSEM relies on a priori knowledge of the system that can be used to rationalize expected relationships in the data.
+The benefits of piecewiseSEM is that it allows for non-independence of the data and allows for each relationship to be evaluated independently using
+
+cc ~ mean water temp; 
+size ~ mean water temp; 
+prevalence ~ cc + size + mean water temp
+
 ```{r}
@@
 ```
+
+Covariance-based SEM
+
+Traditional SEM creates a single variance-covariance matrix between all variables and relationships between variables are estimated simultaneously from that matrix.
+This method assumes independence and normality of errors which are probably not things we can assume about our dataset.
+I also haven't yet figured out how to implement this method of SEM.
+
+```{r}
+
+```

### 2023-01-31 scripts/workshop/sem.Rmd
add text and knit to html

@@
 Make lagged data - I'll use this later to compare results using lagged and not lagged data
+
 ```{r}
@@
 clean <- order %>% group_by(WetAltID) %>% filter(duplicated(WetAltID) | n()==1)
+```
 
+```{r}
+comm_data %<>% filter(!is.na(MeanWaterTempPredC)) #remove missing water temp data
+clean %<>% filter(!is.na(MeanWaterTempPredC)) 
 ```
@@
 
-How to model nonlinear effect of temperature?
-- hump-shaped relationship may be expected with temperature
-
 # SEMs
@@
 piecewiseSEM relies on a priori knowledge of the system that can be used to rationalize expected relationships in the data.
-The benefits of piecewiseSEM is that it allows for non-independence of the data and allows for each relationship to be evaluated independently using
+The benefits of piecewiseSEM is that it allows for non-independence of the data and allows for each relationship to be evaluated independently using a more flexible range of modeling options (e.g.
+GLM and GAM).
+I develop two sets of SEMs here.
+One assumes prevalence is only influenced directly by cc and size and the other assumes that water temp may also have a direct effect on prevalence.
 
@@
 size ~ mean water temp; 
+prevalence ~ cc
 prevalence ~ cc + size + mean water temp
@@
 
+How to model nonlinear effect of temperature?
+- hump-shaped relationship may be expected with temperature
+
+I attempt to use a gam to model the relationship between water temp and cc and a glm to model the effects of cc, size, and temp on infection prevalence but I keep getting errors and can't figure out how to fix it.
+For some reason, I am losing all of the data during this process.
+
 ```{r}
-comm_data %<>% filter(MeanWaterTempPredC>0)
-
 cc_temp <- gam(log10(cc) ~ s(MeanWaterTempPredC), data = comm_data)
@@
 #fix error
-#sem <- psem(cc_temp, size_temp, prev_cc)
-#sem_1 <- psem(cc_temp, size_temp, prev_temp_cc)
+#Error in data.frame(..., check.names = FALSE) :
+#arguments imply differing number of rows: 0, 83
+# sem <- psem(cc_temp, size_temp, prev_cc)
+# sem_1 <- psem(cc_temp, size_temp, prev_temp_cc)
 ```
 
-Using lagged data
+Same relationships using lagged data
+
 ```{r}
@@
 ```
+
+The following chunks goes through the piecewiseSEM steps individually.
 
@@
 
+The following sections shows the same information as before but using summary() instead of doing it individually. I also plot all of the SEMs that were previously evaluated.
+
 ```{r}
@@
 
+In the unlagged data, there are no significant relationships between any variables and prevalence.
+Also, the model that does not include a direct relationship between temperature and prevalence has a slightly lower (35.971 vs.
+36.545) AIC than the other.
+
 ```{r}
@@
 ```
+
+In the lagged data, there is no significant direct relationship between temperature and prevalence but there is a signficant relationship between temperature and cc and cc and prevalence.
+This would mean that there is an indirect relationship between temperature and prevalence that is mediated by cc.
+Temperature has a negative effect on cc but cc has a positive effect on prevalence.
 

### 2023-01-31 scripts/workshop/sem.Rmd
add text

@@
 ```
+
+# Intro
 
@@
   labs(title = "Prev ~ temp")
-comm_data %>% ggplot(.,aes(x=log10(cc), y = Prevalence)) + geom_point() + geom_smooth(method = "lm") + 
+comm_data %>% ggplot(.,aes(x=cc, y = Prevalence)) + geom_point() + geom_smooth(method = "lm") + 
   labs(title = "Prev ~ log(cc)")
@@
 
-# SEMs
+# SEMs intro
 
@@
 Each relationship is modeled as a linear regression and is evaluated separately.
-The 'piecewiseSEM' package generates a test statistic (Fisher's C) and and AIC score through a method known as "tests of directed seperation".
+The 'piecewiseSEM' package generates a test statistic (Fisher's C) and and AIC score through a method known as "tests of directed separation".
 This method evalutes the relationships between variables that are *un*specified in the model and asks whether these relationships are truly independent or whether some causal relationship may exist.
@@
 prev_cc <- lm(Prevalence ~ log10(cc) + log10(size), data = comm_data)
-prev_temp_cc <- glm(Prevalence ~ log10(cc) + log10(size) + MeanWaterTempPredC, data = comm_data)
+prev_temp_cc <- lm(Prevalence ~ log10(cc) + log10(size) + MeanWaterTempPredC, data = comm_data)
 
@@
 lagged_cc_temp <- lm(lag_cc ~ lag_temp, data = clean)
-lagged_size_temp <- lm(log10(lag_size) ~ lag_temp + lag_cc, data = clean)
+lagged_size_temp <- lm(log10(lag_size) ~ lag_temp, data = clean)
 lagged_prev_cc <- lm(Prevalence ~ lag_cc + log10(lag_size), data = clean)
@@
 ```
+
+# piecewiseSEM
 
@@
 ```
+
+# SEM results and plots
 
@@
 
-Lefcheck example
+Both SEMs are unable to do the tests of directed separation. I think this may be due to there not being enough variables in the model for this test to work.
+
+# Next Steps
+
+ - Try to incorporate nonlinear regression into model for temperature relationships
+  - I want to use a GAM with a smooth term for temperature
+ - Check to see if there is a problem without having the directed separation test
+  - Still not sure why it says there are no independence claims present when it seems like there is?
+  - There seems to be a significant relationship between log10(size) ~ cc that isn't being specified but is showing up under the independence claims result
+
+# Lefcheck example
 This is example data and code that I got from Lefcheck's tutorial
@@
 
-Covariance-based SEM
+# Covariance-based SEM
 

### 2023-01-31 scripts/workshop/sem.Rmd
update list formatting

@@
 
- - Try to incorporate nonlinear regression into model for temperature relationships
-  - I want to use a GAM with a smooth term for temperature
- - Check to see if there is a problem without having the directed separation test
-  - Still not sure why it says there are no independence claims present when it seems like there is?
-  - There seems to be a significant relationship between log10(size) ~ cc that isn't being specified but is showing up under the independence claims result
+* Try to incorporate nonlinear regression into model for temperature relationships
+  + I want to use a GAM with a smooth term for temperature
+* Check to see if there is a problem without having the directed separation test
+  + Still not sure why it says there are no independence claims present when it seems like there is?
+  + There seems to be a significant relationship between log10(size) ~ cc that isn't being specified but is showing up under the independence claims result
 

### 2023-02-07 scripts/workshop/sem.Rmd
minor edit
Update sem.Rmd

@@
 library(here)
-
 library(tidyverse)
@@
 
+Example with GAMs from Lefcheck textbook: https://jslefche.github.io/sem_book/local-estimation.html#extensions-to-non-linear-models
+
+```{r}
+set.seed(100)
+n <- 100
+x1 <- rchisq(n, 7)
+mu2 <- 10*x1/(5 + x1)
+x2 <- rnorm(n, mu2, 1)
+x2[x2 <= 0] <- 0.1
+x3 <- rpois(n, lambda = (0.5*x2))
+x4 <- rpois(n, lambda = (0.5*x2))
+p.x5 <- exp(-0.5*x3 + 0.5*x4)/(1 + exp(-0.5*x3 + 0.5*x4))
+x5 <- rbinom(n, size = 1, prob = p.x5)
+dat2 <- data.frame(x1 = x1, x2 = x2, x3 = x3, x4 = x4, x5 = x5)
+```
+
+```{r}
+shipley_psem2 <- psem(
+  lm(x2 ~ x1, data = dat2),
+  lm(x3 ~ x2, data = dat2),
+  lm(x4 ~ x2, data = dat2),
+  lm(x5 ~ x3 + x4, data = dat2)
+)
+
+#LLchisq(shipley_psem2)
+```
+
+```{r}
+
+m1 <- gam(x2 ~ s(x1), data = dat2, family = gaussian)
+m2 <- glm(x3 ~ x2, data = dat2, family = poisson)
+m3 <- gam(x4 ~ x2, data = dat2, family = poisson)
+m4 <- glm(x5 ~ x3 + x4, data = dat2, family = binomial)
+
+shipley_psem3 <- psem(m1, m2, m3, m4)
+
+#LLchisq(shipley_psem3)
+```
+
 # Covariance-based SEM

### 2023-02-09 scripts/workshop/sem.Rmd
use dev version of piecewiseSEM

@@
 
+I fixed this error by installing the development version of piecewiseSEM
+
+```{r}
+#devtools::install_github("jslefche/piecewiseSEM@devel")
+```
+
+Caveat: GAMs can't be used for determining direct and indirect effects because these are not linear coefficients but rather smoothing terms.
+
 ```{r}
@@
 #arguments imply differing number of rows: 0, 83
-# sem <- psem(cc_temp, size_temp, prev_cc)
-# sem_1 <- psem(cc_temp, size_temp, prev_temp_cc)
+sem <- psem(cc_temp, size_temp, prev_cc)
+sem_1 <- psem(cc_temp, size_temp, prev_temp_cc)
 ```
@@
 shipley_psem3 <- psem(m1, m2, m3, m4)
-
 #LLchisq(shipley_psem3)

### 2023-02-09 scripts/workshop/sem.Rmd
add more models with gam

@@
 #arguments imply differing number of rows: 0, 83
-sem <- psem(cc_temp, size_temp, prev_cc)
-sem_1 <- psem(cc_temp, size_temp, prev_temp_cc)
+sem_2 <- psem(cc_temp, size_temp, prev_cc)
+sem_3 <- psem(cc_temp, size_temp, prev_temp_cc)
 ```
@@
 ```{r}
+summary(sem_2)
+```
+
+```{r}
+summary(sem_3)
+```
+
+```{r}
 summary(sem_lagged)
@@
 Both SEMs are unable to do the tests of directed separation. I think this may be due to there not being enough variables in the model for this test to work.
+
+AIC
+```{r}
+AIC(sem, sem_1, sem_2, sem_3, sem_lagged, sem_1_lagged)
+```
 

### 2023-02-09 scripts/workshop/sem.Rmd
add some text

@@
 * Try to incorporate nonlinear regression into model for temperature relationships
-  + I want to use a GAM with a smooth term for temperature
+  + I want to use a GAM with a smooth term for temperature - DONE
+  + I just had to use the development version of piecewiseSEM
 * Check to see if there is a problem without having the directed separation test
@@
   + There seems to be a significant relationship between log10(size) ~ cc that isn't being specified but is showing up under the independence claims result
+* Currently we are running all our models with siteXmonth-level data but should we run these models with individual-level data?
+  + i.e. the qPCR results?
 

### 2023-02-09 scripts/workshop/sem.Rmd
playing around with gam

@@
 ```{r}
-comm_data %>% ggplot(.,aes(x=MeanWaterTempPredC, y = cc)) + geom_point() + geom_smooth() + 
+comm_data %>% ggplot(.,aes(x=MeanWaterTempPredC, y = cc)) + geom_point() + geom_smooth(method = "lm") + 
   labs(title = "cc ~ temp")
-comm_data %>% ggplot(.,aes(x=MeanWaterTempPredC, y = log10(size))) + geom_point() + geom_smooth() + 
+comm_data %>% ggplot(.,aes(x=MeanWaterTempPredC, y = log10(size))) + geom_point() + geom_smooth(method = "lm") + 
   labs(title = "log(size) ~ temp")
-comm_data %>% ggplot(.,aes(x=MeanWaterTempPredC, y = Prevalence)) + geom_point() + geom_smooth() + 
+comm_data %>% ggplot(.,aes(x=MeanWaterTempPredC, y = Prevalence)) + geom_point() + geom_smooth(method = "lm") + 
   labs(title = "Prev ~ temp")
@@
 ```{r}
-cc_temp <- gam(log10(cc) ~ s(MeanWaterTempPredC), data = comm_data)
-size_temp <- lm(log10(size) ~ (MeanWaterTempPredC), data = comm_data)
+cc_temp <- gam(log10(cc) ~ s(MeanWaterTempPredC), data = comm_data, method = "REML")
+size_temp <- gam(log10(size) ~ s(MeanWaterTempPredC), data = comm_data, method = "REML")
 prev_cc <- lm(Prevalence ~ log10(cc) + log10(size), data = comm_data)
@@
 sem_1_lagged <- psem(lagged_cc_temp, lagged_size_temp, lagged_prev_temp_cc)
+```
+
+# Playing around with GAM
+
+```{r}
+gam_1 <- gam(Prevalence ~ s(MeanWaterTempPredC) + s(log10(cc)) + s(log10(size)), data = (comm_data %>% filter(MeanWaterTempPredC>0)), method = "REML", family = "betar")
 ```

### 2023-02-14 scripts/workshop/sem.Rmd
latent variable modeling

@@
 library(dagitty)
+library(lavaan)
 ```
@@
 ```
+
+# Latent Variable Modeling using lavaan
+
+```{r}
+potential <- comm_data %>% mutate(log_cc = log10(cc), log_size = log10(size)) %>% select(cc, log_cc, size, log_size, MeanWaterTempPredC, Prevalence) %>% drop_na() %>% filter(MeanWaterTempPredC > 0)
+cor(potential[,c(2,4,5)])
+```
+
+```{r}
+potential_formula <- "potential =~ log_cc + log_size + MeanWaterTempPredC"
+potential_model <- sem(potential_formula, potential)
+```
+
+```{r}
+summary(potential_model)
+```
+
+```{r}
+potential_formula_1 <- "
+potential =~ log_cc + log_size + MeanWaterTempPredC
+
+Prevalence ~ potential
+"
+
+potential_model_1 <- sem(potential_formula_1, potential)
+```
+
+```{r}
+summary(potential_model_1)
+```

### 2023-02-14 scripts/workshop/sem.Rmd
latent variable modeling attempt

@@
 
+First step is to measure reliability. There should be a sufficient amount of reliability (at least >0.5 but >0.8 or >0.9 is better)
 ```{r}
-potential <- comm_data %>% mutate(log_cc = log10(cc), log_size = log10(size)) %>% select(cc, log_cc, size, log_size, MeanWaterTempPredC, Prevalence) %>% drop_na() %>% filter(MeanWaterTempPredC > 0)
-cor(potential[,c(2,4,5)])
+potential <- comm_data %>% mutate(log_cc = log10(cc), log_size = log10(size), inv_temp = MeanWaterTempPredC*-1) %>% select(cc, log_cc, size, log_size, MeanWaterTempPredC, inv_temp, Prevalence) %>% drop_na() %>% filter(MeanWaterTempPredC > 0)
+indicators <- potential %>% select(log_cc, log_size, inv_temp)
+cor(indicators)
+
+#find average correlation among all variables
+
+r <- mean(c((cor(indicators)[2,1]), (cor(indicators)[3,1]), (cor(indicators)[3,2])))
+sqrt(r)
 ```
@@
 ```{r}
-potential_formula <- "potential =~ log_cc + log_size + MeanWaterTempPredC"
+potential_formula <- "potential =~ log_cc + log_size + inv_temp"
 potential_model <- sem(potential_formula, potential)
@@
 ```{r}
+print(modindices(potential_model, na.remove=F))
+```
+
+```{r}
 potential_formula_1 <- "
-potential =~ log_cc + log_size + MeanWaterTempPredC
-
+potential =~ log_cc + log_size + inv_temp 
 Prevalence ~ potential

### 2023-02-14 scripts/workshop/sem.Rmd
latent model w/ lagged data

@@
 ```{r}
-potential_formula_1 <- "
-potential =~ log_cc + log_size + inv_temp 
-Prevalence ~ potential
-"
+potential_formula_1 <- "\n potential =~ log_cc + log_size + inv_temp \n Prevalence ~ potential"
 
@@
 ```
+
+Again with lagged data
+
+```{r}
+lag_potential <- clean %>% ungroup() %>% mutate(log_cc = log10(lag_cc), log_size = log10(lag_size), inv_temp = lag_temp*-1) %>% select(log_cc, log_size, inv_temp, Prevalence) %>% drop_na() %>% filter(inv_temp < 0)
+indicators <- lag_potential %>% select(log_cc, log_size, inv_temp)
+cor(indicators)
+
+#find average correlation among all variables
+
+r <- mean(c((cor(indicators)[2,1]), (cor(indicators)[3,1]), (cor(indicators)[3,2])))
+sqrt(r)
+```
+
+```{r}
+potential_formula_2 <- "\n potential =~ log_cc + log_size + inv_temp \n Prevalence ~ potential"
+
+potential_model_2 <- sem(potential_formula_1, lag_potential)
+```
+
+```{r}
+summary(potential_model_2)
+```

### 2023-02-14 scripts/workshop/community_final.Rmd
update pathname

--- 2020-02-24 scripts/community_final.Rmd
@@
 ```{r, Ranavirus Summary}
-data <- read_csv("competence.csv")
+data <- read_csv("/Users/dcsuh/Documents/GitHub/rana-competence/data/raw_data/competence.csv")
 
@@
 ```
+
 Same summary as before but with a different measure of competency. These measures are calculated using only RV positive individuals
+
 ```{r, Ranavirus Summary 2}

### 2023-02-14 scripts/workshop/workshop.Rmd
incidence and prevalence plots

@@
 ```
+
+```{r}
+incidence <- data %>% group_by(WetAltID, Month.1) %>% summarize(incidence = sum(RV.Status))
+
+incidence %>% ggplot(.,aes(x=Month.1, y=incidence)) + geom_col() + facet_grid(vars(WetAltID))
+```
+
+```{r}
+comm_data <- readRDS(here("processed_data","comm_data.rds"))
+comm_data %>% ggplot(.,aes(x=Month.1, y=Prevalence)) + geom_point() + facet_wrap(~WetAltID, nrow = 5, ncol = 4)
+```

### 2023-02-14 scripts/workshop/workshop.Rmd
add lines

--- 2022-10-03 scripts/workshop/workshop.Rmd
@@
 ```
+
+```{r}
+incidence <- data %>% group_by(WetAltID, Month.1) %>% summarize(incidence = sum(RV.Status))
+
+incidence %>% ggplot(.,aes(x=Month.1, y=incidence)) + geom_col() + facet_grid(vars(WetAltID))
+incidence %>% ggplot(.,aes(x=Month.1, y=incidence, group = 1)) + geom_point() + geom_line() + facet_wrap(~WetAltID, nrow = 5, ncol = 4)
+```
+
+```{r}
+comm_data <- readRDS(here("processed_data","comm_data.rds"))
+comm_data %>% ungroup() %>% ggplot(.,aes(x=Month.1, y=Prevalence, group = 1)) + geom_point() + geom_line() + facet_wrap(~WetAltID, nrow = 5, ncol = 4)
+```

### 2023-02-15 scripts/workshop/workshop.Rmd
calculating standard deviations

@@
 ```
+
+```{r}
+comm_data %>% ggplot(.,aes(x=size)) + geom_histogram() + labs(title = "Untransformed abundance")
+comm_data %>% ggplot(.,aes(x=log10(size))) + geom_histogram() + labs(title = "log10-transformed abundance")
+comm_data %>% ggplot(.,aes(x=cc)) + geom_histogram() + labs(title = "Untransformed cc")
+comm_data %>% ggplot(.,aes(x=log10(cc))) + geom_histogram() + labs(title = "log10-transformed cc")
+comm_data %>% ggplot(.,aes(x=MeanWaterTempPredC)) + geom_histogram() + labs(title = "Mean Water Temp")
+```
+
+effect sizes
+```{r}
+comm_data <- readRDS(here("processed_data","comm_data.rds"))
+comm_data %>% mutate(mean = mean(size), sd = sd(size), var = mean(size)-size, size = size)
+comm_effect <- comm_data %>% filter(MeanWaterTempPredC>0) %>% mutate(size_sds = (mean(size)-size)/sd(size),
+                                    cc_sds = (mean(cc)-cc)/sd(cc),
+                                    temp_sds = ((mean(MeanWaterTempPredC)-MeanWaterTempPredC)/sd(MeanWaterTempPredC))*-1) #multiply by -1 to make it opposite because
+
+effect_summ <- comm_effect %>% group_by(WetAltID) %>% summarize(size_over = sum(size_sds>0),
+                                                                size_sum = sum(size_sds),
+                                                                cc_over = sum(cc_sds>0),
+                                                                cc_sum = sum(cc_sds),
+                                                                temp_over = sum(temp_sds>0),
+                                                                temp_sum = sum(temp_sds))
+```

### 2023-02-15 scripts/workshop/dynamics.Rmd
make script for numerical solutions to ODE's

--- 2022-09-14 scripts/workshop/alt_fig1.Rmd
@@
 ---
-title: "alternate fig 1"
+title: "dynamics"
 author: "Daniel Suh"
-date: "2/22/2022"
+date: "2023-02-15"
 output: html_document
@@
 
-```{r, message = F}
+```{r}
 library(tidyverse)
@@
 library(ggnewscale)
-
+library(here)
 ```
@@
 }
+
+reference <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = 1/45,mort2 = 1/45,degr = 1/1.947799)
+
+composition <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = 1/60,mort2 = 1/30,degr = 1/1.947799)
+
+size <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = c(1/52.5),mort2 = c(1/52.5),degr = 1/1.947799)
+
+halflife <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = c(1/40),mort2 = 1/40,degr = 1/3.895598)
+
+combined <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = c(1/70),mort2 = c(1/35),degr = 1/3.895598)
+
 ```
-
-```{r}
-reference <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0,trans_3_max = 0.001,
-                     mort1 = 1/45,mort2 = 1/45,degr = 1/1.947799)
-
-composition <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0,trans_3_max = 0.001,
-                       mort1 = 1/60,mort2 = 1/30,degr = 1/1.947799)
-
-size <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0,trans_3_max = 0.001,
-                mort1 = c(1/52.5),mort2 = c(1/52.5),degr = 1/1.947799)
-
-halflife <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0,trans_3_max = 0.001,
-                    mort1 = c(1/40),mort2 = 1/40,degr = 1/3.895598)
-
-combined <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0,trans_3_max = 0.001,
-                    mort1 = c(1/70),mort2 = c(1/35),degr = 1/3.895598)
-```
-
-```{r}
-axis_text_size = 8
-plot_label_size = 20
-
-p1 <- reference %>% ggplot(.,aes(x=prop3,y=prop1))+
-  geom_contour(aes(x=prop3,y=prop1,z=eigen,colour = factor(..level.. == 1,levels = c(F,T)),group=mort1), show.legend = FALSE)+
-  theme_classic()+
-  scale_colour_manual(na.value = NA, values = c(NA, "red")) + 
-  labs(x = "Environmental", y = "Contact") +
-  xlim(1,8) +
-  ylim(1,14) + 
-  labs(title = "Reference")
-
-p2 <- composition %>% ggplot(.,aes(x=prop3,y=prop1))+
-  geom_contour(aes(x=prop3,y=prop1,z=eigen,colour = factor(..level.. == 1,levels = c(F,T)),group=mort1), show.legend = FALSE)+
-  theme_classic()+
-  scale_colour_manual(na.value = NA, values = c(NA, "red")) + 
-  new_scale_color()+
-  geom_contour(data = reference, 
-               aes(x=prop3,y=prop1,z=eigen,colour = factor(..level.. == 1,levels = c(F,T)),group=mort1), show.legend = F)+
-  scale_colour_manual(na.value = NA, values = c(NA, "gray")) +
-  labs(x = "", y = "") +
-  xlim(1,8) +
-  ylim(1,14) + 
-  labs(title = "A") + 
-  theme(axis.title = element_text(size=axis_text_size),
-        plot.title = element_text(size=plot_label_size))
-#p2
-
-p3 <- size %>% filter(tot != 150) %>% 
-  ggplot(.,aes(x=prop3,y=prop1))+
-  geom_contour(aes(x=prop3,y=prop1,z=eigen,colour = factor(..level.. == 1,levels = c(F,T)),group=mort1), show.legend = FALSE)+
-  theme_classic()+
-  scale_colour_manual(na.value = NA, values = c(NA, "orange")) + 
-  new_scale_color()+
-  geom_contour(data = reference, 
-               aes(x=prop3,y=prop1,z=eigen,colour = factor(..level.. == 1,levels = c(F,T)),group=mort1), show.legend = F)+
-  scale_colour_manual(na.value = NA, values = c(NA, "gray")) +
-  labs(x = "", y = "") +
-  xlim(1,8) +
-  ylim(1,14) + 
-  labs(title = "B") + 
-  theme(plot.title = element_text(size=plot_label_size))
-#p3
-
-p4 <- halflife %>% ggplot(.,aes(x=prop3,y=prop1))+
-  geom_contour(aes(x=prop3,y=prop1,z=eigen,colour = factor(..level.. == 1,levels = c(F,T)),group=mort1), show.legend = FALSE)+
-  scale_colour_manual(na.value = NA, values = c(NA, "forestgreen")) + 
-  new_scale_color()+
-  geom_contour(data = reference, 
-               aes(x=prop3,y=prop1,z=eigen,colour = factor(..level.. == 1,levels = c(F,T)),group=mort1), show.legend = F)+
-  scale_colour_manual(na.value = NA, values = c(NA, "gray")) +
-  labs(x="Relative Environmental", y = "                                                               Relative Contact Transmission Rate", title = "C") +
-  xlim(1,8) +
-  ylim(1,14) + 
-  theme_classic() +
-  theme(axis.title = element_text(size=axis_text_size),
-        plot.title = element_text(size=plot_label_size),
-        axis.title.x=element_text(hjust=1.05))
-#p4
-
-p5 <- combined %>%
-  ggplot(.,aes(x=prop3,y=prop1)) +
-  geom_contour(aes(x=prop3,y=prop1,z=eigen,colour = factor(..level.. == 1,levels = c(F,T)),group=mort1), show.legend = FALSE) +
-  scale_colour_manual(na.value = NA, values = c(NA, "purple")) + 
-  new_scale_color()+
-  geom_contour(data = reference, 
-               aes(x=prop3,y=prop1,z=eigen,colour = factor(..level.. == 1,levels = c(F,T)),group=mort1), show.legend = F) +
-  scale_colour_manual(na.value = NA, values = c(NA, "gray")) +
-  labs(x = "Transmission Rate", y = "", title = "D") +
-  xlim(1,8) +
-  ylim(1,14) + 
-  theme_classic() +
-  theme(axis.title = element_text(size=axis_text_size),
-        plot.title = element_text(size=plot_label_size),
-        axis.title.x=element_text(hjust=-0.625))
-#p5
-```
-
-```{r}
-figure_1 <- (p2 | p3)/(p4 | p5)
-figure_1 + theme(title = "Original fig w/ lines for R0==1")
-```
-
-```{r}
-#heatmaps and contour plots of difference between control and treatment
-
-ref_sel <- reference %>% select(prop3,prop1,eigen) %>% rename(.,ref_eigen = eigen)
-comp_sel <- composition %>% select(prop3,prop1,eigen) %>% rename(.,comp_eigen = eigen)
-size_sel <- size %>% select(prop3,prop1,eigen) %>% rename(.,size_eigen = eigen)
-halflife_sel <- halflife %>% select(prop3,prop1,eigen) %>% rename(.,half_eigen = eigen)
-combined_sel <- combined %>% select(prop3,prop1,eigen) %>% rename(.,all_eigen = eigen)
-
-ref_comp <- inner_join(ref_sel,comp_sel, by = c("prop3","prop1"))
-ref_comp %<>% inner_join(.,size_sel, by =c("prop3","prop1"))
-ref_comp %<>% inner_join(.,halflife_sel, by =c("prop3","prop1"))
-ref_comp %<>% inner_join(.,combined_sel, by =c("prop3","prop1"))
-
-ref_comp %<>% mutate(.,comp_diff = comp_eigen-ref_eigen,
-                     size_diff = size_eigen-ref_eigen,
-                     half_diff = half_eigen-ref_eigen,
-                     all_diff = all_eigen-ref_eigen)
-
-#heatmaps with contours
-
-comp_fig <- ref_comp %>% ggplot(.,aes(x=prop3,y=prop1,fill=comp_diff)) + 
-  geom_tile() +
-  scale_fill_gradient2() + 
-  geom_contour(aes(z=comp_diff))
-
-size_fig <- ref_comp %>% ggplot(.,aes(x=prop3,y=prop1,fill=size_diff)) + 
-  geom_tile() +
-  scale_fill_gradient2() + 
-  geom_contour(aes(z=size_diff))
-
-half_fig <- ref_comp %>% ggplot(.,aes(x=prop3,y=prop1,fill=half_diff)) + 
-  geom_tile() +
-  scale_fill_gradient2() + 
-  geom_contour(aes(z=half_diff))
-
-combined_fig <- ref_comp %>% ggplot(.,aes(x=prop3,y=prop1,fill=all_diff)) + 
-  geom_tile() +
-  scale_fill_gradient2() + 
-  geom_contour(aes(z=all_diff))
-```
-
-```{r}
-(comp_fig | size_fig)/(half_fig | combined_fig) + theme(title = "heatmap of difference between treatment and reference")
-```
-
-```{r}
-#filled contour with standardized breaks
-
-breaks <- c(-0.2,-0.1,0,0.1,0.2,0.3,0.4,0.5,2)
-
-comp_fig <- ref_comp %>% ggplot(.,aes(x=prop3,y=prop1)) + 
-  geom_contour_filled(aes(z=comp_diff),breaks = breaks) +
-  scale_fill_viridis_d(drop=F) +
-  theme(legend.position = "none") + 
-  labs(x="environmental transmission", y = "contact transmission", title = "composition")
-
-size_fig <- ref_comp %>% ggplot(.,aes(x=prop3,y=prop1)) + 
-  geom_contour_filled(aes(z=size_diff),breaks = breaks) +
-  scale_fill_viridis_d(drop=F) +
-  theme(legend.position = "none") + 
-  labs(x="environmental transmission", y = "contact transmission", title = "abundance")
-
-half_fig <- ref_comp %>% ggplot(.,aes(x=prop3,y=prop1)) + 
-  geom_contour_filled(aes(z=half_diff),breaks = breaks) +
-  scale_fill_viridis_d(drop=F) +
-  theme(legend.position = "none") + 
-  labs(x="environmental transmission", y = "contact transmission", title = "halflife")
-
-combined_fig <- ref_comp %>% ggplot(.,aes(x=prop3,y=prop1)) + 
-  geom_contour_filled(aes(z=all_diff),breaks = breaks) +
-  scale_fill_viridis_d(drop=F) +   
-  labs(x="environmental transmission", y = "contact transmission", title = "combined")
-```
-
-```{r}
-(comp_fig | size_fig)/(half_fig | combined_fig) + theme("filled contours of diff between treatment and reference")
-```
-
-```{r}
-#comparison between combined treatment and the additive effect
-
-ref_comp %<>% rowwise %>% mutate(diff_sum = comp_diff + size_diff + half_diff)
-ref_comp %<>% rowwise %>% mutate(effect = all_diff-diff_sum)
-
-# ref_comp %>% ggplot(.,aes(x=prop3,y=prop1,fill=effect)) + 
-#   geom_tile() +
-#   scale_fill_gradient2() + geom_contour(aes(z=effect))
-# 
-# ref_comp %>% ggplot(.,aes(x=prop3,y=prop1)) + 
-#   geom_contour_filled(aes(z=effect))
-
-breaks <- c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9)
-
-comb <- ref_comp %>% ggplot(.,aes(x=prop3,y=prop1)) + 
-  geom_contour_filled(aes(z=all_diff),breaks = breaks) +
-  scale_fill_viridis_d(drop=F) +   
-  theme(legend.position = "none") + 
-  labs(x="", y = "", title = "combined")
-
-sum <- ref_comp %>% ggplot(.,aes(x=prop3,y=prop1)) + 
-  geom_contour_filled(aes(z=diff_sum),breaks = breaks) +
-  scale_fill_viridis_d(drop=F) + 
-  labs(x = "", y = "contact transmission", title = "sum of its parts")
-
-diff <- ref_comp %>% ggplot(.,aes(x=prop3,y=prop1)) + 
-  geom_contour_filled(aes(z=effect),breaks = breaks) +
-  scale_fill_viridis_d(drop=F) + 
-  theme(legend.position = "none") +
-  labs(x = "environmental", y = "", title = "difference (combined - sum of its parts)")
-
-comb / sum / diff 
-```
-
-```{r}
-#comparison between combined treament and additive effect by percent change
-
-ref_percent <- ref_comp %>% mutate(., comp_per = comp_diff/ref_eigen,
-                                   size_per = size_diff/ref_eigen,
-                                   half_per = half_diff/ref_eigen,
-                                   all_per = all_eigen/ref_eigen,
-                                   sum_per = diff_sum/ref_eigen)
-
-breaks <- c(0:20)*0.1
-
-comb_per <- ref_percent %>% ggplot(.,aes(x=prop3,y=prop1)) + 
-  geom_contour_filled(aes(z=all_per),breaks = breaks) +
-  scale_fill_viridis_d(drop=F) +   
-  theme(legend.position = "none") + 
-  labs(x="", y = "", title = "combined")
-
-sum_per <- ref_percent %>% ggplot(.,aes(x=prop3,y=prop1)) + 
-  geom_contour_filled(aes(z=sum_per),breaks = breaks) +
-  scale_fill_viridis_d(drop=F) + 
-  labs(x = "", y = "contact transmission", title = "sum of its parts")
-
-diff_per <- ref_percent %>% ggplot(.,aes(x=prop3,y=prop1)) + 
-  geom_contour_filled(aes(z=all_per-sum_per),breaks = breaks) +
-  scale_fill_viridis_d(drop=F) + 
-  theme(legend.position = "none") +
-  labs(x = "environmental", y = "", title = "difference (combined - sum of its parts)")
-
-comb_per / sum_per / diff_per 
-```

### 2023-02-15 scripts/workshop/workshop.Rmd
make script for numerical solutions to ODE's

@@
 
-effect sizes
+Standard deviations from the mean
+This chunk attempts to characterize each wetland in terms of how exceptional they are compared to other wetlands in terms of abundance, cc, and temp
+The idea is that if certain wetlands tend to be highly abundant with high community competence and lower temperatures, then we might expect them to be a perfect environment for ranavirus to thrive. 
 ```{r}

### 2023-02-15 scripts/workshop/dynamics.Rmd
knit it

@@
 library(here)
+library(deSolve)
 ```
@@
 
-halflife <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = c(1/40),mort2 = 1/40,degr = 1/3.895598)
+halflife <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = c(1/45),mort2 = 1/45,degr = 1/3.895598)
 
@@
 ```
+
+```{r}
+rm0<-function(t,x,params){
+  Sa <- x[1]  
+  Ia <- x[2]
+  Ra <- x[3]
+  Sb <- x[4]  
+  Ib <- x[5]
+  Rb <- x[6]
+  V <- x[7]
+  with(as.list(params),{
+    dSa <- -beta*Sa*Ia - betab*Sa*Ib - phi*Sa*V + lambda - mu_a*Sa
+    dIa <- beta*Sa*Ia + betab*Sa*Ib + phi*Sa*V - alpha*Ia - mu_a*Ia
+    dRa <- alpha*Ia - mu_a*Ra
+
+    dSb <- -betab*Sb*Ib - beta*Sb*Ia - phi*Sb*V + lambda - mu_b*Sb
+    dIb <- betab*Sb*Ib + beta*Sb*Ia + phi*Sb*V - alpha*Ib - mu_b*Ib
+    dRb <- alpha*Ib - mu_b*Rb
+
+    dV <- sigma*Ia + sigma*Ib - epsilon*V
+    res<-c(dSa,dIa,dRa,dSb,dIb,dRb,dV)
+    list(res)
+  })}
+maxTime <- 500.0 # time is in years - run model for this time
+times<-seq(0,maxTime,by=1) # how long we run the model for
+# notes on params
+# mu =     <- density-independent mortality rate
+# beta =       <- contact transmission rate by infected A
+# betab =       <- contact transmission rate by infected B
+# phi =      <- environmental transmission rate
+# lambda =       <- density-dependent growth rate
+# alpha =     <- recovery rate
+# sigma =    <- shedding rate
+# epsilon =       <- viral degradation rate
+params<-c(mu_a=1/50,
+          mu_b=1/50,
+          beta=0.001,
+          betab=0.0001,
+          phi=0.001,
+          lambda=5/3,
+          alpha=1/10,
+          sigma=1/2,
+          epsilon=1/2)  # model parameters
+
+xstart<-c(Sa=116,
+          Ia=1,
+          Ra=0,
+          Sb=83,
+          Ib=1,
+          Rb=0,
+          V=0)  # initial conditions
+
+output<-as.data.frame(lsoda(xstart,times,rm0,params)) # tells computer to solve (integrate) equations
+
+output %>% ggplot(.,aes(x=time))+
+  geom_line(aes(y=Sa,col="Susceptible A"))+
+  geom_line(aes(y=Ia,col="Infected A"))+
+  geom_line(aes(y=Ra,col="Recovered A"))+
+  geom_line(aes(y=Sb,col="Susceptible B"))+
+  geom_line(aes(y=Ib,col="Infected B"))+
+  geom_line(aes(y=Rb,col="Recovered B"))+
+  geom_line(aes(y=V,col="Virus"))+
+  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
+  theme(legend.position = c(0.85,0.5))+
+  labs(y="N",x="Time",col="Population")+
+  ggtitle("2-Species Transmission Model for Ranavirus in an Amphibian Wetland")+theme(plot.title = element_text(hjust=0, size=10))
+```
+
+notes on params
+mu =     <- density-independent mortality rate
+beta =       <- contact transmission rate by infected A
+betab =       <- contact transmission rate by infected B
+phi =      <- environmental transmission rate
+lambda =       <- density-dependent growth rate
+alpha =     <- recovery rate
+sigma =    <- shedding rate
+epsilon =       <- viral degradation rate
+
+transmission parameter values
+species 1 is held constant at 0.0001
+second species is more competent at 0.0004
+env. trans rate is 0.00058
+
+```{r}
+spec_1 <- 0.0001
+spec_2 <- 0.00055
+env_rate <- 0.00065
+```
+
+Neutral
+```{r}
+params_n <- c(mu_a=1/50,
+          mu_b=1/50,
+          beta=spec_1,
+          betab=spec_2,
+          phi=env_rate,
+          lambda=5/3,
+          alpha=1/10,
+          sigma=1/2,
+          epsilon=1/1.947799)  # model parameters
+
+xstart_n <- c(Sa=params_n[[6]]/params_n[[1]],
+          Ia=1,
+          Ra=0,
+          Sb=params_n[[6]]/params_n[[2]],
+          Ib=1,
+          Rb=0,
+          V=0)  # initial conditions
+
+output_n <- as.data.frame(lsoda(xstart_n,times,rm0,params_n)) # tells computer to solve (integrate) equations
+```
+
+```{r}
+output_n %>% ggplot(.,aes(x=time))+
+  geom_line(aes(y=Sa,col="Susceptible A"))+
+  geom_line(aes(y=Ia,col="Infected A"))+
+  geom_line(aes(y=Ra,col="Recovered A"))+
+  geom_line(aes(y=Sb,col="Susceptible B"))+
+  geom_line(aes(y=Ib,col="Infected B"))+
+  geom_line(aes(y=Rb,col="Recovered B"))+
+  geom_line(aes(y=V,col="Virus"))+
+  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
+  theme(legend.position = c(0.85,0.5))+
+  labs(y="N",x="Time",col="Population")+
+  ggtitle("2-Species Transmission Model for Ranavirus in an Amphibian Wetland - Neutral")+theme(plot.title = element_text(hjust=0, size=10))
+```
+
+Community Composition
+```{r}
+params_c <- c(mu_a=1/60,
+          mu_b=1/30,
+          beta=spec_1,
+          betab=spec_2,
+          phi=env_rate,
+          lambda=5/3,
+          alpha=1/10,
+          sigma=1/2,
+          epsilon=1/1.947799)  # model parameters
+
+output_c <- as.data.frame(lsoda(xstart_n,times,rm0,params_c)) # tells computer to solve (integrate) equations
+
+```
+
+```{r}
+output_c %>% ggplot(.,aes(x=time))+
+  geom_line(aes(y=Sa,col="Susceptible A"))+
+  geom_line(aes(y=Ia,col="Infected A"))+
+  geom_line(aes(y=Ra,col="Recovered A"))+
+  geom_line(aes(y=Sb,col="Susceptible B"))+
+  geom_line(aes(y=Ib,col="Infected B"))+
+  geom_line(aes(y=Rb,col="Recovered B"))+
+  geom_line(aes(y=V,col="Virus"))+
+  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
+  theme(legend.position = c(0.85,0.5))+
+  labs(y="N",x="Time",col="Population")+
+  ggtitle("2-Species Transmission Model for Ranavirus in an Amphibian Wetland - Community Composition")+theme(plot.title = element_text(hjust=0, size=10))
+```
+
+Abundance
+```{r}
+params_a <- c(mu_a=1/52.5,
+          mu_b=1/52.5,
+          beta=spec_1,
+          betab=spec_2,
+          phi=env_rate,
+          lambda=5/3,
+          alpha=1/10,
+          sigma=1/2,
+          epsilon=1/1.947799)  # model parameters
+
+output_a <- as.data.frame(lsoda(xstart_n,times,rm0,params_a)) # tells computer to solve (integrate) equations
+```
+
+```{r}
+output_a %>% ggplot(.,aes(x=time))+
+  geom_line(aes(y=Sa,col="Susceptible A"))+
+  geom_line(aes(y=Ia,col="Infected A"))+
+  geom_line(aes(y=Ra,col="Recovered A"))+
+  geom_line(aes(y=Sb,col="Susceptible B"))+
+  geom_line(aes(y=Ib,col="Infected B"))+
+  geom_line(aes(y=Rb,col="Recovered B"))+
+  geom_line(aes(y=V,col="Virus"))+
+  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
+  theme(legend.position = c(0.85,0.5))+
+  labs(y="N",x="Time",col="Population")+
+  ggtitle("2-Species Transmission Model for Ranavirus in an Amphibian Wetland - Abundance")+theme(plot.title = element_text(hjust=0, size=10))
+```
+
+Half-life
+```{r}
+params_h <- c(mu_a=1/45,
+          mu_b=1/45,
+          beta=spec_1,
+          betab=spec_2,
+          phi=env_rate,
+          lambda=5/3,
+          alpha=1/10,
+          sigma=1/2,
+          epsilon=1/3.895598)  # model parameters
+
+output_h <- as.data.frame(lsoda(xstart_n,times,rm0,params_h)) # tells computer to solve (integrate) equations
+```
+
+```{r}
+output_h %>% ggplot(.,aes(x=time))+
+  geom_line(aes(y=Sa,col="Susceptible A"))+
+  geom_line(aes(y=Ia,col="Infected A"))+
+  geom_line(aes(y=Ra,col="Recovered A"))+
+  geom_line(aes(y=Sb,col="Susceptible B"))+
+  geom_line(aes(y=Ib,col="Infected B"))+
+  geom_line(aes(y=Rb,col="Recovered B"))+
+  geom_line(aes(y=V,col="Virus"))+
+  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
+  theme(legend.position = c(0.85,0.5))+
+  labs(y="N",x="Time",col="Population")+
+  ggtitle("2-Species Transmission Model for Ranavirus in an Amphibian Wetland - Half-life")+theme(plot.title = element_text(hjust=0, size=10))
+```
+
+Combined
+```{r}
+params_x <- c(mu_a=1/70,
+          mu_b=1/35,
+          beta=spec_1,
+          betab=spec_2,
+          phi=env_rate,
+          lambda=5/3,
+          alpha=1/10,
+          sigma=1/2,
+          epsilon=1/3.895598)  # model parameters
+
+output_x <- as.data.frame(lsoda(xstart_n,times,rm0,params_x)) # tells computer to solve (integrate) equations
+```
+
+```{r}
+output_x %>% ggplot(.,aes(x=time))+
+  geom_line(aes(y=Sa,col="Susceptible A"))+
+  geom_line(aes(y=Ia,col="Infected A"))+
+  geom_line(aes(y=Ra,col="Recovered A"))+
+  geom_line(aes(y=Sb,col="Susceptible B"))+
+  geom_line(aes(y=Ib,col="Infected B"))+
+  geom_line(aes(y=Rb,col="Recovered B"))+
+  geom_line(aes(y=V,col="Virus"))+
+  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
+  theme(legend.position = c(0.85,0.5))+
+  labs(y="N",x="Time",col="Population")+
+  ggtitle("2-Species Transmission Model for Ranavirus in an Amphibian Wetland - Combined")+theme(plot.title = element_text(hjust=0, size=10))
+```

### 2023-02-15 scripts/workshop/dynamics.Rmd
updated details

--- 2023-02-15 scripts/workshop/dynamics.Rmd
@@
 ```
+
+```{r}
+rm0<-function(t,x,params){
+  Sa <- x[1]  
+  Ia <- x[2]
+  Ra <- x[3]
+  Sb <- x[4]  
+  Ib <- x[5]
+  Rb <- x[6]
+  V <- x[7]
+  with(as.list(params),{
+    dSa <- -beta*Sa*Ia - betab*Sa*Ib - phi*Sa*V + lambda - mu_a*Sa
+    dIa <- beta*Sa*Ia + betab*Sa*Ib + phi*Sa*V - alpha*Ia - mu_a*Ia
+    dRa <- alpha*Ia - mu_a*Ra
+
+    dSb <- -betab*Sb*Ib - beta*Sb*Ia - phi*Sb*V + lambda - mu_b*Sb
+    dIb <- betab*Sb*Ib + beta*Sb*Ia + phi*Sb*V - alpha*Ib - mu_b*Ib
+    dRb <- alpha*Ib - mu_b*Rb
+
+    dV <- sigma*Ia + sigma*Ib - epsilon*V
+    res<-c(dSa,dIa,dRa,dSb,dIb,dRb,dV)
+    list(res)
+  })}
+maxTime <- 300.0 # time is in years - run model for this time
+times<-seq(0,maxTime,by=1) # how long we run the model for
+# notes on params
+# mu =     <- density-independent mortality rate
+# beta =       <- contact transmission rate by infected A
+# betab =       <- contact transmission rate by infected B
+# phi =      <- environmental transmission rate
+# lambda =       <- density-dependent growth rate
+# alpha =     <- recovery rate
+# sigma =    <- shedding rate
+# epsilon =       <- viral degradation rate
+params<-c(mu_a=1/50,
+          mu_b=1/50,
+          beta=0.001,
+          betab=0.0001,
+          phi=0.001,
+          lambda=5/3,
+          alpha=1/10,
+          sigma=1/2,
+          epsilon=1/2)  # model parameters
+
+xstart<-c(Sa=116,
+          Ia=1,
+          Ra=0,
+          Sb=83,
+          Ib=1,
+          Rb=0,
+          V=0)  # initial conditions
+
+output<-as.data.frame(lsoda(xstart,times,rm0,params)) # tells computer to solve (integrate) equations
+
+output %>% ggplot(.,aes(x=time))+
+  geom_line(aes(y=Sa,col="Susceptible A"))+
+  geom_line(aes(y=Ia,col="Infected A"))+
+  geom_line(aes(y=Ra,col="Recovered A"))+
+  geom_line(aes(y=Sb,col="Susceptible B"))+
+  geom_line(aes(y=Ib,col="Infected B"))+
+  geom_line(aes(y=Rb,col="Recovered B"))+
+  geom_line(aes(y=V,col="Virus"))+
+  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
+  theme(legend.position = c(0.85,0.5))+
+  labs(y="N",x="Time",col="Population")+
+  ggtitle("2-Species Transmission Model for Ranavirus in an Amphibian Wetland")+theme(plot.title = element_text(hjust=0, size=10))
+```
+
+notes on params
+mu =     <- density-independent mortality rate
+beta =       <- contact transmission rate by infected A
+betab =       <- contact transmission rate by infected B
+phi =      <- environmental transmission rate
+lambda =       <- density-dependent growth rate
+alpha =     <- recovery rate
+sigma =    <- shedding rate
+epsilon =       <- viral degradation rate
+
+transmission parameter values
+species 1 is held constant at 0.0001
+second species is more competent at 0.0004
+env. trans rate is 0.00058
+
+```{r}
+spec_1 <- 0.0001
+spec_2 <- 0.00055
+env_rate <- 0.00065
+param_num <- which(reference$eigen>1.103065 & reference$eigen<1.103192)
+```
+
+Neutral
+```{r}
+params_n <- c(mu_a=1/50,
+          mu_b=1/50,
+          beta=spec_1,
+          betab=spec_2,
+          phi=env_rate,
+          lambda=5/3,
+          alpha=1/10,
+          sigma=1/2,
+          epsilon=1/1.947799)  # model parameters
+
+xstart_n <- c(Sa=params_n[[6]]/params_n[[1]],
+          Ia=1,
+          Ra=0,
+          Sb=params_n[[6]]/params_n[[2]],
+          Ib=1,
+          Rb=0,
+          V=0)  # initial conditions
+
+output_n <- as.data.frame(lsoda(xstart_n,times,rm0,params_n)) # tells computer to solve (integrate) equations
+
+reference_eigen <- reference$eigen[param_num]
+```
+
+```{r}
+output_n %>% ggplot(.,aes(x=time))+
+  geom_line(aes(y=Sa,col="Susceptible A"))+
+  geom_line(aes(y=Ia,col="Infected A"))+
+  geom_line(aes(y=Ra,col="Recovered A"))+
+  geom_line(aes(y=Sb,col="Susceptible B"))+
+  geom_line(aes(y=Ib,col="Infected B"))+
+  geom_line(aes(y=Rb,col="Recovered B"))+
+  geom_line(aes(y=V,col="Virus"))+
+  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
+  theme(legend.position = c(0.85,0.5))+
+  labs(y="N",x="Time",col="Population")+
+  ylim(0,95)+
+  ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland - Neutral - ", reference_eigen))+theme(plot.title = element_text(hjust=0, size=10))
+```
+
+Community Composition
+```{r}
+params_c <- c(mu_a=1/60,
+          mu_b=1/30,
+          beta=spec_1,
+          betab=spec_2,
+          phi=env_rate,
+          lambda=5/3,
+          alpha=1/10,
+          sigma=1/2,
+          epsilon=1/1.947799)  # model parameters
+
+output_c <- as.data.frame(lsoda(xstart_n,times,rm0,params_c)) # tells computer to solve (integrate) equations
+
+community_eigen <- composition$eigen[param_num]
+```
+
+```{r}
+output_c %>% ggplot(.,aes(x=time))+
+  geom_line(aes(y=Sa,col="Susceptible A"))+
+  geom_line(aes(y=Ia,col="Infected A"))+
+  geom_line(aes(y=Ra,col="Recovered A"))+
+  geom_line(aes(y=Sb,col="Susceptible B"))+
+  geom_line(aes(y=Ib,col="Infected B"))+
+  geom_line(aes(y=Rb,col="Recovered B"))+
+  geom_line(aes(y=V,col="Virus"))+
+  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
+  theme(legend.position = c(0.85,0.5))+
+  labs(y="N",x="Time",col="Population")+
+  ylim(0,95)+
+  ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland - Community - ", community_eigen))+theme(plot.title = element_text(hjust=0, size=10))
+```
+
+Abundance
+```{r}
+params_a <- c(mu_a=1/52.5,
+          mu_b=1/52.5,
+          beta=spec_1,
+          betab=spec_2,
+          phi=env_rate,
+          lambda=5/3,
+          alpha=1/10,
+          sigma=1/2,
+          epsilon=1/1.947799)  # model parameters
+
+output_a <- as.data.frame(lsoda(xstart_n,times,rm0,params_a)) # tells computer to solve (integrate) equations
+
+abundance_eigen <- size$eigen[param_num]
+```
+
+```{r}
+output_a %>% ggplot(.,aes(x=time))+
+  geom_line(aes(y=Sa,col="Susceptible A"))+
+  geom_line(aes(y=Ia,col="Infected A"))+
+  geom_line(aes(y=Ra,col="Recovered A"))+
+  geom_line(aes(y=Sb,col="Susceptible B"))+
+  geom_line(aes(y=Ib,col="Infected B"))+
+  geom_line(aes(y=Rb,col="Recovered B"))+
+  geom_line(aes(y=V,col="Virus"))+
+  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
+  theme(legend.position = c(0.85,0.5))+
+  labs(y="N",x="Time",col="Population")+
+  ylim(0,95)+
+  ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland - Abundance - ", abundance_eigen))+theme(plot.title = element_text(hjust=0, size=10))
+```
+
+Half-life
+```{r}
+params_h <- c(mu_a=1/45,
+          mu_b=1/45,
+          beta=spec_1,
+          betab=spec_2,
+          phi=env_rate,
+          lambda=5/3,
+          alpha=1/10,
+          sigma=1/2,
+          epsilon=1/3.895598)  # model parameters
+
+output_h <- as.data.frame(lsoda(xstart_n,times,rm0,params_h)) # tells computer to solve (integrate) equations
+
+halflife_eigen <- halflife$eigen[param_num]
+```
+
+```{r}
+output_h %>% ggplot(.,aes(x=time))+
+  geom_line(aes(y=Sa,col="Susceptible A"))+
+  geom_line(aes(y=Ia,col="Infected A"))+
+  geom_line(aes(y=Ra,col="Recovered A"))+
+  geom_line(aes(y=Sb,col="Susceptible B"))+
+  geom_line(aes(y=Ib,col="Infected B"))+
+  geom_line(aes(y=Rb,col="Recovered B"))+
+  geom_line(aes(y=V,col="Virus"))+
+  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
+  theme(legend.position = c(0.85,0.5))+
+  labs(y="N",x="Time",col="Population")+
+  ylim(0,95)+
+  ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland - halflife - ", halflife_eigen))+theme(plot.title = element_text(hjust=0, size=10))
+```
+
+Combined
+```{r}
+params_x <- c(mu_a=1/70,
+          mu_b=1/35,
+          beta=spec_1,
+          betab=spec_2,
+          phi=env_rate,
+          lambda=5/3,
+          alpha=1/10,
+          sigma=1/2,
+          epsilon=1/3.895598)  # model parameters
+
+output_x <- as.data.frame(lsoda(xstart_n,times,rm0,params_x)) # tells computer to solve (integrate) equations
+
+combined_eigen <- combined$eigen[param_num]
+```
+
+```{r}
+output_x %>% ggplot(.,aes(x=time))+
+  geom_line(aes(y=Sa,col="Susceptible A"))+
+  geom_line(aes(y=Ia,col="Infected A"))+
+  geom_line(aes(y=Ra,col="Recovered A"))+
+  geom_line(aes(y=Sb,col="Susceptible B"))+
+  geom_line(aes(y=Ib,col="Infected B"))+
+  geom_line(aes(y=Rb,col="Recovered B"))+
+  geom_line(aes(y=V,col="Virus"))+
+  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
+  theme(legend.position = c(0.85,0.5))+
+  labs(y="N",x="Time",col="Population")+
+  ylim(0,95)+
+  ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland - Combined - ", combined_eigen))+theme(plot.title = element_text(hjust=0, size=10))
+```
+
+Make figure with infected peaks from everything
+```{r}
+output_a %<>% mutate(I = Ia+Ib)
+output_c %<>% mutate(I = Ia+Ib)
+output_h %<>% mutate(I = Ia+Ib)
+output_n %<>% mutate(I = Ia+Ib)
+output_x %<>% mutate(I = Ia+Ib)
+```
+
+```{r}
+color_x <- paste("Combined\nR0=", combined_eigen)
+color_c <- paste("Composition\nR0=", community_eigen)
+color_a <- paste("Abundance\nR0=", abundance_eigen)
+color_h <- paste("Halflife\nR0=", halflife_eigen)
+color_n <- paste("Reference\nR0=", reference_eigen)
+
+```
+
+```{r}
+ggplot()+
+  geom_line(data=output_x,mapping = aes(y=I,x=time, color=color_x))+
+  scale_color_manual(name="",values = "black")+
+  new_scale_color()+
+  geom_line(data=output_c, mapping=aes(y=I,x=time, color=color_c))+
+  scale_color_manual(name="",values = "red")+
+  new_scale_color()+
+  geom_line(output_a, mapping=aes(y=I,x=time, color=color_a))+
+  scale_color_manual(name="",values = "yellow")+
+  new_scale_color()+
+  geom_line(output_h, mapping=aes(y=I,x=time, color=color_h))+
+  scale_color_manual(name="",values = "purple")+
+  new_scale_color()+
+  geom_line(output_n, mapping=aes(y=I,x=time, color=color_n))+
+  scale_color_manual(name="",values = "gray")+
+  labs(y="N",x="Time")+
+  ylim(0,30)+
+  ggtitle(paste("Total Infected Individuals (2 Species) Over Time Across Varying Conditions"))+
+  theme_bw()
+```

### 2023-02-15 scripts/workshop/dynamics.Rmd
add figure and update ggplot

@@
   geom_line(output_a, mapping=aes(y=I,x=time, color=color_a))+
-  scale_color_manual(name="",values = "yellow")+
+  scale_color_manual(name="",values = "orange")+
   new_scale_color()+
@@
   geom_line(output_n, mapping=aes(y=I,x=time, color=color_n))+
-  scale_color_manual(name="",values = "gray")+
+  scale_color_manual(name="",values = "forestgreen")+
   labs(y="N",x="Time")+

### 2023-02-15 scripts/workshop/workshop.Rmd
add figure and update ggplot

@@
 incidence %>% ggplot(.,aes(x=Month.1, y=incidence)) + geom_col() + facet_grid(vars(WetAltID))
-incidence %>% ggplot(.,aes(x=Month.1, y=incidence, group = 1)) + geom_point() + geom_line() + facet_wrap(~WetAltID, nrow = 5, ncol = 4)
+incidence %>% ggplot(.,aes(x=Month.1, y=incidence, group = 1)) + geom_point() + geom_line() + facet_wrap(~WetAltID, nrow = 5, ncol = 4) + labs(title="Incidence over time")
 ```
@@
 comm_data <- readRDS(here("processed_data","comm_data.rds"))
-comm_data %>% ungroup() %>% ggplot(.,aes(x=Month.1, y=Prevalence, group = 1)) + geom_point() + geom_line() + facet_wrap(~WetAltID, nrow = 5, ncol = 4)
+comm_data %>% ungroup() %>% ggplot(.,aes(x=Month.1, y=Prevalence, group = 1)) + geom_point() + geom_line() + facet_wrap(~WetAltID, nrow = 5, ncol = 4) + labs(title="Prevalence over time")
+```
+
+```{r}
+comm_data %>% ungroup() %>% ggplot(.,aes(x=Month.1, y=log10(size), group = 1)) + geom_point() + geom_line() + facet_wrap(~WetAltID, nrow = 5, ncol = 4) + labs(title="log10(Abundance) over time")
 ```

### 2023-02-20 scripts/workshop/workshop.Rmd
change bin width

@@
 comm_data %>% ggplot(.,aes(x=log10(cc))) + geom_histogram() + labs(title = "log10-transformed cc")
-comm_data %>% ggplot(.,aes(x=MeanWaterTempPredC)) + geom_histogram() + labs(title = "Mean Water Temp")
+comm_data %>% ggplot(.,aes(x=MeanWaterTempPredC)) + geom_histogram(binwidth=2) + labs(title = "Mean Water Temp")
 ```

### 2023-02-20 base/setup.R
update proj theme

@@ -21,7 +21,7 @@ lapply(pckg_names,
 
 
 # setting the plot theme for the project
-proj_theme <- theme_set(theme_bw(base_size = 13))
+theme_set(theme_bw(base_size = 13)) #theme set will set plot theme without having to specify within ggplot
 
 
 ## Intended target dims

### 2023-02-21 scripts/workshop/workshop.Rmd
update titles
try 3d model
another 3d plot

@@
 
+# Incidence
+
 ```{r}
@@
 
+# Prevalence
+
 ```{r}
@@
 
+# Abundance
+
 ```{r}
@@
 ```
+
+# Histograms
 
@@
 ```
+
+# Effect Sizes
 
@@
 ```
+
+```{r}
+library(plotly)
+```
+
+```{r}
+plot_ly(data = effect_summ, x=effect_summ$temp_over, y=effect_summ$size_over, z=effect_summ$cc_over, 
+        type="scatter3d", mode="markers", color=as.factor(effect_summ$WetAltID))
+
+plot_ly(data = effect_summ, x=effect_summ$temp_sum, y=effect_summ$size_sum, z=effect_summ$cc_sum, 
+        type="scatter3d", mode="markers", color=as.factor(effect_summ$WetAltID))
+```

### 2023-02-21 scripts/workshop/workshop.Rmd
add ridgeplot

@@
 library(reshape2)
+library(ggridges)
 ```
@@
 incidence %>% ggplot(.,aes(x=Month.1, y=incidence)) + geom_col() + facet_grid(vars(WetAltID))
-incidence %>% ggplot(.,aes(x=Month.1, y=incidence, group = 1)) + geom_point() + geom_line() + facet_wrap(~WetAltID, nrow = 5, ncol = 4) + labs(title="Incidence over time")
+incidence %>% ggplot(.,aes(x=Month.1, y=incidence, group = 1)) + geom_point() + geom_line() + 
+  facet_wrap(~WetAltID, nrow = 5, ncol = 4) + labs(title="Incidence over time") + theme(axis.text.x = element_text(angle = 75, vjust = 0.5, hjust=1))
+
+data %>% filter(RV.Status==1) %>% 
+  ggplot(., aes(x=Month.1,
+                y=WetAltID,
+                group=WetAltID,
+                fill=as.factor(WetAltID))) + 
+  geom_density_ridges_gradient(show.legend=F) +
+  labs(title="# of RV cases per site over time")
+
+incidence %>% ggplot(., aes(x=Month.1,
+                            y=WetAltID,
+                            group=WetAltID,
+                            fill=as.factor(WetAltID),
+                            height = incidence)) +
+  geom_density_ridges_gradient(show.legend = F, stat = "identity")
 ```

### 2023-02-21 scripts/workshop/workshop.Rmd
add notes

@@
 ```
+
+# Up to Peak
+
+Redo analysis (effect size) with only the data from months before and including the peak.
+Rationale is that we are most interested in the features of a wetland that make it good or bad for an epidemic to occur.
+After an epidemic begins and peaks, the conditions of the wetland during the decline of the epidemic may not be as important since there are less susceptibles and parasite may be dying out.
+We'll see

### 2023-02-24 scripts/workshop/workshop.Rmd
add text
try to filter out data after epidemic
filter out months after peak prev

@@
                                     cc_sds = (mean(cc)-cc)/sd(cc),
-                                    temp_sds = ((mean(MeanWaterTempPredC)-MeanWaterTempPredC)/sd(MeanWaterTempPredC))*-1) #multiply by -1 to make it opposite because
+                                    temp_sds = ((mean(MeanWaterTempPredC)-MeanWaterTempPredC)/sd(MeanWaterTempPredC))*-1) #multiply by -1 to make it opposite
 
@@
 After an epidemic begins and peaks, the conditions of the wetland during the decline of the epidemic may not be as important since there are less susceptibles and parasite may be dying out.
-We'll see
+We'll see.
+
+```{r}
+comm_data_1 <- readRDS(here("processed_data","comm_data.rds"))
+comm_data_1$month.n <- as.numeric(gsub("Month","",comm_data_1$Month.1))
+comm_data_1 %<>% group_by(WetAltID) %>% mutate(max_mo = ifelse(Prevalence == max(Prevalence),month.n,0)) 
+comm_data_1 %<>% group_by(WetAltID) %>% mutate(max_mo=max(max_mo))
+
+comm_data_2 <- comm_data_1 %>% group_by(WetAltID) %>% filter(!month.n>max_mo)
+```
+
+```{r}
+comm_data_2 %>% mutate(mean = mean(size), sd = sd(size), var = mean(size)-size, size = size)
+comm_effect_2 <- comm_data_2 %>% filter(MeanWaterTempPredC>0) %>% mutate(size_sds = (mean(size)-size)/sd(size),
+                                    cc_sds = (mean(cc)-cc)/sd(cc),
+                                    temp_sds = ((mean(MeanWaterTempPredC)-MeanWaterTempPredC)/sd(MeanWaterTempPredC))*-1) #multiply by -1 to make it opposite
+
+effect_summ_2 <- comm_effect_2 %>% group_by(WetAltID) %>% summarize(size_over = sum(size_sds>0),
+                                                                size_sum = sum(size_sds),
+                                                                cc_over = sum(cc_sds>0),
+                                                                cc_sum = sum(cc_sds),
+                                                                temp_over = sum(temp_sds>0),
+                                                                temp_sum = sum(temp_sds))
+```
+
+```{r}
+plot_ly(data = effect_summ_2, x=effect_summ_2$temp_over, y=effect_summ_2$size_over, z=effect_summ_2$cc_over, 
+        type="scatter3d", mode="markers", color=as.factor(effect_summ_2$WetAltID))
+
+plot_ly(data = effect_summ_2, x=effect_summ_2$temp_sum, y=effect_summ_2$size_sum, z=effect_summ_2$cc_sum, 
+        type="scatter3d", mode="markers", color=as.factor(effect_summ_2$WetAltID))
+```
+
+Also, rerun dynamics without demography.

### 2023-02-28 scripts/workshop/dynamics.Rmd
knit it

@@
   })}
+#no demography
+rm1<-function(t,x,params){
+  Sa <- x[1]  
+  Ia <- x[2]
+  Ra <- x[3]
+  Sb <- x[4]  
+  Ib <- x[5]
+  Rb <- x[6]
+  V <- x[7]
+  with(as.list(params),{
+    dSa <- -beta*Sa*Ia - betab*Sa*Ib - phi*Sa*V
+    dIa <- beta*Sa*Ia + betab*Sa*Ib + phi*Sa*V - alpha*Ia
+    dRa <- alpha*Ia
+
+    dSb <- -betab*Sb*Ib - beta*Sb*Ia - phi*Sb*V
+    dIb <- betab*Sb*Ib + beta*Sb*Ia + phi*Sb*V - alpha*Ib
+    dRb <- alpha*Ib
+
+    dV <- sigma*Ia + sigma*Ib - epsilon*V
+    res<-c(dSa,dIa,dRa,dSb,dIb,dRb,dV)
+    list(res)
+  })}
+
 maxTime <- 300.0 # time is in years - run model for this time
@@
 output_n <- as.data.frame(lsoda(xstart_n,times,rm0,params_n)) # tells computer to solve (integrate) equations
+output_n1<- as.data.frame(lsoda(xstart_n,times,rm1,params_n)) # tells computer to solve (integrate) equations
 
@@
   ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland - Neutral - ", reference_eigen))+theme(plot.title = element_text(hjust=0, size=10))
+
+output_n1 %>% ggplot(.,aes(x=time))+
+  geom_line(aes(y=Sa,col="Susceptible A"))+
+  geom_line(aes(y=Ia,col="Infected A"))+
+  geom_line(aes(y=Ra,col="Recovered A"))+
+  geom_line(aes(y=Sb,col="Susceptible B"))+
+  geom_line(aes(y=Ib,col="Infected B"))+
+  geom_line(aes(y=Rb,col="Recovered B"))+
+  geom_line(aes(y=V,col="Virus"))+
+  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
+  theme(legend.position = c(0.85,0.5))+
+  labs(y="N",x="Time",col="Population")+
+  ylim(0,95)+
+  ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland w/o demo - Neutral - ", reference_eigen))+theme(plot.title = element_text(hjust=0, size=10))
 ```
@@
 output_c <- as.data.frame(lsoda(xstart_n,times,rm0,params_c)) # tells computer to solve (integrate) equations
+output_c1 <- as.data.frame(lsoda(xstart_n,times,rm1,params_c)) # tells computer to solve (integrate) equations
 
@@
   ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland - Community - ", community_eigen))+theme(plot.title = element_text(hjust=0, size=10))
+
+output_c1 %>% ggplot(.,aes(x=time))+
+  geom_line(aes(y=Sa,col="Susceptible A"))+
+  geom_line(aes(y=Ia,col="Infected A"))+
+  geom_line(aes(y=Ra,col="Recovered A"))+
+  geom_line(aes(y=Sb,col="Susceptible B"))+
+  geom_line(aes(y=Ib,col="Infected B"))+
+  geom_line(aes(y=Rb,col="Recovered B"))+
+  geom_line(aes(y=V,col="Virus"))+
+  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
+  theme(legend.position = c(0.85,0.5))+
+  labs(y="N",x="Time",col="Population")+
+  ylim(0,95)+
+  ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland w/o demo - Community - ", community_eigen))+theme(plot.title = element_text(hjust=0, size=10))
 ```
@@
 output_a <- as.data.frame(lsoda(xstart_n,times,rm0,params_a)) # tells computer to solve (integrate) equations
+output_a1 <- as.data.frame(lsoda(xstart_n,times,rm1,params_a)) # tells computer to solve (integrate) equations
 
@@
   ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland - Abundance - ", abundance_eigen))+theme(plot.title = element_text(hjust=0, size=10))
+
+output_a1 %>% ggplot(.,aes(x=time))+
+  geom_line(aes(y=Sa,col="Susceptible A"))+
+  geom_line(aes(y=Ia,col="Infected A"))+
+  geom_line(aes(y=Ra,col="Recovered A"))+
+  geom_line(aes(y=Sb,col="Susceptible B"))+
+  geom_line(aes(y=Ib,col="Infected B"))+
+  geom_line(aes(y=Rb,col="Recovered B"))+
+  geom_line(aes(y=V,col="Virus"))+
+  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
+  theme(legend.position = c(0.85,0.5))+
+  labs(y="N",x="Time",col="Population")+
+  ylim(0,95)+
+  ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland w/o demo - Abundance - ", abundance_eigen))+theme(plot.title = element_text(hjust=0, size=10))
 ```
@@
 output_h <- as.data.frame(lsoda(xstart_n,times,rm0,params_h)) # tells computer to solve (integrate) equations
+output_h1 <- as.data.frame(lsoda(xstart_n,times,rm1,params_h)) # tells computer to solve (integrate) equations
 
@@
   ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland - halflife - ", halflife_eigen))+theme(plot.title = element_text(hjust=0, size=10))
+
+output_h1 %>% ggplot(.,aes(x=time))+
+  geom_line(aes(y=Sa,col="Susceptible A"))+
+  geom_line(aes(y=Ia,col="Infected A"))+
+  geom_line(aes(y=Ra,col="Recovered A"))+
+  geom_line(aes(y=Sb,col="Susceptible B"))+
+  geom_line(aes(y=Ib,col="Infected B"))+
+  geom_line(aes(y=Rb,col="Recovered B"))+
+  geom_line(aes(y=V,col="Virus"))+
+  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
+  theme(legend.position = c(0.85,0.5))+
+  labs(y="N",x="Time",col="Population")+
+  ylim(0,95)+
+  ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland w/o demo - halflife - ", halflife_eigen))+theme(plot.title = element_text(hjust=0, size=10))
 ```
@@
 output_x <- as.data.frame(lsoda(xstart_n,times,rm0,params_x)) # tells computer to solve (integrate) equations
+output_x1 <- as.data.frame(lsoda(xstart_n,times,rm1,params_x)) # tells computer to solve (integrate) equations
 
@@
 output_x %>% ggplot(.,aes(x=time))+
+  geom_line(aes(y=Sa,col="Susceptible A"))+
+  geom_line(aes(y=Ia,col="Infected A"))+
+  geom_line(aes(y=Ra,col="Recovered A"))+
+  geom_line(aes(y=Sb,col="Susceptible B"))+
+  geom_line(aes(y=Ib,col="Infected B"))+
+  geom_line(aes(y=Rb,col="Recovered B"))+
+  geom_line(aes(y=V,col="Virus"))+
+  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
+  theme(legend.position = c(0.85,0.5))+
+  labs(y="N",x="Time",col="Population")+
+  ylim(0,95)+
+  ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland - Combined - ", combined_eigen))+theme(plot.title = element_text(hjust=0, size=10))
+
+output_x1 %>% ggplot(.,aes(x=time))+
   geom_line(aes(y=Sa,col="Susceptible A"))+
@@
 output_x %<>% mutate(I = Ia+Ib)
+
+output_a1 %<>% mutate(I = Ia+Ib)
+output_c1 %<>% mutate(I = Ia+Ib)
+output_h1 %<>% mutate(I = Ia+Ib)
+output_n1 %<>% mutate(I = Ia+Ib)
+output_x1 %<>% mutate(I = Ia+Ib)
 ```
@@
 ```
+
+```{r}
+ggplot()+
+  geom_line(data=output_x1,mapping = aes(y=I,x=time, color=color_x))+
+  scale_color_manual(name="",values = "black")+
+  new_scale_color()+
+  geom_line(data=output_c1, mapping=aes(y=I,x=time, color=color_c))+
+  scale_color_manual(name="",values = "red")+
+  new_scale_color()+
+  geom_line(output_a1, mapping=aes(y=I,x=time, color=color_a))+
+  scale_color_manual(name="",values = "orange")+
+  new_scale_color()+
+  geom_line(output_h1, mapping=aes(y=I,x=time, color=color_h))+
+  scale_color_manual(name="",values = "purple")+
+  new_scale_color()+
+  geom_line(output_n1, mapping=aes(y=I,x=time, color=color_n))+
+  scale_color_manual(name="",values = "forestgreen")+
+  labs(y="N",x="Time")+
+  ylim(0,30)+
+  ggtitle(paste("Total Infected Individuals (2 Species) Over Time Across Varying Conditions w/o demo"))+
+  theme_bw()
+```

### 2023-02-28 scripts/workshop/dynamics.Rmd
fix start params and knit

@@
 
-output_c <- as.data.frame(lsoda(xstart_n,times,rm0,params_c)) # tells computer to solve (integrate) equations
-output_c1 <- as.data.frame(lsoda(xstart_n,times,rm1,params_c)) # tells computer to solve (integrate) equations
+xstart_c <- c(Sa=params_c[[6]]/params_c[[1]],
+          Ia=1,
+          Ra=0,
+          Sb=params_c[[6]]/params_c[[2]],
+          Ib=1,
+          Rb=0,
+          V=0)  # initial conditions
+
+output_c <- as.data.frame(lsoda(xstart_c,times,rm0,params_c)) # tells computer to solve (integrate) equations
+output_c1 <- as.data.frame(lsoda(xstart_c,times,rm1,params_c)) # tells computer to solve (integrate) equations
 
@@
 
-output_a <- as.data.frame(lsoda(xstart_n,times,rm0,params_a)) # tells computer to solve (integrate) equations
-output_a1 <- as.data.frame(lsoda(xstart_n,times,rm1,params_a)) # tells computer to solve (integrate) equations
+xstart_a <- c(Sa=params_a[[6]]/params_a[[1]],
+          Ia=1,
+          Ra=0,
+          Sb=params_a[[6]]/params_a[[2]],
+          Ib=1,
+          Rb=0,
+          V=0)  # initial conditions
+
+output_a <- as.data.frame(lsoda(xstart_a,times,rm0,params_a)) # tells computer to solve (integrate) equations
+output_a1 <- as.data.frame(lsoda(xstart_a,times,rm1,params_a)) # tells computer to solve (integrate) equations
 
@@
 
-output_h <- as.data.frame(lsoda(xstart_n,times,rm0,params_h)) # tells computer to solve (integrate) equations
-output_h1 <- as.data.frame(lsoda(xstart_n,times,rm1,params_h)) # tells computer to solve (integrate) equations
+xstart_h <- c(Sa=params_h[[6]]/params_h[[1]],
+          Ia=1,
+          Ra=0,
+          Sb=params_h[[6]]/params_h[[2]],
+          Ib=1,
+          Rb=0,
+          V=0)  # initial conditions
+
+output_h <- as.data.frame(lsoda(xstart_h,times,rm0,params_h)) # tells computer to solve (integrate) equations
+output_h1 <- as.data.frame(lsoda(xstart_h,times,rm1,params_h)) # tells computer to solve (integrate) equations
 
@@
 
-output_x <- as.data.frame(lsoda(xstart_n,times,rm0,params_x)) # tells computer to solve (integrate) equations
-output_x1 <- as.data.frame(lsoda(xstart_n,times,rm1,params_x)) # tells computer to solve (integrate) equations
+xstart_x <- c(Sa=params_x[[6]]/params_x[[1]],
+          Ia=1,
+          Ra=0,
+          Sb=params_x[[6]]/params_x[[2]],
+          Ib=1,
+          Rb=0,
+          V=0)  # initial conditions
+
+output_x <- as.data.frame(lsoda(xstart_x,times,rm0,params_x)) # tells computer to solve (integrate) equations
+output_x1 <- as.data.frame(lsoda(xstart_x,times,rm1,params_x)) # tells computer to solve (integrate) equations
 
@@
   labs(y="N",x="Time")+
-  ylim(0,30)+
+  ylim(0,40)+
   ggtitle(paste("Total Infected Individuals (2 Species) Over Time Across Varying Conditions"))+
@@
   labs(y="N",x="Time")+
-  ylim(0,30)+
+  ylim(0,40)+
   ggtitle(paste("Total Infected Individuals (2 Species) Over Time Across Varying Conditions w/o demo"))+

### 2023-03-01 scripts/workshop/dynamics.Rmd
update date
add code folding
knit it

@@
 author: "Daniel Suh"
-date: "2023-02-15"
-output: html_document
+date: "`r format(Sys.time(), '%d %B, %Y')`"
+output: 
+  html_document:
+    code_folding: hide
 ---
@@
 
-```{r}
+```{r, message = F}
 library(tidyverse)

### 2023-03-01 scripts/workshop/workshop.Rmd
knit it
add code folding

@@
 date: "`r format(Sys.time(), '%d %B, %Y')`"
-output: html_document
+output: 
+  html_document:
+    code_folding: hide
 ---
@@
 
-effect_summ_2 <- comm_effect_2 %>% group_by(WetAltID) %>% summarize(size_over = sum(size_sds>0),
+effect_summ_2 <- comm_effect_2 %>% group_by(WetAltID) %>% summarize(n = n(),
+                                                                size_over = sum(size_sds>0),
                                                                 size_sum = sum(size_sds),
@@
                                                                 temp_sum = sum(temp_sds))
+effect_summ_2
 ```
@@
 Also, rerun dynamics without demography.
+
+```{r}
+
+```

### 2023-03-01 scripts/workshop/dynamics.Rmd
fix some errors and knit

@@
 ```{r}
-params_n <- c(mu_a=1/50,
-          mu_b=1/50,
+params_n <- c(mu_a=1/45,
+          mu_b=1/45,
           beta=spec_1,

### 2023-03-01 scripts/workshop/dynamics.Rmd
try more dynamics

@@
 combined <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = c(1/70),mort2 = c(1/35),degr = 1/3.895598)
+#abundance and half life
+combined_1 <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = c(1/45),mort2 = c(1/45),degr = 1/3.895598)
+#abundance and composition
+combined_2 <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = c(1/70),mort2 = c(1/35),degr = 1/1.947799)
+#composition and half life
+combined_3 <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = c(1/52.5),mort2 = c(1/52.5),degr=1/3.895598)
 
@@
 
+#abundance and halflife
+params_x1 <- c(mu_a=1/45,
+          mu_b=1/45,
+          beta=spec_1,
+          betab=spec_2,
+          phi=env_rate,
+          lambda=5/3,
+          alpha=1/10,
+          sigma=1/2,
+          epsilon=1/3.895598)  # model parameters
+
+#abundance and composition
+params_x2 <- c(mu_a=1/52.5,
+          mu_b=1/52.5,
+          beta=spec_1,
+          betab=spec_2,
+          phi=env_rate,
+          lambda=5/3,
+          alpha=1/10,
+          sigma=1/2,
+          epsilon=1/1.947799)  # model parameters
+
+#composition and halflife
+params_x3 <- c(mu_a=1/60,
+          mu_b=1/30,
+          beta=spec_1,
+          betab=spec_2,
+          phi=env_rate,
+          lambda=5/3,
+          alpha=1/10,
+          sigma=1/2,
+          epsilon=1/3.895598)  # model parameters
+
 xstart_x <- c(Sa=params_x[[6]]/params_x[[1]],
@@
 
+xstart_x1 <- c(Sa=params_x1[[6]]/params_x1[[1]],
+          Ia=1,
+          Ra=0,
+          Sb=params_x1[[6]]/params_x1[[2]],
+          Ib=1,
+          Rb=0,
+          V=0)  # initial conditions
+
+xstart_x2 <- c(Sa=params_x2[[6]]/params_x2[[1]],
+          Ia=1,
+          Ra=0,
+          Sb=params_x2[[6]]/params_x2[[2]],
+          Ib=1,
+          Rb=0,
+          V=0)  # initial conditions
+
+xstart_x3 <- c(Sa=params_x3[[6]]/params_x3[[1]],
+          Ia=1,
+          Ra=0,
+          Sb=params_x3[[6]]/params_x3[[2]],
+          Ib=1,
+          Rb=0,
+          V=0)  # initial conditions
+
 output_x <- as.data.frame(lsoda(xstart_x,times,rm0,params_x)) # tells computer to solve (integrate) equations
 output_x1 <- as.data.frame(lsoda(xstart_x,times,rm1,params_x)) # tells computer to solve (integrate) equations
+output_x2 <- as.data.frame(lsoda(xstart_x1,times,rm1,params_x1)) # tells computer to solve (integrate) equations
+output_x3 <- as.data.frame(lsoda(xstart_x2,times,rm1,params_x2)) # tells computer to solve (integrate) equations
+output_x4 <- as.data.frame(lsoda(xstart_x3,times,rm1,params_x3)) # tells computer to solve (integrate) equations
 
 combined_eigen <- combined$eigen[param_num]
+combined_eigen_1 <- combined_1$eigen[param_num]
+combined_eigen_2 <- combined_2$eigen[param_num]
+combined_eigen_3 <- combined_3$eigen[param_num]
+
 ```
@@
 output_x1 %<>% mutate(I = Ia+Ib)
+output_x2 %<>% mutate(I = Ia+Ib)
+output_x3 %<>% mutate(I = Ia+Ib)
+output_x4 %<>% mutate(I = Ia+Ib)
 ```
@@
 color_x <- paste("Combined\nR0=", combined_eigen)
+color_x2 <- paste("Abun_Half\nR0=", combined_eigen_1)
+color_x3 <- paste("Abun_Comp\nR0=", combined_eigen_2)
+color_x4 <- paste("Comp_Half\nR0=", combined_eigen_3)
+
 color_c <- paste("Composition\nR0=", community_eigen)
@@
 ```
+
+```{r}
+ggplot()+
+  geom_line(data=output_x1,mapping = aes(y=I,x=time, color=color_x))+
+  scale_color_manual(name="",values = "black")+
+  new_scale_color()+
+  geom_line(data=output_x2, mapping=aes(y=I,x=time, color=color_x2))+
+  scale_color_manual(name="",values = "red")+
+  new_scale_color()+
+  geom_line(output_x3, mapping=aes(y=I,x=time, color=color_x3))+
+  scale_color_manual(name="",values = "orange")+
+  new_scale_color()+
+  geom_line(output_x4, mapping=aes(y=I,x=time, color=color_x4))+
+  scale_color_manual(name="",values = "purple")+
+  labs(y="N",x="Time")+
+  ylim(0,40)+
+  ggtitle(paste("Total Infected Individuals (2 Species) Over Time Across Varying Conditions w/o demo"))+
+  theme_bw()
+```

### 2023-03-01 scripts/workshop/dynamics.Rmd
add reference

@@
   scale_color_manual(name="",values = "purple")+
+  new_scale_color()+
+  geom_line(output_n1, mapping=aes(y=I,x=time, color=color_n))+
+  scale_color_manual(name="",values = "forestgreen")+
   labs(y="N",x="Time")+
@@
 ```
+
+```{r}
+ggplot()+
+  geom_line(data=output_x1,mapping = aes(y=V,x=time, color=color_x))+
+  scale_color_manual(name="",values = "black")+
+  new_scale_color()+
+  geom_line(data=output_x2, mapping=aes(y=V,x=time, color=color_x2))+
+  scale_color_manual(name="",values = "red")+
+  new_scale_color()+
+  geom_line(output_x3, mapping=aes(y=V,x=time, color=color_x3))+
+  scale_color_manual(name="",values = "orange")+
+  new_scale_color()+
+  geom_line(output_x4, mapping=aes(y=V,x=time, color=color_x4))+
+  scale_color_manual(name="",values = "purple")+
+  labs(y="N",x="Time")+
+  ylim(0,75)+
+  ggtitle(paste("Total Virions Over Time Across Varying Conditions w/o demo"))+
+  theme_bw()
+```

### 2023-03-01 scripts/workshop/dynamics.Rmd
lots of plots

@@
 output_x4 %<>% mutate(I = Ia+Ib)
+
+output_a1 %<>% mutate(X = I + V)
+output_c1 %<>% mutate(X = I + V)
+output_h1 %<>% mutate(X = I + V)
+output_n1 %<>% mutate(X = I + V)
+output_x1 %<>% mutate(X = I + V)
+output_x2 %<>% mutate(X = I + V)
+output_x3 %<>% mutate(X = I + V)
+output_x4 %<>% mutate(X = I + V)
 ```
@@
 ggplot()+
+  geom_line(data=output_x1,mapping = aes(y=V,x=time, color=color_x))+
+  scale_color_manual(name="",values = "black")+
+  new_scale_color()+
+  geom_line(data=output_c1, mapping=aes(y=V,x=time, color=color_c))+
+  scale_color_manual(name="",values = "red")+
+  new_scale_color()+
+  geom_line(output_a1, mapping=aes(y=V,x=time, color=color_a))+
+  scale_color_manual(name="",values = "orange")+
+  new_scale_color()+
+  geom_line(output_h1, mapping=aes(y=V,x=time, color=color_h))+
+  scale_color_manual(name="",values = "purple")+
+  new_scale_color()+
+  geom_line(output_n1, mapping=aes(y=V,x=time, color=color_n))+
+  scale_color_manual(name="",values = "forestgreen")+
+  labs(y="N",x="Time")+
+  ylim(0,75)+
+  ggtitle(paste("Total Virions Over Time Across Varying Conditions w/o demo"))+
+  theme_bw()
+```
+
+```{r}
+ggplot()+
+  geom_line(data=output_x1,mapping = aes(y=X,x=time, color=color_x))+
+  scale_color_manual(name="",values = "black")+
+  new_scale_color()+
+  geom_line(data=output_c1, mapping=aes(y=X,x=time, color=color_c))+
+  scale_color_manual(name="",values = "red")+
+  new_scale_color()+
+  geom_line(output_a1, mapping=aes(y=X,x=time, color=color_a))+
+  scale_color_manual(name="",values = "orange")+
+  new_scale_color()+
+  geom_line(output_h1, mapping=aes(y=X,x=time, color=color_h))+
+  scale_color_manual(name="",values = "purple")+
+  new_scale_color()+
+  geom_line(output_n1, mapping=aes(y=X,x=time, color=color_n))+
+  scale_color_manual(name="",values = "forestgreen")+
+  labs(y="N",x="Time")+
+  ylim(0,110)+
+  ggtitle(paste("Total Infective Classes Over Time Across Varying Conditions w/o demo"))+
+  theme_bw()
+```
+
+```{r}
+ggplot()+
   geom_line(data=output_x1,mapping = aes(y=I,x=time, color=color_x))+
@@
 ```
+```{r}
+ggplot()+
+  geom_line(data=output_x1,mapping = aes(y=X,x=time, color=color_x))+
+  scale_color_manual(name="",values = "black")+
+  new_scale_color()+
+  geom_line(data=output_x2, mapping=aes(y=X,x=time, color=color_x2))+
+  scale_color_manual(name="",values = "red")+
+  new_scale_color()+
+  geom_line(output_x3, mapping=aes(y=X,x=time, color=color_x3))+
+  scale_color_manual(name="",values = "orange")+
+  new_scale_color()+
+  geom_line(output_x4, mapping=aes(y=X,x=time, color=color_x4))+
+  scale_color_manual(name="",values = "purple")+
+  new_scale_color()+
+  geom_line(output_n1, mapping=aes(y=X,x=time, color=color_n))+
+  scale_color_manual(name="",values = "forestgreen")+
+  labs(y="N",x="Time")+
+  ylim(0,110)+
+  ggtitle(paste("Total Infective Classes Over Time Across Varying Conditions w/o demo"))+
+  theme_bw()
+```

### 2023-03-14 scripts/workshop/workshop.Rmd
fix some errors and knit
host abundances over time

@@
 
-comm_data_2 <- comm_data_1 %>% group_by(WetAltID) %>% filter(!month.n>max_mo)
+comm_data_2 <- comm_data_1 %>% group_by(WetAltID) %>% filter(!month.n>max_mo) %>% ungroup()
 ```
@@
 ```
+
+# Phenology
+
+Let's look at amphibian phenology from the data
+ - when do amphibian larvae first appear
+ - when are amphibian larvae no longer found (implies emergence)
+ - abundance (average clutch size?)
+
+I want each row to be one month in one wetland and have columns for each species abundance
+```{r}
+ph <- data %>% select(-c(X1, ID, Species, RV.Status, HostDensity, MinWaterTemp, MaxWaterTemp, MeanWaterT, CanopyCover, Stage, SQMean, LogSQMean, ArcSQMean, LogSQMean.1, Date.Collected, SVL, Zscore, Species.4:Species.42, log10_SQ, ln_SQ, log1p_SQ)) %>% filter(Stage2=="L") %>% distinct()
+```
+
+condensing this down was more difficult than expected
+some things to ask Stacey about
+ - some of this might be explained by measurements for one sampling period being taken on separate days
+ - discrepancies exist for a single site_month
+ - - 19_month4: CanopyCover
+ - - 7_month6: HostDensity
+ - lots of sites had the same MeanWaterTempPredC but different MeanWaterT
+ - - these sites also had different MinWaterTemp and MaxWaterTemp
+
+Also I need to reconsider stage with other analyses. There are about 90 adults. I am just going to filter them out for now and assume the values we have are for larvae only?
+ - need to check on this. Do data include adults as well?
+
+Data are now ready. ph only has one of every site-month combo
+```{r}
+ph_ab <- ph %>% pivot_longer(.,cols = c(AB2:AB42))
+ph_ab %>% ggplot(.,aes(x=Month.1, y=value, color=name, group = name)) + geom_point() + geom_line() + facet_wrap(vars(WetAltID))
+```
+
+```{r}
+ph_ra <- ph %>% pivot_longer(.,cols = c(RA2:RA42))
+ph_ra %>% ggplot(.,aes(x=Month.1, y=value, color=name, group=name)) + geom_point() + geom_line() + facet_wrap(vars(WetAltID), ncol=1)
+```

### 2023-03-14 scripts/workshop/workshop.Rmd
looking at hosts over time

@@
 ```{r}
-ph_ab <- ph %>% pivot_longer(.,cols = c(AB2:AB42))
+ph_ab <- ph %>% select(-c(ABAmb,ABSal)) %>% pivot_longer(.,cols = c(AB2:AB42))
 ph_ab %>% ggplot(.,aes(x=Month.1, y=value, color=name, group = name)) + geom_point() + geom_line() + facet_wrap(vars(WetAltID))
@@
 ph_ra <- ph %>% pivot_longer(.,cols = c(RA2:RA42))
-ph_ra %>% ggplot(.,aes(x=Month.1, y=value, color=name, group=name)) + geom_point() + geom_line() + facet_wrap(vars(WetAltID), ncol=1)
+ph_ra %>% ggplot(.,aes(x=Month.1, y=value, color=name, group=name)) + geom_point() + geom_line() + facet_wrap(vars(WetAltID))
 ```
+
+```{r}
+ph_pa <- ph %>% pivot_longer(.,cols = c(PA2:PA42))
+ph_pa %>% ggplot(.,aes(x=Month.1, y=value, color=name, fill=name, group=name)) + geom_col() + facet_wrap(vars(WetAltID))
+```

### 2023-03-14 scripts/workshop/workshop.Rmd
more host stuff

@@
 
+vl %<>% mutate(ln_of_mean = log(mean))
+
 supp1_raw <- vl %>% arrange(mean) %>% 
@@
 
+supp1_ln <- vl %>% arrange(ln_of_mean) %>% 
+  mutate(abb_name = factor(abb_name, levels = abb_name)) %>% 
+  ggplot(.,aes(x=abb_name,y=ln_of_mean)) +
+  geom_point() + 
+  xlab("Species") + ylab("ln(Viral Load)") + 
+  geom_hline(yintercept = 1) + 
+  labs(title = "Viral Loads") + 
+  theme_classic() + 
+  theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
+  scale_y_continuous(breaks=seq(0,6,1))
+
 supp1_raw
 supp1_log10
+supp1_ln
 ```
@@
 ```
+
+First and last sighting of each at each wetland
+```{r}
+ph_range <- ph_pa %>% filter(value==1) %>% group_by(name, WetAltID) %>% summarize(min_mo=min(Month.1), max_mo = max(Month.1))
+ph_range %<>% mutate(first = as.numeric(gsub("Month","",min_mo)),
+                     last = as.numeric(gsub("Month","",max_mo)))
+ph_range_summ <- ph_range %>% group_by(name) %>% summarize(mean_first = mean(first),
+                                                           mean_last = mean(last))
+
+ph_range_summ %<>% mutate(Species = as.numeric(gsub("PA","",name)))
+
+ph_range_summ %<>% left_join(.,vl)
+
+ph_range_summ %>% arrange(mean) %>% mutate(name=factor(name, levels=name)) %>% filter(!is.na(species_name)) %>% 
+  ggplot(.,aes(y = name, color=ln_of_mean)) +
+  geom_linerange(aes(xmin=mean_first,xmax=mean_last,linewidth=1)) + 
+  scale_color_viridis_c()
+```
+
+Do this instead with abundance
+```{r}
+ph_ab_summ <- ph_ab %>% mutate(log1p_value = log1p(value)) %>% group_by(name,Month.1) %>% summarize(log1p_ab_mean = mean(log1p_value),
+                                                                                                    ab_mean = mean(value)) %>% ungroup()
+
+ph_ab_summ %<>% mutate(Species = as.numeric(gsub("AB","",name)))
+
+ph_ab_summ %<>% left_join(.,vl)
+
+ph_ab_summ %>% arrange(mean) %>% mutate(name=factor(name, levels=unique(name))) %>% filter(!is.na(species_name)) %>% 
+              ggplot(., aes(x=Month.1,
+                            y=name,
+                            group=name,
+                            fill=as.factor(mean),
+                            height = ab_mean)) +
+  geom_density_ridges_gradient(show.legend = T, stat = "identity") + labs(x = "Absolute Abundance")
+
+ph_ab_summ %>% arrange(mean) %>% mutate(name=factor(name, levels=unique(name))) %>% filter(!is.na(species_name)) %>% 
+              ggplot(., aes(x=Month.1,
+                            y=name,
+                            group=name,
+                            fill=as.factor(mean),
+                            height = log1p_ab_mean)) +
+  geom_density_ridges_gradient(show.legend = T, stat = "identity") + labs(x = "log1p(Abundance)")
+
+ph_ab_summ %>% arrange(desc(mean)) %>% mutate(name=factor(name, levels=unique(name))) %>% filter(!is.na(species_name)) %>% 
+              ggplot(., aes(x=Month.1,
+                            y=log1p_ab_mean,
+                            fill=as.factor(ln_of_mean))) +
+  geom_col() + labs(x = "log1p(Abundance)") + facet_wrap(vars(name), ncol=1) + theme(strip.background = element_blank(),
+                                                                                     strip.text.x = element_blank())
+```

### 2023-03-15 scripts/workshop/workshop.Rmd
knit it

@@
 
-vl %<>% mutate(ln_of_mean = log(mean))
+vl %<>% mutate(ln_of_mean = log1p(mean))
 
@@
 
+ph_range_summ %<>% rowwise() %>% mutate(median_mo = median(c(mean_first, mean_last)))
+
 ph_range_summ %>% arrange(mean) %>% mutate(name=factor(name, levels=name)) %>% filter(!is.na(species_name)) %>% 
-  ggplot(.,aes(y = name, color=ln_of_mean)) +
+  ggplot(.,aes(y = name, color=ln_of_mean)) + geom_point(aes(x=median_mo, size = 2)) + 
   geom_linerange(aes(xmin=mean_first,xmax=mean_last,linewidth=1)) + 
-  scale_color_viridis_c()
+  scale_color_viridis_c() + guides(size=F, linewidth=F)
+
+ph_range_summ %>% arrange(mean) %>% mutate(abb_name=factor(abb_name, levels=abb_name)) %>% filter(!is.na(species_name)) %>% 
+  ggplot(.,aes(y = abb_name, color=ln_of_mean)) + geom_point(aes(x=median_mo, size = 2)) + 
+  geom_linerange(aes(xmin=mean_first,xmax=mean_last,linewidth=1)) + 
+  scale_color_continuous() + 
+  guides(size=F, linewidth=F, color=guide_legend(title="log1p(viral load)")) + 
+  theme(axis.text.y = element_text(face = "italic", angle = 15)) + 
+  labs(x="Months", y="Species", title = "Average Range of Species Presence")
+
 ```

### 2023-03-29 area_UC.Rmd
Create area_UC.Rmd

---
title: "Area Under Curve"
author: "Daniel Suh"
date: "`r Sys.Date()`"
output: html_document
---

```{r, message = F}
library(here)

source(here("base","src.R"))

library(pracma)
```

Look at area under curve for all variables for each community (i.e. site-month)
```{r}
comm_data <- readRDS(here("processed_data","comm_data.rds"))
```

```{r}
comm_data %<>% mutate(month_n = as.numeric(gsub("Month","",Month.1)),
                      WetAltID = as.factor(WetAltID))
areas <- comm_data %>% group_by(WetAltID) %>% summarize(abund = trapz(month_n, size),
                                                        temp = trapz(month_n, MeanWaterTempPredC),
                                                        cc = trapz(month_n, cc),
                                                        prev = trapz(month_n, Prevalence))
```

```{r}
areas %>% ggplot(.,aes(x=abund,y=prev)) + geom_point() + geom_smooth(method = "lm")
areas %>% ggplot(.,aes(x=cc,y=prev)) + geom_point() + geom_smooth(method = "lm")
areas %>% ggplot(.,aes(x=temp,y=prev)) + geom_point() + geom_smooth(method = "lm")
```

```{r}
abund_lm <- lm(prev ~ log(abund), areas)
cc_lm <- lm(prev ~ cc, areas)
temp_lm <- lm(prev ~ temp, areas)
```

### 2023-03-29 area_UC.Rmd
Update area_UC.Rmd

@@
 library(pracma)
+library(ggridges)
 ```
@@
                       WetAltID = as.factor(WetAltID))
-areas <- comm_data %>% group_by(WetAltID) %>% summarize(abund = trapz(month_n, size),
+areas <- comm_data %>% filter(!(WetAltID %in% c(2,18,19,20))) %>% 
+  group_by(WetAltID) %>% summarize(abund = trapz(month_n, size), 
                                    temp = trapz(month_n, MeanWaterTempPredC),
@@
 ```{r}
-areas %>% ggplot(.,aes(x=abund,y=prev)) + geom_point() + geom_smooth(method = "lm")
+comm_data %>% ggplot(., aes(x=Month.1,
+                            y=WetAltID,
+                            group=WetAltID,
+                            fill=as.factor(WetAltID),
+                            height = size)) +
+  geom_density_ridges_gradient(show.legend = F, stat = "identity") + 
+  labs(title = "Abundance over time")
+
+comm_data %>% ggplot(., aes(x=Month.1,
+                            y=WetAltID,
+                            group=WetAltID,
+                            fill=as.factor(WetAltID),
+                            height = cc)) +
+  geom_density_ridges_gradient(show.legend = F, stat = "identity") + 
+  labs(title = "CC over time")
+
+comm_data %>% ggplot(., aes(x=Month.1,
+                            y=WetAltID,
+                            group=WetAltID,
+                            fill=as.factor(WetAltID),
+                            height = MeanWaterTempPredC)) +
+  geom_density_ridges_gradient(show.legend = F, stat = "identity") + 
+  labs(title = "Temp over time")
+
+comm_data %>% ggplot(., aes(x=Month.1,
+                            y=WetAltID,
+                            group=WetAltID,
+                            fill=as.factor(WetAltID),
+                            height = Prevalence)) +
+  geom_density_ridges_gradient(show.legend = F, stat = "identity") + 
+  labs(title = "Prevalence over time")
+```
+
+```{r}
+areas %>% ggplot(.,aes(x=log(abund),y=prev)) + geom_point() + geom_smooth(method = "lm")
 areas %>% ggplot(.,aes(x=cc,y=prev)) + geom_point() + geom_smooth(method = "lm")
@@
 temp_lm <- lm(prev ~ temp, areas)
+summary(abund_lm)
+summary(cc_lm)
+summary(temp_lm)
 ```

### 2023-06-02 scripts/workshop/sensitivity.Rmd
Create sensitivity.Rmd

--- 2023-03-01 scripts/workshop/dynamics.Rmd
@@
 ---
-title: "dynamics"
+title: "sensitivity"
 author: "Daniel Suh"
-date: "`r format(Sys.time(), '%d %B, %Y')`"
-output: 
-  html_document:
-    code_folding: hide
+date: "`r Sys.Date()`"
+output: html_document
 ---
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
 
@@
 ```
-
-```{r}
-#define function contour(): inputs are parameters and output is values for grid search for R0
-contour <- function(trans_1_min, trans_1_max, trans_3_min, trans_3_max, mort1, mort2, degr) {
-  #expand data
-  data <- expand_grid("trans1" = seq(from = trans_1_min, to = trans_1_max, by = 0.00001),
-                      "trans2" = c(0.0001),
-                      "trans3" = seq(from = trans_3_min, to = trans_3_max, by = 0.00001),
-                      "shed1" = c(rep(1/2, length("trans1"))),
-                      "shed2" = c(rep(1/2, length("trans1"))),
-                      "reco1" = c(rep(1/10, length("trans1"))),
-                      "reco2" = c(rep(1/10, length("trans1"))),
-                      "mort1" = c(mort1), 
-                      "mort2" = c(mort2), 
-                      "degr" = c(degr), 
-                      "birth1" = c(rep(5/3, length("trans1"))),
-                      "birth2" = c(rep(5/3, length("trans1"))))
-
-  #calculate disease-free equilibria for use in next-gen matrix
-  data %<>% add_column(dfe1 = (data$birth1/data$mort1),dfe2 = (data$birth2/data$mort2), eigen = NA) 
-  #make matrix
-  mat <- matrix(1, nrow = 3, ncol = 3)
-
-  m <- nrow(data)
-
-  #calculate eigenvalue using next-gen matrix. max eigenvalue is R0
-  for(n in 1:m){
-    mat[1] = (data$trans1[n]*data$dfe1[n])/(data$reco1[n]+data$mort1[n])
-    mat[2] = (data$trans1[n]*data$dfe2[n])/(data$reco2[n]+data$mort2[n])
-    mat[3] = (data$shed1[n]/data$degr[n])
-    mat[4] = (data$trans2[n]*data$dfe1[n])/(data$reco1[n]+data$mort1[n])
-    mat[5] = (data$trans2[n]*data$dfe2[n])/(data$reco2[n]+data$mort2[n])
-    mat[6] = (data$shed2[n]/data$degr[n])
-    mat[7] = (data$trans3[n]*data$dfe1[n])/(data$reco1[n]+data$mort1[n])
-    mat[8] = (data$trans3[n]*data$dfe2[n])/(data$reco2[n]+data$mort2[n])
-    mat[9] = 0
-    eigen <- max(eigen(mat)$values)
-    data$eigen[n] <- eigen
-  }
-
-  data$span1 <- 1/data$mort1 #lifespan for species 1
-  data$span2 <- 1/data$mort2 #lifespan for species 2
-  data$half <- log(1/2)/-data$degr #viral half-life
-  data$epid <- data$eigen > 1 #invasion potential (i.e. R0 > 1)
-  data$prop1 <- data$trans1/data$trans2 #proportion of transmission of species 2 over species 1
-  data$prop3 <- data$trans3/data$trans2 #proportion of environmental transmission over species 1
-  data$prop_mort <- data$mort1/data$mort2 #relative mortality for species 1 and 2
-  data$cap1 <- data$birth1/data$mort1 #carrying capacity for species 1
-  data$cap2 <- data$birth2/data$mort2 #carrying capacity for species 2
-  data$cap_prop <- data$cap1/data$cap2 #relative carrying capacities for species 1 and 2
-  data$tot <- data$cap1 + data$cap2 #total abundance for species 1 and 2
-  data$cc <- data$trans1*(data$cap1/data$tot) + data$trans2*(data$cap2/data$tot) #cc defined as transmission of a species multiplied by its relative abundance
-  output <- data %>% dplyr::select(trans1, trans2, prop1, prop3, trans3, eigen, prop_mort, mort1, mort2, degr, half, cap1, cap2, cap_prop, tot, cc) %>% distinct()
-  return(output)
-}
-
-reference <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = 1/45,mort2 = 1/45,degr = 1/1.947799)
-
-composition <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = 1/60,mort2 = 1/30,degr = 1/1.947799)
-
-size <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = c(1/52.5),mort2 = c(1/52.5),degr = 1/1.947799)
-
-halflife <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = c(1/45),mort2 = 1/45,degr = 1/3.895598)
-
-combined <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = c(1/70),mort2 = c(1/35),degr = 1/3.895598)
-#abundance and half life
-combined_1 <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = c(1/45),mort2 = c(1/45),degr = 1/3.895598)
-#abundance and composition
-combined_2 <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = c(1/70),mort2 = c(1/35),degr = 1/1.947799)
-#composition and half life
-combined_3 <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = c(1/52.5),mort2 = c(1/52.5),degr=1/3.895598)
-
-```
-
-```{r}
-rm0<-function(t,x,params){
-  Sa <- x[1]  
-  Ia <- x[2]
-  Ra <- x[3]
-  Sb <- x[4]  
-  Ib <- x[5]
-  Rb <- x[6]
-  V <- x[7]
-  with(as.list(params),{
-    dSa <- -beta*Sa*Ia - betab*Sa*Ib - phi*Sa*V + lambda - mu_a*Sa
-    dIa <- beta*Sa*Ia + betab*Sa*Ib + phi*Sa*V - alpha*Ia - mu_a*Ia
-    dRa <- alpha*Ia - mu_a*Ra
-
-    dSb <- -betab*Sb*Ib - beta*Sb*Ia - phi*Sb*V + lambda - mu_b*Sb
-    dIb <- betab*Sb*Ib + beta*Sb*Ia + phi*Sb*V - alpha*Ib - mu_b*Ib
-    dRb <- alpha*Ib - mu_b*Rb
-
-    dV <- sigma*Ia + sigma*Ib - epsilon*V
-    res<-c(dSa,dIa,dRa,dSb,dIb,dRb,dV)
-    list(res)
-  })}
-#no demography
-rm1<-function(t,x,params){
-  Sa <- x[1]  
-  Ia <- x[2]
-  Ra <- x[3]
-  Sb <- x[4]  
-  Ib <- x[5]
-  Rb <- x[6]
-  V <- x[7]
-  with(as.list(params),{
-    dSa <- -beta*Sa*Ia - betab*Sa*Ib - phi*Sa*V
-    dIa <- beta*Sa*Ia + betab*Sa*Ib + phi*Sa*V - alpha*Ia
-    dRa <- alpha*Ia
-
-    dSb <- -betab*Sb*Ib - beta*Sb*Ia - phi*Sb*V
-    dIb <- betab*Sb*Ib + beta*Sb*Ia + phi*Sb*V - alpha*Ib
-    dRb <- alpha*Ib
-
-    dV <- sigma*Ia + sigma*Ib - epsilon*V
-    res<-c(dSa,dIa,dRa,dSb,dIb,dRb,dV)
-    list(res)
-  })}
-
-maxTime <- 300.0 # time is in years - run model for this time
-times<-seq(0,maxTime,by=1) # how long we run the model for
-# notes on params
-# mu =     <- density-independent mortality rate
-# beta =       <- contact transmission rate by infected A
-# betab =       <- contact transmission rate by infected B
-# phi =      <- environmental transmission rate
-# lambda =       <- density-dependent growth rate
-# alpha =     <- recovery rate
-# sigma =    <- shedding rate
-# epsilon =       <- viral degradation rate
-params<-c(mu_a=1/50,
-          mu_b=1/50,
-          beta=0.001,
-          betab=0.0001,
-          phi=0.001,
-          lambda=5/3,
-          alpha=1/10,
-          sigma=1/2,
-          epsilon=1/2)  # model parameters
-
-xstart<-c(Sa=116,
-          Ia=1,
-          Ra=0,
-          Sb=83,
-          Ib=1,
-          Rb=0,
-          V=0)  # initial conditions
-
-output<-as.data.frame(lsoda(xstart,times,rm0,params)) # tells computer to solve (integrate) equations
-
-output %>% ggplot(.,aes(x=time))+
-  geom_line(aes(y=Sa,col="Susceptible A"))+
-  geom_line(aes(y=Ia,col="Infected A"))+
-  geom_line(aes(y=Ra,col="Recovered A"))+
-  geom_line(aes(y=Sb,col="Susceptible B"))+
-  geom_line(aes(y=Ib,col="Infected B"))+
-  geom_line(aes(y=Rb,col="Recovered B"))+
-  geom_line(aes(y=V,col="Virus"))+
-  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
-  theme(legend.position = c(0.85,0.5))+
-  labs(y="N",x="Time",col="Population")+
-  ggtitle("2-Species Transmission Model for Ranavirus in an Amphibian Wetland")+theme(plot.title = element_text(hjust=0, size=10))
-```
-
-notes on params
-mu =     <- density-independent mortality rate
-beta =       <- contact transmission rate by infected A
-betab =       <- contact transmission rate by infected B
-phi =      <- environmental transmission rate
-lambda =       <- density-dependent growth rate
-alpha =     <- recovery rate
-sigma =    <- shedding rate
-epsilon =       <- viral degradation rate
-
-transmission parameter values
-species 1 is held constant at 0.0001
-second species is more competent at 0.0004
-env. trans rate is 0.00058
-
-```{r}
-spec_1 <- 0.0001
-spec_2 <- 0.00055
-env_rate <- 0.00065
-param_num <- which(reference$eigen>1.103065 & reference$eigen<1.103192)
-```
-
-Neutral
-```{r}
-params_n <- c(mu_a=1/45,
-          mu_b=1/45,
-          beta=spec_1,
-          betab=spec_2,
-          phi=env_rate,
-          lambda=5/3,
-          alpha=1/10,
-          sigma=1/2,
-          epsilon=1/1.947799)  # model parameters
-
-xstart_n <- c(Sa=params_n[[6]]/params_n[[1]],
-          Ia=1,
-          Ra=0,
-          Sb=params_n[[6]]/params_n[[2]],
-          Ib=1,
-          Rb=0,
-          V=0)  # initial conditions
-
-output_n <- as.data.frame(lsoda(xstart_n,times,rm0,params_n)) # tells computer to solve (integrate) equations
-output_n1<- as.data.frame(lsoda(xstart_n,times,rm1,params_n)) # tells computer to solve (integrate) equations
-
-reference_eigen <- reference$eigen[param_num]
-```
-
-```{r}
-output_n %>% ggplot(.,aes(x=time))+
-  geom_line(aes(y=Sa,col="Susceptible A"))+
-  geom_line(aes(y=Ia,col="Infected A"))+
-  geom_line(aes(y=Ra,col="Recovered A"))+
-  geom_line(aes(y=Sb,col="Susceptible B"))+
-  geom_line(aes(y=Ib,col="Infected B"))+
-  geom_line(aes(y=Rb,col="Recovered B"))+
-  geom_line(aes(y=V,col="Virus"))+
-  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
-  theme(legend.position = c(0.85,0.5))+
-  labs(y="N",x="Time",col="Population")+
-  ylim(0,95)+
-  ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland - Neutral - ", reference_eigen))+theme(plot.title = element_text(hjust=0, size=10))
-
-output_n1 %>% ggplot(.,aes(x=time))+
-  geom_line(aes(y=Sa,col="Susceptible A"))+
-  geom_line(aes(y=Ia,col="Infected A"))+
-  geom_line(aes(y=Ra,col="Recovered A"))+
-  geom_line(aes(y=Sb,col="Susceptible B"))+
-  geom_line(aes(y=Ib,col="Infected B"))+
-  geom_line(aes(y=Rb,col="Recovered B"))+
-  geom_line(aes(y=V,col="Virus"))+
-  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
-  theme(legend.position = c(0.85,0.5))+
-  labs(y="N",x="Time",col="Population")+
-  ylim(0,95)+
-  ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland w/o demo - Neutral - ", reference_eigen))+theme(plot.title = element_text(hjust=0, size=10))
-```
-
-Community Composition
-```{r}
-params_c <- c(mu_a=1/60,
-          mu_b=1/30,
-          beta=spec_1,
-          betab=spec_2,
-          phi=env_rate,
-          lambda=5/3,
-          alpha=1/10,
-          sigma=1/2,
-          epsilon=1/1.947799)  # model parameters
-
-xstart_c <- c(Sa=params_c[[6]]/params_c[[1]],
-          Ia=1,
-          Ra=0,
-          Sb=params_c[[6]]/params_c[[2]],
-          Ib=1,
-          Rb=0,
-          V=0)  # initial conditions
-
-output_c <- as.data.frame(lsoda(xstart_c,times,rm0,params_c)) # tells computer to solve (integrate) equations
-output_c1 <- as.data.frame(lsoda(xstart_c,times,rm1,params_c)) # tells computer to solve (integrate) equations
-
-community_eigen <- composition$eigen[param_num]
-```
-
-```{r}
-output_c %>% ggplot(.,aes(x=time))+
-  geom_line(aes(y=Sa,col="Susceptible A"))+
-  geom_line(aes(y=Ia,col="Infected A"))+
-  geom_line(aes(y=Ra,col="Recovered A"))+
-  geom_line(aes(y=Sb,col="Susceptible B"))+
-  geom_line(aes(y=Ib,col="Infected B"))+
-  geom_line(aes(y=Rb,col="Recovered B"))+
-  geom_line(aes(y=V,col="Virus"))+
-  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
-  theme(legend.position = c(0.85,0.5))+
-  labs(y="N",x="Time",col="Population")+
-  ylim(0,95)+
-  ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland - Community - ", community_eigen))+theme(plot.title = element_text(hjust=0, size=10))
-
-output_c1 %>% ggplot(.,aes(x=time))+
-  geom_line(aes(y=Sa,col="Susceptible A"))+
-  geom_line(aes(y=Ia,col="Infected A"))+
-  geom_line(aes(y=Ra,col="Recovered A"))+
-  geom_line(aes(y=Sb,col="Susceptible B"))+
-  geom_line(aes(y=Ib,col="Infected B"))+
-  geom_line(aes(y=Rb,col="Recovered B"))+
-  geom_line(aes(y=V,col="Virus"))+
-  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
-  theme(legend.position = c(0.85,0.5))+
-  labs(y="N",x="Time",col="Population")+
-  ylim(0,95)+
-  ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland w/o demo - Community - ", community_eigen))+theme(plot.title = element_text(hjust=0, size=10))
-```
-
-Abundance
-```{r}
-params_a <- c(mu_a=1/52.5,
-          mu_b=1/52.5,
-          beta=spec_1,
-          betab=spec_2,
-          phi=env_rate,
-          lambda=5/3,
-          alpha=1/10,
-          sigma=1/2,
-          epsilon=1/1.947799)  # model parameters
-
-xstart_a <- c(Sa=params_a[[6]]/params_a[[1]],
-          Ia=1,
-          Ra=0,
-          Sb=params_a[[6]]/params_a[[2]],
-          Ib=1,
-          Rb=0,
-          V=0)  # initial conditions
-
-output_a <- as.data.frame(lsoda(xstart_a,times,rm0,params_a)) # tells computer to solve (integrate) equations
-output_a1 <- as.data.frame(lsoda(xstart_a,times,rm1,params_a)) # tells computer to solve (integrate) equations
-
-abundance_eigen <- size$eigen[param_num]
-```
-
-```{r}
-output_a %>% ggplot(.,aes(x=time))+
-  geom_line(aes(y=Sa,col="Susceptible A"))+
-  geom_line(aes(y=Ia,col="Infected A"))+
-  geom_line(aes(y=Ra,col="Recovered A"))+
-  geom_line(aes(y=Sb,col="Susceptible B"))+
-  geom_line(aes(y=Ib,col="Infected B"))+
-  geom_line(aes(y=Rb,col="Recovered B"))+
-  geom_line(aes(y=V,col="Virus"))+
-  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
-  theme(legend.position = c(0.85,0.5))+
-  labs(y="N",x="Time",col="Population")+
-  ylim(0,95)+
-  ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland - Abundance - ", abundance_eigen))+theme(plot.title = element_text(hjust=0, size=10))
-
-output_a1 %>% ggplot(.,aes(x=time))+
-  geom_line(aes(y=Sa,col="Susceptible A"))+
-  geom_line(aes(y=Ia,col="Infected A"))+
-  geom_line(aes(y=Ra,col="Recovered A"))+
-  geom_line(aes(y=Sb,col="Susceptible B"))+
-  geom_line(aes(y=Ib,col="Infected B"))+
-  geom_line(aes(y=Rb,col="Recovered B"))+
-  geom_line(aes(y=V,col="Virus"))+
-  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
-  theme(legend.position = c(0.85,0.5))+
-  labs(y="N",x="Time",col="Population")+
-  ylim(0,95)+
-  ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland w/o demo - Abundance - ", abundance_eigen))+theme(plot.title = element_text(hjust=0, size=10))
-```
-
-Half-life
-```{r}
-params_h <- c(mu_a=1/45,
-          mu_b=1/45,
-          beta=spec_1,
-          betab=spec_2,
-          phi=env_rate,
-          lambda=5/3,
-          alpha=1/10,
-          sigma=1/2,
-          epsilon=1/3.895598)  # model parameters
-
-xstart_h <- c(Sa=params_h[[6]]/params_h[[1]],
-          Ia=1,
-          Ra=0,
-          Sb=params_h[[6]]/params_h[[2]],
-          Ib=1,
-          Rb=0,
-          V=0)  # initial conditions
-
-output_h <- as.data.frame(lsoda(xstart_h,times,rm0,params_h)) # tells computer to solve (integrate) equations
-output_h1 <- as.data.frame(lsoda(xstart_h,times,rm1,params_h)) # tells computer to solve (integrate) equations
-
-halflife_eigen <- halflife$eigen[param_num]
-```
-
-```{r}
-output_h %>% ggplot(.,aes(x=time))+
-  geom_line(aes(y=Sa,col="Susceptible A"))+
-  geom_line(aes(y=Ia,col="Infected A"))+
-  geom_line(aes(y=Ra,col="Recovered A"))+
-  geom_line(aes(y=Sb,col="Susceptible B"))+
-  geom_line(aes(y=Ib,col="Infected B"))+
-  geom_line(aes(y=Rb,col="Recovered B"))+
-  geom_line(aes(y=V,col="Virus"))+
-  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
-  theme(legend.position = c(0.85,0.5))+
-  labs(y="N",x="Time",col="Population")+
-  ylim(0,95)+
-  ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland - halflife - ", halflife_eigen))+theme(plot.title = element_text(hjust=0, size=10))
-
-output_h1 %>% ggplot(.,aes(x=time))+
-  geom_line(aes(y=Sa,col="Susceptible A"))+
-  geom_line(aes(y=Ia,col="Infected A"))+
-  geom_line(aes(y=Ra,col="Recovered A"))+
-  geom_line(aes(y=Sb,col="Susceptible B"))+
-  geom_line(aes(y=Ib,col="Infected B"))+
-  geom_line(aes(y=Rb,col="Recovered B"))+
-  geom_line(aes(y=V,col="Virus"))+
-  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
-  theme(legend.position = c(0.85,0.5))+
-  labs(y="N",x="Time",col="Population")+
-  ylim(0,95)+
-  ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland w/o demo - halflife - ", halflife_eigen))+theme(plot.title = element_text(hjust=0, size=10))
-```
-
-Combined
-```{r}
-params_x <- c(mu_a=1/70,
-          mu_b=1/35,
-          beta=spec_1,
-          betab=spec_2,
-          phi=env_rate,
-          lambda=5/3,
-          alpha=1/10,
-          sigma=1/2,
-          epsilon=1/3.895598)  # model parameters
-
-#abundance and halflife
-params_x1 <- c(mu_a=1/45,
-          mu_b=1/45,
-          beta=spec_1,
-          betab=spec_2,
-          phi=env_rate,
-          lambda=5/3,
-          alpha=1/10,
-          sigma=1/2,
-          epsilon=1/3.895598)  # model parameters
-
-#abundance and composition
-params_x2 <- c(mu_a=1/52.5,
-          mu_b=1/52.5,
-          beta=spec_1,
-          betab=spec_2,
-          phi=env_rate,
-          lambda=5/3,
-          alpha=1/10,
-          sigma=1/2,
-          epsilon=1/1.947799)  # model parameters
-
-#composition and halflife
-params_x3 <- c(mu_a=1/60,
-          mu_b=1/30,
-          beta=spec_1,
-          betab=spec_2,
-          phi=env_rate,
-          lambda=5/3,
-          alpha=1/10,
-          sigma=1/2,
-          epsilon=1/3.895598)  # model parameters
-
-xstart_x <- c(Sa=params_x[[6]]/params_x[[1]],
-          Ia=1,
-          Ra=0,
-          Sb=params_x[[6]]/params_x[[2]],
-          Ib=1,
-          Rb=0,
-          V=0)  # initial conditions
-
-xstart_x1 <- c(Sa=params_x1[[6]]/params_x1[[1]],
-          Ia=1,
-          Ra=0,
-          Sb=params_x1[[6]]/params_x1[[2]],
-          Ib=1,
-          Rb=0,
-          V=0)  # initial conditions
-
-xstart_x2 <- c(Sa=params_x2[[6]]/params_x2[[1]],
-          Ia=1,
-          Ra=0,
-          Sb=params_x2[[6]]/params_x2[[2]],
-          Ib=1,
-          Rb=0,
-          V=0)  # initial conditions
-
-xstart_x3 <- c(Sa=params_x3[[6]]/params_x3[[1]],
-          Ia=1,
-          Ra=0,
-          Sb=params_x3[[6]]/params_x3[[2]],
-          Ib=1,
-          Rb=0,
-          V=0)  # initial conditions
-
-output_x <- as.data.frame(lsoda(xstart_x,times,rm0,params_x)) # tells computer to solve (integrate) equations
-output_x1 <- as.data.frame(lsoda(xstart_x,times,rm1,params_x)) # tells computer to solve (integrate) equations
-output_x2 <- as.data.frame(lsoda(xstart_x1,times,rm1,params_x1)) # tells computer to solve (integrate) equations
-output_x3 <- as.data.frame(lsoda(xstart_x2,times,rm1,params_x2)) # tells computer to solve (integrate) equations
-output_x4 <- as.data.frame(lsoda(xstart_x3,times,rm1,params_x3)) # tells computer to solve (integrate) equations
-
-combined_eigen <- combined$eigen[param_num]
-combined_eigen_1 <- combined_1$eigen[param_num]
-combined_eigen_2 <- combined_2$eigen[param_num]
-combined_eigen_3 <- combined_3$eigen[param_num]
-
-```
-
-```{r}
-output_x %>% ggplot(.,aes(x=time))+
-  geom_line(aes(y=Sa,col="Susceptible A"))+
-  geom_line(aes(y=Ia,col="Infected A"))+
-  geom_line(aes(y=Ra,col="Recovered A"))+
-  geom_line(aes(y=Sb,col="Susceptible B"))+
-  geom_line(aes(y=Ib,col="Infected B"))+
-  geom_line(aes(y=Rb,col="Recovered B"))+
-  geom_line(aes(y=V,col="Virus"))+
-  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
-  theme(legend.position = c(0.85,0.5))+
-  labs(y="N",x="Time",col="Population")+
-  ylim(0,95)+
-  ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland - Combined - ", combined_eigen))+theme(plot.title = element_text(hjust=0, size=10))
-
-output_x1 %>% ggplot(.,aes(x=time))+
-  geom_line(aes(y=Sa,col="Susceptible A"))+
-  geom_line(aes(y=Ia,col="Infected A"))+
-  geom_line(aes(y=Ra,col="Recovered A"))+
-  geom_line(aes(y=Sb,col="Susceptible B"))+
-  geom_line(aes(y=Ib,col="Infected B"))+
-  geom_line(aes(y=Rb,col="Recovered B"))+
-  geom_line(aes(y=V,col="Virus"))+
-  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
-  theme(legend.position = c(0.85,0.5))+
-  labs(y="N",x="Time",col="Population")+
-  ylim(0,95)+
-  ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland - Combined - ", combined_eigen))+theme(plot.title = element_text(hjust=0, size=10))
-```
-
-Make figure with infected peaks from everything
-```{r}
-output_a %<>% mutate(I = Ia+Ib)
-output_c %<>% mutate(I = Ia+Ib)
-output_h %<>% mutate(I = Ia+Ib)
-output_n %<>% mutate(I = Ia+Ib)
-output_x %<>% mutate(I = Ia+Ib)
-
-output_a1 %<>% mutate(I = Ia+Ib)
-output_c1 %<>% mutate(I = Ia+Ib)
-output_h1 %<>% mutate(I = Ia+Ib)
-output_n1 %<>% mutate(I = Ia+Ib)
-output_x1 %<>% mutate(I = Ia+Ib)
-output_x2 %<>% mutate(I = Ia+Ib)
-output_x3 %<>% mutate(I = Ia+Ib)
-output_x4 %<>% mutate(I = Ia+Ib)
-```
-
-```{r}
-color_x <- paste("Combined\nR0=", combined_eigen)
-color_x2 <- paste("Abun_Half\nR0=", combined_eigen_1)
-color_x3 <- paste("Abun_Comp\nR0=", combined_eigen_2)
-color_x4 <- paste("Comp_Half\nR0=", combined_eigen_3)
-
-color_c <- paste("Composition\nR0=", community_eigen)
-color_a <- paste("Abundance\nR0=", abundance_eigen)
-color_h <- paste("Halflife\nR0=", halflife_eigen)
-color_n <- paste("Reference\nR0=", reference_eigen)
-
-```
-
-```{r}
-ggplot()+
-  geom_line(data=output_x,mapping = aes(y=I,x=time, color=color_x))+
-  scale_color_manual(name="",values = "black")+
-  new_scale_color()+
-  geom_line(data=output_c, mapping=aes(y=I,x=time, color=color_c))+
-  scale_color_manual(name="",values = "red")+
-  new_scale_color()+
-  geom_line(output_a, mapping=aes(y=I,x=time, color=color_a))+
-  scale_color_manual(name="",values = "orange")+
-  new_scale_color()+
-  geom_line(output_h, mapping=aes(y=I,x=time, color=color_h))+
-  scale_color_manual(name="",values = "purple")+
-  new_scale_color()+
-  geom_line(output_n, mapping=aes(y=I,x=time, color=color_n))+
-  scale_color_manual(name="",values = "forestgreen")+
-  labs(y="N",x="Time")+
-  ylim(0,40)+
-  ggtitle(paste("Total Infected Individuals (2 Species) Over Time Across Varying Conditions"))+
-  theme_bw()
-```
-
-```{r}
-ggplot()+
-  geom_line(data=output_x1,mapping = aes(y=I,x=time, color=color_x))+
-  scale_color_manual(name="",values = "black")+
-  new_scale_color()+
-  geom_line(data=output_c1, mapping=aes(y=I,x=time, color=color_c))+
-  scale_color_manual(name="",values = "red")+
-  new_scale_color()+
-  geom_line(output_a1, mapping=aes(y=I,x=time, color=color_a))+
-  scale_color_manual(name="",values = "orange")+
-  new_scale_color()+
-  geom_line(output_h1, mapping=aes(y=I,x=time, color=color_h))+
-  scale_color_manual(name="",values = "purple")+
-  new_scale_color()+
-  geom_line(output_n1, mapping=aes(y=I,x=time, color=color_n))+
-  scale_color_manual(name="",values = "forestgreen")+
-  labs(y="N",x="Time")+
-  ylim(0,40)+
-  ggtitle(paste("Total Infected Individuals (2 Species) Over Time Across Varying Conditions w/o demo"))+
-  theme_bw()
-```
-
-```{r}
-ggplot()+
-  geom_line(data=output_x1,mapping = aes(y=I,x=time, color=color_x))+
-  scale_color_manual(name="",values = "black")+
-  new_scale_color()+
-  geom_line(data=output_x2, mapping=aes(y=I,x=time, color=color_x2))+
-  scale_color_manual(name="",values = "red")+
-  new_scale_color()+
-  geom_line(output_x3, mapping=aes(y=I,x=time, color=color_x3))+
-  scale_color_manual(name="",values = "orange")+
-  new_scale_color()+
-  geom_line(output_x4, mapping=aes(y=I,x=time, color=color_x4))+
-  scale_color_manual(name="",values = "purple")+
-  labs(y="N",x="Time")+
-  ylim(0,40)+
-  ggtitle(paste("Total Infected Individuals (2 Species) Over Time Across Varying Conditions w/o demo"))+
-  theme_bw()
-```

### 2023-06-02 scripts/workshop/sensitivity.Rmd
Update sensitivity.Rmd

@@
 
-```{r, message = F}
+```{r library, message = F}
 library(tidyverse)
@@
 ```
+
+```{r model}
+rm0<-function(t,x,params){
+  Sa <- x[1]  
+  Ia <- x[2]
+  Ra <- x[3]
+  Sb <- x[4]  
+  Ib <- x[5]
+  Rb <- x[6]
+  V <- x[7]
+  with(as.list(params),{
+    dSa <- -beta*Sa*Ia - betab*Sa*Ib - phi*Sa*V + lambda - mu_a*Sa
+    dIa <- beta*Sa*Ia + betab*Sa*Ib + phi*Sa*V - alpha*Ia - mu_a*Ia
+    dRa <- alpha*Ia - mu_a*Ra
+
+    dSb <- -betab*Sb*Ib - beta*Sb*Ia - phi*Sb*V + lambda - mu_b*Sb
+    dIb <- betab*Sb*Ib + beta*Sb*Ia + phi*Sb*V - alpha*Ib - mu_b*Ib
+    dRb <- alpha*Ib - mu_b*Rb
+
+    dV <- sigma*Ia + sigma*Ib - epsilon*V
+    res<-c(dSa,dIa,dRa,dSb,dIb,dRb,dV)
+    list(res)
+  })}
+#no demography
+rm1<-function(t,x,params){
+  Sa <- x[1]  
+  Ia <- x[2]
+  Ra <- x[3]
+  Sb <- x[4]  
+  Ib <- x[5]
+  Rb <- x[6]
+  V <- x[7]
+  with(as.list(params),{
+    dSa <- -beta*Sa*Ia - betab*Sa*Ib - phi*Sa*V
+    dIa <- beta*Sa*Ia + betab*Sa*Ib + phi*Sa*V - alpha*Ia
+    dRa <- alpha*Ia
+
+    dSb <- -betab*Sb*Ib - beta*Sb*Ia - phi*Sb*V
+    dIb <- betab*Sb*Ib + beta*Sb*Ia + phi*Sb*V - alpha*Ib
+    dRb <- alpha*Ib
+
+    dV <- sigma*Ia + sigma*Ib - epsilon*V
+    res<-c(dSa,dIa,dRa,dSb,dIb,dRb,dV)
+    list(res)
+  })}
+
+maxTime <- 300.0 # time is in years - run model for this time
+times<-seq(0,maxTime,by=1) # how long we run the model for
+# notes on params
+# mu =     <- density-independent mortality rate
+# beta =       <- contact transmission rate by infected A
+# betab =       <- contact transmission rate by infected B
+# phi =      <- environmental transmission rate
+# lambda =       <- density-dependent growth rate
+# alpha =     <- recovery rate
+# sigma =    <- shedding rate
+# epsilon =       <- viral degradation rate
+params<-c(mu_a=1/50,
+          mu_b=1/50,
+          beta=0.001,
+          betab=0.0001,
+          phi=0.001,
+          lambda=5/3,
+          alpha=1/10,
+          sigma=1/2,
+          epsilon=1/2)  # model parameters
+
+xstart<-c(Sa=116,
+          Ia=1,
+          Ra=0,
+          Sb=83,
+          Ib=1,
+          Rb=0,
+          V=0)  # initial conditions
+
+output<-as.data.frame(lsoda(xstart,times,rm0,params)) # tells computer to solve (integrate) equations
+```
+
+```{r lhs}
+h <- 1000               #choose number of points
+set.seed(8878896)
+lhs<-maximinLHS(h,6)   #simulate
+```
+
+```{r}
+mu_a_min <- 1/50
+mu_a_max <- 1/50
+mu_b_min <- 1/50
+mu_b_max <- 1/50
+beta_min <- 0.001
+beta_max <- 0.001
+betab_min <- 0.0001
+betab_max <- 0.0001
+phi_min <- 0.001
+phi_max <- 0.001
+lambda_min <- 5/3
+lambda_max <- 5/3
+alpha_min <- 1/10
+alpha_max <- 1/10
+sigma_min <- 1/2
+sigma_max <- 1/2
+epsilon_min <- 1/2
+epsilon_max <- 1/2
+
+params.set <- cbind(
+  mu_a <- lhs[,1]*(mu_a_max - mu_a_min) + mu_a_min,
+  mu_b <- lhs[,2]*(mu_b_max - mu_b_min) + mu_b_min,
+  beta <- lhs[,3]*(beta_max - beta_min) + beta_min,
+  betab <- lhs[,4]*(betab_max - betab_min) + betab_min,
+  phi <- lhs[,5]*(phi_max - phi_min) + phi_min,
+  lambda <- lhs[,6]*(lambda_max - lambda_min) + lambda_min,
+  alpha <- lhs[,7]*(alpha_max - alpha_min) + alpha_min,
+  sigma <- lhs[,8]*(sigma_max - sigma_min) + sigma_min,
+  epsilon <- lhs[,9]*(epsilon_max - epsilon_min) + epsilon_min,
+)
+
+```
+
+```{r}
+
+```

### 2023-06-02 scripts/workshop/sensitivity.Rmd
Update sensitivity.Rmd

@@
 library(deSolve)
+library(lhs)
 ```
@@
 set.seed(8878896)
-lhs<-maximinLHS(h,6)   #simulate
+lhs<-maximinLHS(h,9)   #simulate
 ```
@@
 ```{r}
-mu_a_min <- 1/50
-mu_a_max <- 1/50
-mu_b_min <- 1/50
-mu_b_max <- 1/50
-beta_min <- 0.001
-beta_max <- 0.001
-betab_min <- 0.0001
-betab_max <- 0.0001
-phi_min <- 0.001
-phi_max <- 0.001
-lambda_min <- 5/3
-lambda_max <- 5/3
-alpha_min <- 1/10
-alpha_max <- 1/10
-sigma_min <- 1/2
-sigma_max <- 1/2
-epsilon_min <- 1/2
-epsilon_max <- 1/2
+mu_a_min <- 1/50*0.8
+mu_a_max <- 1/50*1.2
+mu_b_min <- 1/50*0.8
+mu_b_max <- 1/50*1.2
+beta_min <- 0.001*0.8
+beta_max <- 0.001*1.2
+betab_min <- 0.0001*0.8
+betab_max <- 0.0001*1.2
+phi_min <- 0.001*0.8
+phi_max <- 0.001*1.2
+lambda_min <- 5/3*0.8
+lambda_max <- 5/3*1.2
+alpha_min <- 1/10*0.8
+alpha_max <- 1/10*1.2
+sigma_min <- 1/2*0.8
+sigma_max <- 1/2*1.2
+epsilon_min <- 1/2*0.8
+epsilon_max <- 1/2*1.2
 
@@
   sigma <- lhs[,8]*(sigma_max - sigma_min) + sigma_min,
-  epsilon <- lhs[,9]*(epsilon_max - epsilon_min) + epsilon_min,
+  epsilon <- lhs[,9]*(epsilon_max - epsilon_min) + epsilon_min
 )
@@
 ```{r}
+levels <- 11
 
+h2 <- 250
+
+j <- 1
+
+data <- data.frame(matrix(rep(NA,levels*h2*14),nrow=levels*h2))
+for(i in 1:h2){
+
+    data[j,1:9] <- as.list(c(params.set[i,]))
+    params <- params.set[i,]
+    out <- as.data.frame(lsoda(xstart, times, rm0, params))
+    data[j,10] <- tail(out$Ia, 1)
+    data[k,11] <- tail(out$Ib, 1)s
+    j <- j+1
+
+}
 ```

### 2023-06-02 scripts/workshop/sensitivity.Rmd
fix fig 5 alternate

@@
     data[j,10] <- tail(out$Ia, 1)
-    data[k,11] <- tail(out$Ib, 1)s
+    data[k,11] <- tail(out$Ib, 1)
     j <- j+1

### 2023-06-02 scripts/figures/fig5_alternate.R

@@ -45,9 +45,9 @@ order %<>% group_by(WetAltID) %>% mutate(max_prev = max(Prevalence)) %>% ungroup
 
 for(n in 2:nrow(order)){
   if (order$Prevalence[n] > order$Prevalence[n-1]) {
-    order$prev_ratio[n-1] <- (order$Prevalence[n] - order$Prevalence[n-1]) / order$max_prev[n-1]
+    order$prev_ratio[n-1] <- (order$Prevalence[n] - order$Prevalence[n-1]) / (order$max_prev[n-1]-order$Prevalence[n-1])
   } else if (order$Prevalence[n] < order$Prevalence[n-1]) {
-    order$prev_ratio[n-1] <- 1 - ((order$Prevalence[n-1] - order$Prevalence[n]) / order$max_prev[n-1])
+    order$prev_ratio[n-1] <- 1 - ((order$Prevalence[n-1] - order$Prevalence[n]) / order$max_prev[n-1]-order$Prevalence[n])
   } else {
     order$prev_ratio[n-1] <- 0
   }

### 2023-06-02 scripts/workshop/dynamics.Rmd
make faceted plot for each species

@@
 ```{r}
-ggplot()+
+both <- ggplot()+
   geom_line(data=output_x1,mapping = aes(y=I,x=time, color=color_x))+
@@
   ylim(0,40)+
-  ggtitle(paste("Total Infected Individuals (2 Species) Over Time Across Varying Conditions w/o demo"))+
+  ggtitle(paste("Both"))+
   theme_bw()
@@
   ylim(0,40)+
-  ggtitle(paste("Total Infected Individuals (2 Species) Over Time Across Varying Conditions w/o demo"))+
+  xlim(0,250)+
+  ggtitle(paste(""))+
   theme_bw()
@@
 ```
+
+```{r}
+spec_A <- ggplot()+
+  geom_line(data=output_x1,mapping = aes(y=Ia,x=time, color=color_x))+
+  scale_color_manual(name="",values = "black")+
+  new_scale_color()+
+  geom_line(data=output_c1, mapping=aes(y=Ia,x=time, color=color_c))+
+  scale_color_manual(name="",values = "red")+
+  new_scale_color()+
+  geom_line(output_a1, mapping=aes(y=Ia,x=time, color=color_a))+
+  scale_color_manual(name="",values = "orange")+
+  new_scale_color()+
+  geom_line(output_h1, mapping=aes(y=Ia,x=time, color=color_h))+
+  scale_color_manual(name="",values = "purple")+
+  new_scale_color()+
+  geom_line(output_n1, mapping=aes(y=Ia,x=time, color=color_n))+
+  scale_color_manual(name="",values = "forestgreen")+
+  labs(y="N",x="Time")+
+  ylim(0,40)+
+  ggtitle(paste("Species A"))+
+  theme_bw()+
+  theme(legend.position = "none")
+
+```
+
+```{r}
+spec_B <- ggplot()+
+  geom_line(data=output_x1,mapping = aes(y=Ib,x=time, color=color_x))+
+  scale_color_manual(name="",values = "black")+
+  new_scale_color()+
+  geom_line(data=output_c1, mapping=aes(y=Ib,x=time, color=color_c))+
+  scale_color_manual(name="",values = "red")+
+  new_scale_color()+
+  geom_line(output_a1, mapping=aes(y=Ib,x=time, color=color_a))+
+  scale_color_manual(name="",values = "orange")+
+  new_scale_color()+
+  geom_line(output_h1, mapping=aes(y=Ib,x=time, color=color_h))+
+  scale_color_manual(name="",values = "purple")+
+  new_scale_color()+
+  geom_line(output_n1, mapping=aes(y=Ib,x=time, color=color_n))+
+  scale_color_manual(name="",values = "forestgreen")+
+  labs(y="N",x="Time")+
+  ylim(0,40)+
+  ggtitle(paste("Species B"))+
+  theme_bw()+
+  theme(legend.position = "none")
+```
+
+```{r}
+dynamics <- spec_A + spec_B + both
+```

### 2023-06-05 scripts/workshop/dynamics.Rmd
trying out no demo changes

@@
 ```{r}
+params_c1 <- c(mu_a=1/45,
+          mu_b=1/45,
+          beta=spec_1,
+          betab=spec_2,
+          phi=env_rate,
+          lambda=5/3,
+          alpha=1/10,
+          sigma=1/2,
+          epsilon=1/1.947799)  # model parameters
+
+xstart_c1 <- c(Sa=100,
+          Ia=1,
+          Ra=0,
+          Sb=50,
+          Ib=1,
+          Rb=0,
+          V=0)  # initial conditions
+
+output_c2 <- as.data.frame(lsoda(xstart_c1,times,rm1,params_c1)) # tells computer to solve (integrate) equations
+```
+
+```{r}
 output_c %>% ggplot(.,aes(x=time))+
@@
 output_c1 %<>% mutate(I = Ia+Ib)
+output_c2 %<>% mutate(I = Ia+Ib)
 output_h1 %<>% mutate(I = Ia+Ib)
@@
 output_c1 %<>% mutate(X = I + V)
+output_c2 %<>% mutate(X = I + V)
 output_h1 %<>% mutate(X = I + V)

### 2023-06-22 scripts/workshop/dynamics.Rmd
make dynamics figure

@@
   ylim(0,75)+
-  ggtitle(paste("Total Virions Over Time Across Varying Conditions w/o demo"))+
-  theme_bw()
+  ggtitle(paste("Total Virions Over Time Across Varying Conditions w/o demo"))
 ```

### 2023-06-22 scripts/figures/fig1_a.R

@@ -1,0 +1,372 @@
+#Daniel Suh and Andrew Park
+#6/22/23
+
+#Figure 1a in rv_cc manuscript
+
+
+library(here)
+
+source(here("base", "src.R"))
+library(deSolve)
+
+#define function contour(): inputs are parameters and output is values for grid search for R0
+contour <- function(trans_1_min, trans_1_max, trans_3_min, trans_3_max, mort1, mort2, degr) {
+  #expand data
+  data <- expand_grid("trans1" = seq(from = trans_1_min, to = trans_1_max, by = 0.00001),
+                      "trans2" = c(0.0001),
+                      "trans3" = seq(from = trans_3_min, to = trans_3_max, by = 0.00001),
+                      "shed1" = c(rep(1/2, length("trans1"))),
+                      "shed2" = c(rep(1/2, length("trans1"))),
+                      "reco1" = c(rep(1/10, length("trans1"))),
+                      "reco2" = c(rep(1/10, length("trans1"))),
+                      "mort1" = c(mort1), 
+                      "mort2" = c(mort2), 
+                      "degr" = c(degr), 
+                      "birth1" = c(rep(5/3, length("trans1"))),
+                      "birth2" = c(rep(5/3, length("trans1"))))
+  
+  #calculate disease-free equilibria for use in next-gen matrix
+  data %<>% add_column(dfe1 = (data$birth1/data$mort1),dfe2 = (data$birth2/data$mort2), eigen = NA) 
+  #make matrix
+  mat <- matrix(1, nrow = 3, ncol = 3)
+  
+  m <- nrow(data)
+  
+  #calculate eigenvalue using next-gen matrix. max eigenvalue is R0
+  for(n in 1:m){
+    mat[1] = (data$trans1[n]*data$dfe1[n])/(data$reco1[n]+data$mort1[n])
+    mat[2] = (data$trans1[n]*data$dfe2[n])/(data$reco2[n]+data$mort2[n])
+    mat[3] = (data$shed1[n]/data$degr[n])
+    mat[4] = (data$trans2[n]*data$dfe1[n])/(data$reco1[n]+data$mort1[n])
+    mat[5] = (data$trans2[n]*data$dfe2[n])/(data$reco2[n]+data$mort2[n])
+    mat[6] = (data$shed2[n]/data$degr[n])
[…]

### 2023-06-22 scripts/figures/fig1_code.R

@@ -155,3 +155,11 @@ figure_1 <- (p2 | p3)/(p4 | p5)
 figure_1
 ggsave("fig_1.png",plot=figure_1+ proj_theme,width = outwidth[1], height = outwidth[1]/golden,device="png",path=here("figures"))
 
+
+
+
+
+
+
+
+

### 2023-06-22 scripts/figures/supp6_code.R

@@ -58,7 +58,7 @@ y %<>% filter(abund>0) #shouldn't we filter out any species that were absent? (D
 for (k in 1:dim(y)[1]){ #loop through each row
   #get mean pd and min pd for each row
   tmp <- y %>% filter(siteID==y$siteID[k]) #filter for single site-month
-  pd.row <- which(rownames(pd)==y$Species[k]) #select rows from pd matching site-month
+  pd.row <- which(rownames(pd) %in% y$Species[k]) #select rows from pd matching site-month
   #select species that are not the species from that row
   otherSpp <- tmp %>% filter(Species!=y$Species[k]) %>% select(Species) %>% pull() 
   pd.cols <- which(rownames(pd) %in% otherSpp) #select pd for not focal species
@@ -107,7 +107,7 @@ y %>% filter(relAbund>0) %>% ggplot(.,aes(x=min.pdAbund,y=relAbund))+
   geom_point()+geom_smooth(method="loess",span=1.4)
 
 
-y %>% mutate(bin = cut_width(mpdAbund, width = 50)) %>% ggplot(.,aes(x=bin, y=relAbund)) + geom_boxplot() + labs(x = "Distance to Closest Neighbor", y = "Relative Abundance")
+y %>% filter(!is.infinite(min.pd)) %>% mutate(bin = cut_width(min.pd, width = 50)) %>% ggplot(.,aes(x=bin, y=relAbund)) + geom_boxplot() + labs(x = "Distance to Closest Neighbor", y = "Relative Abundance")
 
 
 

### 2023-08-21 base/setup.R
update figs

@@ -21,7 +21,7 @@ lapply(pckg_names,
 
 
 # setting the plot theme for the project
-proj_theme <- theme_set(theme_bw(base_size = 13)) #theme set will set plot theme without having to specify within ggplot
+proj_theme <- theme_set(theme_bw(base_size = 18)) #theme set will set plot theme without having to specify within ggplot
 
 
 ## Intended target dims

### 2023-08-21 scripts/figures/fig5_alternate.R

@@ -45,9 +45,9 @@ order %<>% group_by(WetAltID) %>% mutate(max_prev = max(Prevalence)) %>% ungroup
 
 for(n in 2:nrow(order)){
   if (order$Prevalence[n] > order$Prevalence[n-1]) {
-    order$prev_ratio[n-1] <- (order$Prevalence[n] - order$Prevalence[n-1]) / (order$max_prev[n-1]-order$Prevalence[n-1])
+    order$prev_ratio[n-1] <- (order$Prevalence[n] - order$Prevalence[n-1]) / (order$max_prev[n-1])
   } else if (order$Prevalence[n] < order$Prevalence[n-1]) {
-    order$prev_ratio[n-1] <- 1 - ((order$Prevalence[n-1] - order$Prevalence[n]) / order$max_prev[n-1]-order$Prevalence[n])
+    order$prev_ratio[n-1] <- 1 - ((order$Prevalence[n-1] - order$Prevalence[n]) / order$max_prev[n-1])
   } else {
     order$prev_ratio[n-1] <- 0
   }

### 2023-08-21 scripts/figures/supp6_code.R

@@ -93,8 +93,10 @@ y %>% filter(relAbund>0) %>% ggplot(.,aes(x=mpd,y=relAbund))+
 #above line gives us the relative abundance of each species and the mean pd for their site-month
 #hosts in the same site-month should be stacked vertically
 supp6 <- y %>% filter(!is.na(mpd)) %>% ggplot(.,aes(x=min.pd,y=relAbund))+
-  geom_point()+geom_smooth(method="loess",span=1.4)+
-  theme_classic()+labs(x="Distance to Closest Neighbor", y = "Relative Abundance")
+  geom_smooth(method="loess",span=1.4)+
+  theme_classic()+
+  labs(x="Distance to Closest Neighbor", y = "Relative Abundance")+
+  theme(text = element_text(size = 18))
 supp6
 #supp6 gives us the relative abundance for each species and the pd for their closest phylogenetic neighbor in their site-month
 #so each vertical stack should be an instance where two species co-occurred
@@ -112,4 +114,4 @@ y %>% filter(!is.infinite(min.pd)) %>% mutate(bin = cut_width(min.pd, width = 50
 
 
 
-#ggsave("supp6.png",plot=supp6,device="png",path=here("figures"))
+ggsave("supp6.png",plot=supp6,device="png",path=here("figures"))

### 2023-09-22 scripts/figures/fig5_alternate.R
update equations

@@ -43,25 +43,25 @@ order %<>% group_by(WetAltID) %>% mutate(max_prev = max(Prevalence)) %>% ungroup
 #   }
 # }
 
-for(n in 2:nrow(order)){
-  if (order$Prevalence[n] > order$Prevalence[n-1]) {
-    order$prev_ratio[n-1] <- (order$Prevalence[n] - order$Prevalence[n-1]) / (order$max_prev[n-1])
-  } else if (order$Prevalence[n] < order$Prevalence[n-1]) {
-    order$prev_ratio[n-1] <- 1 - ((order$Prevalence[n-1] - order$Prevalence[n]) / order$max_prev[n-1])
-  } else {
-    order$prev_ratio[n-1] <- 0
-  }
-}
-
 # for(n in 2:nrow(order)){
 #   if (order$Prevalence[n] > order$Prevalence[n-1]) {
-#     order$prev_ratio[n-1] <- (order$Prevalence[n] - order$Prevalence[n-1]) / 1
+#     order$prev_ratio[n-1] <- (order$Prevalence[n] - order$Prevalence[n-1]) / (order$max_prev[n-1])
 #   } else if (order$Prevalence[n] < order$Prevalence[n-1]) {
-#     order$prev_ratio[n-1] <- 1 - ((order$Prevalence[n-1] - order$Prevalence[n]) / 1)
+#     order$prev_ratio[n-1] <- 1 - ((order$Prevalence[n-1] - order$Prevalence[n]) / order$max_prev[n-1])
 #   } else {
 #     order$prev_ratio[n-1] <- 0
 #   }
 # }
+
+for(n in 2:nrow(order)){
+  if (order$Prevalence[n] > order$Prevalence[n-1]) {
+    order$prev_ratio[n-1] <- (order$Prevalence[n] - order$Prevalence[n-1]) / ((order$max_prev[n-1]) - order$Prevalence[n-1])
+  } else if (order$Prevalence[n] < order$Prevalence[n-1]) {
+    order$prev_ratio[n-1] <- 1 - ((order$Prevalence[n-1] - order$Prevalence[n]) / (order$Prevalence[n-1]))
+  } else {
+    order$prev_ratio[n-1] <- 0
+  }
+}
 
 
 

### 2023-11-07 scripts/figures/fig1.R
update eqs

@@ -94,7 +94,7 @@ size_corr <- clean %>% ggplot(.,aes(x=log10(size), y=prev_ratio)) +
         axis.text = element_text(size=axis_text_size)) +
   ylim(0,0.65)
 
-m2 <- cor.test(clean$size,clean$prev_ratio,method="spearman")
+m2 <- cor.test(log10(clean$size),clean$prev_ratio,method="spearman")
 
 
 temp_corr <- clean %>% ggplot(.,aes(x=MeanWaterTempPredC, y=prev_ratio)) +
@@ -110,6 +110,8 @@ m3 <- cor.test(clean$MeanWaterTempPredC,clean$prev_ratio,method="spearman")
 
 #multiple comparisons test
 pvals <- c(m1$p.value,m2$p.value,m3$p.value)
+rho <- c(m1$estimate, m2$estimate, m3$estimate)
+rho
 pvals
 p.adjust(pvals,method="holm")
 

### 2023-11-07 scripts/figures/supp6_code.R

@@ -1,6 +1,6 @@
 #Script Originator: Daniel Suh
 
-#This script creates the inset for figure 3
+#This script creates the inset for figure 4
 #Objective: A plot that models the cophenetic distance of a community and relative abundances of species
 #This provides a look at patterns related to species abundance and phylogenetic relatedness
 #We hypothesize that the species in our data are most abundant when they are moderately related to their neighbor

### 2023-11-07 scripts/figures/fig1.R
rename figs

@@ -1,138 +1,492 @@
 #Daniel Suh and Andrew Park
 
-#Figure 1 in rv_cc manuscript
+#Figure 2 in rv_cc manuscript
+
 
 library(here)
 
-source(here("base","src.R"))
+source(here("base", "src.R"))
+library(deSolve)
 
-
-comm_data <- readRDS(here("processed_data","comm_data.rds"))
-
-axis_title_size = 15
-axis_text_size = 12
-
-prev_cc <- comm_data
-
-order <- prev_cc[order(prev_cc$WetAltID, prev_cc$Month.1),]
-
-#this removes months out of sequence
-order <- order[-c(4,14,47,90),]
-
-order %<>% add_column(prev_ratio = NA)
-
-order %<>% group_by(WetAltID) %>% mutate(max_prev = max(Prevalence)) %>% ungroup()
-
-#create new column that is a ratio between this and last month's prevalence and scales from 0 to 1
-#if next month prev is higher, then range is (0,1)
-#if next month prev is lower, then range is still (0,1)
-# for(n in 2:nrow(order)){
-#   if (order$Prevalence[n] > order$Prevalence[n-1] && order$Prevalence[n-1] > 0) {
-#     order$prev_ratio[n-1] <- (order$Prevalence[n] - order$Prevalence[n-1]) / order$Prevalence[n-1]
-#   } else if (order$Prevalence[n-1] == 0 && order$Prevalence[n] > 0) {
-#     order$prev_ratio[n-1] <- order$Prevalence[n]
-#   } else if (order$Prevalence[n] < order$Prevalence[n-1] && order$Prevalence[n] > 0) {
-#     order$prev_ratio[n-1] <- (order$Prevalence[n-1] - order$Prevalence[n]) / order$Prevalence[n]
-#   } else if (order$Prevalence[n] == 0 && order$Prevalence[n-1] > 0) {
-#     order$prev_ratio[n-1] <- 1-order$Prevalence[n-1]
-#   } else if (order$Prevalence[n-1] > 0 && order$Prevalence[n] == order$Prevalence[n-1]) {
-#     order$prev_ratio[n-1] <- 1
-#   } else {
-#     order$prev_ratio[n-1] <- 0
-#   }
-# }
-
-# for(n in 2:nrow(order)){
-#   if (order$Prevalence[n] > order$Prevalence[n-1]) {
-#     order$prev_ratio[n-1] <- (order$Prevalence[n] - order$Prevalence[n-1]) / (order$max_prev[n-1])
[…]

### 2023-11-07 scripts/figures/fig2.R

@@ -1,492 +1,92 @@
-#Daniel Suh and Andrew Park
+#Script Originator: Daniel Suh
 
-#Figure 2 in rv_cc manuscript
+#This script creates figure 3
 
 
 library(here)
-
-source(here("base", "src.R"))
-library(deSolve)
-
-#define function contour(): inputs are parameters and output is values for grid search for R0
-contour <- function(trans_1_min, trans_1_max, trans_3_min, trans_3_max, mort1, mort2, degr) {
-  #expand data
-  data <- expand_grid("trans1" = seq(from = trans_1_min, to = trans_1_max, by = 0.00001),
-                      "trans2" = c(0.0001),
-                      "trans3" = seq(from = trans_3_min, to = trans_3_max, by = 0.00001),
-                      "shed1" = c(rep(1/2, length("trans1"))),
-                      "shed2" = c(rep(1/2, length("trans1"))),
-                      "reco1" = c(rep(1/10, length("trans1"))),
-                      "reco2" = c(rep(1/10, length("trans1"))),
-                      "mort1" = c(mort1), 
-                      "mort2" = c(mort2), 
-                      "degr" = c(degr), 
-                      "birth1" = c(rep(5/3, length("trans1"))),
-                      "birth2" = c(rep(5/3, length("trans1"))))
-  
-  #calculate disease-free equilibria for use in next-gen matrix
-  data %<>% add_column(dfe1 = (data$birth1/data$mort1),dfe2 = (data$birth2/data$mort2), eigen = NA) 
-  #make matrix
-  mat <- matrix(1, nrow = 3, ncol = 3)
-  
-  m <- nrow(data)
-  
-  #calculate eigenvalue using next-gen matrix. max eigenvalue is R0
-  for(n in 1:m){
-    mat[1] = (data$trans1[n]*data$dfe1[n])/(data$reco1[n]+data$mort1[n])
-    mat[2] = (data$trans1[n]*data$dfe2[n])/(data$reco2[n]+data$mort2[n])
-    mat[3] = (data$shed1[n]/data$degr[n])
-    mat[4] = (data$trans2[n]*data$dfe1[n])/(data$reco1[n]+data$mort1[n])
-    mat[5] = (data$trans2[n]*data$dfe2[n])/(data$reco2[n]+data$mort2[n])
[…]

### 2023-11-07 scripts/figures/fig3.R

@@ -1,92 +1,0 @@
-#Script Originator: Daniel Suh
-
-#This script creates figure 3
-
-
-library(here)
-library("ggtree")
-
-source(here("base","src.R"))
-
-comm_data <- readRDS(here("processed_data","comm_data.rds"))
-vl <- readRDS(here("processed_data","vl.rds"))
-pruned_tree <- readRDS(here("processed_data","tree.rds"))
-
-
-#format data for phylo tree
-vl %<>% filter(tnrs_name!="Dryophytes avivoca") # as doesn't appear sampled in any communities? (AP)
-#H. avivoca (ID=32) has no abundance measured but has one recording for viral load
-#viral load for H. avivoca is actually the highest but should not be included because there is no abundance data
-
-vl %<>% filter(m>3)
-
-pruned_tree$tip.label <- gsub("_"," ",pruned_tree$tip.label)
-
-
-vl %<>% select(species_name,mean,log1p_mean) %>% mutate(.,label = species_name) %>% select(-species_name)
-vl$label <- gsub("_", " ", vl$label)
-vl$label <- gsub("Anaxyrus terrestris","Bufo terrestris",vl$label)
-vl$label <- gsub("Lithobates","Rana",vl$label)
-vl$label <- gsub("catesbeianus","catesbeiana",vl$label)
-vl$label <- gsub("sphenocephalus","sphenocephala",vl$label)
-vl$label <- gsub("clamitans","clamitans",vl$label)
-
-
-vl %<>% mutate(labelAlt=label)
-top4sampled <- c("Pseudacris crucifer","Rana sphenocephala","Bufo terrestris","Notophthalmus viridescens")
-vl %<>% mutate(labelAlt=as_factor(if_else(labelAlt %in% top4sampled,labelAlt,"Low competence")))
-vl %<>% mutate(labelAlt=fct_relevel(labelAlt,c(top4sampled,"Low competence")))
-
-vl %<>% relocate(label) #make label first column
-vl_tree <- pruned_tree %>% ggtree(.,) %<+% vl #integrate vl data into tree and make new tree object
-
-
-#format data for bar plots
-comm_data$siteID <- reorder(comm_data$siteID, -comm_data$cc)
-z <- comm_data
-z %<>% select(siteID,ABLow,AB26,AB42,AB21,AB9,J,size,cc)
-z %<>% mutate(Low=ABLow)
[…]

### 2023-11-07 scripts/figures/supp1.R

@@ -1,0 +1,138 @@
+#Daniel Suh and Andrew Park
+
+#Figure 1 in rv_cc manuscript
+
+library(here)
+
+source(here("base","src.R"))
+
+
+comm_data <- readRDS(here("processed_data","comm_data.rds"))
+
+axis_title_size = 15
+axis_text_size = 12
+
+prev_cc <- comm_data
+
+order <- prev_cc[order(prev_cc$WetAltID, prev_cc$Month.1),]
+
+#this removes months out of sequence
+order <- order[-c(4,14,47,90),]
+
+order %<>% add_column(prev_ratio = NA)
+
+order %<>% group_by(WetAltID) %>% mutate(max_prev = max(Prevalence)) %>% ungroup()
+
+#create new column that is a ratio between this and last month's prevalence and scales from 0 to 1
+#if next month prev is higher, then range is (0,1)
+#if next month prev is lower, then range is still (0,1)
+# for(n in 2:nrow(order)){
+#   if (order$Prevalence[n] > order$Prevalence[n-1] && order$Prevalence[n-1] > 0) {
+#     order$prev_ratio[n-1] <- (order$Prevalence[n] - order$Prevalence[n-1]) / order$Prevalence[n-1]
+#   } else if (order$Prevalence[n-1] == 0 && order$Prevalence[n] > 0) {
+#     order$prev_ratio[n-1] <- order$Prevalence[n]
+#   } else if (order$Prevalence[n] < order$Prevalence[n-1] && order$Prevalence[n] > 0) {
+#     order$prev_ratio[n-1] <- (order$Prevalence[n-1] - order$Prevalence[n]) / order$Prevalence[n]
+#   } else if (order$Prevalence[n] == 0 && order$Prevalence[n-1] > 0) {
+#     order$prev_ratio[n-1] <- 1-order$Prevalence[n-1]
+#   } else if (order$Prevalence[n-1] > 0 && order$Prevalence[n] == order$Prevalence[n-1]) {
+#     order$prev_ratio[n-1] <- 1
+#   } else {
+#     order$prev_ratio[n-1] <- 0
+#   }
+# }
+
+# for(n in 2:nrow(order)){
+#   if (order$Prevalence[n] > order$Prevalence[n-1]) {
+#     order$prev_ratio[n-1] <- (order$Prevalence[n] - order$Prevalence[n-1]) / (order$max_prev[n-1])
[…]

### 2023-11-07 scripts/figures/fig2.R
update fig2

@@ -85,8 +85,8 @@ newTree <- vl_tree +
   guides(color=F)+xlim(0,350)
 
 
-figure_3 <- cc_plot/RA_plot/newTree
-figure_3
+figure_2 <- cc_plot/RA_plot/newTree
+figure_2
 
-#ggsave("fig3.png",plot=figure_3,width = outwidth[1], height = outwidth[1]/golden,device="png",path=here("figures"))
+ggsave("fig2.png",plot=figure_2,width = outwidth[1], height = outwidth[1]/golden,device="png",path=here("figures"))
 

### 2023-11-07 scripts/figures/fig2_inset.R

@@ -92,12 +92,12 @@ y %>% filter(relAbund>0) %>% ggplot(.,aes(x=mpd,y=relAbund))+
   geom_point()+geom_smooth(method="loess",span=1.4)
 #above line gives us the relative abundance of each species and the mean pd for their site-month
 #hosts in the same site-month should be stacked vertically
-fig3_inset <- y %>% filter(!is.na(mpd)) %>% ggplot(.,aes(x=min.pd,y=relAbund))+
+fig2_inset <- y %>% filter(!is.na(mpd)) %>% ggplot(.,aes(x=min.pd,y=relAbund))+
   geom_smooth(method="loess",span=1.4)+
   theme_classic()+
   labs(x="Phylogenetic Distance to Closest Neighbor", y = "Relative Abundance")+
   theme(text = element_text(size = 18))
-fig3_inset
+fig2_inset
 #supp6 gives us the relative abundance for each species and the pd for their closest phylogenetic neighbor in their site-month
 #so each vertical stack should be an instance where two species co-occurred
 #the species included are limited to those that were sampled for ranavirus
@@ -114,4 +114,4 @@ y %>% filter(!is.infinite(min.pd)) %>% mutate(bin = cut_width(min.pd, width = 50
 
 
 
-#ggsave("fig3_inset.png",plot=fig3_inset,device="png",path=here("figures"))
+ggsave("fig2_inset.png",plot=fig2_inset,device="png",path=here("figures"))

### 2023-11-17 case_studies.Rmd
finding case study examples

---
title: "case_studies"
output: html_document
date: "`r Sys.Date()`"
---

Let's take a look at other viruses from GMPD and Verena databases that fit the pattern we are interested in — infects multiple host species and has both contact and environmental transmission.

```{r}
library(here)
library(tidyverse)
library(magrittr)
library(rvest)
```

```{r}
gmpd <- read_csv(here("data/GMPD/GMPD_main.csv"))
gmpd_traits <- read_csv(here("data/GMPD/GMPD_parasite_traits.csv"))
clover <- read_csv(here("data/clover/CLOVER_1.0_Viruses_AssociationsFlatFile.csv"))
```

```{r}
gmpd_num_hosts <- gmpd %>% filter(ParType=="Virus") %>% dplyr::select(HostCorrectedName, ParasiteCorrectedName) %>% distinct() %>% group_by(ParasiteCorrectedName) %>% summarize(num_hosts = n())
gmpd_generalists <- gmpd_num_hosts %>% filter(num_hosts>1)

gmpd_close_nonclose <- gmpd_traits %>% filter(close+nonclose==2)

gmpd_select <- inner_join(gmpd_close_nonclose, gmpd_generalists, by = "ParasiteCorrectedName")
```

clover doesn't include data on transmission mode so will need to get those data from somewhere else if we want to use that

for now, we can take a look at the parasites that are positive from GMPD

Stealing this google scholar scraper from another project
```{r}
names <- tibble(species = gmpd_select$ParasiteCorrectedName)

names$results <- NA

#names$species <- tolower(names$species) 
names$species <- str_replace_all(string = names$species, pattern = " ", replacement = "+")

for (i in 1:nrow(names)){
  tmp <- pull(names[i,1])

  link <- str_replace(string="https://scholar.google.com/scholar?hl=en&as_sdt=0%2C11&q=REPLACEME&btnG=", pattern="REPLACEME", replacement=tmp)

  page <- read_html(link)

  summaries_css <- page %>% html_elements(css = ".gs_ab_mdw")

  text <- capture.output(print(summaries_css[2]))

  line <- stringr::str_extract(text, "About.+results")

  line_a <- str_replace(line, ",", "")

  line_b <- line_a[2]

  number <- as.numeric(str_extract(line_b, "[:digit:]+"))

  names$results[i] <- number
}

```

### 2023-11-20 scripts/workshop/case_studies.Rmd
move file
update case studies

--- 2023-11-17 case_studies.Rmd
@@
 
-Stealing this google scholar scraper from another project
+Stealing this google scholar scraper from another project (doesn't always work)
 ```{r}
@@
 
+write_csv(names, here("data/scholar_results.csv"))
 ```

### 2024-04-19 scripts/figures/fig1.R
update fig 1 axis labels and update eqs and prev ratio

@@ -152,7 +152,8 @@ p4 <- halflife %>% ggplot(.,aes(x=prop3,y=prop1))+
   geom_point(aes(x=env_prop,y=cont_prop))+
   scale_colour_manual(na.value = NA, values = c(NA, "gray")) +
   labs(x="", 
-       y = "                                      Relative Contact Transmission Rate", title = "Half-life") +
+       y = expression(paste("                               Relative Contact\n                            Transmission Rate ", "(",
+                            frac(Beta[B], Beta[A]),")")), title = "Half-life") +
   xlim(2,7) +
   ylim(2,10) +
   theme_classic() +
@@ -174,7 +175,7 @@ p5 <- combined %>%
                linewidth=line_width) +
   geom_point(aes(x=env_prop,y=cont_prop))+
   scale_colour_manual(na.value = NA, values = c(NA, "gray")) +
-  labs(x = "            Relative Environmental\n               Transmission Rate", y = "", title = "Combined") +
+  labs(x = expression(paste("         Relative Environmental\n            Transmission Rate", "(", frac(Phi, Beta[A]), ")")), y = "", title = "Combined") +
   xlim(2,7) +
   ylim(2,10) +
   theme_classic() +
@@ -480,7 +481,7 @@ spec_B <- ggplot()+
 
 figure_1 <- p2 / p3 / p4 / p5 | both
 figure_1
-#ggsave("fig1.png",plot=figure_1,width = outwidth[1], height = outwidth[1]/golden,device="png",path=here("figures"))
+ggsave("fig1.png",plot=figure_1,width = outwidth[1], height = outwidth[1]/golden,device="png",path=here("figures"))
 
 
 

### 2024-04-19 scripts/figures/supp2.R

@@ -7,6 +7,13 @@ library(here)
 source(here("base","src.R"))
 
 
+#Notes
+
+# 1. need to figure out how to deal with months where prevalence is decreasing before the peak. There don't seem to be any where the prevalence is increasing after the peak
+
+# 2. Also not sure how to deal with months with just very little data
+
+
 comm_data <- readRDS(here("processed_data","comm_data.rds"))
 
 axis_title_size = 15
@@ -23,34 +30,8 @@ order %<>% add_column(prev_ratio = NA)
 
 order %<>% group_by(WetAltID) %>% mutate(max_prev = max(Prevalence)) %>% ungroup()
 
-#create new column that is a ratio between this and last month's prevalence and scales from 0 to 1
-#if next month prev is higher, then range is (0,1)
-#if next month prev is lower, then range is still (0,1)
-# for(n in 2:nrow(order)){
-#   if (order$Prevalence[n] > order$Prevalence[n-1] && order$Prevalence[n-1] > 0) {
-#     order$prev_ratio[n-1] <- (order$Prevalence[n] - order$Prevalence[n-1]) / order$Prevalence[n-1]
-#   } else if (order$Prevalence[n-1] == 0 && order$Prevalence[n] > 0) {
-#     order$prev_ratio[n-1] <- order$Prevalence[n]
-#   } else if (order$Prevalence[n] < order$Prevalence[n-1] && order$Prevalence[n] > 0) {
-#     order$prev_ratio[n-1] <- (order$Prevalence[n-1] - order$Prevalence[n]) / order$Prevalence[n]
-#   } else if (order$Prevalence[n] == 0 && order$Prevalence[n-1] > 0) {
-#     order$prev_ratio[n-1] <- 1-order$Prevalence[n-1]
-#   } else if (order$Prevalence[n-1] > 0 && order$Prevalence[n] == order$Prevalence[n-1]) {
-#     order$prev_ratio[n-1] <- 1
-#   } else {
-#     order$prev_ratio[n-1] <- 0
-#   }
-# }
+#order %<>% filter(!WetAltID %in% c(8,11,20))
 
-# for(n in 2:nrow(order)){
-#   if (order$Prevalence[n] > order$Prevalence[n-1]) {
[…]
